---
title: "betaVR_WAAL_widowing"
author: "Sarah Becker"
date: '2024-01-06'
output: html_document
---


```{r, echo= FALSE, include =FALSE}
library(here)
library(tidyverse)
library(RColorBrewer)

#set up various plotting themes
themeo <-theme_classic()+
  theme(strip.background = element_blank(),
        panel.grid.major = element_line(colour = "transparent"), 
        panel.grid.minor = element_line(colour = "transparent"),
        axis.line = element_blank(),
        axis.title = element_text(color = "black", size = 10),
        axis.ticks.length=unit(-0.1, "cm"),
        axis.ticks = element_line(color = "black", size = .25),
        axis.ticks.x = element_line(color = "black"),
        axis.ticks.y = element_line(color = "black"),
        panel.border = element_rect(colour = "black", fill=NA, size=.5),
        panel.spacing = unit(1, "lines"),
        legend.position = "bottom",
        legend.title = element_text(colour = "black", size = 10, face = "bold"),
        legend.text = element_text(color = "black", size =8)) #default legend position
```

**set vital rates. Try first without widowing**
```{r}
#survival rates - drawn from Weimerskirch 1998
s_c <- 0.64 #chick+ fledgling survival year 0-1
s_cc <- 0.7244736 #conditional chick survival (disregarding parental survival) *calculated from other VRs, not directly from literature
s_j <- 0.85 #juvenile survival year 1-3
s_i <- 0.92 #immature survival year 3-7
s_s <- 0.94 #subadult survival year 7-10
s_a <- 0.94 #adult survival year 10+
#males show senescence but femailes don't, so not including it here, though perhaps I should since it impacts breeding success?

#fecundity building blocks
#breeding success - drawn from Weimerskirch 1998 &Rackete 2021
bs_a <- 0.76 #adult breeding success

#breeding probability (biennial breeders) - adult rate drawn from Rackete 2021, subadult downscaled slightly from that
bp_a <- 0.45 #adult breeding probablity (biennial breeding, most breeding by age 10)

c <- 0.5# # of female chicks per year

#fecundity 
f_a <- bs_a*bp_a*c
  
#add widowing
#assume 25% prob year1 -> 100% prob year 4?
#Sun et al 2022 - takes 4 years to reach next breeding attempt, breeding prob with new mate after widowhood = ~0.25 (fig5)
w <- 0.25
w4 <- 1
```

**create matrix**
```{r}
#set up age-based matrix based on vital rates

#Age classes (years)  0-1        1-2    2-3    3-4    4-5    5-6    6-7    7-8   8-9     9-10     10+         nw,        w1,             w2,              w3,             w4 
m <- matrix(data = c(0,          0,    0,      0,     0,     0,     0,     0,      0,      0,       s_a*f_a,    s_a*f_a,    s_a*f_a*w,    s_a*f_a*w,     s_a*f_a*w,     s_a*f_a*w4,        #0-1
                    s_cc*(s_a^2),0,    0,      0,     0,     0,     0,     0,      0,      0,       0,          0,          0,             0,              0,              0,              #1-2
                    0,           s_j,  0,      0,     0,     0,     0,     0,      0,      0,       0,          0,          0,             0,              0,              0,              #2-3
                    0,           0,    s_j,    0,     0,     0,     0,     0,      0,      0,       0,          0,          0,             0,              0,              0,              #3-4
                    0,           0,    0,      s_i,   0,     0,     0,     0,      0,      0,       0,          0,          0,             0,              0,              0,              #4-5
                    0,           0,    0,      0,     s_i,   0,     0,     0,      0,      0,       0,          0,          0,             0,              0,              0,              #5-6
                    0,           0,    0,      0,     0,     s_i,   0,     0,      0,      0,       0,          0,          0,             0,              0,              0,              #6-7
                    0,           0,    0,      0,     0,     0,     s_i,   0,      0,      0,       0,          0,          0,             0,              0,              0,              #7-8
                    0,           0,    0,      0,     0,     0,     0,     s_s,    0,      0,       0,          0,          0,             0,              0,              0,              #8-9
                    0,           0,    0,      0,     0,     0,     0,     0,      s_s,    0,       0,          0,          0,             0,              0,              0,              #9-10
                    0,           0,    0,      0,     0,     0,     0,     0,      0,      s_s,     0,          0,          0,             0,              0,              0,              #10
                    0,           0,    0,      0,     0,     0,     0,     0,      0,      0,       s_a*s_a,    s_a*s_a,    s_a*s_a*w,     s_a*s_a*w,      s_a*s_a*w,      s_a*s_a*w4,     #nw
                    0,           0,    0,      0,     0,     0,     0,     0,      0,      0,       s_a*(1-s_a),s_a*(1-s_a),s_a*(1-s_a)*w, s_a*(1-s_a)*w,  s_a*(1-s_a)*w,  s_a*(1-s_a)*w4, #w1
                    0,           0,    0,      0,     0,     0,     0,     0,      0,      0,       0,          0,          s_a*(1-w),     0,              0,              0,              #w2 
                    0,           0,    0,      0,     0,     0,     0,     0,      0,      0,       0,          0,          0,             s_a*(1-w),      0,              0,              #w3 
                    0,           0,    0,      0,     0,     0,     0,     0,      0,      0,       0,          0,          0,             0,              s_a*(1-w),      0),             #w4 
                   nrow = 16, byrow = TRUE) 
m
```

**extract lambda, SSDs, and RVs from eigenvalues**
```{r}
#deterministic growth rate (lambda)
#extract eigenvalues and lambda from the matrix 
m_eig <- eigen(m)
lambda <- as.numeric(m_eig$values[1])
lambda # this matches the ideal population growth rate in the table

#Stable state distribution
#extract and scale the right eigenvector 
SSD <- round(as.numeric(m_eig$vectors[,1]/sum(m_eig$vectors[,1])),3)
SSD

#extract reproductive values
RV <- as.numeric(eigen(t(m))$vectors[,1]/eigen(t(m))$vectors[1,1])
RV
```

**identify mortality rate, mitigation efficacy**
```{r}
#bycatch mortality rates
# ~25 - 35% of mortality has been attributed to bycatch (Cleeland et al 2021), 
# so I calculated the baseline mortality and divided by 4 (testing 33% now - /3)
m_j <- (1-s_j)/3 
m_i <- (1-s_i)/3
m_s <- (1-s_s)/3
m_a <- (1-s_a)/3

mitigation_eff <- 0.8
```

**loop through different bycatch reductions scenarios, find lambda for each**
```{r}
bycatch_reductions <- data.frame(reduction = seq(0, 1, by =0.01), lambda = NA)

for(i in 1:nrow(bycatch_reductions)){
  #New_surv = 1-((baseline_bycatch_mort*(1-mit_eff*mit_amount))+baseline_non_bycatch_mort)
  #new_surv = 1-(modified bycatch mortality + baseline non-bycatch mortality)
  n_s_j <- 1-((m_j*(1-(mitigation_eff*bycatch_reductions$reduction[i])))+(m_j*2))
  n_s_i <- 1-((m_i*(1-(mitigation_eff*bycatch_reductions$reduction[i])))+(m_i*2))
  n_s_s <- 1-((m_s*(1-(mitigation_eff*bycatch_reductions$reduction[i])))+(m_s*2))
  n_s_a <- 1-((m_a*(1-(mitigation_eff*bycatch_reductions$reduction[i])))+(m_s*2))
  
  #Age classes (years)  0-1           1-2    2-3     3-4    4-5    5-6    6-7    7-8    8-9    9-10   10+              nw,              w1,                w2,                 w3,                 w4 
  n_m <- matrix(data = c(0,           0,     0,      0,     0,     0,     0,     0,     0,     0,     n_s_a*f_a,       n_s_a*f_a,       n_s_a*f_a*w,       n_s_a*f_a*w,        n_s_a*f_a*w,        n_s_a*f_a*1,        #0-1
                       s_cc*(n_s_a^2),0,     0,      0,     0,     0,     0,     0,     0,     0,     0,               0,               0,                 0,                  0,                  0,                  #1-2
                       0,             n_s_j, 0,      0,     0,     0,     0,     0,     0,     0,     0,               0,               0,                 0,                  0,                  0,                  #2-3
                       0,             0,     n_s_j,  0,     0,     0,     0,     0,     0,     0,     0,               0,               0,                 0,                  0,                  0,                  #3-4
                       0,             0,     0,      n_s_i, 0,     0,     0,     0,     0,     0,     0,               0,               0,                 0,                  0,                  0,                  #4-5
                       0,             0,     0,      0,     n_s_i, 0,     0,     0,     0,     0,     0,               0,               0,                 0,                  0,                  0,                  #5-6
                       0,             0,     0,      0,     0,     n_s_i, 0,     0,     0,     0,     0,               0,               0,                 0,                  0,                  0,                  #6-7
                       0,             0,     0,      0,     0,     0,     n_s_i, 0,     0,     0,     0,               0,               0,                 0,                  0,                  0,                  #7-8
                       0,             0,     0,      0,     0,     0,     0,     n_s_s, 0,     0,     0,               0,               0,                 0,                  0,                  0,                  #8-9
                       0,             0,     0,      0,     0,     0,     0,     0,     n_s_s, 0,     0,               0,               0,                 0,                  0,                  0,                  #9-10
                       0,             0,     0,      0,     0,     0,     0,     0,     0,     n_s_s, 0,               0,               0,                 0,                  0,                  0,                  #10
                       0,             0,     0,      0,     0,     0,     0,     0,     0,     0,     n_s_a*n_s_a,     n_s_a*n_s_a,     n_s_a*n_s_a*w,     n_s_a*n_s_a*w,      n_s_a*n_s_a*w,      n_s_a*n_s_a*w4,     #nw
                       0,             0,     0,      0,     0,     0,     0,     0,     0,     0,     n_s_a*(1-n_s_a), n_s_a*(1-n_s_a), n_s_a*(1-n_s_a)*w, n_s_a*(1-n_s_a)*w,  n_s_a*(1-n_s_a)*w,  n_s_a*(1-n_s_a)*w4, #w1
                       0,             0,     0,      0,     0,     0,     0,     0,     0,     0,     0,               0,               n_s_a*(1-w),       0,                  0,                  0,                  #w2 
                       0,             0,     0,      0,     0,     0,     0,     0,     0,     0,     0,               0,               0,                 n_s_a*(1-w),        0,                  0,                  #w3 
                       0,             0,     0,      0,     0,     0,     0,     0,     0,     0,     0,               0,               0,                 0,                  n_s_a*(1-w),        0),                 #w4 
                      nrow = 16, byrow = TRUE) 
  
  #extract eigenvalues and lambda from the matrix 
  n_m_eig <- eigen(n_m)
  lambda_n <- as.numeric(n_m_eig$values[1])
  bycatch_reductions$lambda[i] <- lambda_n # this matches the ideal population growth rate in the table

}
```

**plot bycatch reduction against changes to lambda**
```{r}
ggplot(bycatch_reductions)+
  geom_line(aes(x=reduction, y=as.numeric(lambda)))+
  geom_hline(aes(yintercept = 1), color = "red", alpha =0.5)+ 
  # geom_vline(aes(xintercept = 0.66), color = "red", alpha =0.5)+ #illegal vs legal fishing transition
  # geom_vline(aes(xintercept = 0.33), color = "red", alpha =0.5)+ #illegal vs legal fishing transition
  labs(x="% fisheries with bycatch mitigation", y="lambda")+
  themeo+
  coord_cartesian(expand=FALSE)
```

**add uncertainty - vary mitigation efficacy**
```{r}
bycatch_reductions2 <- seq(0, 1, by =0.01)
mitigation_eff_range <- seq(0.5,1, by  = 0.1)
output <- matrix(nrow = length(bycatch_reductions2), ncol = length(mitigation_eff_range)) %>% as.data.frame() 

for(k in 1:length(mitigation_eff_range)){
  mit_eff <- mitigation_eff_range[k]
  
  for(i in 1:length(bycatch_reductions2)){
  
  n_s_j <- 1-((m_j*(1-(mit_eff*bycatch_reductions2[i])))+(m_j*2))
  n_s_i <- 1-((m_i*(1-(mit_eff*bycatch_reductions2[i])))+(m_i*2))
  n_s_s <- 1-((m_s*(1-(mit_eff*bycatch_reductions2[i])))+(m_s*2))
  n_s_a <- 1-((m_a*(1-(mit_eff*bycatch_reductions2[i])))+(m_s*2))
  
  #Age classes (years)  0-1           1-2    2-3     3-4    4-5    5-6    6-7    7-8    8-9    9-10   10+              nw,              w1,                w2,                 w3,                 w4 
  n_m <- matrix(data = c(0,           0,     0,      0,     0,     0,     0,     0,     0,     0,     n_s_a*f_a,       n_s_a*f_a,       n_s_a*f_a*w,       n_s_a*f_a*w,        n_s_a*f_a*w,        n_s_a*f_a*w4,       #0-1
                       s_cc*(n_s_a^2),0,     0,      0,     0,     0,     0,     0,     0,     0,     0,               0,               0,                 0,                  0,                  0,                  #1-2
                       0,             n_s_j, 0,      0,     0,     0,     0,     0,     0,     0,     0,               0,               0,                 0,                  0,                  0,                  #2-3
                       0,             0,     n_s_j,  0,     0,     0,     0,     0,     0,     0,     0,               0,               0,                 0,                  0,                  0,                  #3-4
                       0,             0,     0,      n_s_i, 0,     0,     0,     0,     0,     0,     0,               0,               0,                 0,                  0,                  0,                  #4-5
                       0,             0,     0,      0,     n_s_i, 0,     0,     0,     0,     0,     0,               0,               0,                 0,                  0,                  0,                  #5-6
                       0,             0,     0,      0,     0,     n_s_i, 0,     0,     0,     0,     0,               0,               0,                 0,                  0,                  0,                  #6-7
                       0,             0,     0,      0,     0,     0,     n_s_i, 0,     0,     0,     0,               0,               0,                 0,                  0,                  0,                  #7-8
                       0,             0,     0,      0,     0,     0,     0,     n_s_s, 0,     0,     0,               0,               0,                 0,                  0,                  0,                  #8-9
                       0,             0,     0,      0,     0,     0,     0,     0,     n_s_s, 0,     0,               0,               0,                 0,                  0,                  0,                  #9-10
                       0,             0,     0,      0,     0,     0,     0,     0,     0,     n_s_s, 0,               0,               0,                 0,                  0,                  0,                  #10
                       0,             0,     0,      0,     0,     0,     0,     0,     0,     0,     n_s_a*n_s_a,     n_s_a*n_s_a,     n_s_a*n_s_a*w,     n_s_a*n_s_a*w,      n_s_a*n_s_a*w,      n_s_a*n_s_a*w4,     #nw
                       0,             0,     0,      0,     0,     0,     0,     0,     0,     0,     n_s_a*(1-n_s_a), n_s_a*(1-n_s_a), n_s_a*(1-n_s_a)*w, n_s_a*(1-n_s_a)*w,  n_s_a*(1-n_s_a)*w,  n_s_a*(1-n_s_a)*w4, #w1
                       0,             0,     0,      0,     0,     0,     0,     0,     0,     0,     0,               0,               n_s_a*(1-w),       0,                  0,                  0,                  #w2 
                       0,             0,     0,      0,     0,     0,     0,     0,     0,     0,     0,               0,               0,                 n_s_a*(1-w),        0,                  0,                  #w3 
                       0,             0,     0,      0,     0,     0,     0,     0,     0,     0,     0,               0,               0,                 0,                  n_s_a*(1-w),        0),                 #w4 
                      nrow = 16, byrow = TRUE) 
  
  #extract eigenvalues and lambda from the matrix 
  n_m_eig <- eigen(n_m)
  lambda_n <- as.numeric(n_m_eig$values[1])
  output[i,k] <- lambda_n # this matches the ideal population growth rate in the table

}
}

```

*format output*
```{r}
oldnames = c("V1","V2","V3","V4" ,"V5", "V6")
newnames = c("0.5", "0.6", "0.7", "0.8", "0.9", "1")

output<- output %>% 
  rename_with(~ newnames[which(oldnames == .x)], .cols = oldnames)

output$bycatch_reductions <- bycatch_reductions2

output_l <- output %>% 
  pivot_longer(cols = !bycatch_reductions,names_to = "mit_eff", values_to = "lambda")

output_l$mit_eff <- as.numeric(output_l$mit_eff)
```

```{r}
ggplot(output_l)+
  geom_line(aes(x=bycatch_reductions, y=lambda, group = factor(mit_eff), color = factor(mit_eff)))+
  geom_hline(aes(yintercept = 1), color = "darkgrey", alpha =0.5)+ 
  # geom_vline(aes(xintercept = 0.66), color = "red", alpha =0.5)+ #illegal vs legal fishing transition
  # geom_vline(aes(xintercept = 0.33), color = "red", alpha =0.5)+ #illegal vs legal fishing transition
  labs(x="% fisheries with bycatch mitigation", y="lambda", color = "% mitigation efficacy")+
  themeo+
  guides(colour = guide_legend(nrow = 1))+
  coord_cartesian(expand=FALSE)+
  scale_color_manual(values=c("#c7e9b4", "#7fcdbb", "#41b6c4", "#1d91c0", "#225ea8", "#253494"))
```

**add vital rate uncertainty (simulate 100 runs with varying range of vital rates (randomly draw with replacement from values found in the literature))** 
**mortality = 1/3 bycatch, 2/3 other causes**

**first, set VR uncertainty from beta dist**
```{r}
#set mean, CI for each vital rate, use these to pull VR rates within beta distribution
#s_c2
s_c2mean <-0.64
s_c2upper_CI <- 0.65
s_c2lower_CI <- 0.63
  
s_c2SE <- (s_c2upper_CI - s_c2lower_CI) / 3.92

s_c2var <- s_c2SE^2

s_c2alpha = s_c2mean * ((1 - s_c2mean) / s_c2var - 1 / s_c2mean)
s_c2beta = (1 - s_c2mean) * ((1 -s_c2mean) /s_c2var - 1 / (1 - s_c2mean))

#s_cc2
s_cc2mean <-0.7244736
s_cc2upper_CI <- 0.73
s_cc2lower_CI <- 0.72
  
s_cc2SE <- (s_cc2upper_CI - s_cc2lower_CI) / 3.92

s_cc2var <- s_cc2SE^2

s_cc2alpha = s_cc2mean * ((1 - s_cc2mean) / s_cc2var - 1 / s_cc2mean)
s_cc2beta = (1 - s_cc2mean) * ((1 -s_cc2mean) /s_cc2var - 1 / (1 - s_cc2mean))

#s_j2
#s_j2mean <-0.825
s_j2mean <-0.85
s_j2upper_CI <- 0.85
s_j2lower_CI <- 0.8
  
s_j2SE <- (s_j2upper_CI - s_j2lower_CI) / 3.92

s_j2var <- s_j2SE^2

s_j2alpha = s_j2mean * ((1 - s_j2mean) / s_j2var - 1 / s_j2mean)
s_j2beta = (1 - s_j2mean) * ((1 -s_j2mean) /s_j2var - 1 / (1 - s_j2mean))

# s_i2
 #s_i2mean <-0.925
 s_i2mean <-0.92
 s_i2upper_CI <- 0.94
 s_i2lower_CI <- 0.91
  
 s_i2SE <- ( s_i2upper_CI -  s_i2lower_CI) / 3.92

 s_i2var <-  s_i2SE^2

 s_i2alpha =  s_i2mean * ((1 -  s_i2mean) /  s_i2var - 1 /  s_i2mean)
 s_i2beta = (1 -  s_i2mean) * ((1 - s_i2mean) / s_i2var - 1 / (1 -  s_i2mean))
 
# s_s2
 #s_s2mean <-0.944
 s_s2mean <-0.94
 s_s2upper_CI <- 0.95
 s_s2lower_CI <- 0.94
  
 s_s2SE <- ( s_s2upper_CI -  s_s2lower_CI) / 3.92

 s_s2var <-  s_s2SE^2

 s_s2alpha =  s_s2mean * ((1 -  s_s2mean) /  s_s2var - 1 /  s_s2mean)
 s_s2beta = (1 -  s_s2mean) * ((1 - s_s2mean) / s_s2var - 1 / (1 -  s_s2mean))
 
 # s_a2
 #s_a2mean <-0.945
  s_a2mean <-0.94
 s_a2upper_CI <- 0.95
 s_a2lower_CI <- 0.94
  
 s_a2SE <- ( s_a2upper_CI -  s_a2lower_CI) / 3.92

 s_a2var <-  s_a2SE^2

 s_a2alpha =  s_a2mean * ((1 -  s_a2mean) /  s_a2var - 1 /  s_a2mean)
 s_a2beta = (1 -  s_a2mean) * ((1 - s_a2mean) / s_a2var - 1 / (1 -  s_a2mean))

 
  # bs_a2
 bs_a2mean <-0.76
 bs_a2upper_CI <- 0.8
 bs_a2lower_CI <- 0.68
  
 bs_a2SE <- ( bs_a2upper_CI -  bs_a2lower_CI) / 3.92

 bs_a2var <-  bs_a2SE^2

 bs_a2alpha =  bs_a2mean * ((1 -  bs_a2mean) /  bs_a2var - 1 /  bs_a2mean)
 bs_a2beta = (1 -  bs_a2mean) * ((1 - bs_a2mean) / bs_a2var - 1 / (1 -  bs_a2mean))
 
 
   # bp_a2
 bp_a2mean <-0.45
 bp_a2upper_CI <- 0.7
 bp_a2lower_CI <- 0.4
  
 bp_a2SE <- ( bp_a2upper_CI -  bp_a2lower_CI) / 3.92

 bp_a2var <-  bp_a2SE^2

 bp_a2alpha =  bp_a2mean * ((1 -  bp_a2mean) /  bp_a2var - 1 /  bp_a2mean)
 bp_a2beta = (1 -  bp_a2mean) * ((1 - bp_a2mean) / bp_a2var - 1 / (1 -  bp_a2mean))
 
```

*run VR uncertainty loop*
```{r}
full_output <- vector("list", 6)
bycatch_reductions2 <- seq(0, 1, by =0.01)
mitigation_eff_range <- seq(0.5,1, by  = 0.1)
sims = 1:100

set.seed(123)

for(k in 1:length(mitigation_eff_range)){
  mit_eff <- mitigation_eff_range[k]
  
  output2 <- matrix(nrow = length(bycatch_reductions2), ncol = length(sims)) %>% as.data.frame() 
  
for(j in 1:length(sims)){
  
    set.seed(j)
    
    s_c2 <- random_values <- rbeta(n = 1, shape1 = s_c2alpha, shape2 = s_c2beta) #chick+ fledgling survival year 0-1 - 0.64 
    s_j2 <- random_values <- rbeta(n = 1, shape1 = s_j2alpha, shape2 = s_j2beta)#juvenile survival year 1-3 -  0.8 
    s_i2 <- random_values <- rbeta(n = 1, shape1 = s_i2alpha, shape2 = s_i2beta) #immature survival year 3-7 -  0.92
    s_s2 <- random_values <- rbeta(n = 1, shape1 = s_s2alpha, shape2 = s_s2beta) #subadult survival year 7-10 - 0.94
    s_a2 <- random_values <- rbeta(n = 1, shape1 = s_a2alpha, shape2 = s_a2beta) #adult survival year 10+ - 0.94

    bs_a2 <- random_values <- rbeta(n = 1, shape1 = bs_a2alpha, shape2 = bs_a2beta) #adult breeding success - 0.8 
    bp_a2 <- random_values <- rbeta(n = 1, shape1 = bp_a2alpha, shape2 = bp_a2beta)  #adult breeding probablity (biennial breeding, most breeding by age 10) - 0.45
    c2 <- 0.5# # of female chicks per year
    f_a2 <- bs_a*bp_a*c #fecundity
  
    # not sure why i decided to do this so removing it for now, and reverting matrix to old widowing rates (w and w4)
    # w2 <- sample(seq(0.4,0.5, by =0.01), size=1, replace = TRUE) 
    # w42 <- 1
    
    which1 <- as.numeric(sample(c(3,4), size=1, replace = TRUE))  #select whether to partition mortality into 3 sections (33% = bycatch) or 4 (25%=bycatch)
    m_j2 <- (1-s_j2)/which1 # divide total moratlity by 3 or 4 to get bycatch mortality
    o_m_j2 <- m_j2*(which1-1) #multiply bycatch by 1-(3 or 4) to get non-bycatch mortality

    which2 <- as.numeric(sample(c(3,4), size=1, replace = TRUE))
    m_i2 <- (1-s_i2)/which2
    o_m_i2 <- m_i2*(which2-1)

    which3 <- as.numeric(sample(c(3,4), size=1, replace = TRUE))
    m_s2 <- (1-s_s2)/which3
    o_m_s2 <- m_s2*(which3-1)

    which4 <- as.numeric(sample(c(3,4), size=1, replace = TRUE))
    m_a2 <- (1-s_a2)/which4
    o_m_a2 <- m_a2*(which4-1)
    
    
for(i in 1:length(bycatch_reductions2)){
  
  #calculate new survival rates = 1-(bycatch mortality*(1-efficacy of intervention action * proportion of fishery implementing intervention actions))+mortality remainder)
  n_s_j2 <- 1-((m_j2*(1-(mit_eff*bycatch_reductions2[i])))+(o_m_j2))
  n_s_i2 <- 1-((m_i2*(1-(mit_eff*bycatch_reductions2[i])))+(o_m_i2))
  n_s_s2 <- 1-((m_s2*(1-(mit_eff*bycatch_reductions2[i])))+(o_m_s2))
  n_s_a2 <- 1-((m_a2*(1-(mit_eff*bycatch_reductions2[i])))+(o_m_a2))
  
  #Age classes (years)  0-1             1-2    2-3    3-4    4-5    5-6    6-7    7-8    8-9    9-10   10+               nw,               w1,                  w2,                  w3,                  w4 
  n_m2 <- matrix(data=c(0,              0,     0,     0,     0,     0,     0,     0,     0,     0,     n_s_a2*f_a2,      n_s_a2*f_a2,      n_s_a2*f_a2*w,      n_s_a2*f_a2*w,      n_s_a2*f_a2*w,      n_s_a2*f_a2*w4,          #0-1
                        s_cc*(n_s_a2^2),0,     0,     0,     0,     0,     0,     0,     0,     0,     0,                0,                0,                   0,                   0,                   0,                    #1-2
                        0,              n_s_j2,0,     0,     0,     0,     0,     0,     0,     0,     0,                0,                0,                   0,                   0,                   0,                    #2-3
                        0,              0,     n_s_j2,0,     0,     0,     0,     0,     0,     0,     0,                0,                0,                   0,                   0,                   0,                    #3-4
                        0,              0,     0,     n_s_i2,0,     0,     0,     0,     0,     0,     0,                0,                0,                   0,                   0,                   0,                    #4-5
                        0,              0,     0,     0,     n_s_i2,0,     0,     0,     0,     0,     0,                0,                0,                   0,                   0,                   0,                    #5-6
                        0,              0,     0,     0,     0,     n_s_i2,0,     0,     0,     0,     0,                0,                0,                   0,                   0,                   0,                    #6-7
                        0,              0,     0,     0,     0,     0,     n_s_i2,0,     0,     0,     0,                0,                0,                   0,                   0,                   0,                    #7-8
                        0,              0,     0,     0,     0,     0,     0,     n_s_s2,0,     0,     0,                0,                0,                   0,                   0,                   0,                    #8-9
                        0,              0,     0,     0,     0,     0,     0,     0,     n_s_s2,0,     0,                0,                0,                   0,                   0,                   0,                    #9-10
                        0,              0,     0,     0,     0,     0,     0,     0,     0,     n_s_s2,0,                0,                0,                   0,                   0,                   0,                    #10
                        0,              0,     0,     0,     0,     0,     0,     0,     0,     0,     n_s_a2*n_s_a2,    n_s_a2*n_s_a2,    n_s_a2*n_s_a2*w,     n_s_a2*n_s_a2*w,     n_s_a2*n_s_a2*w,     n_s_a2*n_s_a2*w4,     #nw
                        0,              0,     0,     0,     0,     0,     0,     0,     0,     0,     n_s_a2*(1-n_s_a2),n_s_a2*(1-n_s_a2),n_s_a2*(1-n_s_a2)*w, n_s_a2*(1-n_s_a2)*w, n_s_a2*(1-n_s_a2)*w, n_s_a2*(1-n_s_a2)*w4, #w1
                        0,              0,     0,     0,     0,     0,     0,     0,     0,     0,     0,                0,                n_s_a2*(1-w),        0,                   0,                   0,                    #w2 
                        0,              0,     0,     0,     0,     0,     0,     0,     0,     0,     0,                0,                0,                   n_s_a2*(1-w),        0,                   0,                    #w3 
                        0,              0,     0,     0,     0,     0,     0,     0,     0,     0,     0,                0,                0,                   0,                   n_s_a2*(1-w),        0),                   #w4 
                       nrow = 16, byrow = TRUE) 
  
  #extract eigenvalues and lambda from the matrix 
  n_m_eig2 <- eigen(n_m2)
  lambda_n2 <- as.numeric(n_m_eig2$values[1])
  output2[i,j] <- lambda_n2 # this matches the ideal population growth rate in the table
  output2 <- as.data.frame(output2)
  output2$mitigation_efficacy <- mit_eff
  output2$bycatch_reduction <- bycatch_reductions2
    
 full_output[[k]] <- output2

}
}
}

full_output[[1]]

sim_results <- full_output[[1]]
sim_results$quantile5 <-  apply(sim_results[1:100], 1, quantile, (probs =c(0.05)))
sim_results$quantile95 <-  apply(sim_results[1:100], 1, quantile, (probs =c(0.95)))
```
*format output*
```{r}
mitigation_eff_range #0.5 0.6 0.7 0.8 0.9 1.0

mit05_l <- full_output[[1]] %>% 
  pivot_longer(cols = starts_with("V"), names_to = "simrun", values_to = "lambda")
mit06_l <- full_output[[2]] %>% 
  pivot_longer(cols = starts_with("V"), names_to = "simrun", values_to = "lambda")
mit07_l <- full_output[[3]] %>% 
  pivot_longer(cols = starts_with("V"), names_to = "simrun", values_to = "lambda")
mit08_l <- full_output[[4]] %>% 
  pivot_longer(cols = starts_with("V"), names_to = "simrun", values_to = "lambda")
mit09_l <- full_output[[5]] %>% 
  pivot_longer(cols = starts_with("V"), names_to = "simrun", values_to = "lambda")
mit1_l <- full_output[[6]] %>% 
  pivot_longer(cols = starts_with("V"), names_to = "simrun", values_to = "lambda")

fulldat <- rbind(mit05_l,mit06_l,mit07_l,mit08_l,mit09_l,mit1_l)
fulldat$mitsim <- paste(fulldat$mitigation_efficacy, fulldat$simrun)

mit05 <- full_output[[1]] 
mit06 <- full_output[[2]] 
mit07 <- full_output[[3]] 
mit08 <- full_output[[4]] 
mit09 <- full_output[[5]] 
mit1 <- full_output[[6]] 

full_dat_w <- rbind(mit05,mit06,mit07,mit08,mit09,mit1)
full_dat_w$quantile5 <-  apply(full_dat_w[1:100], 1, quantile, (probs =c(0.05)))
full_dat_w$quantile95 <-  apply(full_dat_w[1:100], 1, quantile, (probs =c(0.95)))
full_dat_w$mitsim <- paste(full_dat_w$mitigation_efficacy, full_dat_w$simrun)

```

```{r}

output_l$mitigation_efficacy <- output_l$mit_eff
fulldat <- fulldat %>% group_by(mitigation_efficacy, bycatch_reduction) %>% mutate(mean_lambda = mean(lambda)) %>% ungroup()

ggplot(fulldat)+
  geom_line(aes(x=bycatch_reduction, y=lambda, group = simrun), color = "grey")+
  geom_line(data = output_l, aes(x=bycatch_reductions, y=lambda))+
  geom_hline(aes(yintercept = 1), color = "red", alpha =0.5)+ 
  geom_line(aes(x=bycatch_reduction, y=mean_lambda, group = simrun), color = "blue")+
  # geom_vline(aes(xintercept = 0.66), color = "red", alpha =0.5)+ #illegal vs legal fisheries transition
  # geom_vline(aes(xintercept = 0.33), color = "red", alpha =0.5)+ #illegal vs legal fisheries transition
  labs(x="% fisheries with bycatch mitigation", y="lambda")+
  facet_wrap(~mitigation_efficacy)+
  themeo+
  coord_cartesian(expand=FALSE)


```

```{r}
fig2_widowing <- ggplot(full_dat_w)+
  geom_line(aes(x=bycatch_reduction, y=quantile5), color = "grey")+
  geom_line(aes(x=bycatch_reduction, y=quantile95), color = "grey")+
  geom_ribbon(aes(x=bycatch_reduction, ymin=quantile5, ymax=quantile95), color = "lightgrey", alpha =0.25)+
  geom_line(data = fulldat, aes(x=bycatch_reduction, y=mean_lambda, group = simrun), color = "blue")+ #mean VR
  # geom_line(data = output_l, aes(x=bycatch_reductions, y=lambda))+ #"best" VR
  geom_hline(aes(yintercept = 1), color = "red", alpha =0.5)+ 
  # geom_vline(aes(xintercept = 0.66), color = "red", alpha =0.5)+ #illegal vs legal fisheries transition
  # geom_vline(aes(xintercept = 0.33), color = "red", alpha =0.5)+ #illegal vs legal fisheries transition
  labs(x="% fisheries with bycatch mitigation", y="lambda")+
  facet_wrap(~mitigation_efficacy)+
  themeo+
  coord_cartesian(expand=FALSE)

 # ggsave(fig2_widowing, path ="/Users/sarahbecker/Dropbox/CU_ENVS/Classes/Spring2022/ENVS_5100_QuantConsBiol/project/w_albatross_demo_wd",file="fig2_widowing.png")
```


```{r}
fig2_widowingb <- ggplot(full_dat_w)+
  geom_line(aes(x=bycatch_reduction, y=quantile5), color = "grey")+
  geom_line(aes(x=bycatch_reduction, y=quantile95), color = "grey")+
  geom_ribbon(aes(x=bycatch_reduction, ymin=quantile5, ymax=quantile95,color = factor(mitigation_efficacy),fill  = factor(mitigation_efficacy)), alpha =0.25)+
  # geom_line(data = output_l, aes(x=bycatch_reductions, y=lambda))+ #"best" VR
  geom_line(data = fulldat, aes(x=bycatch_reduction, y=mean_lambda, group = simrun), color = "blue")+ #mean VR
  geom_hline(aes(yintercept = 1), color = "#525252", alpha =0.5)+ 
  labs(x="% fisheries with bycatch mitigation", y="lambda", fill = "mitigation efficacy", color = "mitigation efficacy")+
  geom_vline(aes(xintercept = 0.66), color = "#a50f15", alpha =0.5)+  #illegal vs legal fisheries transition
  # geom_vline(aes(xintercept = 0.6), color = "#a50f15", alpha =0.25)+  #variation in illegal vs legal fisheries transition
  # geom_vline(aes(xintercept = 0.72), color = "#a50f15", alpha =0.25)+ #varitation in illegal vs legal fisheries transition
  facet_wrap(~mitigation_efficacy, nrow =1)+
  themeo+
  guides(colour = guide_legend(nrow = 1))+
  coord_cartesian(expand=FALSE)+
  scale_x_continuous(breaks = c(0, 0.25, 0.5, 0.75, 1), labels=(c("0", "0.25", "0.5", "0.75", "1")))+
  scale_color_manual(values=c("#c7e9b4", "#7fcdbb", "#41b6c4", "#1d91c0", "#225ea8", "#253494"))+
  scale_fill_manual(values=c("#c7e9b4", "#7fcdbb", "#41b6c4", "#1d91c0", "#225ea8", "#253494"))

 # ggsave(fig2_widowingb, path ="/Users/sarahbecker/Dropbox/CU_ENVS/Classes/Spring2022/ENVS_5100_QuantConsBiol/project/w_albatross_demo_wd", file="fig2_widowingb.png", width = 14, height = 3.5, units = "in")
```


```{r}
 fig2_widowingc <- ggplot(full_dat_w)+
  geom_line(aes(x=bycatch_reduction, y=quantile5), color = "grey")+
  geom_line(aes(x=bycatch_reduction, y=quantile95), color = "grey")+
  geom_ribbon(aes(x=bycatch_reduction, ymin=quantile5, ymax=quantile95,color = factor(mitigation_efficacy),fill  = factor(mitigation_efficacy)), alpha =0.25)+
  # geom_line(data = output_l, aes(x=bycatch_reductions, y=lambda))+ #"best" VR
  geom_line(data = fulldat, aes(x=bycatch_reduction, y=mean_lambda, group = simrun), color = "blue")+ #mean VR
  geom_hline(aes(yintercept = 1), color = "#525252", alpha =0.5)+
  labs(x="% fisheries with bycatch mitigation", y="lambda")+
  facet_wrap(~mitigation_efficacy, nrow = 1)+
  themeo+
  guides(colour = guide_legend(nrow = 1))+
  coord_cartesian(expand=FALSE)+
  scale_color_manual(values=c("#c7e9b4", "#7fcdbb", "#41b6c4", "#1d91c0", "#225ea8", "#253494"))+
  scale_fill_manual(values=c("#c7e9b4", "#7fcdbb", "#41b6c4", "#1d91c0", "#225ea8", "#253494"))

 # ggsave(fig2_widowingc, path ="/Users/sarahbecker/Dropbox/CU_ENVS/Classes/Spring2022/ENVS_5100_QuantConsBiol/project/w_albatross_demo_wd", file="fig2_widowingc.png",width = 14, height = 3.5, units = "in")

```

 
```{r}
fig2_widowingd <- ggplot(full_dat_w)+
  geom_line(aes(x=bycatch_reduction, y=quantile5), color = "grey")+
  geom_line(aes(x=bycatch_reduction, y=quantile95), color = "grey")+
  geom_ribbon(aes(x=bycatch_reduction, ymin=quantile5, ymax=quantile95,color = factor(mitigation_efficacy),fill  = factor(mitigation_efficacy)), alpha =0.25)+
  # geom_line(data = output_l, aes(x=bycatch_reductions, y=lambda))+ #"best" VR
  geom_line(data = fulldat, aes(x=bycatch_reduction, y=mean_lambda, group = simrun), color = "blue")+ #mean VR
  geom_hline(aes(yintercept = 1), color = "#525252", alpha =0.5)+ 
  labs(x="% fisheries with bycatch mitigation", y="lambda")+
  geom_vline(aes(xintercept = 0.33), color = "#a50f15", alpha =0.5)+
  facet_wrap(~mitigation_efficacy, nrow = 1)+
  themeo+
  guides(colour = guide_legend(nrow = 1))+
  coord_cartesian(expand=FALSE)+
  scale_color_manual(values=c("#c7e9b4", "#7fcdbb", "#41b6c4", "#1d91c0", "#225ea8", "#253494"))+
  scale_fill_manual(values=c("#c7e9b4", "#7fcdbb", "#41b6c4", "#1d91c0", "#225ea8", "#253494"))

 # ggsave(fig2_widowingd, path ="/Users/sarahbecker/Dropbox/CU_ENVS/Classes/Spring2022/ENVS_5100_QuantConsBiol/project/w_albatross_demo_wd", file="fig2_widowingd.png",width = 14, height = 3.5, units = "in")
```

```{r}
fig1_widowing <- ggplot(full_dat_w)+
  geom_ribbon(aes(x=bycatch_reduction, ymin=quantile5, ymax=quantile95, color = factor(mitigation_efficacy),fill  = factor(mitigation_efficacy)), alpha =0.25)+
  geom_line(data = bycatch_reductions, aes(x=reduction, y=as.numeric(lambda)))+
  geom_hline(aes(yintercept = 1), color = "#525252")+ 
  labs(x="% fisheries with bycatch mitigation", y="lambda", color = "mitigation efficacy", fill = "mitigation efficacy")+
  themeo+
  guides(colour = guide_legend(nrow = 1))+
  coord_cartesian(expand=FALSE)+
  scale_color_manual(values=c("#c7e9b4", "#7fcdbb", "#41b6c4", "#1d91c0", "#225ea8", "#253494"))+
  scale_fill_manual(values=c("#c7e9b4", "#7fcdbb", "#41b6c4", "#1d91c0", "#225ea8", "#253494"))

#  ggsave(fig1_widowing, path ="/Users/sarahbecker/Dropbox/CU_ENVS/Classes/Spring2022/ENVS_5100_QuantConsBiol/project/w_albatross_demo_wd",file="fig1_widowing.png")
```

```{r}
fig1_widowing_illegal <- ggplot(full_dat_w)+
  geom_ribbon(aes(x=bycatch_reduction, ymin=quantile5, ymax=quantile95, color = factor(mitigation_efficacy),fill  = factor(mitigation_efficacy)), alpha =0.25)+
  geom_line(data = bycatch_reductions, aes(x=reduction, y=as.numeric(lambda)))+
  geom_hline(aes(yintercept = 1), color = "#525252")+ 
  geom_vline(aes(xintercept = 0.34), color = "#a50f15", alpha =0.5)+ #illegal vs legal fisheries transition
  geom_vline(aes(xintercept = 0.28), color = "#a50f15", alpha =0.25)+ #variation in illegal vs legal fisheries transition
  geom_vline(aes(xintercept = 0.40), color = "#a50f15", alpha =0.25)+ #variation in illegal vs legal fisheries transition
  labs(x="% fisheries with bycatch mitigation", y="lambda", color = "mitigation efficacy", fill = "mitigation efficacy")+
  themeo+
  guides(colour = guide_legend(nrow = 1))+
  coord_cartesian(expand=FALSE)+
  scale_color_manual(values=c("#c7e9b4", "#7fcdbb", "#41b6c4", "#1d91c0", "#225ea8", "#253494"))+
  scale_fill_manual(values=c("#c7e9b4", "#7fcdbb", "#41b6c4", "#1d91c0", "#225ea8", "#253494"))

# ggsave(fig1_widowing_illegal, path ="/Users/sarahbecker/Dropbox/CU_ENVS/Classes/Spring2022/ENVS_5100_QuantConsBiol/project/w_albatross_demo_wd", file="fig1_widowing_illegal.png")
```

```{r}
fig1_widowing_legal <- ggplot(full_dat_w)+
  geom_ribbon(aes(x=bycatch_reduction, ymin=quantile5, ymax=quantile95, color = factor(mitigation_efficacy),fill  = factor(mitigation_efficacy)), alpha =0.25)+
  geom_line(data = bycatch_reductions, aes(x=reduction, y=as.numeric(lambda)))+
  geom_hline(aes(yintercept = 1), color = "#525252")+
  geom_vline(aes(xintercept = 0.66), color = "#a50f15", alpha =0.5)+ #illegal vs legal fisheries transition
  geom_vline(aes(xintercept = 0.6), color = "#a50f15", alpha =0.25)+ #variation in illegal vs legal fisheries transition
  geom_vline(aes(xintercept = 0.72), color = "#a50f15", alpha =0.25)+ #variation in illegal vs legal fisheries transition
  labs(x="% fisheries with bycatch mitigation", y="lambda", color = "mitigation efficacy", fill = "mitigation efficacy")+
  themeo+
  guides(colour = guide_legend(nrow = 1))+
  coord_cartesian(expand=FALSE)+
  scale_color_manual(values=c("#c7e9b4", "#7fcdbb", "#41b6c4", "#1d91c0", "#225ea8", "#253494"))+
  scale_fill_manual(values=c("#c7e9b4", "#7fcdbb", "#41b6c4", "#1d91c0", "#225ea8", "#253494"))

# ggsave(fig1_widowing_legal, path ="/Users/sarahbecker/Dropbox/CU_ENVS/Classes/Spring2022/ENVS_5100_QuantConsBiol/project/w_albatross_demo_wd", file="fig1_widowing_legal.png")
```


```{r}
ggplot(fulldat)+
  geom_line(aes(x=bycatch_reduction, y=lambda, group = mitsim, color = factor(mitigation_efficacy)), alpha =0.5)+
  geom_line(data = bycatch_reductions, aes(x=reduction, y=as.numeric(lambda)))+
  geom_hline(aes(yintercept = 1), color = "darkgrey", alpha =0.5)+ 
  # geom_vline(aes(xintercept = 0.66), color = "red", alpha =0.5)+ #illegal vs legal fisheries transition
  # geom_vline(aes(xintercept = 0.33), color = "red", alpha =0.5)+ #illegal vs legal fisheries transition
  labs(x="% fisheries with bycatch mitigation", y="lambda")+
  themeo
```

