---
title: "beta_widowing_updated_matrix_mortality"
author: "Sarah Becker"
date: '2024-02-08'
output: html_document
---


```{r, echo= FALSE, include =FALSE}
library(here)
library(tidyverse)
library(RColorBrewer)

#set up various plotting themes
themeo <-theme_classic()+
  theme(strip.background = element_blank(),
        panel.grid.major = element_line(colour = "transparent"), 
        panel.grid.minor = element_line(colour = "transparent"),
        axis.line = element_blank(),
        axis.title = element_text(color = "black", size = 10),
        axis.ticks.length=unit(-0.1, "cm"),
        axis.ticks = element_line(color = "black", size = .25),
        axis.ticks.x = element_line(color = "black"),
        axis.ticks.y = element_line(color = "black"),
        panel.border = element_rect(colour = "black", fill=NA, size=.5),
        panel.spacing = unit(1, "lines"),
        legend.position = "bottom",
        legend.title = element_text(colour = "black", size = 10, face = "bold"),
        legend.text = element_text(color = "black", size =8)) #default legend position
```

**set vital rates. Try first without widowing**
```{r}
#survival rates - drawn from Weimerskirch 1998
s_c <- 0.64 #chick+ fledgling survival year 0-1
s_cc <- 0.7244736 #conditional chick survival (disregarding parental survival) *calculted from other VR, not directly from literature
s_j <- 0.85 #juvenile survival year 1-3
s_i <- 0.92 #immature survival year 3-7
s_s <- 0.94 #subadult survival year 7-10
s_a <- 0.94 #adult survival year 10+
#males show senescence but femailes don't, so not including it here, though perhaps I should since it impacts breeding success?

#fecundity building blocks
#breeding success - drawn from Weimerskirch 1998 &Rackete 2021
bs_a <- 0.76 #adult breeding success

#breeding probability (biennial breeders) - adult rate drawn from Rackete 2021, subadult downscaled slightly from that
bp_a <- 0.45 #adult breeding probablity (biennial breeding, most breeding by age 10)

c <- 0.5# # of female chicks per year

#fecundity 
#f_a <- bs_a*bp_a*c
f_a <- bp_a*c*s_cc
  
#add widowing
#assume 25% prob year1 -> 100% prob year 4?
#Sun et al 2022 - takes 4 years to reach next breeding attempt, breeding prob with new mate after widowhood = ~0.25 (fig5)

w <- 1-0.1^.25
w4 <- 1
```

**create matrix**
```{r}
#set up age-based matrix based on vital rates

#Age classes (years)  0-1        1-2    2-3    3-4    4-5    5-6    6-7    7-8   8-9    nw,          w1,            w2,             w3,             w4 
m <- matrix(data = c(0,          0,    0,      0,     0,     0,     0,     0,      0,   s_a^2*f_a,   s_a^2*f_a*w,   s_a^2*f_a*w,    s_a^2*f_a*w,    s_a^2*f_a*w4,   #0-1
                    s_j,         0,    0,      0,     0,     0,     0,     0,      0,   0,           0,             0,              0,              0,              #1-2
                    0,           s_j,  0,      0,     0,     0,     0,     0,      0,   0,           0,             0,              0,              0,              #2-3
                    0,           0,    s_j,    0,     0,     0,     0,     0,      0,   0,           0,             0,              0,              0,              #3-4
                    0,           0,    0,      s_i,   0,     0,     0,     0,      0,   0,           0,             0,              0,              0,              #4-5
                    0,           0,    0,      0,     s_i,   0,     0,     0,      0,   0,           0,             0,              0,              0,              #5-6
                    0,           0,    0,      0,     0,     s_i,   0,     0,      0,   0,           0,             0,              0,              0,              #6-7
                    0,           0,    0,      0,     0,     0,     s_i,   0,      0,   0,           0,             0,              0,              0,              #7-8
                    0,           0,    0,      0,     0,     0,     0,     s_s,    0,   0,           0,             0,              0,              0,              #8-9
                    0,           0,    0,      0,     0,     0,     0,     0,      s_s, s_a^2,       s_a^2*w,       s_a^2*w,        s_a^2*w,        s_a^2*w4,       #nw
                    0,           0,    0,      0,     0,     0,     0,     0,      0,   s_a*(1-s_a), s_a*(1-s_a)*w, s_a*(1-s_a)*w,  s_a*(1-s_a)*w,  s_a*(1-s_a)*w4, #w1
                    0,           0,    0,      0,     0,     0,     0,     0,      0,   0,           s_a*(1-w),     0,              0,              0,              #w2 
                    0,           0,    0,      0,     0,     0,     0,     0,      0,   0,           0,             s_a*(1-w),      0,              0,              #w3 
                    0,           0,    0,      0,     0,     0,     0,     0,      0,   0,           0,             0,              s_a*(1-w),      0),             #w4 
                   nrow = 14, byrow = TRUE) 
m
```

**extract lambda, SSDs, and RVs from eigenvalues**
```{r}
#deterministic growth rate (lambda)
#extract eigenvalues and lambda from the matrix 
m_eig <- eigen(m)
lambda <- as.numeric(m_eig$values[1])
lambda # this matches the ideal population growth rate in the table

#Stable state distribution
#extract and scale the right eigenvector 
SSD <- round(as.numeric(m_eig$vectors[,1]/sum(m_eig$vectors[,1])),3)
SSD

#extract reproductive values
RV <- as.numeric(eigen(t(m))$vectors[,1]/eigen(t(m))$vectors[1,1])
RV
```

*per capita mortality per hook*
*first, get total population estimates*
```{r}
MI_SDD <- sum(SSD[10:14]) #proportion mature individuals
JI_SDD <- sum(SSD[1:9]) #proportion juvenile/immature/subadult

#total fledged population from MI estimates: 
#from iucn redlist:https://www.iucnredlist.org/species/22698305/132640680#population
# In 1998, the total annual breeding population was estimated at 8,500 pairs, equivalent to c. 28,000 mature individuals (Gales 1998). However, current estimates are 1,553 pairs on South Georgia (Georgias del Sur) (Poncet et al. 2006), 1,800 pairs on Prince Edward Island (2008, Ryan et al. 2009), c. 1,900 pairs on Marion Island (2013, ACAP 2009), c. 340 pairs on Iles Crozet (CNRS Chinzè Monitoring Database 2010), c. 354 pairs in Iles Kerguelen (CNRS Chinzè Monitoring Database 2011), and 4 pairs on Macquarie Island (DPIWPE 2010, unpublished data), making a total of c. 6,000 annual breeding pairs. Using the same ratio as Gales (1998) for estimating the number of mature individuals, this would equate to approximately 20,100 mature individuals.

br_mi_rate <- 20100/6000

#South georgia mature individuals from reported breeding pairs
SG_MI <- 1553*br_mi_rate

#get total pop of fledged birds from MI  numbers and stable state distribution
#MI= MI_SDD*total pop
#total pop = MI/MI_SDD
SG_pop <- SG_MI/MI_SDD

#check against juvenile pop and SDD
juv_pop <- SG_pop-SG_MI #calc juv pop
juv_pop/SG_pop # prop to full fledged pop
JI_SDD #compare to SDD
```
*now, per capita hook mortality*
```{r}
BPUE_1000 <- ((0.5460*0.0108)+(0.3276*(0.0066))+(0.4169*0.1111)+(0.0120*0)+(0.2690*0.044)+(0.01060*0.044))/6 
                                    # #bycatch rate per 1000 hooks for South Atlantic (all south atlantic studies averaged) (Klaer 2012, table 2.5) #0.01111316
#BPUE_1000 <- 0.0308*0.105 #bycatch rate per 1000 hooks for South Atlantic (below 25S) (Klaer 2012, table 2.5) #0.003234
hooks <- 6710000 #average annual number of hooks in South West Atlantic from 1999-2009(Jimenez et al 2016, table 1)
#overlap <- 1466368 #average annual number of hooks in South West Atlantic from 1999-2009 that birds are exposed to (Jimenez et al 2016, table 1)
BPUE_1 <- BPUE_1000/1000 #convert to per 1 hook rate

bycatch_mortality <- BPUE_1*hooks #multiply by average annual hooks

per_cap_bycatch_mortality <- bycatch_mortality/SG_pop
```

*partition mortality between natural and fishery mortality*
```{r}
#surv = natural mortality - fishing mortality
#surv - fishing mortality = natural mortality

#fishing mortality = per capita calculation above
fm <- per_cap_bycatch_mortality

#natural mortality
nm_j <- (1-s_j)-fm
nm_i <- (1-s_i)-fm
nm_s <- (1-s_s)-fm
nm_a <- (1-s_a)-fm
```



**identify mortality rate, mitigation efficacy**
```{r}
#bycatch mortality rates
# ~25 - 35% of mortality has been attributed to bycatch (Cleeland et al 2021), 
# so I calculated the baseline mortality and divided by 4 (testing 33% now - /3)
# m_j <- (1-s_j)/3
# m_i <- (1-s_i)/3
# m_s <- (1-s_s)/3
# m_a <- (1-s_a)/3

mitigation_eff <- 0.8
```


**loop through different bycatch reductions scenarios, find lambda for each**
```{r}
bycatch_reductions <- data.frame(reduction = seq(0, 1, by =0.01), lambda = NA)

for(i in 1:nrow(bycatch_reductions)){
  # 
  # n_s_j <- 1-((m_j*(1-(mitigation_eff*bycatch_reductions$reduction[i])))+(m_j*2))
  # n_s_i <- 1-((m_i*(1-(mitigation_eff*bycatch_reductions$reduction[i])))+(m_i*2))
  # n_s_s <- 1-((m_s*(1-(mitigation_eff*bycatch_reductions$reduction[i])))+(m_s*2))
  # n_s_a <- 1-((m_a*(1-(mitigation_eff*bycatch_reductions$reduction[i])))+(m_s*2))


  n_s_j <- 1-((fm*(1-(mitigation_eff*bycatch_reductions$reduction[i])))+(nm_j))
  n_s_i <- 1-((fm*(1-(mitigation_eff*bycatch_reductions$reduction[i])))+(nm_i))
  n_s_s <- 1-((fm*(1-(mitigation_eff*bycatch_reductions$reduction[i])))+(nm_s))
  n_s_a <- 1-((fm*(1-(mitigation_eff*bycatch_reductions$reduction[i])))+(nm_s))

  
#Age classes (years)   0-1     1-2     2-3     3-4     4-5    5-6    6-7    7-8   8-9      nw,              w1,                w2,                w3,                w4 
n_m <- matrix(data = c(0,      0,      0,      0,      0,     0,     0,     0,      0,     n_s_a^2*f_a,     n_s_a^2*f_a*w,     n_s_a^2*f_a*w,     n_s_a^2*f_a*w,     n_s_a^2*f_a*w4,      #0-1
                       n_s_j,  0,      0,      0,      0,     0,     0,     0,      0,     0,               0,                 0,                 0,                 0,                   #1-2
                       0,      n_s_j,  0,      0,      0,     0,     0,     0,      0,     0,               0,                 0,                 0,                 0,                   #2-3
                       0,      0,      n_s_j,  0,      0,     0,     0,     0,      0,     0,               0,                 0,                 0,                 0,                   #3-4
                       0,      0,      0,      n_s_i,  0,     0,     0,     0,      0,     0,               0,                 0,                 0,                 0,                   #4-5
                       0,      0,      0,      0,      n_s_i, 0,     0,     0,      0,     0,               0,                 0,                 0,                 0,                   #5-6
                       0,      0,      0,      0,      0,     n_s_i, 0,     0,      0,     0,               0,                 0,                 0,                 0,                   #6-7
                       0,      0,      0,      0,      0,     0,     n_s_i, 0,      0,     0,               0,                 0,                 0,                 0,                   #7-8
                       0,      0,      0,      0,      0,     0,     0,     n_s_s,  0,     0,               0,                 0,                 0,                 0,                   #8-9
                       0,      0,      0,      0,      0,     0,     0,     0,      n_s_s, n_s_a^2,         n_s_a^2*w,         n_s_a^2*w,         n_s_a^2*w,         n_s_a^2*w4,          #nw
                       0,      0,      0,      0,      0,     0,     0,     0,      0,     n_s_a*(1-n_s_a), n_s_a*(1-n_s_a)*w, n_s_a*(1-n_s_a)*w, n_s_a*(1-n_s_a)*w, n_s_a*(1-n_s_a)*w4,  #w1
                       0,      0,      0,      0,      0,     0,     0,     0,      0,     0,               n_s_a*(1-w),       0,                 0,                 0,                   #w2 
                       0,      0,      0,      0,      0,     0,     0,     0,      0,     0,               0,                 n_s_a*(1-w),       0,                 0,                   #w3 
                       0,      0,      0,      0,      0,     0,     0,     0,      0,     0,               0,                 0,                 n_s_a*(1-w),       0),                  #w4 
                    nrow = 14, byrow = TRUE) 
n_m
  
  #extract eigenvalues and lambda from the matrix 
  n_m_eig <- eigen(n_m)
  lambda_n <- as.numeric(n_m_eig$values[1])
  bycatch_reductions$lambda[i] <- lambda_n 

}
```

**plot bycatch reduction against changes to lambda**
```{r}
ggplot(bycatch_reductions)+
  geom_line(aes(x=reduction, y=as.numeric(lambda)))+
  geom_hline(aes(yintercept = 1), color = "red", alpha =0.5)+ 
  # geom_vline(aes(xintercept = 0.66), color = "red", alpha =0.5)+ #illegal vs legal fisheries transition
  # geom_vline(aes(xintercept = 0.33), color = "red", alpha =0.5)+ #illegal vs legal fisheries transition
  labs(x="% fisheries with bycatch mitigation", y="lambda")+
  themeo+
  coord_cartesian(expand=FALSE)
```

**add uncertainty - vary mitigation efficacy**
```{r}
bycatch_reductions2 <- seq(0, 1, by =0.01)
mitigation_eff_range <- seq(0.5,1, by  = 0.1)
output <- matrix(nrow = length(bycatch_reductions2), ncol = length(mitigation_eff_range)) %>% as.data.frame() 

for(k in 1:length(mitigation_eff_range)){
  mit_eff <- mitigation_eff_range[k]
  
  for(i in 1:length(bycatch_reductions2)){
  
  # n_s_j <- 1-((m_j*(1-(mit_eff*bycatch_reductions2[i])))+(m_j*2))
  # n_s_i <- 1-((m_i*(1-(mit_eff*bycatch_reductions2[i])))+(m_i*2))
  # n_s_s <- 1-((m_s*(1-(mit_eff*bycatch_reductions2[i])))+(m_s*2))
  # n_s_a <- 1-((m_a*(1-(mit_eff*bycatch_reductions2[i])))+(m_s*2))
  # 
  
  n_s_j <- 1-((fm*(1-(mit_eff*bycatch_reductions$reduction[i])))+(nm_j))
  n_s_i <- 1-((fm*(1-(mit_eff*bycatch_reductions$reduction[i])))+(nm_i))
  n_s_s <- 1-((fm*(1-(mit_eff*bycatch_reductions$reduction[i])))+(nm_s))
  n_s_a <- 1-((fm*(1-(mit_eff*bycatch_reductions$reduction[i])))+(nm_s))
  
  
 #Age classes (years)   0-1     1-2     2-3     3-4     4-5    5-6    6-7    7-8   8-9      nw,              w1,                w2,                w3,                w4 
n_m <- matrix(data = c(0,      0,      0,      0,      0,     0,     0,     0,      0,     n_s_a^2*f_a,     n_s_a^2*f_a*w,     n_s_a^2*f_a*w,     n_s_a^2*f_a*w,     n_s_a^2*f_a*w4,      #0-1
                       n_s_j,  0,      0,      0,      0,     0,     0,     0,      0,     0,               0,                 0,                 0,                 0,                   #1-2
                       0,      n_s_j,  0,      0,      0,     0,     0,     0,      0,     0,               0,                 0,                 0,                 0,                   #2-3
                       0,      0,      n_s_j,  0,      0,     0,     0,     0,      0,     0,               0,                 0,                 0,                 0,                   #3-4
                       0,      0,      0,      n_s_i,  0,     0,     0,     0,      0,     0,               0,                 0,                 0,                 0,                   #4-5
                       0,      0,      0,      0,      n_s_i, 0,     0,     0,      0,     0,               0,                 0,                 0,                 0,                   #5-6
                       0,      0,      0,      0,      0,     n_s_i, 0,     0,      0,     0,               0,                 0,                 0,                 0,                   #6-7
                       0,      0,      0,      0,      0,     0,     n_s_i, 0,      0,     0,               0,                 0,                 0,                 0,                   #7-8
                       0,      0,      0,      0,      0,     0,     0,     n_s_s,  0,     0,               0,                 0,                 0,                 0,                   #8-9
                       0,      0,      0,      0,      0,     0,     0,     0,      n_s_s, n_s_a^2,         n_s_a^2*w,         n_s_a^2*w,         n_s_a^2*w,         n_s_a^2*w4,          #nw
                       0,      0,      0,      0,      0,     0,     0,     0,      0,     n_s_a*(1-n_s_a), n_s_a*(1-n_s_a)*w, n_s_a*(1-n_s_a)*w, n_s_a*(1-n_s_a)*w, n_s_a*(1-n_s_a)*w4,  #w1
                       0,      0,      0,      0,      0,     0,     0,     0,      0,     0,               n_s_a*(1-w),       0,                 0,                 0,                   #w2 
                       0,      0,      0,      0,      0,     0,     0,     0,      0,     0,               0,                 n_s_a*(1-w),       0,                 0,                   #w3 
                       0,      0,      0,      0,      0,     0,     0,     0,      0,     0,               0,                 0,                 n_s_a*(1-w),       0),                  #w4 
                    nrow = 14, byrow = TRUE) 
n_m
  
  #extract eigenvalues and lambda from the matrix 
  n_m_eig <- eigen(n_m)
  lambda_n <- as.numeric(n_m_eig$values[1])
  output[i,k] <- lambda_n # this matches the ideal population growth rate in the table

}
}

```

*format output*
```{r}
oldnames = c("V1","V2","V3","V4" ,"V5", "V6")
newnames = c("0.5", "0.6", "0.7", "0.8", "0.9", "1")

output<- output %>%
  rename_with(~ newnames[which(oldnames == .x)], .cols = oldnames)

output$bycatch_reductions <- bycatch_reductions2

output_l <- output %>% 
  pivot_longer(cols = !bycatch_reductions,names_to = "mit_eff", values_to = "lambda")

output_l$mit_eff <- as.numeric(output_l$mit_eff)
```

```{r}
ggplot(output_l)+
  geom_line(aes(x=bycatch_reductions, y=lambda, group = factor(mit_eff), color = factor(mit_eff)))+
  geom_hline(aes(yintercept = 1), color = "darkgrey", alpha =0.5)+ 
  # geom_vline(aes(xintercept = 0.66), color = "red", alpha =0.5)+ #illegal vs legal fisheries transition
  # geom_vline(aes(xintercept = 0.33), color = "red", alpha =0.5)+ #illegal vs legal fisheries transition
  labs(x="% fisheries with bycatch mitigation", y="lambda", color = "% mitigation efficacy")+
  themeo+
  guides(colour = guide_legend(nrow = 1))+
  coord_cartesian(expand=FALSE)+
  scale_color_manual(values=c("#c7e9b4", "#7fcdbb", "#41b6c4", "#1d91c0", "#225ea8", "#253494"))
```

**add vital rate uncertainty, simulate 1000 runs with varying range of vital rates drawn from beta distributions based on variability found in the literature** 
**mortality = 1/3 bycatch, 2/3 other causes**

**first, set VR uncertainty from beta dist**
```{r}
#set mean, CI for each vital rate, use these to pull VR rates within beta distribution
#s_c2
s_c2mean <-0.64
s_c2upper_CI <- 0.65
s_c2lower_CI <- 0.63
  
s_c2SE <- (s_c2upper_CI - s_c2lower_CI) / 3.92

s_c2var <- s_c2SE^2

s_c2alpha = s_c2mean * ((1 - s_c2mean) / s_c2var - 1 / s_c2mean)
s_c2beta = (1 - s_c2mean) * ((1 -s_c2mean) /s_c2var - 1 / (1 - s_c2mean))

#s_cc2
s_cc2mean <-0.7244736
s_cc2upper_CI <- 0.73
s_cc2lower_CI <- 0.72
  
s_cc2SE <- (s_cc2upper_CI - s_cc2lower_CI) / 3.92

s_cc2var <- s_cc2SE^2

s_cc2alpha = s_cc2mean * ((1 - s_cc2mean) / s_cc2var - 1 / s_cc2mean)
s_cc2beta = (1 - s_cc2mean) * ((1 -s_cc2mean) /s_cc2var - 1 / (1 - s_cc2mean))

#s_j2
#s_j2mean <-0.825
s_j2mean <-0.85
s_j2upper_CI <- 0.85
# s_j2upper_CI <- 0.89
s_j2lower_CI <- 0.8
  
s_j2SE <- (s_j2upper_CI - s_j2lower_CI) / 3.92

s_j2var <- s_j2SE^2

s_j2alpha = s_j2mean * ((1 - s_j2mean) / s_j2var - 1 / s_j2mean)
s_j2beta = (1 - s_j2mean) * ((1 -s_j2mean) /s_j2var - 1 / (1 - s_j2mean))

# s_i2
 #s_i2mean <-0.925
 s_i2mean <-0.92
 s_i2upper_CI <- 0.94
 s_i2lower_CI <- 0.91
  
 s_i2SE <- ( s_i2upper_CI -  s_i2lower_CI) / 3.92

 s_i2var <-  s_i2SE^2

 s_i2alpha =  s_i2mean * ((1 -  s_i2mean) /  s_i2var - 1 /  s_i2mean)
 s_i2beta = (1 -  s_i2mean) * ((1 - s_i2mean) / s_i2var - 1 / (1 -  s_i2mean))
 
# s_s2
 #s_s2mean <-0.944
 s_s2mean <-0.94
 s_s2upper_CI <- 0.95
 s_s2lower_CI <- 0.94
  
 s_s2SE <- ( s_s2upper_CI -  s_s2lower_CI) / 3.92

 s_s2var <-  s_s2SE^2

 s_s2alpha =  s_s2mean * ((1 -  s_s2mean) /  s_s2var - 1 /  s_s2mean)
 s_s2beta = (1 -  s_s2mean) * ((1 - s_s2mean) / s_s2var - 1 / (1 -  s_s2mean))
 
 # s_a2
 #s_a2mean <-0.945
  s_a2mean <-0.94
 s_a2upper_CI <- 0.95
 s_a2lower_CI <- 0.94
  
 s_a2SE <- ( s_a2upper_CI -  s_a2lower_CI) / 3.92

 s_a2var <-  s_a2SE^2

 s_a2alpha =  s_a2mean * ((1 -  s_a2mean) /  s_a2var - 1 /  s_a2mean)
 s_a2beta = (1 -  s_a2mean) * ((1 - s_a2mean) / s_a2var - 1 / (1 -  s_a2mean))

 
  # bs_a2
 bs_a2mean <-0.76
 bs_a2upper_CI <- 0.8
 bs_a2lower_CI <- 0.68
  
 bs_a2SE <- ( bs_a2upper_CI -  bs_a2lower_CI) / 3.92

 bs_a2var <-  bs_a2SE^2

 bs_a2alpha =  bs_a2mean * ((1 -  bs_a2mean) /  bs_a2var - 1 /  bs_a2mean)
 bs_a2beta = (1 -  bs_a2mean) * ((1 - bs_a2mean) / bs_a2var - 1 / (1 -  bs_a2mean))
 
 
   # bp_a2
 bp_a2mean <-0.45
 bp_a2upper_CI <- 0.7
 bp_a2lower_CI <- 0.4
  
 bp_a2SE <- ( bp_a2upper_CI -  bp_a2lower_CI) / 3.92

 bp_a2var <-  bp_a2SE^2

 bp_a2alpha =  bp_a2mean * ((1 -  bp_a2mean) /  bp_a2var - 1 /  bp_a2mean)
 bp_a2beta = (1 -  bp_a2mean) * ((1 - bp_a2mean) / bp_a2var - 1 / (1 -  bp_a2mean))
 
 #fm2
 
# BPUE_1000 <- ((0.5460*0.0108)+(0.3276*0.0066)+(0.4169*0.1111)+(0.0120*0)+(0.2690*0.044)+(0.01060*0.044))/6 
# hooks <- 6710000 #average annual number of hooks in South West Atlantic from 1999-2009(Jimenez et al 2016, table 1)
# BPUE_1 <- BPUE_1000/1000 #convert to per 1 hook rate
# bycatch_mortality <- BPUE_1*hooks #multiply by average annual hooks
# per_cap_bycatch_mortality <- bycatch_mortality/SG_pop


fm2_mean <- fm
fm2_upperCI <- (((0.4169*0.1111)/1000)*hooks)/SG_pop
fm2_lowerCI <- (((0.01060*0.044)/1000)*hooks)/SG_pop

fm2_SE <- ( fm2_upperCI -  fm2_lowerCI) / 3.92

fm2_var <-  fm2_SE^2

 fm2_alpha =  fm2_mean * ((1 -  fm2_mean) /  fm2_var - 1 /  fm2_mean)
 fm2_beta = (1 -  fm2_mean) * ((1 - fm2_mean) / fm2_var - 1 / (1 -  fm2_mean))

```


*look at distributions of each variable*
```{r}
 s_c2_rv <-  rbeta(n = 1000, shape1 = s_c2alpha, shape2 = s_c2beta) #chick+ fledgling survival year 0-1 - 0.64 
    p_s_c2<- ggplot(as.data.frame(s_c2_rv))+geom_histogram(aes(x=s_c2_rv))+geom_vline(data = as.data.frame(s_c), aes(xintercept = s_c), color = "red")+theme_bw()
      
 s_cc2_rv <- rbeta(n = 1000, shape1 = s_cc2alpha, shape2 = s_cc2beta) # conditional chick+ fledgling survival year 0-1 - 0.73
    p_s_cc2 <- ggplot(as.data.frame(s_cc2_rv))+geom_histogram(aes(x=s_cc2_rv))+geom_vline(data = as.data.frame(s_cc), aes(xintercept = s_cc), color = "red")+theme_bw()
   
 s_j2_rv <-  rbeta(n = 1000, shape1 = s_j2alpha, shape2 = s_j2beta)#juvenile survival year 1-3 -  0.8 
     p_s_j2 <- ggplot(as.data.frame(s_j2_rv))+geom_histogram(aes(x=s_j2_rv))+geom_vline(data = as.data.frame(s_j), aes(xintercept = s_j), color = "red")+theme_bw()
      
 s_i2_rv <-  rbeta(n = 1000, shape1 = s_i2alpha, shape2 = s_i2beta) #immature survival year 3-7 -  0.92
     p_s_i2 <- ggplot(as.data.frame(s_i2_rv))+geom_histogram(aes(x=s_i2_rv))+geom_vline(data = as.data.frame(s_i), aes(xintercept = s_i), color = "red")+theme_bw()
     
 s_s2_rv <-  rbeta(n = 1000, shape1 = s_s2alpha, shape2 = s_s2beta) #subadult survival year 7-10 - 0.94
     p_s_s2 <- ggplot(as.data.frame(s_s2_rv))+geom_histogram(aes(x=s_s2_rv))+geom_vline(data = as.data.frame(s_s), aes(xintercept = s_s), color = "red")+theme_bw()
     
 s_a2_rv <-  rbeta(n = 1000, shape1 = s_a2alpha, shape2 = s_a2beta) #adult survival year 10+ - 0.94
     p_s_a2 <- ggplot(as.data.frame(s_a2_rv))+geom_histogram(aes(x=s_a2_rv))+geom_vline(data = as.data.frame(s_a), aes(xintercept = s_a), color = "red")+theme_bw()
    
 bs_a2_rv <- rbeta(n = 1000, shape1 = bs_a2alpha, shape2 = bs_a2beta) #adult breeding success - 0.8 
    p_bs_a2 <- ggplot(as.data.frame(bs_a2_rv))+geom_histogram(aes(x=bs_a2_rv))+geom_vline(data = as.data.frame(bs_a), aes(xintercept = bs_a), color = "red")+theme_bw()
       
bp_a2_rv <-  rbeta(n = 1000, shape1 = bp_a2alpha, shape2 = bp_a2beta)  #adult breeding probablity (biennial breeding, most breeding by age 10) - 0.45
    p_bp_a2 <- ggplot(as.data.frame(bp_a2_rv))+geom_histogram(aes(x=bp_a2_rv))+geom_vline(data = as.data.frame(bp_a), aes(xintercept = bp_a), color = "red")+theme_bw()
    
   fm2_rv <-  rbeta(n = 1000, shape1 = fm2_alpha, shape2 = fm2_beta)  #adult breeding probablity (biennial breeding, most breeding by age 10) - 0.45
    p_fm2 <- ggplot(as.data.frame(fm2_rv))+geom_histogram(aes(x=fm2_rv))+geom_vline(data = as.data.frame(bp_a), aes(xintercept = fm), color = "red")+theme_bw()
```

*run VR uncertainty loop*
```{r}
full_output <- vector("list", 6)
bycatch_reductions2 <- seq(0, 1, by =0.01)
mitigation_eff_range <- seq(0.5,1, by  = 0.1)
sims = 1:1000

set.seed(123)

for(k in 1:length(mitigation_eff_range)){
  mit_eff <- mitigation_eff_range[k]
  
  output2 <- matrix(nrow = length(bycatch_reductions2), ncol = length(sims)) %>% as.data.frame() 
  
for(j in 1:length(sims)){
    
  set.seed(j)
  
    s_c2 <-    rbeta(n = 1, shape1 = s_c2alpha, shape2 = s_c2beta) #chick+ fledgling survival year 0-1 - 0.64 
    s_cc2 <-  rbeta(n = 1, shape1 = s_cc2alpha, shape2 = s_cc2beta) # conditional chick+ fledgling survival year 0-1 - 0.64 
    s_j2 <- rbeta(n = 1, shape1 = s_j2alpha, shape2 = s_j2beta)#juvenile survival year 1-3 -  0.8 
    s_i2 <-  rbeta(n = 1, shape1 = s_i2alpha, shape2 = s_i2beta) #immature survival year 3-7 -  0.92
    s_s2 <- rbeta(n = 1, shape1 = s_s2alpha, shape2 = s_s2beta) #subadult survival year 7-10 - 0.94
    s_a2 <-  rbeta(n = 1, shape1 = s_a2alpha, shape2 = s_a2beta) #adult survival year 10+ - 0.94
    bs_a2 <-  rbeta(n = 1, shape1 = bs_a2alpha, shape2 = bs_a2beta) #adult breeding success - 0.8 
    bp_a2 <-  rbeta(n = 1, shape1 = bp_a2alpha, shape2 = bp_a2beta)  #adult breeding probablity (biennial breeding, most breeding by age 10) - 0.45
    c2 <- 0.5# # of female chicks per year
    f_a2 <- bp_a*c2*s_cc2 #fecundity
  
    # not sure why i decided to do this so removing it for now, and reverting matrix to old widowing rates (w and w4)
    # w2 <- sample(seq(0.4,0.5, by =0.01), size=1, replace = TRUE) 
    # w42 <- 1
    
  #  which1 <- as.numeric(sample(c(3,4), size=1, replace = TRUE))  #select whether to partition mortality into 3 sections (33% = bycatch) or 4 (25%=bycatch)
    # m_j2 <- (1-s_j2)/which1 # divide total moratlity by 3 or 4 to get bycatch mortality
    # o_m_j2 <- m_j2*(which1-1) #multiply bycatch by 1-(3 or 4) to get non-bycatch mortality
    # 
    # which2 <- as.numeric(sample(c(3,4), size=1, replace = TRUE))
    # m_i2 <- (1-s_i2)/which2
    # o_m_i2 <- m_i2*(which2-1)
    # 
    # which3 <- as.numeric(sample(c(3,4), size=1, replace = TRUE))
    # m_s2 <- (1-s_s2)/which3
    # o_m_s2 <- m_s2*(which3-1)
    # 
    # which4 <- as.numeric(sample(c(3,4), size=1, replace = TRUE))
    # m_a2 <- (1-s_a2)/which4
    # o_m_a2 <- m_a2*(which4-1)
    
    #fishing mortality = per capita calculation above
   # fm2 <- fm #set bycatch mortality
    fm2 <-  rbeta(n = 1, shape1 = fm2_alpha, shape2 = fm2_beta)  #variable bycatch mortality
    
    #natural mortality
    nm_j2 <- (1-s_j2)-fm2
    nm_i2 <- (1-s_i2)-fm2
    nm_s2 <- (1-s_s2)-fm2
    nm_a2 <- (1-s_a2)-fm2

    
for(i in 1:length(bycatch_reductions2)){
  
  #calculate new survival rates = 1-(bycatch mortality*(1-efficacy of intervention action * proportion of fishery implementing intervention actions))+mortality remainder)
  # n_s_j2 <- 1-((m_j2*(1-(mit_eff*bycatch_reductions2[i])))+(o_m_j2))
  # n_s_i2 <- 1-((m_i2*(1-(mit_eff*bycatch_reductions2[i])))+(o_m_i2))
  # n_s_s2 <- 1-((m_s2*(1-(mit_eff*bycatch_reductions2[i])))+(o_m_s2))
  # n_s_a2 <- 1-((m_a2*(1-(mit_eff*bycatch_reductions2[i])))+(o_m_a2))
  # 
    
  n_s_j2 <- 1-((fm2*(1-(mit_eff*bycatch_reductions$reduction[i])))+(nm_j2))
  n_s_i2 <- 1-((fm2*(1-(mit_eff*bycatch_reductions$reduction[i])))+(nm_i2))
  n_s_s2 <- 1-((fm2*(1-(mit_eff*bycatch_reductions$reduction[i])))+(nm_s2))
  n_s_a2 <- 1-((fm2*(1-(mit_eff*bycatch_reductions$reduction[i])))+(nm_s2))
  

   #Age classes (years)    0-1      1-2      2-3     3-4     4-5     5-6     6-7    7-8     8-9      nw,                w1,                  w2,                   w3,                   w4
  n_m2 <- matrix(data = c( 0,       0,       0,      0,      0,      0,      0,      0,      0,      n_s_a2^2*f_a2,     n_s_a2^2*f_a2*w,     n_s_a2^2*f_a2*w,      n_s_a2^2*f_a2*w,      n_s_a2^2*f_a2*w4,      #0-1
                           n_s_j2,  0,       0,      0,      0,      0,      0,      0,      0,      0,                 0,                   0,                    0,                    0,                     #1-2
                           0,       n_s_j2,  0,      0,      0,      0,      0,      0,      0,      0,                 0,                   0,                    0,                    0,                     #2-3
                           0,       0,       n_s_j2, 0,      0,      0,      0,      0,      0,      0,                 0,                   0,                    0,                    0,                     #3-4
                           0,       0,       0,      n_s_i2, 0,      0,      0,      0,      0,      0,                 0,                   0,                    0,                    0,                     #4-5
                           0,       0,       0,      0,      n_s_i2, 0,      0,      0,      0,      0,                 0,                   0,                    0,                    0,                     #5-6
                           0,       0,       0,      0,      0,      n_s_i2, 0,      0,      0,      0,                 0,                   0,                    0,                    0,                     #6-7
                           0,       0,       0,      0,      0,      0,      n_s_i2, 0,      0,      0,                 0,                   0,                    0,                    0,                     #7-8
                           0,       0,       0,      0,      0,      0,      0,      n_s_s2, 0,      0,                 0,                   0,                    0,                    0,                     #8-9
                           0,       0,       0,      0,      0,      0,      0,      0,      n_s_s2, n_s_a2^2,          n_s_a2^2*w,          n_s_a2^2*w,           n_s_a2^2*w,           n_s_a2^2*w4,           #nw
                           0,       0,       0,      0,      0,      0,      0,      0,      0,      n_s_a2*(1-n_s_a2), n_s_a2*(1-n_s_a2)*w, n_s_a2*(1-n_s_a2)*w,  n_s_a2*(1-n_s_a2)*w,  n_s_a2*(1-n_s_a2)*w4,  #w1
                           0,       0,       0,      0,      0,      0,      0,      0,      0,      0,                 n_s_a2*(1-w),        0,                    0,                    0,                     #w2
                           0,       0,       0,      0,      0,      0,      0,      0,      0,      0,                 0,                   n_s_a2*(1-w),         0,                    0,                     #w3
                           0,       0,       0,      0,      0,      0,      0,      0,      0,      0,                 0,                   0,                    n_s_a2*(1-w),         0),                    #w4
                   nrow = 14, byrow = TRUE)
n_m2
  
   
  
  #extract eigenvalues and lambda from the matrix 
  n_m_eig2 <- eigen(n_m2)
  lambda_n2 <- as.numeric(n_m_eig2$values[1])
  output2[i,j] <- lambda_n2 # this matches the ideal population growth rate in the table
  output2 <- as.data.frame(output2)
  output2$mitigation_efficacy <- mit_eff
  output2$bycatch_reduction <- bycatch_reductions2
    
 full_output[[k]] <- output2

}
}
}

full_output[[1]]

 sim_results <- full_output[[1]]
 sim_results$quantile5 <-  apply(sim_results[1:100], 1, quantile, (probs =c(0.05)))
 sim_results$quantile95 <-  apply(sim_results[1:100], 1, quantile, (probs =c(0.95)))
```
*format output*
```{r}
mitigation_eff_range #0.5 0.6 0.7 0.8 0.9 1.0

mit05_l <- full_output[[1]] %>% 
  pivot_longer(cols = starts_with("V"), names_to = "simrun", values_to = "lambda")
mit06_l <- full_output[[2]] %>% 
  pivot_longer(cols = starts_with("V"), names_to = "simrun", values_to = "lambda")
mit07_l <- full_output[[3]] %>% 
  pivot_longer(cols = starts_with("V"), names_to = "simrun", values_to = "lambda")
mit08_l <- full_output[[4]] %>% 
  pivot_longer(cols = starts_with("V"), names_to = "simrun", values_to = "lambda")
mit09_l <- full_output[[5]] %>% 
  pivot_longer(cols = starts_with("V"), names_to = "simrun", values_to = "lambda")
mit1_l <- full_output[[6]] %>% 
  pivot_longer(cols = starts_with("V"), names_to = "simrun", values_to = "lambda")

fulldat <- rbind(mit05_l,mit06_l,mit07_l,mit08_l,mit09_l,mit1_l)
fulldat$mitsim <- paste(fulldat$mitigation_efficacy, fulldat$simrun)

mit05 <- full_output[[1]] 
mit06 <- full_output[[2]] 
mit07 <- full_output[[3]] 
mit08 <- full_output[[4]] 
mit09 <- full_output[[5]] 
mit1 <- full_output[[6]] 

full_dat_w <- rbind(mit05,mit06,mit07,mit08,mit09,mit1)
full_dat_w$quantile5 <-  apply(full_dat_w[1:100], 1, quantile, (probs =c(0.05)))
full_dat_w$quantile95 <-  apply(full_dat_w[1:100], 1, quantile, (probs =c(0.95)))
full_dat_w$quantile25 <-  apply(full_dat_w[1:100], 1, quantile, (probs =c(0.25)))
full_dat_w$quantile75 <-  apply(full_dat_w[1:100], 1, quantile, (probs =c(0.75)))
full_dat_w$mitsim <- paste(full_dat_w$mitigation_efficacy, full_dat_w$simrun)

```

```{r}

output_l$mitigation_efficacy <- output_l$mit_eff
fulldat <- fulldat %>% group_by(mitigation_efficacy, bycatch_reduction) %>% mutate(mean_lambda = mean(lambda)) %>% ungroup()

ggplot(fulldat)+
  geom_line(aes(x=bycatch_reduction, y=lambda, group = simrun), color = "grey", alpha = 0.1)+
 # geom_line(data = output_l, aes(x=bycatch_reductions, y=lambda))+ #"best" VR
  geom_hline(aes(yintercept = 1), color = "red", alpha =0.5)+ 
  geom_line(aes(x=bycatch_reduction, y=mean_lambda, group = simrun), color = "blue")+ #mean VR
  # geom_vline(aes(xintercept = 0.66), color = "red", alpha =0.5)+ #illegal vs legal fisheries transition
  # geom_vline(aes(xintercept = 0.33), color = "red", alpha =0.5)+ #illegal vs legal fisheries transition
   geom_vline(aes(xintercept = 0.66), color = "#a50f15", alpha =0.25)+ #illegal vs legal fisheries transition - weimerskirch
   geom_vline(aes(xintercept = 0.8), color = "#a50f15", alpha =0.5)+ #mean illegal vs legal fisheries transition - welch &weimerskirch
  geom_vline(aes(xintercept = 0.94), color = "#a50f15", alpha =0.25)+ #illegal vs legal fisheries transition - welch 
  labs(x="% fisheries with bycatch mitigation", y="lambda")+
  facet_wrap(~mitigation_efficacy, nrow = 1)+
  themeo+
  coord_cartesian(expand=FALSE)
```

```{r}
fig2_widowing <- ggplot(full_dat_w)+
  geom_line(aes(x=bycatch_reduction, y=quantile5), color = "grey")+
  geom_line(aes(x=bycatch_reduction, y=quantile95), color = "grey")+
  geom_ribbon(aes(x=bycatch_reduction, ymin=quantile5, ymax=quantile95), color = "lightgrey", alpha =0.25)+
    geom_ribbon(aes(x=bycatch_reduction, ymin=quantile25, ymax=quantile75), color = "lightgrey", alpha =0.25)+
  geom_line(data = fulldat, aes(x=bycatch_reduction, y=mean_lambda, group = simrun), color = "blue")+ #mean VR
  # geom_line(data = output_l, aes(x=bycatch_reductions, y=lambda))+ #"best" VR
  geom_hline(aes(yintercept = 1), color = "red", alpha =0.5)+ 
  # geom_vline(aes(xintercept = 0.66), color = "red", alpha =0.5)+ #illegal vs legal fisheries transition
  # geom_vline(aes(xintercept = 0.33), color = "red", alpha =0.5)+ #illegal vs legal fisheries transition
  labs(x="% fisheries with bycatch mitigation", y="lambda")+
  facet_wrap(~mitigation_efficacy)+
  themeo+
  coord_cartesian(expand=FALSE)

 # ggsave(fig2_widowing, path ="/Users/sarahbecker/Dropbox/CU_ENVS/Classes/Spring2022/ENVS_5100_QuantConsBiol/project/w_albatross_demo_wd", file="fig2_widowing.png")
```


```{r}
fig2_widowingb <- ggplot(full_dat_w)+
  geom_line(aes(x=bycatch_reduction, y=quantile5), color = "grey")+
  geom_line(aes(x=bycatch_reduction, y=quantile95), color = "grey")+
  geom_ribbon(aes(x=bycatch_reduction, ymin=quantile5, ymax=quantile95,color = factor(mitigation_efficacy),fill  = factor(mitigation_efficacy)), alpha =0.25)+
  geom_ribbon(aes(x=bycatch_reduction, ymin=quantile25, ymax=quantile75,color = factor(mitigation_efficacy),fill  = factor(mitigation_efficacy)), alpha =0.25)+
  #geom_line(data = output_l, aes(x=bycatch_reductions, y=lambda))+ #"best" VR
  geom_line(data = fulldat, aes(x=bycatch_reduction, y=mean_lambda, group = simrun), color = "blue")+ #mean VR
  geom_hline(aes(yintercept = 1), color = "#525252", alpha =0.5)+ 
  labs(x="% fisheries with bycatch mitigation", y="lambda", fill = "mitigation efficacy", color = "mitigation efficacy")+
geom_vline(aes(xintercept = 0.66), color = "#a50f15", alpha =0.25)+ #illegal vs legal fisheries transition - weimerskirch
   geom_vline(aes(xintercept = 0.8), color = "#a50f15", alpha =0.5)+ #mean illegal vs legal fisheries transition - welch &weimerskirch
  geom_vline(aes(xintercept = 0.94), color = "#a50f15", alpha =0.25)+ #illegal vs legal fisheries transition - welch 
  # geom_vline(aes(xintercept = 0.6), color = "#a50f15", alpha =0.25)+ #variation in illegal vs legal fisheries transition
  # geom_vline(aes(xintercept = 0.72), color = "#a50f15", alpha =0.25)+ #variation in illegal vs legal fisheries transition
  facet_wrap(~mitigation_efficacy, nrow =1)+
  themeo+
  guides(colour = guide_legend(nrow = 1))+
  coord_cartesian(expand=FALSE)+
  scale_x_continuous(breaks = c(0, 0.25, 0.5, 0.75, 1), labels=(c("0", "0.25", "0.5", "0.75", "1")))+
  scale_color_manual(values=c("#c7e9b4", "#7fcdbb", "#41b6c4", "#1d91c0", "#225ea8", "#253494"))+
  scale_fill_manual(values=c("#c7e9b4", "#7fcdbb", "#41b6c4", "#1d91c0", "#225ea8", "#253494"))

 # ggsave(fig2_widowingb, path ="/Users/sarahbecker/Dropbox/CU_ENVS/Classes/Spring2022/ENVS_5100_QuantConsBiol/project/w_albatross_demo_wd", file="fig2_widowingb.png", width = 14, height = 3.5, units = "in")
```


```{r}
fig2_widowingc <- ggplot(full_dat_w)+
  geom_line(aes(x=bycatch_reduction, y=quantile5), color = "grey")+
  geom_line(aes(x=bycatch_reduction, y=quantile95), color = "grey")+
  geom_ribbon(aes(x=bycatch_reduction, ymin=quantile5, ymax=quantile95,color = factor(mitigation_efficacy),fill  = factor(mitigation_efficacy)), alpha =0.25)+
  # geom_line(data = output_l, aes(x=bycatch_reductions, y=lambda))+ #"best" VR
  geom_line(data = fulldat, aes(x=bycatch_reduction, y=mean_lambda, group = simrun), color = "blue")+ #mean VR
  geom_hline(aes(yintercept = 1), color = "#525252", alpha =0.5)+ 
  labs(x="% fisheries with bycatch mitigation", y="lambda")+
  facet_wrap(~mitigation_efficacy, nrow = 1)+
  themeo+
  guides(colour = guide_legend(nrow = 1))+
  coord_cartesian(expand=FALSE)+
  scale_color_manual(values=c("#c7e9b4", "#7fcdbb", "#41b6c4", "#1d91c0", "#225ea8", "#253494"))+
  scale_fill_manual(values=c("#c7e9b4", "#7fcdbb", "#41b6c4", "#1d91c0", "#225ea8", "#253494"))

# ggsave(fig2_widowingc, path ="/Users/sarahbecker/Dropbox/CU_ENVS/Classes/Spring2022/ENVS_5100_QuantConsBiol/project/w_albatross_demo_wd", file="fig2_widowingc.png",width = 14, height = 3.5, units = "in")
```

 
```{r}
fig2_widowingd <- ggplot(full_dat_w)+
  geom_line(aes(x=bycatch_reduction, y=quantile5), color = "grey")+
  geom_line(aes(x=bycatch_reduction, y=quantile95), color = "grey")+
  geom_ribbon(aes(x=bycatch_reduction, ymin=quantile5, ymax=quantile95,color = factor(mitigation_efficacy),fill  = factor(mitigation_efficacy)), alpha =0.25)+
  # geom_line(data = output_l, aes(x=bycatch_reductions, y=lambda))+ #"best" VR
  geom_line(data = fulldat, aes(x=bycatch_reduction, y=mean_lambda, group = simrun), color = "blue")+ #mean VR
  geom_hline(aes(yintercept = 1), color = "#525252", alpha =0.5)+ 
  labs(x="% fisheries with bycatch mitigation", y="lambda")+
  geom_vline(aes(xintercept = 0.33), color = "#a50f15", alpha =0.5)+ #illegal vs legal fisheries transition
  facet_wrap(~mitigation_efficacy, nrow = 1)+
  themeo+
  guides(colour = guide_legend(nrow = 1))+
  coord_cartesian(expand=FALSE)+
  scale_color_manual(values=c("#c7e9b4", "#7fcdbb", "#41b6c4", "#1d91c0", "#225ea8", "#253494"))+
  scale_fill_manual(values=c("#c7e9b4", "#7fcdbb", "#41b6c4", "#1d91c0", "#225ea8", "#253494"))

 # ggsave(fig2_widowingd, path ="/Users/sarahbecker/Dropbox/CU_ENVS/Classes/Spring2022/ENVS_5100_QuantConsBiol/project/w_albatross_demo_wd",  file="fig2_widowingd.png",width = 14, height = 3.5, units = "in")
```

```{r}
fig1_widowing <- ggplot(full_dat_w)+
  geom_ribbon(aes(x=bycatch_reduction, ymin=quantile5, ymax=quantile95, color = factor(mitigation_efficacy),fill  = factor(mitigation_efficacy)), alpha =0.25)+
  geom_line(data = bycatch_reductions, aes(x=reduction, y=as.numeric(lambda)))+
  geom_hline(aes(yintercept = 1), color = "#525252")+ 
  labs(x="% fisheries with bycatch mitigation", y="lambda", color = "mitigation efficacy", fill = "mitigation efficacy")+
  themeo+
  guides(colour = guide_legend(nrow = 1))+
  coord_cartesian(expand=FALSE)+
  scale_color_manual(values=c("#c7e9b4", "#7fcdbb", "#41b6c4", "#1d91c0", "#225ea8", "#253494"))+
  scale_fill_manual(values=c("#c7e9b4", "#7fcdbb", "#41b6c4", "#1d91c0", "#225ea8", "#253494"))

#  ggsave(fig1_widowing, path ="/Users/sarahbecker/Dropbox/CU_ENVS/Classes/Spring2022/ENVS_5100_QuantConsBiol/project/w_albatross_demo_wd",  file="fig1_widowing.png")
```

```{r}
fig1_widowing_illegal <- ggplot(full_dat_w)+
  geom_ribbon(aes(x=bycatch_reduction, ymin=quantile5, ymax=quantile95, color = factor(mitigation_efficacy),fill  = factor(mitigation_efficacy)), alpha =0.25)+
  geom_line(data = bycatch_reductions, aes(x=reduction, y=as.numeric(lambda)))+
  geom_hline(aes(yintercept = 1), color = "#525252")+ 
  geom_vline(aes(xintercept = 0.34), color = "#a50f15", alpha =0.5)+ #illegal vs legal fisheries transition
  geom_vline(aes(xintercept = 0.28), color = "#a50f15", alpha =0.25)+ #variation in illegal vs legal fisheries transition
  geom_vline(aes(xintercept = 0.40), color = "#a50f15", alpha =0.25)+ #variation in illegal vs legal fisheries transition
  labs(x="% fisheries with bycatch mitigation", y="lambda", color = "mitigation efficacy", fill = "mitigation efficacy")+
  themeo+
  guides(colour = guide_legend(nrow = 1))+
  coord_cartesian(expand=FALSE)+
  scale_color_manual(values=c("#c7e9b4", "#7fcdbb", "#41b6c4", "#1d91c0", "#225ea8", "#253494"))+
  scale_fill_manual(values=c("#c7e9b4", "#7fcdbb", "#41b6c4", "#1d91c0", "#225ea8", "#253494"))

# ggsave(fig1_widowing_illegal, path ="/Users/sarahbecker/Dropbox/CU_ENVS/Classes/Spring2022/ENVS_5100_QuantConsBiol/project/w_albatross_demo_wd", file="fig1_widowing_illegal.png")
```

```{r}
fig1_widowing_legal <- ggplot(full_dat_w)+
  geom_ribbon(aes(x=bycatch_reduction, ymin=quantile5, ymax=quantile95, color = factor(mitigation_efficacy),fill  = factor(mitigation_efficacy)), alpha =0.25)+
  geom_line(data = bycatch_reductions, aes(x=reduction, y=as.numeric(lambda)))+
  geom_hline(aes(yintercept = 1), color = "#525252")+ 
  geom_vline(aes(xintercept = 0.66), color = "#a50f15", alpha =0.5)+ #illegal vs legal fisheries transition
  geom_vline(aes(xintercept = 0.6), color = "#a50f15", alpha =0.25)+ #variation in illegal vs legal fisheries transition
  geom_vline(aes(xintercept = 0.72), color = "#a50f15", alpha =0.25)+ #variation in illegal vs legal fisheries transition
  labs(x="% fisheries with bycatch mitigation", y="lambda", color = "mitigation efficacy", fill = "mitigation efficacy")+
  themeo+
  guides(colour = guide_legend(nrow = 1))+
  coord_cartesian(expand=FALSE)+
  scale_color_manual(values=c("#c7e9b4", "#7fcdbb", "#41b6c4", "#1d91c0", "#225ea8", "#253494"))+
  scale_fill_manual(values=c("#c7e9b4", "#7fcdbb", "#41b6c4", "#1d91c0", "#225ea8", "#253494"))

# ggsave(fig1_widowing_legal, path ="/Users/sarahbecker/Dropbox/CU_ENVS/Classes/Spring2022/ENVS_5100_QuantConsBiol/project/w_albatross_demo_wd", file="fig1_widowing_legal.png")
```


```{r}
ggplot(fulldat)+
  geom_line(aes(x=bycatch_reduction, y=lambda, group = mitsim, color = factor(mitigation_efficacy)), alpha =0.5)+
  geom_line(data = bycatch_reductions, aes(x=reduction, y=as.numeric(lambda)))+
  geom_hline(aes(yintercept = 1), color = "darkgrey", alpha =0.5)+ 
  # geom_vline(aes(xintercept = 0.66), color = "red", alpha =0.5)+ #illegal vs legal fisheries transition
  # geom_vline(aes(xintercept = 0.33), color = "red", alpha =0.5)+ #illegal vs legal fisheries transition
  labs(x="% fisheries with bycatch mitigation", y="lambda")+
  themeo
```

