---
title: "WAAL Population Projections — 3/3 Mitigation Scenario"
output: html_document
---

```{r setup, echo = FALSE, include = FALSE}
library(here)
library(tidyverse)
library(RColorBrewer)

# Must match the date suffix used when 06a was last run
date_sfx      <- "2026-02-25"
target_period <- "1990-1994"

themeo <- theme_classic() +
  theme(strip.background = element_blank(),
        panel.grid.major = element_line(colour = "transparent"),
        panel.grid.minor = element_line(colour = "transparent"),
        axis.line = element_blank(),
        axis.title = element_text(color = "black", size = 10),
        axis.ticks.length = unit(-0.1, "cm"),
        axis.ticks = element_line(color = "black", size = .25),
        axis.ticks.x = element_line(color = "black"),
        axis.ticks.y = element_line(color = "black"),
        panel.border = element_rect(colour = "black", fill = NA, size = .5),
        panel.spacing = unit(1, "lines"),
        legend.position = "bottom",
        legend.title = element_text(colour = "black", size = 10, face = "bold"),
        legend.text = element_text(color = "black", size = 8))
```

*POPULATION PROJECTIONS — 3/3 MITIGATION SCENARIO*

Two analyses:
  1. Immediate full 3/3 compliance — how long for the population to recover?
  2. Delayed implementation (5, 10, 20 year phase-in) — what is the cost of waiting?

Both use the scalar projection N(t+1) = lambda * N(t), following the homework structure.
For Part 1, lambda is fixed (one value per scenario).
For Part 2, lambda changes each year as coverage ramps up — computed explicitly and stored
in a coverage schedule table before the projection runs, so every input can be inspected.

---

**vital rates**
```{r vital-rates}
# Stage-specific survival rates (Pardo et al., 1980-2012)
s_Juv <- 0.846
s_EF  <- 0.913
s_ENB <- 0.943
s_ES  <- 0.895
s_IF  <- 0.920
s_Imm <- 0.921
s_IS  <- 0.892
s_PF  <- 0.935

# Stage-specific behavioral rates
b_EF  <- 0.746;  b_ENB <- 0.589;  b_ES  <- 0.016
b_IF  <- 0.744;  b_Imm <- 0.101;  b_IS  <- 0.011;  b_PF  <- 0.941

k_EF  <- 0.631;  k_ENB <- 0.675;  k_ES  <- 0.500
k_IF  <- 0.627;  k_Imm <- 0.585;  k_IS  <- 0.627;  k_PF  <- 0.791

r_EF  <- 0.973;  r_ENB <- 0.953;  r_ES  <- 0.292
r_IF  <- 0.948;  r_Imm <- 0.692;  r_IS  <- 0.459;  r_PF  <- 0.972

# Mitigation efficacy: 3/3 combined (tori lines + night setting + line weighting)
#numbers pulled from the literature and will need citations and confirmation
eff_tori_lines  <- 0.75
eff_night_set   <- 0.60
eff_line_weight <- 0.65
eff_3of3 <- 1 - (1 - eff_tori_lines) * (1 - eff_night_set) * (1 - eff_line_weight)

```

**load per-capita fishing mortality and natural mortality from 06a output**
```{r load-data}
# Per-capita FM and natural mortality saved by 06a (mortality_partition CSV)
mort_part <- read_csv(
  here(paste0("output/mortality_partition_", target_period, "_", date_sfx, ".csv")),
  show_col_types = FALSE)

mort_part

# Extract per-capita DLL and PLL fishing mortality by stage group
dem_fm_juv <- mort_part %>% filter(stage == "Juv") %>% pull(fm_dll)
pel_fm_juv <- mort_part %>% filter(stage == "Juv") %>% pull(fm_pll)
dem_fm_imm <- mort_part %>% filter(stage == "Imm") %>% pull(fm_dll)
pel_fm_imm <- mort_part %>% filter(stage == "Imm") %>% pull(fm_pll)
dem_fm_fb  <- mort_part %>% filter(stage == "EF")  %>% pull(fm_dll)  # same for IF
pel_fm_fb  <- mort_part %>% filter(stage == "EF")  %>% pull(fm_pll)
dem_fm_sb  <- mort_part %>% filter(stage == "ES")  %>% pull(fm_dll)  # same for IS
pel_fm_sb  <- mort_part %>% filter(stage == "ES")  %>% pull(fm_pll)
dem_fm_nb  <- mort_part %>% filter(stage == "ENB") %>% pull(fm_dll)  # same for PF
pel_fm_nb  <- mort_part %>% filter(stage == "ENB") %>% pull(fm_pll)

# Natural mortality by stage (total mortality minus fishing mortality)
nm_juv <- mort_part %>% filter(stage == "Juv") %>% pull(nm)
nm_imm <- mort_part %>% filter(stage == "Imm") %>% pull(nm)
nm_efb <- mort_part %>% filter(stage == "EF")  %>% pull(nm)
nm_esb <- mort_part %>% filter(stage == "ES")  %>% pull(nm)
nm_enb <- mort_part %>% filter(stage == "ENB") %>% pull(nm)
nm_ifb <- mort_part %>% filter(stage == "IF")  %>% pull(nm)
nm_isb <- mort_part %>% filter(stage == "IS")  %>% pull(nm)
nm_pf  <- mort_part %>% filter(stage == "PF")  %>% pull(nm)

# Starting population: total Bird Island birds in the target period
dem_props <- read_csv(here("data/WAAL_dist_Clay2019/Dem_props copy.csv"),
                      show_col_types = FALSE) %>%
  mutate(Dem_class = str_replace(Dem_class, "J2\\+3", "J2J3"))

pop_period <- dem_props %>%
  filter(Period == target_period, Dem_class != "J1") %>%
  select(Dem_class, Pop)

N0 <- sum(pop_period$Pop)

```

**build matrices and extract lambdas**
```{r build-matrices}
# --- No mitigation matrix ---
# Survival = observed Pardo rates, which already encode fishing mortality
# as it existed in the baseline period (1990-1994)
# Stages: J2 J3 Imm4 Imm5 PB IS IF ES EF ENB Sabb
m_base <- matrix(data = c(
  0,     0,      0,      0,      0,                                          0.5*s_Juv,              0,                       0.5*s_Juv,              0,                       0,                         0,
  s_Juv, 0,      0,      0,      0,                                          0,                       0,                       0,                       0,                       0,                         0,
  0,     s_Juv,  0,      0,      0,                                          0,                       0,                       0,                       0,                       0,                         0,
  0,     0,      s_Imm,  0,      0,                                          0,                       0,                       0,                       0,                       0,                         0,
  0,     0,      0,      s_Imm,  s_Imm*r_Imm*(1-b_Imm)+s_Imm*(1-r_Imm),    0,                       0,                       0,                       0,                       0,                         0,
  0,     0,      0,      0,      s_Imm*r_Imm*b_Imm*k_Imm,                   0,                       0,                       0,                       0,                       0,                         0,
  0,     0,      0,      0,      s_Imm*r_Imm*b_Imm*(1-k_Imm),               0,                       0,                       0,                       0,                       0,                         0,
  0,     0,      0,      0,      0,                                          s_IS*r_IS*b_IS*k_IS,     s_IF*r_IF*b_IF*k_IF,     s_ES*r_ES*b_ES*k_ES,     s_EF*r_EF*b_EF*k_EF,     s_ENB*r_ENB*b_ENB*k_ENB,   s_PF*r_PF*b_PF*k_PF,
  0,     0,      0,      0,      0,                                          s_IS*r_IS*b_IS*(1-k_IS), s_IF*r_IF*b_IF*(1-k_IF), s_ES*r_ES*b_ES*(1-k_ES), s_EF*r_EF*b_EF*(1-k_EF), s_ENB*r_ENB*b_ENB*(1-k_ENB),s_PF*r_PF*b_PF*(1-k_PF),
  0,     0,      0,      0,      0,                                          s_IS*r_IS*(1-b_IS),      s_IF*r_IF*(1-b_IF),      s_ES*r_ES*(1-b_ES),      s_EF*r_EF*(1-b_EF),      s_ENB*r_ENB*(1-b_ENB),     s_PF*r_PF*(1-b_PF),
  0,     0,      0,      0,      0,                                          s_IS*(1-r_IS),           s_IF*(1-r_IF),           s_ES*(1-r_ES),           s_EF*(1-r_EF),           s_ENB*(1-r_ENB),           s_PF*(1-r_PF)
), nrow = 11, byrow = TRUE)

lambda_base <- as.numeric(eigen(m_base)$values[1])
SSD         <- as.numeric(eigen(m_base)$vectors[, 1] / sum(eigen(m_base)$vectors[, 1]))
n0_vec      <- SSD * N0

# --- Full 3/3 mitigation matrix ---
# Survival rates with fishing mortality reduced by eff_3of3 (tori lines + night setting + line weighting)
# ns_* = 1 - (FM_total * (1 - eff_3of3) + natural_mortality)
# FM values from 06a mortality partition output
ns_Juv <- 1 - ((dem_fm_juv + pel_fm_juv) * (1 - eff_3of3) + nm_juv)
ns_Imm <- 1 - ((dem_fm_imm + pel_fm_imm) * (1 - eff_3of3) + nm_imm)
ns_EF  <- 1 - ((dem_fm_fb  + pel_fm_fb)  * (1 - eff_3of3) + nm_efb)
ns_ES  <- 1 - ((dem_fm_sb  + pel_fm_sb)  * (1 - eff_3of3) + nm_esb)
ns_ENB <- 1 - ((dem_fm_nb  + pel_fm_nb)  * (1 - eff_3of3) + nm_enb)
ns_IF  <- 1 - ((dem_fm_fb  + pel_fm_fb)  * (1 - eff_3of3) + nm_ifb)
ns_IS  <- 1 - ((dem_fm_sb  + pel_fm_sb)  * (1 - eff_3of3) + nm_isb)
ns_PF  <- 1 - ((dem_fm_nb  + pel_fm_nb)  * (1 - eff_3of3) + nm_pf)

m_3of3 <- matrix(data = c(
  0,      0,       0,       0,       0,                                           0.5*ns_Juv,               0,                        0.5*ns_Juv,               0,                        0,                          0,
  ns_Juv, 0,       0,       0,       0,                                           0,                        0,                        0,                        0,                        0,                          0,
  0,      ns_Juv,  0,       0,       0,                                           0,                        0,                        0,                        0,                        0,                          0,
  0,      0,       ns_Imm,  0,       0,                                           0,                        0,                        0,                        0,                        0,                          0,
  0,      0,       0,       ns_Imm,  ns_Imm*r_Imm*(1-b_Imm)+ns_Imm*(1-r_Imm),   0,                        0,                        0,                        0,                        0,                          0,
  0,      0,       0,       0,       ns_Imm*r_Imm*b_Imm*k_Imm,                   0,                        0,                        0,                        0,                        0,                          0,
  0,      0,       0,       0,       ns_Imm*r_Imm*b_Imm*(1-k_Imm),               0,                        0,                        0,                        0,                        0,                          0,
  0,      0,       0,       0,       0,                                           ns_IS*r_IS*b_IS*k_IS,     ns_IF*r_IF*b_IF*k_IF,     ns_ES*r_ES*b_ES*k_ES,     ns_EF*r_EF*b_EF*k_EF,     ns_ENB*r_ENB*b_ENB*k_ENB,   ns_PF*r_PF*b_PF*k_PF,
  0,      0,       0,       0,       0,                                           ns_IS*r_IS*b_IS*(1-k_IS), ns_IF*r_IF*b_IF*(1-k_IF), ns_ES*r_ES*b_ES*(1-k_ES), ns_EF*r_EF*b_EF*(1-k_EF), ns_ENB*r_ENB*b_ENB*(1-k_ENB),ns_PF*r_PF*b_PF*(1-k_PF),
  0,      0,       0,       0,       0,                                           ns_IS*r_IS*(1-b_IS),      ns_IF*r_IF*(1-b_IF),      ns_ES*r_ES*(1-b_ES),      ns_EF*r_EF*(1-b_EF),      ns_ENB*r_ENB*(1-b_ENB),     ns_PF*r_PF*(1-b_PF),
  0,      0,       0,       0,       0,                                           ns_IS*(1-r_IS),           ns_IF*(1-r_IF),           ns_ES*(1-r_ES),           ns_EF*(1-r_EF),           ns_ENB*(1-r_ENB),           ns_PF*(1-r_PF)
), nrow = 11, byrow = TRUE)

lambda_3of3 <- as.numeric(eigen(m_3of3)$values[1])

```

---

*PART 1 — IMMEDIATE FULL COMPLIANCE*

How long does it take the population to recover if 3/3 mitigation is implemented
across all fisheries immediately at 100% coverage?

**Step 1: set up output storage**
```{r part1-setup}
years <- 100
yearslong  <- 500

# One column for N, rows = years + 1 (year 0 through year 100)
output_base <- matrix(nrow = years + 1, ncol = 1) %>% as.data.frame()
output_imm  <- matrix(nrow = years + 1, ncol = 1) %>% as.data.frame()

# Year 0: starting population
output_base[1, 1] <- N0
output_imm[1, 1]  <- N0

# Long-run storage
output_base_long <- matrix(nrow = yearslong + 1, ncol = 1) %>% as.data.frame()
output_imm_long  <- matrix(nrow = yearslong + 1, ncol = 1) %>% as.data.frame()
output_base_long[1, 1] <- N0
output_imm_long[1, 1]  <- N0

```

**Step 3: project — N(t+1) = lambda * N(t)**
```{r part1-project}
# No mitigation: same lambda every year
for (k in 1:years) {
  output_base[k + 1, 1] <- lambda_base * output_base[k, 1]
}

# Immediate full 3/3: same lambda every year
for (k in 1:years) {
  output_imm[k + 1, 1] <- lambda_3of3 * output_imm[k, 1]
}

# Combine and label for future plotting— short projection (100 yr)
proj_part1 <- bind_rows(
  output_base %>% mutate(scenario = "No mitigation (baseline)"),
  output_imm  %>% mutate(scenario = "Immediate 3/3 (100% coverage)")
) %>%
  mutate(year     = rep(0:years, 2),
         N        = V1,
         scenario = factor(scenario,
           levels = c("No mitigation (baseline)", "Immediate 3/3 (100% coverage)"))) %>%
  select(year, N, scenario)
 
# Long projection loops
for (k in 1:yearslong) {
  output_base_long[k + 1, 1] <- lambda_base  * output_base_long[k, 1]
}
for (k in 1:yearslong) {
  output_imm_long[k + 1, 1]  <- lambda_3of3  * output_imm_long[k, 1]
}

# Combine and label for future plotting — long projection (yearslong yr)
proj_part1_long <- bind_rows(
  output_base_long %>% mutate(scenario = "No mitigation (baseline)"),
  output_imm_long  %>% mutate(scenario = "Immediate 3/3 (100% coverage)")
) %>%
  mutate(year     = rep(0:yearslong, 2),
         N        = V1,
         scenario = factor(scenario,
           levels = c("No mitigation (baseline)", "Immediate 3/3 (100% coverage)"))) %>%
  select(year, N, scenario)

# Inspect first few years to verify
print(head(proj_part1 %>% filter(scenario == "No mitigation (baseline)"), 6))
print(head(proj_part1 %>% filter(scenario == "Immediate 3/3 (100% coverage)"), 6))
```

**recovery milestones**
```{r part1-milestones}
# Years for the immediate-3/3 mean trajectory to reach growth milestones above N0
imm_traj <- proj_part1 %>% filter(scenario == "Immediate 3/3 (100% coverage)")

# helper: return NA_integer_ if milestone not reached within projection horizon
first_yr <- function(traj, threshold) {
  x <- traj %>% filter(N >= threshold) %>% slice(1) %>% pull(year)
  if (length(x) == 0) NA_integer_ else x
}

fmt_yr      <- function(x) ifelse(is.na(x), paste0(">", years),     as.character(x))
fmt_yr_long <- function(x) ifelse(is.na(x), paste0(">", yearslong), as.character(x))

milestone_10  <- first_yr(imm_traj, N0 * 1.10)
milestone_25  <- first_yr(imm_traj, N0 * 1.25)
milestone_50  <- first_yr(imm_traj, N0 * 1.50)
milestone_dbl <- first_yr(imm_traj, N0 * 2.00)


# Milestones for long projection — check against full yearslong horizon
imm_traj_long <- proj_part1_long %>% filter(scenario == "Immediate 3/3 (100% coverage)")

milestone_10_long  <- first_yr(imm_traj_long, N0 * 1.10)
milestone_25_long  <- first_yr(imm_traj_long, N0 * 1.25)
milestone_50_long  <- first_yr(imm_traj_long, N0 * 1.50)
milestone_dbl_long <- first_yr(imm_traj_long, N0 * 2.00)

```

**observed population totals from dem_props (validation)**
```{r dem-obs}
# 5-year period averages from Clay 2019 demographic data
# Excluding J1 to match N0 calculation; plotted at start year of each period
dem_obs <- dem_props %>%
  filter(Dem_class != "J1",
         Period %in% c("1990-1994", "1995-1999", "2000-2004", "2005-2009")) %>%
  group_by(Period) %>%
  summarise(N_obs = sum(Pop), .groups = "drop") %>%
  mutate(year = c(0, 5, 10, 15))

# Adults = SB + FB + NB;  Immatures = IMM + J2+3
dem_obs_split <- dem_props %>%
  filter(Dem_class != "J1",
         Period %in% c("1990-1994", "1995-1999", "2000-2004", "2005-2009")) %>%
  mutate(group = ifelse(Dem_class %in% c("IMM", "J2+3"), "Immatures", "Adults"),
         year  = case_when(
           Period == "1990-1994" ~ 0,
           Period == "1995-1999" ~ 5,
           Period == "2000-2004" ~ 10,
           Period == "2005-2009" ~ 15
         )) %>%
  group_by(Period, year, group) %>%
  summarise(N_obs = sum(Pop), .groups = "drop")

print(dem_obs_split)

# Bird Island breeding pair counts (Poncet et al. survey data)
# Years are end-year of austral season; x-axis is years since 1990 (start of baseline period)
# Scale factor converts breeding pairs to total N for plotting on primary y-axis:
# interpolating between 1984 (1370 pairs) and 2004 (948 pairs) gives ~1200 pairs at 1990,
# so scale_factor = N0 / 1200
pairs_scale <- N0 / 1200

bird_obs <- tibble(
  season = c("1983/84", "2003/04", "2014/15", "2023/24"),
  year   = c(-6, 14, 25, 34),       # years relative to 1990 baseline start
  pairs  = c(1370, 948, 772, 699),
  N_scaled = pairs * pairs_scale     # scaled to primary y-axis for plotting
)

print(bird_obs)
```

**plot Part 1**
```{r plot-part1, fig.width = 9, fig.height = 5}
milestones_df <- tibble(
  year  = c(milestone_10, milestone_25, milestone_50, milestone_dbl),
  label = c("+10%", "+25%", "+50%", "\u00D72"),
  N     = c(N0 * 1.10, N0 * 1.25, N0 * 1.50, N0 * 2.0)
) %>%
  filter(!is.na(year), year <= years)

p_part1 <- ggplot(proj_part1, aes(x = year, y = N,
                                   color = scenario, linetype = scenario)) +
  geom_line(linewidth = 0.9) +
  geom_hline(yintercept = N0, color = "grey50", linetype = "dashed", linewidth = 0.4) +
  geom_point(data = dem_obs,
             aes(x = year, y = N_obs),
             inherit.aes = FALSE, color = "black", size = 3, shape = 17) +
  geom_text(data = dem_obs,
            aes(x = year, y = N_obs, label = Period),
            inherit.aes = FALSE, color = "black", size = 2.5, vjust = -0.8) +
  geom_point(data = bird_obs,
             aes(x = year, y = N_scaled),
             inherit.aes = FALSE, color = "#e6550d", size = 3, shape = 16) +
  geom_line(data = bird_obs,
            aes(x = year, y = N_scaled),
            inherit.aes = FALSE, color = "#e6550d", linewidth = 0.5, linetype = "dotted") +
  geom_text(data = bird_obs,
            aes(x = year, y = N_scaled, label = season),
            inherit.aes = FALSE, color = "#e6550d", size = 2.5, vjust = -0.8) +
  geom_point(data = milestones_df,
             aes(x = year, y = N),
             inherit.aes = FALSE, color = "#253494", size = 2.5) +
  geom_text(data = milestones_df,
            aes(x = year, y = N,
                label = paste0(label, "\n(yr ", year, ")")),
            inherit.aes = FALSE,
            color = "#253494", size = 2.8, hjust = -0.15, vjust = 0.5) +
  scale_color_manual(values = c("No mitigation (baseline)"      = "#d73027",
                                 "Immediate 3/3 (100% coverage)" = "#253494")) +
  scale_linetype_manual(values = c("No mitigation (baseline)"      = "dashed",
                                    "Immediate 3/3 (100% coverage)" = "solid")) +
  scale_x_continuous(breaks = seq(-10, years, 10), limits = c(-10, years)) +
  scale_y_continuous(labels = scales::comma,
                     breaks = scales::pretty_breaks(n = 6),
                     sec.axis = sec_axis(~ . / pairs_scale,
                                         name = "Breeding pairs (Bird Island)",
                                         labels = scales::comma)) +
  labs(x = "Years since 1990 baseline",
       y = "Population (N individuals)",
       color = NULL, linetype = NULL,
       title = paste0("Population projection vs observed breeding pairs — ",
                      target_period, " baseline"),
       subtitle = paste0("\u03BB (no mitigation) = ", round(lambda_base, 5),
                         ";  \u03BB (full 3/3) = ", round(lambda_3of3, 5),
                         ";  N\u2080 = ", scales::comma(round(N0)), " birds")) +
  themeo +
  theme(legend.position = "bottom")
p_part1
```

**plot Part 1 (long run) — breeding pairs, `r yearslong`-year horizon**
```{r plot-part1-long, fig.width = 9, fig.height = 5}
milestones_long_df <- tibble(
  year  = c(milestone_10_long, milestone_25_long, milestone_50_long, milestone_dbl_long),
  label = c("+10%", "+25%", "+50%", "\u00D72"),
  N     = c(N0 * 1.10, N0 * 1.25, N0 * 1.50, N0 * 2.0)
) %>%
  filter(!is.na(year), year <= yearslong)

p_part1_long <- ggplot(proj_part1_long, aes(x = year, y = N,
                                             color = scenario, linetype = scenario)) +
  geom_line(linewidth = 0.9) +
  geom_hline(yintercept = N0, color = "grey50", linetype = "dashed", linewidth = 0.4) +
  geom_point(data = bird_obs,
             aes(x = year, y = N_scaled),
             inherit.aes = FALSE, color = "#e6550d", size = 3, shape = 16) +
  geom_text(data = bird_obs,
            aes(x = year, y = N_scaled, label = season),
            inherit.aes = FALSE, color = "#e6550d", size = 2.5, vjust = -0.8) +
  geom_point(data = milestones_long_df,
             aes(x = year, y = N),
             inherit.aes = FALSE, color = "#253494", size = 2.5) +
  geom_text(data = milestones_long_df,
            aes(x = year, y = N,
                label = paste0(label, "\n(yr ", year, ")")),
            inherit.aes = FALSE,
            color = "#253494", size = 2.8, hjust = -0.15, vjust = 0.5) +
  scale_color_manual(values = c("No mitigation (baseline)"      = "#d73027",
                                 "Immediate 3/3 (100% coverage)" = "#253494")) +
  scale_linetype_manual(values = c("No mitigation (baseline)"      = "dashed",
                                    "Immediate 3/3 (100% coverage)" = "solid")) +
  scale_x_continuous(breaks = seq(0, yearslong, 100), limits = c(-10, yearslong)) +
  scale_y_continuous(labels = scales::comma,
                     breaks = scales::pretty_breaks(n = 6),
                     sec.axis = sec_axis(~ . / pairs_scale,
                                         name = "Breeding pairs (Bird Island)",
                                         labels = scales::comma)) +
  labs(x = paste0("Years since 1990 baseline (", yearslong, "-year horizon)"),
       y = "Population (N individuals)",
       color = NULL, linetype = NULL,
       title = paste0("Long-run population projection vs observed breeding pairs — ",
                      target_period, " baseline"),
       subtitle = paste0("\u03BB (no mitigation) = ", round(lambda_base, 5),
                         ";  \u03BB (full 3/3) = ", round(lambda_3of3, 5),
                         ";  N\u2080 = ", scales::comma(round(N0)), " birds")) +
  themeo +
  theme(legend.position = "bottom")
p_part1_long
```

**plot Part 1b — projection vs dem_props age class totals**
```{r plot-part1-obs, fig.width = 9, fig.height = 5}
# Adults = SB + FB + NB;  Immatures = IMM + J2+3
# Secondary y-axis scaled to breeding pairs equivalent using same pairs_scale
p_part1_obs <- ggplot(proj_part1, aes(x = year, y = N,
                                       color = scenario, linetype = scenario)) +
  geom_line(linewidth = 0.9) +
  geom_hline(yintercept = N0, color = "grey50", linetype = "dashed", linewidth = 0.4) +
  geom_line(data = dem_obs_split,
            aes(x = year, y = N_obs, group = group, linetype = group),
            inherit.aes = FALSE, color = "black", linewidth = 0.7) +
  geom_point(data = dem_obs_split,
             aes(x = year, y = N_obs, shape = group),
             inherit.aes = FALSE, color = "black", size = 3) +
  geom_point(data = dem_obs,
             aes(x = year, y = N_obs),
             inherit.aes = FALSE, color = "black", size = 3, shape = 17) +
  geom_text(data = dem_obs,
            aes(x = year, y = N_obs, label = Period),
            inherit.aes = FALSE, color = "black", size = 2.5, vjust = -0.8) +
  scale_color_manual(values = c("No mitigation (baseline)"      = "#d73027",
                                 "Immediate 3/3 (100% coverage)" = "#253494")) +
  scale_linetype_manual(values = c("No mitigation (baseline)"      = "dashed",
                                    "Immediate 3/3 (100% coverage)" = "solid",
                                    "Adults"                        = "solid",
                                    "Immatures"                     = "dashed"),
                        guide = "none") +
  scale_shape_manual(values = c("Adults" = 16, "Immatures" = 1),
                     name = "Observed age class") +
  scale_x_continuous(breaks = seq(0, years, 10), limits = c(0, years)) +
  scale_y_continuous(labels = scales::comma,
                     breaks = scales::pretty_breaks(n = 6),
                     sec.axis = sec_axis(~ . / N0 * 100,
                                         name = "% of 1990\u20131994 baseline",
                                         labels = function(x) paste0(round(x), "%"))) +
  labs(x = "Years since 1990 baseline",
       y = "Population (N individuals)",
       color = NULL,
       title = paste0("Population projection vs observed age class totals — ",
                      target_period, " baseline"),
       subtitle = "Black triangles = total; filled circles = adults; open circles = immatures (Clay 2019)") +
  themeo +
  theme(legend.position = "bottom")
p_part1_obs
```

**plot Part 1b (long run) — projection vs dem_props age class totals, `r yearslong`-year horizon**
```{r plot-part1-obs-long, fig.width = 9, fig.height = 5}
p_part1_obs_long <- ggplot(proj_part1_long, aes(x = year, y = N,
                                                  color = scenario, linetype = scenario)) +
  geom_line(linewidth = 0.9) +
  geom_hline(yintercept = N0, color = "grey50", linetype = "dashed", linewidth = 0.4) +
  geom_line(data = dem_obs_split,
            aes(x = year, y = N_obs, group = group, linetype = group),
            inherit.aes = FALSE, color = "black", linewidth = 0.7) +
  geom_point(data = dem_obs_split,
             aes(x = year, y = N_obs, shape = group),
             inherit.aes = FALSE, color = "black", size = 3) +
  geom_point(data = dem_obs,
             aes(x = year, y = N_obs),
             inherit.aes = FALSE, color = "black", size = 3, shape = 17) +
  geom_text(data = dem_obs,
            aes(x = year, y = N_obs, label = Period),
            inherit.aes = FALSE, color = "black", size = 2.5, vjust = -0.8) +
  scale_color_manual(values = c("No mitigation (baseline)"      = "#d73027",
                                 "Immediate 3/3 (100% coverage)" = "#253494")) +
  scale_linetype_manual(values = c("No mitigation (baseline)"      = "dashed",
                                    "Immediate 3/3 (100% coverage)" = "solid",
                                    "Adults"                        = "solid",
                                    "Immatures"                     = "dashed"),
                        guide = "none") +
  scale_shape_manual(values = c("Adults" = 16, "Immatures" = 1),
                     name = "Observed age class") +
  scale_x_continuous(breaks = seq(0, yearslong, 100), limits = c(0, yearslong)) +
  scale_y_continuous(labels = scales::comma,
                     breaks = scales::pretty_breaks(n = 6),
                     sec.axis = sec_axis(~ . / N0 * 100,
                                         name = "% of 1990\u20131994 baseline",
                                         labels = function(x) paste0(round(x), "%"))) +
  labs(x = paste0("Years since 1990 baseline (", yearslong, "-year horizon)"),
       y = "Population (N individuals)",
       color = NULL,
       title = paste0("Long-run projection vs observed age class totals — ",
                      target_period, " baseline"),
       subtitle = "Black triangles = total; filled circles = adults; open circles = immatures (Clay 2019)") +
  themeo +
  theme(legend.position = "bottom")
p_part1_obs_long
```

---

*PART 2 — STAGE-STRUCTURED MATRIX PROJECTION*

Full matrix projection tracking all 11 stages. Each time step uses N(t+1) = A %*% N(t),
where A is the 11x11 projection matrix. Both scenarios start from the same initial stage
vector (SSD scaled to N0). Total N should match Part 1 exactly — used as validation.

**set up stage storage and project**
```{r part2-project}
stage_names <- c("J2", "J3", "Imm4", "Imm5", "PB", "IS", "IF", "ES", "EF", "ENB", "Sabb")

# Storage: (years+1) rows x 11 cols, one matrix per scenario
mat_base  <- matrix(0, nrow = years + 1, ncol = 11)
mat_mitig <- matrix(0, nrow = years + 1, ncol = 11)
colnames(mat_base)  <- stage_names
colnames(mat_mitig) <- stage_names

# Initialize year 0 at stable stage distribution scaled to N0
mat_base[1, ]  <- n0_vec
mat_mitig[1, ] <- n0_vec

# Matrix projection: N(t+1) = A %*% N(t)
for (k in 1:years) {
  mat_base[k + 1, ]  <- m_base  %*% mat_base[k, ]
  mat_mitig[k + 1, ] <- m_3of3  %*% mat_mitig[k, ]
}

# Convert to long format with broad stage groupings for plotting
to_long <- function(mat, scenario_label) {
  as.data.frame(mat) %>%
    mutate(year = 0:years, scenario = scenario_label) %>%
    pivot_longer(cols = all_of(stage_names), names_to = "stage", values_to = "N") %>%
    mutate(
      stage = factor(stage, levels = stage_names),
      group = case_when(
        stage %in% c("J2", "J3")              ~ "Juveniles",
        stage %in% c("Imm4", "Imm5", "PB")   ~ "Immatures / pre-breeders",
        stage %in% c("IS", "IF", "ES", "EF") ~ "Breeders",
        stage %in% c("ENB", "Sabb")          ~ "Non-breeders"
      ),
      group = factor(group, levels = c("Juveniles", "Immatures / pre-breeders",
                                        "Breeders", "Non-breeders"))
    )
}

proj_part2 <- bind_rows(
  to_long(mat_base,  "No mitigation (baseline)"),
  to_long(mat_mitig, "Immediate 3/3 (100% coverage)")
) %>%
  mutate(scenario = factor(scenario,
                            levels = c("No mitigation (baseline)",
                                       "Immediate 3/3 (100% coverage)")))

# Total N by year and scenario — should match Part 1 scalar projection
proj_part2_total <- proj_part2 %>%
  group_by(year, scenario) %>%
  summarise(N_total = sum(N), .groups = "drop")
```

**plot Part 2a — total N validation against Part 1**
```{r plot-part2-total, fig.width = 9, fig.height = 5}
p_part2_total <- ggplot(proj_part2_total, aes(x = year, y = N_total,
                                               color = scenario, linetype = scenario)) +
  geom_line(linewidth = 0.9) +
  geom_line(data = proj_part1, aes(x = year, y = N, group = scenario),
            linewidth = 0.4, linetype = "dotted", color = "black", alpha = 0.4,
            inherit.aes = FALSE) +
  geom_hline(yintercept = N0, color = "grey50", linetype = "dashed", linewidth = 0.4) +
  scale_color_manual(values = c("No mitigation (baseline)"      = "#d73027",
                                 "Immediate 3/3 (100% coverage)" = "#253494")) +
  scale_linetype_manual(values = c("No mitigation (baseline)"      = "dashed",
                                    "Immediate 3/3 (100% coverage)" = "solid")) +
  scale_x_continuous(breaks = seq(0, years, 10), limits = c(0, years)) +
  scale_y_continuous(labels = scales::comma, breaks = scales::pretty_breaks(n = 6)) +
  labs(x = "Years since 1990 baseline",
       y = "Total population (N individuals)",
       color = NULL, linetype = NULL,
       title = paste0("Stage-structured projection — total N vs Part 1 scalar (dotted black) — ",
                      target_period, " baseline"),
       subtitle = "Solid/dashed colored lines = matrix projection; dotted black = Part 1 scalar; should overlap exactly") +
  themeo +
  theme(legend.position = "bottom")
p_part2_total
```

**plot Part 2b — stage group trajectories by scenario**
```{r plot-part2-stages, fig.width = 10, fig.height = 6}
# Sum N within broad stage groups for each year and scenario
proj_part2_groups <- proj_part2 %>%
  group_by(year, scenario, group) %>%
  summarise(N = sum(N), .groups = "drop")

p_part2_stages <- ggplot(proj_part2_groups, aes(x = year, y = N,
                                                  color = group, group = group)) +
  geom_line(linewidth = 0.8) +
  geom_hline(yintercept = N0, color = "grey50", linetype = "dashed", linewidth = 0.3) +
  facet_wrap(~ scenario, ncol = 2) +
  scale_color_brewer(palette = "Dark2", name = "Stage group") +
  scale_x_continuous(breaks = seq(0, years, 20)) +
  scale_y_continuous(labels = scales::comma, breaks = scales::pretty_breaks(n = 5)) +
  labs(x = "Years since 1990 baseline",
       y = "N individuals",
       title = paste0("Stage-structured projection by group — ", target_period, " baseline"),
       subtitle = "Each panel is one scenario; colors = broad stage groups") +
  themeo +
  theme(legend.position = "bottom")
p_part2_stages
```

**plot Part 2c — proportional stage composition over time**
```{r plot-part2-composition, fig.width = 10, fig.height = 5}
proj_part2_pct <- proj_part2_groups %>%
  group_by(year, scenario) %>%
  mutate(pct = N / sum(N) * 100) %>%
  ungroup()

p_part2_composition <- ggplot(proj_part2_pct, aes(x = year, y = pct,
                                                    fill = group)) +
  geom_area(alpha = 0.85) +
  facet_wrap(~ scenario, ncol = 2) +
  scale_fill_brewer(palette = "Dark2", name = "Stage group") +
  scale_x_continuous(breaks = seq(0, years, 20)) +
  scale_y_continuous(labels = function(x) paste0(round(x), "%")) +
  labs(x = "Years since 1990 baseline",
       y = "% of total population",
       title = paste0("Stage composition over time — ", target_period, " baseline"),
       subtitle = "Stacked area; each panel is one scenario") +
  themeo +
  theme(legend.position = "bottom")
p_part2_composition
```

---

*PART 3 — 5-YEAR MITIGATION PHASE-IN*

Survival improves in equal steps over 5 years. Each year uses a new matrix until year 6,
when the full 3/3 matrix is used for the remainder of the projection.
Improvement per step = (ns_* - s_*) / 5 for each stage-specific survival rate.

**compute phase-in survival rates**
```{r part3-survival-rates}
# Phase step 1: 20% of total improvement applied
s_Juv_p1 <- s_Juv + (1/5) * (ns_Juv - s_Juv)
s_Imm_p1 <- s_Imm + (1/5) * (ns_Imm - s_Imm)
s_EF_p1  <- s_EF  + (1/5) * (ns_EF  - s_EF)
s_ES_p1  <- s_ES  + (1/5) * (ns_ES  - s_ES)
s_ENB_p1 <- s_ENB + (1/5) * (ns_ENB - s_ENB)
s_IF_p1  <- s_IF  + (1/5) * (ns_IF  - s_IF)
s_IS_p1  <- s_IS  + (1/5) * (ns_IS  - s_IS)
s_PF_p1  <- s_PF  + (1/5) * (ns_PF  - s_PF)

# Phase step 2: 40%
s_Juv_p2 <- s_Juv + (2/5) * (ns_Juv - s_Juv)
s_Imm_p2 <- s_Imm + (2/5) * (ns_Imm - s_Imm)
s_EF_p2  <- s_EF  + (2/5) * (ns_EF  - s_EF)
s_ES_p2  <- s_ES  + (2/5) * (ns_ES  - s_ES)
s_ENB_p2 <- s_ENB + (2/5) * (ns_ENB - s_ENB)
s_IF_p2  <- s_IF  + (2/5) * (ns_IF  - s_IF)
s_IS_p2  <- s_IS  + (2/5) * (ns_IS  - s_IS)
s_PF_p2  <- s_PF  + (2/5) * (ns_PF  - s_PF)

# Phase step 3: 60%
s_Juv_p3 <- s_Juv + (3/5) * (ns_Juv - s_Juv)
s_Imm_p3 <- s_Imm + (3/5) * (ns_Imm - s_Imm)
s_EF_p3  <- s_EF  + (3/5) * (ns_EF  - s_EF)
s_ES_p3  <- s_ES  + (3/5) * (ns_ES  - s_ES)
s_ENB_p3 <- s_ENB + (3/5) * (ns_ENB - s_ENB)
s_IF_p3  <- s_IF  + (3/5) * (ns_IF  - s_IF)
s_IS_p3  <- s_IS  + (3/5) * (ns_IS  - s_IS)
s_PF_p3  <- s_PF  + (3/5) * (ns_PF  - s_PF)

# Phase step 4: 80%
s_Juv_p4 <- s_Juv + (4/5) * (ns_Juv - s_Juv)
s_Imm_p4 <- s_Imm + (4/5) * (ns_Imm - s_Imm)
s_EF_p4  <- s_EF  + (4/5) * (ns_EF  - s_EF)
s_ES_p4  <- s_ES  + (4/5) * (ns_ES  - s_ES)
s_ENB_p4 <- s_ENB + (4/5) * (ns_ENB - s_ENB)
s_IF_p4  <- s_IF  + (4/5) * (ns_IF  - s_IF)
s_IS_p4  <- s_IS  + (4/5) * (ns_IS  - s_IS)
s_PF_p4  <- s_PF  + (4/5) * (ns_PF  - s_PF)

# Phase step 5: 100% — should equal ns_* values (same as m_3of3)
s_Juv_p5 <- s_Juv + (5/5) * (ns_Juv - s_Juv)
s_Imm_p5 <- s_Imm + (5/5) * (ns_Imm - s_Imm)
s_EF_p5  <- s_EF  + (5/5) * (ns_EF  - s_EF)
s_ES_p5  <- s_ES  + (5/5) * (ns_ES  - s_ES)
s_ENB_p5 <- s_ENB + (5/5) * (ns_ENB - s_ENB)
s_IF_p5  <- s_IF  + (5/5) * (ns_IF  - s_IF)
s_IS_p5  <- s_IS  + (5/5) * (ns_IS  - s_IS)
s_PF_p5  <- s_PF  + (5/5) * (ns_PF  - s_PF)

# Print all phase survival rates for transparency
tibble(
  stage = c("Juv", "Imm", "EF", "ES", "ENB", "IF", "IS", "PF"),
  baseline  = c(s_Juv, s_Imm, s_EF, s_ES, s_ENB, s_IF, s_IS, s_PF),
  phase1    = c(s_Juv_p1, s_Imm_p1, s_EF_p1, s_ES_p1, s_ENB_p1, s_IF_p1, s_IS_p1, s_PF_p1),
  phase2    = c(s_Juv_p2, s_Imm_p2, s_EF_p2, s_ES_p2, s_ENB_p2, s_IF_p2, s_IS_p2, s_PF_p2),
  phase3    = c(s_Juv_p3, s_Imm_p3, s_EF_p3, s_ES_p3, s_ENB_p3, s_IF_p3, s_IS_p3, s_PF_p3),
  phase4    = c(s_Juv_p4, s_Imm_p4, s_EF_p4, s_ES_p4, s_ENB_p4, s_IF_p4, s_IS_p4, s_PF_p4),
  full_3of3 = c(s_Juv_p5, s_Imm_p5, s_EF_p5, s_ES_p5, s_ENB_p5, s_IF_p5, s_IS_p5, s_PF_p5)
) %>% print()
```

**build phase-in matrices**
```{r part3-matrices}
# Stages: J2 J3 Imm4 Imm5 PB IS IF ES EF ENB Sabb

m_phase1 <- matrix(data = c(
  0,        0,        0,        0,        0,                                                0.5*s_Juv_p1,                  0,                           0.5*s_Juv_p1,                  0,                           0,                               0,
  s_Juv_p1, 0,        0,        0,        0,                                                0,                             0,                           0,                             0,                           0,                               0,
  0,        s_Juv_p1, 0,        0,        0,                                                0,                             0,                           0,                             0,                           0,                               0,
  0,        0,        s_Imm_p1, 0,        0,                                                0,                             0,                           0,                             0,                           0,                               0,
  0,        0,        0,        s_Imm_p1, s_Imm_p1*r_Imm*(1-b_Imm)+s_Imm_p1*(1-r_Imm),    0,                             0,                           0,                             0,                           0,                               0,
  0,        0,        0,        0,        s_Imm_p1*r_Imm*b_Imm*k_Imm,                       0,                             0,                           0,                             0,                           0,                               0,
  0,        0,        0,        0,        s_Imm_p1*r_Imm*b_Imm*(1-k_Imm),                   0,                             0,                           0,                             0,                           0,                               0,
  0,        0,        0,        0,        0,                                                s_IS_p1*r_IS*b_IS*k_IS,        s_IF_p1*r_IF*b_IF*k_IF,      s_ES_p1*r_ES*b_ES*k_ES,        s_EF_p1*r_EF*b_EF*k_EF,      s_ENB_p1*r_ENB*b_ENB*k_ENB,      s_PF_p1*r_PF*b_PF*k_PF,
  0,        0,        0,        0,        0,                                                s_IS_p1*r_IS*b_IS*(1-k_IS),    s_IF_p1*r_IF*b_IF*(1-k_IF),  s_ES_p1*r_ES*b_ES*(1-k_ES),    s_EF_p1*r_EF*b_EF*(1-k_EF),  s_ENB_p1*r_ENB*b_ENB*(1-k_ENB),  s_PF_p1*r_PF*b_PF*(1-k_PF),
  0,        0,        0,        0,        0,                                                s_IS_p1*r_IS*(1-b_IS),         s_IF_p1*r_IF*(1-b_IF),       s_ES_p1*r_ES*(1-b_ES),         s_EF_p1*r_EF*(1-b_EF),       s_ENB_p1*r_ENB*(1-b_ENB),        s_PF_p1*r_PF*(1-b_PF),
  0,        0,        0,        0,        0,                                                s_IS_p1*(1-r_IS),              s_IF_p1*(1-r_IF),            s_ES_p1*(1-r_ES),              s_EF_p1*(1-r_EF),            s_ENB_p1*(1-r_ENB),              s_PF_p1*(1-r_PF)
), nrow = 11, byrow = TRUE)

m_phase2 <- matrix(data = c(
  0,        0,        0,        0,        0,                                                0.5*s_Juv_p2,                  0,                           0.5*s_Juv_p2,                  0,                           0,                               0,
  s_Juv_p2, 0,        0,        0,        0,                                                0,                             0,                           0,                             0,                           0,                               0,
  0,        s_Juv_p2, 0,        0,        0,                                                0,                             0,                           0,                             0,                           0,                               0,
  0,        0,        s_Imm_p2, 0,        0,                                                0,                             0,                           0,                             0,                           0,                               0,
  0,        0,        0,        s_Imm_p2, s_Imm_p2*r_Imm*(1-b_Imm)+s_Imm_p2*(1-r_Imm),    0,                             0,                           0,                             0,                           0,                               0,
  0,        0,        0,        0,        s_Imm_p2*r_Imm*b_Imm*k_Imm,                       0,                             0,                           0,                             0,                           0,                               0,
  0,        0,        0,        0,        s_Imm_p2*r_Imm*b_Imm*(1-k_Imm),                   0,                             0,                           0,                             0,                           0,                               0,
  0,        0,        0,        0,        0,                                                s_IS_p2*r_IS*b_IS*k_IS,        s_IF_p2*r_IF*b_IF*k_IF,      s_ES_p2*r_ES*b_ES*k_ES,        s_EF_p2*r_EF*b_EF*k_EF,      s_ENB_p2*r_ENB*b_ENB*k_ENB,      s_PF_p2*r_PF*b_PF*k_PF,
  0,        0,        0,        0,        0,                                                s_IS_p2*r_IS*b_IS*(1-k_IS),    s_IF_p2*r_IF*b_IF*(1-k_IF),  s_ES_p2*r_ES*b_ES*(1-k_ES),    s_EF_p2*r_EF*b_EF*(1-k_EF),  s_ENB_p2*r_ENB*b_ENB*(1-k_ENB),  s_PF_p2*r_PF*b_PF*(1-k_PF),
  0,        0,        0,        0,        0,                                                s_IS_p2*r_IS*(1-b_IS),         s_IF_p2*r_IF*(1-b_IF),       s_ES_p2*r_ES*(1-b_ES),         s_EF_p2*r_EF*(1-b_EF),       s_ENB_p2*r_ENB*(1-b_ENB),        s_PF_p2*r_PF*(1-b_PF),
  0,        0,        0,        0,        0,                                                s_IS_p2*(1-r_IS),              s_IF_p2*(1-r_IF),            s_ES_p2*(1-r_ES),              s_EF_p2*(1-r_EF),            s_ENB_p2*(1-r_ENB),              s_PF_p2*(1-r_PF)
), nrow = 11, byrow = TRUE)

m_phase3 <- matrix(data = c(
  0,        0,        0,        0,        0,                                                0.5*s_Juv_p3,                  0,                           0.5*s_Juv_p3,                  0,                           0,                               0,
  s_Juv_p3, 0,        0,        0,        0,                                                0,                             0,                           0,                             0,                           0,                               0,
  0,        s_Juv_p3, 0,        0,        0,                                                0,                             0,                           0,                             0,                           0,                               0,
  0,        0,        s_Imm_p3, 0,        0,                                                0,                             0,                           0,                             0,                           0,                               0,
  0,        0,        0,        s_Imm_p3, s_Imm_p3*r_Imm*(1-b_Imm)+s_Imm_p3*(1-r_Imm),    0,                             0,                           0,                             0,                           0,                               0,
  0,        0,        0,        0,        s_Imm_p3*r_Imm*b_Imm*k_Imm,                       0,                             0,                           0,                             0,                           0,                               0,
  0,        0,        0,        0,        s_Imm_p3*r_Imm*b_Imm*(1-k_Imm),                   0,                             0,                           0,                             0,                           0,                               0,
  0,        0,        0,        0,        0,                                                s_IS_p3*r_IS*b_IS*k_IS,        s_IF_p3*r_IF*b_IF*k_IF,      s_ES_p3*r_ES*b_ES*k_ES,        s_EF_p3*r_EF*b_EF*k_EF,      s_ENB_p3*r_ENB*b_ENB*k_ENB,      s_PF_p3*r_PF*b_PF*k_PF,
  0,        0,        0,        0,        0,                                                s_IS_p3*r_IS*b_IS*(1-k_IS),    s_IF_p3*r_IF*b_IF*(1-k_IF),  s_ES_p3*r_ES*b_ES*(1-k_ES),    s_EF_p3*r_EF*b_EF*(1-k_EF),  s_ENB_p3*r_ENB*b_ENB*(1-k_ENB),  s_PF_p3*r_PF*b_PF*(1-k_PF),
  0,        0,        0,        0,        0,                                                s_IS_p3*r_IS*(1-b_IS),         s_IF_p3*r_IF*(1-b_IF),       s_ES_p3*r_ES*(1-b_ES),         s_EF_p3*r_EF*(1-b_EF),       s_ENB_p3*r_ENB*(1-b_ENB),        s_PF_p3*r_PF*(1-b_PF),
  0,        0,        0,        0,        0,                                                s_IS_p3*(1-r_IS),              s_IF_p3*(1-r_IF),            s_ES_p3*(1-r_ES),              s_EF_p3*(1-r_EF),            s_ENB_p3*(1-r_ENB),              s_PF_p3*(1-r_PF)
), nrow = 11, byrow = TRUE)

m_phase4 <- matrix(data = c(
  0,        0,        0,        0,        0,                                                0.5*s_Juv_p4,                  0,                           0.5*s_Juv_p4,                  0,                           0,                               0,
  s_Juv_p4, 0,        0,        0,        0,                                                0,                             0,                           0,                             0,                           0,                               0,
  0,        s_Juv_p4, 0,        0,        0,                                                0,                             0,                           0,                             0,                           0,                               0,
  0,        0,        s_Imm_p4, 0,        0,                                                0,                             0,                           0,                             0,                           0,                               0,
  0,        0,        0,        s_Imm_p4, s_Imm_p4*r_Imm*(1-b_Imm)+s_Imm_p4*(1-r_Imm),    0,                             0,                           0,                             0,                           0,                               0,
  0,        0,        0,        0,        s_Imm_p4*r_Imm*b_Imm*k_Imm,                       0,                             0,                           0,                             0,                           0,                               0,
  0,        0,        0,        0,        s_Imm_p4*r_Imm*b_Imm*(1-k_Imm),                   0,                             0,                           0,                             0,                           0,                               0,
  0,        0,        0,        0,        0,                                                s_IS_p4*r_IS*b_IS*k_IS,        s_IF_p4*r_IF*b_IF*k_IF,      s_ES_p4*r_ES*b_ES*k_ES,        s_EF_p4*r_EF*b_EF*k_EF,      s_ENB_p4*r_ENB*b_ENB*k_ENB,      s_PF_p4*r_PF*b_PF*k_PF,
  0,        0,        0,        0,        0,                                                s_IS_p4*r_IS*b_IS*(1-k_IS),    s_IF_p4*r_IF*b_IF*(1-k_IF),  s_ES_p4*r_ES*b_ES*(1-k_ES),    s_EF_p4*r_EF*b_EF*(1-k_EF),  s_ENB_p4*r_ENB*b_ENB*(1-k_ENB),  s_PF_p4*r_PF*b_PF*(1-k_PF),
  0,        0,        0,        0,        0,                                                s_IS_p4*r_IS*(1-b_IS),         s_IF_p4*r_IF*(1-b_IF),       s_ES_p4*r_ES*(1-b_ES),         s_EF_p4*r_EF*(1-b_EF),       s_ENB_p4*r_ENB*(1-b_ENB),        s_PF_p4*r_PF*(1-b_PF),
  0,        0,        0,        0,        0,                                                s_IS_p4*(1-r_IS),              s_IF_p4*(1-r_IF),            s_ES_p4*(1-r_ES),              s_EF_p4*(1-r_EF),            s_ENB_p4*(1-r_ENB),              s_PF_p4*(1-r_PF)
), nrow = 11, byrow = TRUE)

# Phase 5 = full 3/3 (s_*_p5 == ns_*); should be identical to m_3of3
m_phase5 <- matrix(data = c(
  0,        0,        0,        0,        0,                                                0.5*s_Juv_p5,                  0,                           0.5*s_Juv_p5,                  0,                           0,                               0,
  s_Juv_p5, 0,        0,        0,        0,                                                0,                             0,                           0,                             0,                           0,                               0,
  0,        s_Juv_p5, 0,        0,        0,                                                0,                             0,                           0,                             0,                           0,                               0,
  0,        0,        s_Imm_p5, 0,        0,                                                0,                             0,                           0,                             0,                           0,                               0,
  0,        0,        0,        s_Imm_p5, s_Imm_p5*r_Imm*(1-b_Imm)+s_Imm_p5*(1-r_Imm),    0,                             0,                           0,                             0,                           0,                               0,
  0,        0,        0,        0,        s_Imm_p5*r_Imm*b_Imm*k_Imm,                       0,                             0,                           0,                             0,                           0,                               0,
  0,        0,        0,        0,        s_Imm_p5*r_Imm*b_Imm*(1-k_Imm),                   0,                             0,                           0,                             0,                           0,                               0,
  0,        0,        0,        0,        0,                                                s_IS_p5*r_IS*b_IS*k_IS,        s_IF_p5*r_IF*b_IF*k_IF,      s_ES_p5*r_ES*b_ES*k_ES,        s_EF_p5*r_EF*b_EF*k_EF,      s_ENB_p5*r_ENB*b_ENB*k_ENB,      s_PF_p5*r_PF*b_PF*k_PF,
  0,        0,        0,        0,        0,                                                s_IS_p5*r_IS*b_IS*(1-k_IS),    s_IF_p5*r_IF*b_IF*(1-k_IF),  s_ES_p5*r_ES*b_ES*(1-k_ES),    s_EF_p5*r_EF*b_EF*(1-k_EF),  s_ENB_p5*r_ENB*b_ENB*(1-k_ENB),  s_PF_p5*r_PF*b_PF*(1-k_PF),
  0,        0,        0,        0,        0,                                                s_IS_p5*r_IS*(1-b_IS),         s_IF_p5*r_IF*(1-b_IF),       s_ES_p5*r_ES*(1-b_ES),         s_EF_p5*r_EF*(1-b_EF),       s_ENB_p5*r_ENB*(1-b_ENB),        s_PF_p5*r_PF*(1-b_PF),
  0,        0,        0,        0,        0,                                                s_IS_p5*(1-r_IS),              s_IF_p5*(1-r_IF),            s_ES_p5*(1-r_ES),              s_EF_p5*(1-r_EF),            s_ENB_p5*(1-r_ENB),              s_PF_p5*(1-r_PF)
), nrow = 11, byrow = TRUE)
```

**project with phase-in matrices**
```{r part3-project}
mat_phasein <- matrix(0, nrow = years + 1, ncol = 11)
colnames(mat_phasein) <- stage_names
mat_phasein[1, ] <- n0_vec

# Years 1-5: use phase matrices; year 6 onward: use m_3of3 (full mitigation)
phase_list <- list(m_phase1, m_phase2, m_phase3, m_phase4, m_phase5)

for (k in 1:years) {
  A <- if (k <= 5) phase_list[[k]] else m_3of3
  mat_phasein[k + 1, ] <- A %*% mat_phasein[k, ]
}

proj_part3_total <- as.data.frame(mat_phasein) %>%
  mutate(year     = 0:years,
         N        = rowSums(across(all_of(stage_names))),
         scenario = "5-year phase-in") %>%
  select(year, N, scenario)
```

**plot Part 3 — phase-in vs immediate vs no mitigation**
```{r plot-part3, fig.width = 9, fig.height = 5}
proj_compare <- bind_rows(proj_part1, proj_part3_total) %>%
  mutate(scenario = factor(scenario,
                            levels = c("No mitigation (baseline)",
                                       "5-year phase-in",
                                       "Immediate 3/3 (100% coverage)")))

p_part3 <- ggplot(proj_compare, aes(x = year, y = N,
                                     color = scenario, linetype = scenario)) +
  geom_line(linewidth = 0.9) +
  geom_hline(yintercept = N0, color = "grey50", linetype = "dashed", linewidth = 0.4) +
  scale_color_manual(values = c("No mitigation (baseline)"      = "#d73027",
                                 "5-year phase-in"               = "#1a9641",
                                 "Immediate 3/3 (100% coverage)" = "#253494")) +
  scale_linetype_manual(values = c("No mitigation (baseline)"      = "dashed",
                                    "5-year phase-in"               = "solid",
                                    "Immediate 3/3 (100% coverage)" = "solid")) +
  scale_x_continuous(breaks = seq(0, years, 10), limits = c(0, years)) +
  scale_y_continuous(labels = scales::comma, breaks = scales::pretty_breaks(n = 6)) +
  labs(x = "Years since 1990 baseline",
       y = "Population (N individuals)",
       color = NULL, linetype = NULL,
       title = paste0("Population projection — immediate vs 5-year phase-in — ",
                      target_period, " baseline"),
       subtitle = paste0("N\u2080 = ", scales::comma(round(N0)),
                         ";  green = 5-yr phase-in;  blue = immediate 3/3;  red = no mitigation")) +
  themeo +
  theme(legend.position = "bottom")
p_part3
```

---

*PART 4 — 10-YEAR PHASE-IN*

**10-year phase-in survival rates**
```{r part4-10yr-survival-rates}
# Each step adds 1/10 of total improvement
s_Juv_10y1  <- s_Juv + (1/10)  * (ns_Juv - s_Juv);  s_Imm_10y1  <- s_Imm + (1/10)  * (ns_Imm - s_Imm)
s_EF_10y1   <- s_EF  + (1/10)  * (ns_EF  - s_EF);   s_ES_10y1   <- s_ES  + (1/10)  * (ns_ES  - s_ES)
s_ENB_10y1  <- s_ENB + (1/10)  * (ns_ENB - s_ENB);  s_IF_10y1   <- s_IF  + (1/10)  * (ns_IF  - s_IF)
s_IS_10y1   <- s_IS  + (1/10)  * (ns_IS  - s_IS);   s_PF_10y1   <- s_PF  + (1/10)  * (ns_PF  - s_PF)

s_Juv_10y2  <- s_Juv + (2/10)  * (ns_Juv - s_Juv);  s_Imm_10y2  <- s_Imm + (2/10)  * (ns_Imm - s_Imm)
s_EF_10y2   <- s_EF  + (2/10)  * (ns_EF  - s_EF);   s_ES_10y2   <- s_ES  + (2/10)  * (ns_ES  - s_ES)
s_ENB_10y2  <- s_ENB + (2/10)  * (ns_ENB - s_ENB);  s_IF_10y2   <- s_IF  + (2/10)  * (ns_IF  - s_IF)
s_IS_10y2   <- s_IS  + (2/10)  * (ns_IS  - s_IS);   s_PF_10y2   <- s_PF  + (2/10)  * (ns_PF  - s_PF)

s_Juv_10y3  <- s_Juv + (3/10)  * (ns_Juv - s_Juv);  s_Imm_10y3  <- s_Imm + (3/10)  * (ns_Imm - s_Imm)
s_EF_10y3   <- s_EF  + (3/10)  * (ns_EF  - s_EF);   s_ES_10y3   <- s_ES  + (3/10)  * (ns_ES  - s_ES)
s_ENB_10y3  <- s_ENB + (3/10)  * (ns_ENB - s_ENB);  s_IF_10y3   <- s_IF  + (3/10)  * (ns_IF  - s_IF)
s_IS_10y3   <- s_IS  + (3/10)  * (ns_IS  - s_IS);   s_PF_10y3   <- s_PF  + (3/10)  * (ns_PF  - s_PF)

s_Juv_10y4  <- s_Juv + (4/10)  * (ns_Juv - s_Juv);  s_Imm_10y4  <- s_Imm + (4/10)  * (ns_Imm - s_Imm)
s_EF_10y4   <- s_EF  + (4/10)  * (ns_EF  - s_EF);   s_ES_10y4   <- s_ES  + (4/10)  * (ns_ES  - s_ES)
s_ENB_10y4  <- s_ENB + (4/10)  * (ns_ENB - s_ENB);  s_IF_10y4   <- s_IF  + (4/10)  * (ns_IF  - s_IF)
s_IS_10y4   <- s_IS  + (4/10)  * (ns_IS  - s_IS);   s_PF_10y4   <- s_PF  + (4/10)  * (ns_PF  - s_PF)

s_Juv_10y5  <- s_Juv + (5/10)  * (ns_Juv - s_Juv);  s_Imm_10y5  <- s_Imm + (5/10)  * (ns_Imm - s_Imm)
s_EF_10y5   <- s_EF  + (5/10)  * (ns_EF  - s_EF);   s_ES_10y5   <- s_ES  + (5/10)  * (ns_ES  - s_ES)
s_ENB_10y5  <- s_ENB + (5/10)  * (ns_ENB - s_ENB);  s_IF_10y5   <- s_IF  + (5/10)  * (ns_IF  - s_IF)
s_IS_10y5   <- s_IS  + (5/10)  * (ns_IS  - s_IS);   s_PF_10y5   <- s_PF  + (5/10)  * (ns_PF  - s_PF)

s_Juv_10y6  <- s_Juv + (6/10)  * (ns_Juv - s_Juv);  s_Imm_10y6  <- s_Imm + (6/10)  * (ns_Imm - s_Imm)
s_EF_10y6   <- s_EF  + (6/10)  * (ns_EF  - s_EF);   s_ES_10y6   <- s_ES  + (6/10)  * (ns_ES  - s_ES)
s_ENB_10y6  <- s_ENB + (6/10)  * (ns_ENB - s_ENB);  s_IF_10y6   <- s_IF  + (6/10)  * (ns_IF  - s_IF)
s_IS_10y6   <- s_IS  + (6/10)  * (ns_IS  - s_IS);   s_PF_10y6   <- s_PF  + (6/10)  * (ns_PF  - s_PF)

s_Juv_10y7  <- s_Juv + (7/10)  * (ns_Juv - s_Juv);  s_Imm_10y7  <- s_Imm + (7/10)  * (ns_Imm - s_Imm)
s_EF_10y7   <- s_EF  + (7/10)  * (ns_EF  - s_EF);   s_ES_10y7   <- s_ES  + (7/10)  * (ns_ES  - s_ES)
s_ENB_10y7  <- s_ENB + (7/10)  * (ns_ENB - s_ENB);  s_IF_10y7   <- s_IF  + (7/10)  * (ns_IF  - s_IF)
s_IS_10y7   <- s_IS  + (7/10)  * (ns_IS  - s_IS);   s_PF_10y7   <- s_PF  + (7/10)  * (ns_PF  - s_PF)

s_Juv_10y8  <- s_Juv + (8/10)  * (ns_Juv - s_Juv);  s_Imm_10y8  <- s_Imm + (8/10)  * (ns_Imm - s_Imm)
s_EF_10y8   <- s_EF  + (8/10)  * (ns_EF  - s_EF);   s_ES_10y8   <- s_ES  + (8/10)  * (ns_ES  - s_ES)
s_ENB_10y8  <- s_ENB + (8/10)  * (ns_ENB - s_ENB);  s_IF_10y8   <- s_IF  + (8/10)  * (ns_IF  - s_IF)
s_IS_10y8   <- s_IS  + (8/10)  * (ns_IS  - s_IS);   s_PF_10y8   <- s_PF  + (8/10)  * (ns_PF  - s_PF)

s_Juv_10y9  <- s_Juv + (9/10)  * (ns_Juv - s_Juv);  s_Imm_10y9  <- s_Imm + (9/10)  * (ns_Imm - s_Imm)
s_EF_10y9   <- s_EF  + (9/10)  * (ns_EF  - s_EF);   s_ES_10y9   <- s_ES  + (9/10)  * (ns_ES  - s_ES)
s_ENB_10y9  <- s_ENB + (9/10)  * (ns_ENB - s_ENB);  s_IF_10y9   <- s_IF  + (9/10)  * (ns_IF  - s_IF)
s_IS_10y9   <- s_IS  + (9/10)  * (ns_IS  - s_IS);   s_PF_10y9   <- s_PF  + (9/10)  * (ns_PF  - s_PF)

# Step 10: 100% — should equal ns_* (same as m_3of3)
s_Juv_10y10 <- s_Juv + (10/10) * (ns_Juv - s_Juv);  s_Imm_10y10 <- s_Imm + (10/10) * (ns_Imm - s_Imm)
s_EF_10y10  <- s_EF  + (10/10) * (ns_EF  - s_EF);   s_ES_10y10  <- s_ES  + (10/10) * (ns_ES  - s_ES)
s_ENB_10y10 <- s_ENB + (10/10) * (ns_ENB - s_ENB);  s_IF_10y10  <- s_IF  + (10/10) * (ns_IF  - s_IF)
s_IS_10y10  <- s_IS  + (10/10) * (ns_IS  - s_IS);   s_PF_10y10  <- s_PF  + (10/10) * (ns_PF  - s_PF)

tibble(
  stage     = c("Juv","Imm","EF","ES","ENB","IF","IS","PF"),
  baseline  = c(s_Juv, s_Imm, s_EF, s_ES, s_ENB, s_IF, s_IS, s_PF),
  y1=c(s_Juv_10y1,s_Imm_10y1,s_EF_10y1,s_ES_10y1,s_ENB_10y1,s_IF_10y1,s_IS_10y1,s_PF_10y1),
  y2=c(s_Juv_10y2,s_Imm_10y2,s_EF_10y2,s_ES_10y2,s_ENB_10y2,s_IF_10y2,s_IS_10y2,s_PF_10y2),
  y3=c(s_Juv_10y3,s_Imm_10y3,s_EF_10y3,s_ES_10y3,s_ENB_10y3,s_IF_10y3,s_IS_10y3,s_PF_10y3),
  y4=c(s_Juv_10y4,s_Imm_10y4,s_EF_10y4,s_ES_10y4,s_ENB_10y4,s_IF_10y4,s_IS_10y4,s_PF_10y4),
  y5=c(s_Juv_10y5,s_Imm_10y5,s_EF_10y5,s_ES_10y5,s_ENB_10y5,s_IF_10y5,s_IS_10y5,s_PF_10y5),
  y6=c(s_Juv_10y6,s_Imm_10y6,s_EF_10y6,s_ES_10y6,s_ENB_10y6,s_IF_10y6,s_IS_10y6,s_PF_10y6),
  y7=c(s_Juv_10y7,s_Imm_10y7,s_EF_10y7,s_ES_10y7,s_ENB_10y7,s_IF_10y7,s_IS_10y7,s_PF_10y7),
  y8=c(s_Juv_10y8,s_Imm_10y8,s_EF_10y8,s_ES_10y8,s_ENB_10y8,s_IF_10y8,s_IS_10y8,s_PF_10y8),
  y9=c(s_Juv_10y9,s_Imm_10y9,s_EF_10y9,s_ES_10y9,s_ENB_10y9,s_IF_10y9,s_IS_10y9,s_PF_10y9),
  full_3of3=c(s_Juv_10y10,s_Imm_10y10,s_EF_10y10,s_ES_10y10,s_ENB_10y10,s_IF_10y10,s_IS_10y10,s_PF_10y10)
) %>% print()
```

**10-year phase-in matrices**
```{r part4-10yr-matrices}
# Stages: J2 J3 Imm4 Imm5 PB IS IF ES EF ENB Sabb
m_10y1 <- matrix(data = c(
  0,0,0,0,0, 0.5*s_Juv_10y1,0,0.5*s_Juv_10y1,0,0,0,
  s_Juv_10y1,0,0,0,0, 0,0,0,0,0,0,
  0,s_Juv_10y1,0,0,0, 0,0,0,0,0,0,
  0,0,s_Imm_10y1,0,0, 0,0,0,0,0,0,
  0,0,0,s_Imm_10y1, s_Imm_10y1*r_Imm*(1-b_Imm)+s_Imm_10y1*(1-r_Imm), 0,0,0,0,0,0,
  0,0,0,0, s_Imm_10y1*r_Imm*b_Imm*k_Imm, 0,0,0,0,0,0,
  0,0,0,0, s_Imm_10y1*r_Imm*b_Imm*(1-k_Imm), 0,0,0,0,0,0,
  0,0,0,0,0, s_IS_10y1*r_IS*b_IS*k_IS, s_IF_10y1*r_IF*b_IF*k_IF, s_ES_10y1*r_ES*b_ES*k_ES, s_EF_10y1*r_EF*b_EF*k_EF, s_ENB_10y1*r_ENB*b_ENB*k_ENB, s_PF_10y1*r_PF*b_PF*k_PF,
  0,0,0,0,0, s_IS_10y1*r_IS*b_IS*(1-k_IS), s_IF_10y1*r_IF*b_IF*(1-k_IF), s_ES_10y1*r_ES*b_ES*(1-k_ES), s_EF_10y1*r_EF*b_EF*(1-k_EF), s_ENB_10y1*r_ENB*b_ENB*(1-k_ENB), s_PF_10y1*r_PF*b_PF*(1-k_PF),
  0,0,0,0,0, s_IS_10y1*r_IS*(1-b_IS), s_IF_10y1*r_IF*(1-b_IF), s_ES_10y1*r_ES*(1-b_ES), s_EF_10y1*r_EF*(1-b_EF), s_ENB_10y1*r_ENB*(1-b_ENB), s_PF_10y1*r_PF*(1-b_PF),
  0,0,0,0,0, s_IS_10y1*(1-r_IS), s_IF_10y1*(1-r_IF), s_ES_10y1*(1-r_ES), s_EF_10y1*(1-r_EF), s_ENB_10y1*(1-r_ENB), s_PF_10y1*(1-r_PF)
), nrow=11, byrow=TRUE)

m_10y2 <- matrix(data = c(
  0,0,0,0,0, 0.5*s_Juv_10y2,0,0.5*s_Juv_10y2,0,0,0,
  s_Juv_10y2,0,0,0,0, 0,0,0,0,0,0,
  0,s_Juv_10y2,0,0,0, 0,0,0,0,0,0,
  0,0,s_Imm_10y2,0,0, 0,0,0,0,0,0,
  0,0,0,s_Imm_10y2, s_Imm_10y2*r_Imm*(1-b_Imm)+s_Imm_10y2*(1-r_Imm), 0,0,0,0,0,0,
  0,0,0,0, s_Imm_10y2*r_Imm*b_Imm*k_Imm, 0,0,0,0,0,0,
  0,0,0,0, s_Imm_10y2*r_Imm*b_Imm*(1-k_Imm), 0,0,0,0,0,0,
  0,0,0,0,0, s_IS_10y2*r_IS*b_IS*k_IS, s_IF_10y2*r_IF*b_IF*k_IF, s_ES_10y2*r_ES*b_ES*k_ES, s_EF_10y2*r_EF*b_EF*k_EF, s_ENB_10y2*r_ENB*b_ENB*k_ENB, s_PF_10y2*r_PF*b_PF*k_PF,
  0,0,0,0,0, s_IS_10y2*r_IS*b_IS*(1-k_IS), s_IF_10y2*r_IF*b_IF*(1-k_IF), s_ES_10y2*r_ES*b_ES*(1-k_ES), s_EF_10y2*r_EF*b_EF*(1-k_EF), s_ENB_10y2*r_ENB*b_ENB*(1-k_ENB), s_PF_10y2*r_PF*b_PF*(1-k_PF),
  0,0,0,0,0, s_IS_10y2*r_IS*(1-b_IS), s_IF_10y2*r_IF*(1-b_IF), s_ES_10y2*r_ES*(1-b_ES), s_EF_10y2*r_EF*(1-b_EF), s_ENB_10y2*r_ENB*(1-b_ENB), s_PF_10y2*r_PF*(1-b_PF),
  0,0,0,0,0, s_IS_10y2*(1-r_IS), s_IF_10y2*(1-r_IF), s_ES_10y2*(1-r_ES), s_EF_10y2*(1-r_EF), s_ENB_10y2*(1-r_ENB), s_PF_10y2*(1-r_PF)
), nrow=11, byrow=TRUE)

m_10y3 <- matrix(data = c(
  0,0,0,0,0, 0.5*s_Juv_10y3,0,0.5*s_Juv_10y3,0,0,0,
  s_Juv_10y3,0,0,0,0, 0,0,0,0,0,0,
  0,s_Juv_10y3,0,0,0, 0,0,0,0,0,0,
  0,0,s_Imm_10y3,0,0, 0,0,0,0,0,0,
  0,0,0,s_Imm_10y3, s_Imm_10y3*r_Imm*(1-b_Imm)+s_Imm_10y3*(1-r_Imm), 0,0,0,0,0,0,
  0,0,0,0, s_Imm_10y3*r_Imm*b_Imm*k_Imm, 0,0,0,0,0,0,
  0,0,0,0, s_Imm_10y3*r_Imm*b_Imm*(1-k_Imm), 0,0,0,0,0,0,
  0,0,0,0,0, s_IS_10y3*r_IS*b_IS*k_IS, s_IF_10y3*r_IF*b_IF*k_IF, s_ES_10y3*r_ES*b_ES*k_ES, s_EF_10y3*r_EF*b_EF*k_EF, s_ENB_10y3*r_ENB*b_ENB*k_ENB, s_PF_10y3*r_PF*b_PF*k_PF,
  0,0,0,0,0, s_IS_10y3*r_IS*b_IS*(1-k_IS), s_IF_10y3*r_IF*b_IF*(1-k_IF), s_ES_10y3*r_ES*b_ES*(1-k_ES), s_EF_10y3*r_EF*b_EF*(1-k_EF), s_ENB_10y3*r_ENB*b_ENB*(1-k_ENB), s_PF_10y3*r_PF*b_PF*(1-k_PF),
  0,0,0,0,0, s_IS_10y3*r_IS*(1-b_IS), s_IF_10y3*r_IF*(1-b_IF), s_ES_10y3*r_ES*(1-b_ES), s_EF_10y3*r_EF*(1-b_EF), s_ENB_10y3*r_ENB*(1-b_ENB), s_PF_10y3*r_PF*(1-b_PF),
  0,0,0,0,0, s_IS_10y3*(1-r_IS), s_IF_10y3*(1-r_IF), s_ES_10y3*(1-r_ES), s_EF_10y3*(1-r_EF), s_ENB_10y3*(1-r_ENB), s_PF_10y3*(1-r_PF)
), nrow=11, byrow=TRUE)

m_10y4 <- matrix(data = c(
  0,0,0,0,0, 0.5*s_Juv_10y4,0,0.5*s_Juv_10y4,0,0,0,
  s_Juv_10y4,0,0,0,0, 0,0,0,0,0,0,
  0,s_Juv_10y4,0,0,0, 0,0,0,0,0,0,
  0,0,s_Imm_10y4,0,0, 0,0,0,0,0,0,
  0,0,0,s_Imm_10y4, s_Imm_10y4*r_Imm*(1-b_Imm)+s_Imm_10y4*(1-r_Imm), 0,0,0,0,0,0,
  0,0,0,0, s_Imm_10y4*r_Imm*b_Imm*k_Imm, 0,0,0,0,0,0,
  0,0,0,0, s_Imm_10y4*r_Imm*b_Imm*(1-k_Imm), 0,0,0,0,0,0,
  0,0,0,0,0, s_IS_10y4*r_IS*b_IS*k_IS, s_IF_10y4*r_IF*b_IF*k_IF, s_ES_10y4*r_ES*b_ES*k_ES, s_EF_10y4*r_EF*b_EF*k_EF, s_ENB_10y4*r_ENB*b_ENB*k_ENB, s_PF_10y4*r_PF*b_PF*k_PF,
  0,0,0,0,0, s_IS_10y4*r_IS*b_IS*(1-k_IS), s_IF_10y4*r_IF*b_IF*(1-k_IF), s_ES_10y4*r_ES*b_ES*(1-k_ES), s_EF_10y4*r_EF*b_EF*(1-k_EF), s_ENB_10y4*r_ENB*b_ENB*(1-k_ENB), s_PF_10y4*r_PF*b_PF*(1-k_PF),
  0,0,0,0,0, s_IS_10y4*r_IS*(1-b_IS), s_IF_10y4*r_IF*(1-b_IF), s_ES_10y4*r_ES*(1-b_ES), s_EF_10y4*r_EF*(1-b_EF), s_ENB_10y4*r_ENB*(1-b_ENB), s_PF_10y4*r_PF*(1-b_PF),
  0,0,0,0,0, s_IS_10y4*(1-r_IS), s_IF_10y4*(1-r_IF), s_ES_10y4*(1-r_ES), s_EF_10y4*(1-r_EF), s_ENB_10y4*(1-r_ENB), s_PF_10y4*(1-r_PF)
), nrow=11, byrow=TRUE)

m_10y5 <- matrix(data = c(
  0,0,0,0,0, 0.5*s_Juv_10y5,0,0.5*s_Juv_10y5,0,0,0,
  s_Juv_10y5,0,0,0,0, 0,0,0,0,0,0,
  0,s_Juv_10y5,0,0,0, 0,0,0,0,0,0,
  0,0,s_Imm_10y5,0,0, 0,0,0,0,0,0,
  0,0,0,s_Imm_10y5, s_Imm_10y5*r_Imm*(1-b_Imm)+s_Imm_10y5*(1-r_Imm), 0,0,0,0,0,0,
  0,0,0,0, s_Imm_10y5*r_Imm*b_Imm*k_Imm, 0,0,0,0,0,0,
  0,0,0,0, s_Imm_10y5*r_Imm*b_Imm*(1-k_Imm), 0,0,0,0,0,0,
  0,0,0,0,0, s_IS_10y5*r_IS*b_IS*k_IS, s_IF_10y5*r_IF*b_IF*k_IF, s_ES_10y5*r_ES*b_ES*k_ES, s_EF_10y5*r_EF*b_EF*k_EF, s_ENB_10y5*r_ENB*b_ENB*k_ENB, s_PF_10y5*r_PF*b_PF*k_PF,
  0,0,0,0,0, s_IS_10y5*r_IS*b_IS*(1-k_IS), s_IF_10y5*r_IF*b_IF*(1-k_IF), s_ES_10y5*r_ES*b_ES*(1-k_ES), s_EF_10y5*r_EF*b_EF*(1-k_EF), s_ENB_10y5*r_ENB*b_ENB*(1-k_ENB), s_PF_10y5*r_PF*b_PF*(1-k_PF),
  0,0,0,0,0, s_IS_10y5*r_IS*(1-b_IS), s_IF_10y5*r_IF*(1-b_IF), s_ES_10y5*r_ES*(1-b_ES), s_EF_10y5*r_EF*(1-b_EF), s_ENB_10y5*r_ENB*(1-b_ENB), s_PF_10y5*r_PF*(1-b_PF),
  0,0,0,0,0, s_IS_10y5*(1-r_IS), s_IF_10y5*(1-r_IF), s_ES_10y5*(1-r_ES), s_EF_10y5*(1-r_EF), s_ENB_10y5*(1-r_ENB), s_PF_10y5*(1-r_PF)
), nrow=11, byrow=TRUE)

m_10y6 <- matrix(data = c(
  0,0,0,0,0, 0.5*s_Juv_10y6,0,0.5*s_Juv_10y6,0,0,0,
  s_Juv_10y6,0,0,0,0, 0,0,0,0,0,0,
  0,s_Juv_10y6,0,0,0, 0,0,0,0,0,0,
  0,0,s_Imm_10y6,0,0, 0,0,0,0,0,0,
  0,0,0,s_Imm_10y6, s_Imm_10y6*r_Imm*(1-b_Imm)+s_Imm_10y6*(1-r_Imm), 0,0,0,0,0,0,
  0,0,0,0, s_Imm_10y6*r_Imm*b_Imm*k_Imm, 0,0,0,0,0,0,
  0,0,0,0, s_Imm_10y6*r_Imm*b_Imm*(1-k_Imm), 0,0,0,0,0,0,
  0,0,0,0,0, s_IS_10y6*r_IS*b_IS*k_IS, s_IF_10y6*r_IF*b_IF*k_IF, s_ES_10y6*r_ES*b_ES*k_ES, s_EF_10y6*r_EF*b_EF*k_EF, s_ENB_10y6*r_ENB*b_ENB*k_ENB, s_PF_10y6*r_PF*b_PF*k_PF,
  0,0,0,0,0, s_IS_10y6*r_IS*b_IS*(1-k_IS), s_IF_10y6*r_IF*b_IF*(1-k_IF), s_ES_10y6*r_ES*b_ES*(1-k_ES), s_EF_10y6*r_EF*b_EF*(1-k_EF), s_ENB_10y6*r_ENB*b_ENB*(1-k_ENB), s_PF_10y6*r_PF*b_PF*(1-k_PF),
  0,0,0,0,0, s_IS_10y6*r_IS*(1-b_IS), s_IF_10y6*r_IF*(1-b_IF), s_ES_10y6*r_ES*(1-b_ES), s_EF_10y6*r_EF*(1-b_EF), s_ENB_10y6*r_ENB*(1-b_ENB), s_PF_10y6*r_PF*(1-b_PF),
  0,0,0,0,0, s_IS_10y6*(1-r_IS), s_IF_10y6*(1-r_IF), s_ES_10y6*(1-r_ES), s_EF_10y6*(1-r_EF), s_ENB_10y6*(1-r_ENB), s_PF_10y6*(1-r_PF)
), nrow=11, byrow=TRUE)

m_10y7 <- matrix(data = c(
  0,0,0,0,0, 0.5*s_Juv_10y7,0,0.5*s_Juv_10y7,0,0,0,
  s_Juv_10y7,0,0,0,0, 0,0,0,0,0,0,
  0,s_Juv_10y7,0,0,0, 0,0,0,0,0,0,
  0,0,s_Imm_10y7,0,0, 0,0,0,0,0,0,
  0,0,0,s_Imm_10y7, s_Imm_10y7*r_Imm*(1-b_Imm)+s_Imm_10y7*(1-r_Imm), 0,0,0,0,0,0,
  0,0,0,0, s_Imm_10y7*r_Imm*b_Imm*k_Imm, 0,0,0,0,0,0,
  0,0,0,0, s_Imm_10y7*r_Imm*b_Imm*(1-k_Imm), 0,0,0,0,0,0,
  0,0,0,0,0, s_IS_10y7*r_IS*b_IS*k_IS, s_IF_10y7*r_IF*b_IF*k_IF, s_ES_10y7*r_ES*b_ES*k_ES, s_EF_10y7*r_EF*b_EF*k_EF, s_ENB_10y7*r_ENB*b_ENB*k_ENB, s_PF_10y7*r_PF*b_PF*k_PF,
  0,0,0,0,0, s_IS_10y7*r_IS*b_IS*(1-k_IS), s_IF_10y7*r_IF*b_IF*(1-k_IF), s_ES_10y7*r_ES*b_ES*(1-k_ES), s_EF_10y7*r_EF*b_EF*(1-k_EF), s_ENB_10y7*r_ENB*b_ENB*(1-k_ENB), s_PF_10y7*r_PF*b_PF*(1-k_PF),
  0,0,0,0,0, s_IS_10y7*r_IS*(1-b_IS), s_IF_10y7*r_IF*(1-b_IF), s_ES_10y7*r_ES*(1-b_ES), s_EF_10y7*r_EF*(1-b_EF), s_ENB_10y7*r_ENB*(1-b_ENB), s_PF_10y7*r_PF*(1-b_PF),
  0,0,0,0,0, s_IS_10y7*(1-r_IS), s_IF_10y7*(1-r_IF), s_ES_10y7*(1-r_ES), s_EF_10y7*(1-r_EF), s_ENB_10y7*(1-r_ENB), s_PF_10y7*(1-r_PF)
), nrow=11, byrow=TRUE)

m_10y8 <- matrix(data = c(
  0,0,0,0,0, 0.5*s_Juv_10y8,0,0.5*s_Juv_10y8,0,0,0,
  s_Juv_10y8,0,0,0,0, 0,0,0,0,0,0,
  0,s_Juv_10y8,0,0,0, 0,0,0,0,0,0,
  0,0,s_Imm_10y8,0,0, 0,0,0,0,0,0,
  0,0,0,s_Imm_10y8, s_Imm_10y8*r_Imm*(1-b_Imm)+s_Imm_10y8*(1-r_Imm), 0,0,0,0,0,0,
  0,0,0,0, s_Imm_10y8*r_Imm*b_Imm*k_Imm, 0,0,0,0,0,0,
  0,0,0,0, s_Imm_10y8*r_Imm*b_Imm*(1-k_Imm), 0,0,0,0,0,0,
  0,0,0,0,0, s_IS_10y8*r_IS*b_IS*k_IS, s_IF_10y8*r_IF*b_IF*k_IF, s_ES_10y8*r_ES*b_ES*k_ES, s_EF_10y8*r_EF*b_EF*k_EF, s_ENB_10y8*r_ENB*b_ENB*k_ENB, s_PF_10y8*r_PF*b_PF*k_PF,
  0,0,0,0,0, s_IS_10y8*r_IS*b_IS*(1-k_IS), s_IF_10y8*r_IF*b_IF*(1-k_IF), s_ES_10y8*r_ES*b_ES*(1-k_ES), s_EF_10y8*r_EF*b_EF*(1-k_EF), s_ENB_10y8*r_ENB*b_ENB*(1-k_ENB), s_PF_10y8*r_PF*b_PF*(1-k_PF),
  0,0,0,0,0, s_IS_10y8*r_IS*(1-b_IS), s_IF_10y8*r_IF*(1-b_IF), s_ES_10y8*r_ES*(1-b_ES), s_EF_10y8*r_EF*(1-b_EF), s_ENB_10y8*r_ENB*(1-b_ENB), s_PF_10y8*r_PF*(1-b_PF),
  0,0,0,0,0, s_IS_10y8*(1-r_IS), s_IF_10y8*(1-r_IF), s_ES_10y8*(1-r_ES), s_EF_10y8*(1-r_EF), s_ENB_10y8*(1-r_ENB), s_PF_10y8*(1-r_PF)
), nrow=11, byrow=TRUE)

m_10y9 <- matrix(data = c(
  0,0,0,0,0, 0.5*s_Juv_10y9,0,0.5*s_Juv_10y9,0,0,0,
  s_Juv_10y9,0,0,0,0, 0,0,0,0,0,0,
  0,s_Juv_10y9,0,0,0, 0,0,0,0,0,0,
  0,0,s_Imm_10y9,0,0, 0,0,0,0,0,0,
  0,0,0,s_Imm_10y9, s_Imm_10y9*r_Imm*(1-b_Imm)+s_Imm_10y9*(1-r_Imm), 0,0,0,0,0,0,
  0,0,0,0, s_Imm_10y9*r_Imm*b_Imm*k_Imm, 0,0,0,0,0,0,
  0,0,0,0, s_Imm_10y9*r_Imm*b_Imm*(1-k_Imm), 0,0,0,0,0,0,
  0,0,0,0,0, s_IS_10y9*r_IS*b_IS*k_IS, s_IF_10y9*r_IF*b_IF*k_IF, s_ES_10y9*r_ES*b_ES*k_ES, s_EF_10y9*r_EF*b_EF*k_EF, s_ENB_10y9*r_ENB*b_ENB*k_ENB, s_PF_10y9*r_PF*b_PF*k_PF,
  0,0,0,0,0, s_IS_10y9*r_IS*b_IS*(1-k_IS), s_IF_10y9*r_IF*b_IF*(1-k_IF), s_ES_10y9*r_ES*b_ES*(1-k_ES), s_EF_10y9*r_EF*b_EF*(1-k_EF), s_ENB_10y9*r_ENB*b_ENB*(1-k_ENB), s_PF_10y9*r_PF*b_PF*(1-k_PF),
  0,0,0,0,0, s_IS_10y9*r_IS*(1-b_IS), s_IF_10y9*r_IF*(1-b_IF), s_ES_10y9*r_ES*(1-b_ES), s_EF_10y9*r_EF*(1-b_EF), s_ENB_10y9*r_ENB*(1-b_ENB), s_PF_10y9*r_PF*(1-b_PF),
  0,0,0,0,0, s_IS_10y9*(1-r_IS), s_IF_10y9*(1-r_IF), s_ES_10y9*(1-r_ES), s_EF_10y9*(1-r_EF), s_ENB_10y9*(1-r_ENB), s_PF_10y9*(1-r_PF)
), nrow=11, byrow=TRUE)

# m_10y10: 100% improvement — should equal m_3of3
m_10y10 <- matrix(data = c(
  0,0,0,0,0, 0.5*s_Juv_10y10,0,0.5*s_Juv_10y10,0,0,0,
  s_Juv_10y10,0,0,0,0, 0,0,0,0,0,0,
  0,s_Juv_10y10,0,0,0, 0,0,0,0,0,0,
  0,0,s_Imm_10y10,0,0, 0,0,0,0,0,0,
  0,0,0,s_Imm_10y10, s_Imm_10y10*r_Imm*(1-b_Imm)+s_Imm_10y10*(1-r_Imm), 0,0,0,0,0,0,
  0,0,0,0, s_Imm_10y10*r_Imm*b_Imm*k_Imm, 0,0,0,0,0,0,
  0,0,0,0, s_Imm_10y10*r_Imm*b_Imm*(1-k_Imm), 0,0,0,0,0,0,
  0,0,0,0,0, s_IS_10y10*r_IS*b_IS*k_IS, s_IF_10y10*r_IF*b_IF*k_IF, s_ES_10y10*r_ES*b_ES*k_ES, s_EF_10y10*r_EF*b_EF*k_EF, s_ENB_10y10*r_ENB*b_ENB*k_ENB, s_PF_10y10*r_PF*b_PF*k_PF,
  0,0,0,0,0, s_IS_10y10*r_IS*b_IS*(1-k_IS), s_IF_10y10*r_IF*b_IF*(1-k_IF), s_ES_10y10*r_ES*b_ES*(1-k_ES), s_EF_10y10*r_EF*b_EF*(1-k_EF), s_ENB_10y10*r_ENB*b_ENB*(1-k_ENB), s_PF_10y10*r_PF*b_PF*(1-k_PF),
  0,0,0,0,0, s_IS_10y10*r_IS*(1-b_IS), s_IF_10y10*r_IF*(1-b_IF), s_ES_10y10*r_ES*(1-b_ES), s_EF_10y10*r_EF*(1-b_EF), s_ENB_10y10*r_ENB*(1-b_ENB), s_PF_10y10*r_PF*(1-b_PF),
  0,0,0,0,0, s_IS_10y10*(1-r_IS), s_IF_10y10*(1-r_IF), s_ES_10y10*(1-r_ES), s_EF_10y10*(1-r_EF), s_ENB_10y10*(1-r_ENB), s_PF_10y10*(1-r_PF)
), nrow=11, byrow=TRUE)

phase_list_10y <- list(m_10y1,m_10y2,m_10y3,m_10y4,m_10y5,
                       m_10y6,m_10y7,m_10y8,m_10y9,m_10y10)
```

---

*PART 5 — 20-YEAR PHASE-IN*

**20-year phase-in survival rates**
```{r part5-20yr-survival-rates}
# Each step adds 1/20 of total improvement
s_Juv_20y1  <- s_Juv + (1/20)  * (ns_Juv - s_Juv);  s_Imm_20y1  <- s_Imm + (1/20)  * (ns_Imm - s_Imm)
s_EF_20y1   <- s_EF  + (1/20)  * (ns_EF  - s_EF);   s_ES_20y1   <- s_ES  + (1/20)  * (ns_ES  - s_ES)
s_ENB_20y1  <- s_ENB + (1/20)  * (ns_ENB - s_ENB);  s_IF_20y1   <- s_IF  + (1/20)  * (ns_IF  - s_IF)
s_IS_20y1   <- s_IS  + (1/20)  * (ns_IS  - s_IS);   s_PF_20y1   <- s_PF  + (1/20)  * (ns_PF  - s_PF)

s_Juv_20y2  <- s_Juv + (2/20)  * (ns_Juv - s_Juv);  s_Imm_20y2  <- s_Imm + (2/20)  * (ns_Imm - s_Imm)
s_EF_20y2   <- s_EF  + (2/20)  * (ns_EF  - s_EF);   s_ES_20y2   <- s_ES  + (2/20)  * (ns_ES  - s_ES)
s_ENB_20y2  <- s_ENB + (2/20)  * (ns_ENB - s_ENB);  s_IF_20y2   <- s_IF  + (2/20)  * (ns_IF  - s_IF)
s_IS_20y2   <- s_IS  + (2/20)  * (ns_IS  - s_IS);   s_PF_20y2   <- s_PF  + (2/20)  * (ns_PF  - s_PF)

s_Juv_20y3  <- s_Juv + (3/20)  * (ns_Juv - s_Juv);  s_Imm_20y3  <- s_Imm + (3/20)  * (ns_Imm - s_Imm)
s_EF_20y3   <- s_EF  + (3/20)  * (ns_EF  - s_EF);   s_ES_20y3   <- s_ES  + (3/20)  * (ns_ES  - s_ES)
s_ENB_20y3  <- s_ENB + (3/20)  * (ns_ENB - s_ENB);  s_IF_20y3   <- s_IF  + (3/20)  * (ns_IF  - s_IF)
s_IS_20y3   <- s_IS  + (3/20)  * (ns_IS  - s_IS);   s_PF_20y3   <- s_PF  + (3/20)  * (ns_PF  - s_PF)

s_Juv_20y4  <- s_Juv + (4/20)  * (ns_Juv - s_Juv);  s_Imm_20y4  <- s_Imm + (4/20)  * (ns_Imm - s_Imm)
s_EF_20y4   <- s_EF  + (4/20)  * (ns_EF  - s_EF);   s_ES_20y4   <- s_ES  + (4/20)  * (ns_ES  - s_ES)
s_ENB_20y4  <- s_ENB + (4/20)  * (ns_ENB - s_ENB);  s_IF_20y4   <- s_IF  + (4/20)  * (ns_IF  - s_IF)
s_IS_20y4   <- s_IS  + (4/20)  * (ns_IS  - s_IS);   s_PF_20y4   <- s_PF  + (4/20)  * (ns_PF  - s_PF)

s_Juv_20y5  <- s_Juv + (5/20)  * (ns_Juv - s_Juv);  s_Imm_20y5  <- s_Imm + (5/20)  * (ns_Imm - s_Imm)
s_EF_20y5   <- s_EF  + (5/20)  * (ns_EF  - s_EF);   s_ES_20y5   <- s_ES  + (5/20)  * (ns_ES  - s_ES)
s_ENB_20y5  <- s_ENB + (5/20)  * (ns_ENB - s_ENB);  s_IF_20y5   <- s_IF  + (5/20)  * (ns_IF  - s_IF)
s_IS_20y5   <- s_IS  + (5/20)  * (ns_IS  - s_IS);   s_PF_20y5   <- s_PF  + (5/20)  * (ns_PF  - s_PF)

s_Juv_20y6  <- s_Juv + (6/20)  * (ns_Juv - s_Juv);  s_Imm_20y6  <- s_Imm + (6/20)  * (ns_Imm - s_Imm)
s_EF_20y6   <- s_EF  + (6/20)  * (ns_EF  - s_EF);   s_ES_20y6   <- s_ES  + (6/20)  * (ns_ES  - s_ES)
s_ENB_20y6  <- s_ENB + (6/20)  * (ns_ENB - s_ENB);  s_IF_20y6   <- s_IF  + (6/20)  * (ns_IF  - s_IF)
s_IS_20y6   <- s_IS  + (6/20)  * (ns_IS  - s_IS);   s_PF_20y6   <- s_PF  + (6/20)  * (ns_PF  - s_PF)

s_Juv_20y7  <- s_Juv + (7/20)  * (ns_Juv - s_Juv);  s_Imm_20y7  <- s_Imm + (7/20)  * (ns_Imm - s_Imm)
s_EF_20y7   <- s_EF  + (7/20)  * (ns_EF  - s_EF);   s_ES_20y7   <- s_ES  + (7/20)  * (ns_ES  - s_ES)
s_ENB_20y7  <- s_ENB + (7/20)  * (ns_ENB - s_ENB);  s_IF_20y7   <- s_IF  + (7/20)  * (ns_IF  - s_IF)
s_IS_20y7   <- s_IS  + (7/20)  * (ns_IS  - s_IS);   s_PF_20y7   <- s_PF  + (7/20)  * (ns_PF  - s_PF)

s_Juv_20y8  <- s_Juv + (8/20)  * (ns_Juv - s_Juv);  s_Imm_20y8  <- s_Imm + (8/20)  * (ns_Imm - s_Imm)
s_EF_20y8   <- s_EF  + (8/20)  * (ns_EF  - s_EF);   s_ES_20y8   <- s_ES  + (8/20)  * (ns_ES  - s_ES)
s_ENB_20y8  <- s_ENB + (8/20)  * (ns_ENB - s_ENB);  s_IF_20y8   <- s_IF  + (8/20)  * (ns_IF  - s_IF)
s_IS_20y8   <- s_IS  + (8/20)  * (ns_IS  - s_IS);   s_PF_20y8   <- s_PF  + (8/20)  * (ns_PF  - s_PF)

s_Juv_20y9  <- s_Juv + (9/20)  * (ns_Juv - s_Juv);  s_Imm_20y9  <- s_Imm + (9/20)  * (ns_Imm - s_Imm)
s_EF_20y9   <- s_EF  + (9/20)  * (ns_EF  - s_EF);   s_ES_20y9   <- s_ES  + (9/20)  * (ns_ES  - s_ES)
s_ENB_20y9  <- s_ENB + (9/20)  * (ns_ENB - s_ENB);  s_IF_20y9   <- s_IF  + (9/20)  * (ns_IF  - s_IF)
s_IS_20y9   <- s_IS  + (9/20)  * (ns_IS  - s_IS);   s_PF_20y9   <- s_PF  + (9/20)  * (ns_PF  - s_PF)

s_Juv_20y10 <- s_Juv + (10/20) * (ns_Juv - s_Juv);  s_Imm_20y10 <- s_Imm + (10/20) * (ns_Imm - s_Imm)
s_EF_20y10  <- s_EF  + (10/20) * (ns_EF  - s_EF);   s_ES_20y10  <- s_ES  + (10/20) * (ns_ES  - s_ES)
s_ENB_20y10 <- s_ENB + (10/20) * (ns_ENB - s_ENB);  s_IF_20y10  <- s_IF  + (10/20) * (ns_IF  - s_IF)
s_IS_20y10  <- s_IS  + (10/20) * (ns_IS  - s_IS);   s_PF_20y10  <- s_PF  + (10/20) * (ns_PF  - s_PF)

s_Juv_20y11 <- s_Juv + (11/20) * (ns_Juv - s_Juv);  s_Imm_20y11 <- s_Imm + (11/20) * (ns_Imm - s_Imm)
s_EF_20y11  <- s_EF  + (11/20) * (ns_EF  - s_EF);   s_ES_20y11  <- s_ES  + (11/20) * (ns_ES  - s_ES)
s_ENB_20y11 <- s_ENB + (11/20) * (ns_ENB - s_ENB);  s_IF_20y11  <- s_IF  + (11/20) * (ns_IF  - s_IF)
s_IS_20y11  <- s_IS  + (11/20) * (ns_IS  - s_IS);   s_PF_20y11  <- s_PF  + (11/20) * (ns_PF  - s_PF)

s_Juv_20y12 <- s_Juv + (12/20) * (ns_Juv - s_Juv);  s_Imm_20y12 <- s_Imm + (12/20) * (ns_Imm - s_Imm)
s_EF_20y12  <- s_EF  + (12/20) * (ns_EF  - s_EF);   s_ES_20y12  <- s_ES  + (12/20) * (ns_ES  - s_ES)
s_ENB_20y12 <- s_ENB + (12/20) * (ns_ENB - s_ENB);  s_IF_20y12  <- s_IF  + (12/20) * (ns_IF  - s_IF)
s_IS_20y12  <- s_IS  + (12/20) * (ns_IS  - s_IS);   s_PF_20y12  <- s_PF  + (12/20) * (ns_PF  - s_PF)

s_Juv_20y13 <- s_Juv + (13/20) * (ns_Juv - s_Juv);  s_Imm_20y13 <- s_Imm + (13/20) * (ns_Imm - s_Imm)
s_EF_20y13  <- s_EF  + (13/20) * (ns_EF  - s_EF);   s_ES_20y13  <- s_ES  + (13/20) * (ns_ES  - s_ES)
s_ENB_20y13 <- s_ENB + (13/20) * (ns_ENB - s_ENB);  s_IF_20y13  <- s_IF  + (13/20) * (ns_IF  - s_IF)
s_IS_20y13  <- s_IS  + (13/20) * (ns_IS  - s_IS);   s_PF_20y13  <- s_PF  + (13/20) * (ns_PF  - s_PF)

s_Juv_20y14 <- s_Juv + (14/20) * (ns_Juv - s_Juv);  s_Imm_20y14 <- s_Imm + (14/20) * (ns_Imm - s_Imm)
s_EF_20y14  <- s_EF  + (14/20) * (ns_EF  - s_EF);   s_ES_20y14  <- s_ES  + (14/20) * (ns_ES  - s_ES)
s_ENB_20y14 <- s_ENB + (14/20) * (ns_ENB - s_ENB);  s_IF_20y14  <- s_IF  + (14/20) * (ns_IF  - s_IF)
s_IS_20y14  <- s_IS  + (14/20) * (ns_IS  - s_IS);   s_PF_20y14  <- s_PF  + (14/20) * (ns_PF  - s_PF)

s_Juv_20y15 <- s_Juv + (15/20) * (ns_Juv - s_Juv);  s_Imm_20y15 <- s_Imm + (15/20) * (ns_Imm - s_Imm)
s_EF_20y15  <- s_EF  + (15/20) * (ns_EF  - s_EF);   s_ES_20y15  <- s_ES  + (15/20) * (ns_ES  - s_ES)
s_ENB_20y15 <- s_ENB + (15/20) * (ns_ENB - s_ENB);  s_IF_20y15  <- s_IF  + (15/20) * (ns_IF  - s_IF)
s_IS_20y15  <- s_IS  + (15/20) * (ns_IS  - s_IS);   s_PF_20y15  <- s_PF  + (15/20) * (ns_PF  - s_PF)

s_Juv_20y16 <- s_Juv + (16/20) * (ns_Juv - s_Juv);  s_Imm_20y16 <- s_Imm + (16/20) * (ns_Imm - s_Imm)
s_EF_20y16  <- s_EF  + (16/20) * (ns_EF  - s_EF);   s_ES_20y16  <- s_ES  + (16/20) * (ns_ES  - s_ES)
s_ENB_20y16 <- s_ENB + (16/20) * (ns_ENB - s_ENB);  s_IF_20y16  <- s_IF  + (16/20) * (ns_IF  - s_IF)
s_IS_20y16  <- s_IS  + (16/20) * (ns_IS  - s_IS);   s_PF_20y16  <- s_PF  + (16/20) * (ns_PF  - s_PF)

s_Juv_20y17 <- s_Juv + (17/20) * (ns_Juv - s_Juv);  s_Imm_20y17 <- s_Imm + (17/20) * (ns_Imm - s_Imm)
s_EF_20y17  <- s_EF  + (17/20) * (ns_EF  - s_EF);   s_ES_20y17  <- s_ES  + (17/20) * (ns_ES  - s_ES)
s_ENB_20y17 <- s_ENB + (17/20) * (ns_ENB - s_ENB);  s_IF_20y17  <- s_IF  + (17/20) * (ns_IF  - s_IF)
s_IS_20y17  <- s_IS  + (17/20) * (ns_IS  - s_IS);   s_PF_20y17  <- s_PF  + (17/20) * (ns_PF  - s_PF)

s_Juv_20y18 <- s_Juv + (18/20) * (ns_Juv - s_Juv);  s_Imm_20y18 <- s_Imm + (18/20) * (ns_Imm - s_Imm)
s_EF_20y18  <- s_EF  + (18/20) * (ns_EF  - s_EF);   s_ES_20y18  <- s_ES  + (18/20) * (ns_ES  - s_ES)
s_ENB_20y18 <- s_ENB + (18/20) * (ns_ENB - s_ENB);  s_IF_20y18  <- s_IF  + (18/20) * (ns_IF  - s_IF)
s_IS_20y18  <- s_IS  + (18/20) * (ns_IS  - s_IS);   s_PF_20y18  <- s_PF  + (18/20) * (ns_PF  - s_PF)

s_Juv_20y19 <- s_Juv + (19/20) * (ns_Juv - s_Juv);  s_Imm_20y19 <- s_Imm + (19/20) * (ns_Imm - s_Imm)
s_EF_20y19  <- s_EF  + (19/20) * (ns_EF  - s_EF);   s_ES_20y19  <- s_ES  + (19/20) * (ns_ES  - s_ES)
s_ENB_20y19 <- s_ENB + (19/20) * (ns_ENB - s_ENB);  s_IF_20y19  <- s_IF  + (19/20) * (ns_IF  - s_IF)
s_IS_20y19  <- s_IS  + (19/20) * (ns_IS  - s_IS);   s_PF_20y19  <- s_PF  + (19/20) * (ns_PF  - s_PF)

# Step 20: 100% — should equal ns_* (same as m_3of3)
s_Juv_20y20 <- s_Juv + (20/20) * (ns_Juv - s_Juv);  s_Imm_20y20 <- s_Imm + (20/20) * (ns_Imm - s_Imm)
s_EF_20y20  <- s_EF  + (20/20) * (ns_EF  - s_EF);   s_ES_20y20  <- s_ES  + (20/20) * (ns_ES  - s_ES)
s_ENB_20y20 <- s_ENB + (20/20) * (ns_ENB - s_ENB);  s_IF_20y20  <- s_IF  + (20/20) * (ns_IF  - s_IF)
s_IS_20y20  <- s_IS  + (20/20) * (ns_IS  - s_IS);   s_PF_20y20  <- s_PF  + (20/20) * (ns_PF  - s_PF)

tibble(
  stage     = c("Juv","Imm","EF","ES","ENB","IF","IS","PF"),
  baseline  = c(s_Juv, s_Imm, s_EF, s_ES, s_ENB, s_IF, s_IS, s_PF),
  y1 =c(s_Juv_20y1, s_Imm_20y1, s_EF_20y1, s_ES_20y1, s_ENB_20y1, s_IF_20y1, s_IS_20y1, s_PF_20y1),
  y5 =c(s_Juv_20y5, s_Imm_20y5, s_EF_20y5, s_ES_20y5, s_ENB_20y5, s_IF_20y5, s_IS_20y5, s_PF_20y5),
  y10=c(s_Juv_20y10,s_Imm_20y10,s_EF_20y10,s_ES_20y10,s_ENB_20y10,s_IF_20y10,s_IS_20y10,s_PF_20y10),
  y15=c(s_Juv_20y15,s_Imm_20y15,s_EF_20y15,s_ES_20y15,s_ENB_20y15,s_IF_20y15,s_IS_20y15,s_PF_20y15),
  y19=c(s_Juv_20y19,s_Imm_20y19,s_EF_20y19,s_ES_20y19,s_ENB_20y19,s_IF_20y19,s_IS_20y19,s_PF_20y19),
  full_3of3=c(s_Juv_20y20,s_Imm_20y20,s_EF_20y20,s_ES_20y20,s_ENB_20y20,s_IF_20y20,s_IS_20y20,s_PF_20y20)
) %>% print()
```

**20-year phase-in matrices**
```{r part5-20yr-matrices}
# Stages: J2 J3 Imm4 Imm5 PB IS IF ES EF ENB Sabb
m_20y1 <- matrix(data = c(
  0,0,0,0,0, 0.5*s_Juv_20y1,0,0.5*s_Juv_20y1,0,0,0,
  s_Juv_20y1,0,0,0,0, 0,0,0,0,0,0,
  0,s_Juv_20y1,0,0,0, 0,0,0,0,0,0,
  0,0,s_Imm_20y1,0,0, 0,0,0,0,0,0,
  0,0,0,s_Imm_20y1, s_Imm_20y1*r_Imm*(1-b_Imm)+s_Imm_20y1*(1-r_Imm), 0,0,0,0,0,0,
  0,0,0,0, s_Imm_20y1*r_Imm*b_Imm*k_Imm, 0,0,0,0,0,0,
  0,0,0,0, s_Imm_20y1*r_Imm*b_Imm*(1-k_Imm), 0,0,0,0,0,0,
  0,0,0,0,0, s_IS_20y1*r_IS*b_IS*k_IS, s_IF_20y1*r_IF*b_IF*k_IF, s_ES_20y1*r_ES*b_ES*k_ES, s_EF_20y1*r_EF*b_EF*k_EF, s_ENB_20y1*r_ENB*b_ENB*k_ENB, s_PF_20y1*r_PF*b_PF*k_PF,
  0,0,0,0,0, s_IS_20y1*r_IS*b_IS*(1-k_IS), s_IF_20y1*r_IF*b_IF*(1-k_IF), s_ES_20y1*r_ES*b_ES*(1-k_ES), s_EF_20y1*r_EF*b_EF*(1-k_EF), s_ENB_20y1*r_ENB*b_ENB*(1-k_ENB), s_PF_20y1*r_PF*b_PF*(1-k_PF),
  0,0,0,0,0, s_IS_20y1*r_IS*(1-b_IS), s_IF_20y1*r_IF*(1-b_IF), s_ES_20y1*r_ES*(1-b_ES), s_EF_20y1*r_EF*(1-b_EF), s_ENB_20y1*r_ENB*(1-b_ENB), s_PF_20y1*r_PF*(1-b_PF),
  0,0,0,0,0, s_IS_20y1*(1-r_IS), s_IF_20y1*(1-r_IF), s_ES_20y1*(1-r_ES), s_EF_20y1*(1-r_EF), s_ENB_20y1*(1-r_ENB), s_PF_20y1*(1-r_PF)
), nrow=11, byrow=TRUE)

m_20y2 <- matrix(data = c(
  0,0,0,0,0, 0.5*s_Juv_20y2,0,0.5*s_Juv_20y2,0,0,0,
  s_Juv_20y2,0,0,0,0, 0,0,0,0,0,0,
  0,s_Juv_20y2,0,0,0, 0,0,0,0,0,0,
  0,0,s_Imm_20y2,0,0, 0,0,0,0,0,0,
  0,0,0,s_Imm_20y2, s_Imm_20y2*r_Imm*(1-b_Imm)+s_Imm_20y2*(1-r_Imm), 0,0,0,0,0,0,
  0,0,0,0, s_Imm_20y2*r_Imm*b_Imm*k_Imm, 0,0,0,0,0,0,
  0,0,0,0, s_Imm_20y2*r_Imm*b_Imm*(1-k_Imm), 0,0,0,0,0,0,
  0,0,0,0,0, s_IS_20y2*r_IS*b_IS*k_IS, s_IF_20y2*r_IF*b_IF*k_IF, s_ES_20y2*r_ES*b_ES*k_ES, s_EF_20y2*r_EF*b_EF*k_EF, s_ENB_20y2*r_ENB*b_ENB*k_ENB, s_PF_20y2*r_PF*b_PF*k_PF,
  0,0,0,0,0, s_IS_20y2*r_IS*b_IS*(1-k_IS), s_IF_20y2*r_IF*b_IF*(1-k_IF), s_ES_20y2*r_ES*b_ES*(1-k_ES), s_EF_20y2*r_EF*b_EF*(1-k_EF), s_ENB_20y2*r_ENB*b_ENB*(1-k_ENB), s_PF_20y2*r_PF*b_PF*(1-k_PF),
  0,0,0,0,0, s_IS_20y2*r_IS*(1-b_IS), s_IF_20y2*r_IF*(1-b_IF), s_ES_20y2*r_ES*(1-b_ES), s_EF_20y2*r_EF*(1-b_EF), s_ENB_20y2*r_ENB*(1-b_ENB), s_PF_20y2*r_PF*(1-b_PF),
  0,0,0,0,0, s_IS_20y2*(1-r_IS), s_IF_20y2*(1-r_IF), s_ES_20y2*(1-r_ES), s_EF_20y2*(1-r_EF), s_ENB_20y2*(1-r_ENB), s_PF_20y2*(1-r_PF)
), nrow=11, byrow=TRUE)

m_20y3 <- matrix(data = c(
  0,0,0,0,0, 0.5*s_Juv_20y3,0,0.5*s_Juv_20y3,0,0,0,
  s_Juv_20y3,0,0,0,0, 0,0,0,0,0,0,
  0,s_Juv_20y3,0,0,0, 0,0,0,0,0,0,
  0,0,s_Imm_20y3,0,0, 0,0,0,0,0,0,
  0,0,0,s_Imm_20y3, s_Imm_20y3*r_Imm*(1-b_Imm)+s_Imm_20y3*(1-r_Imm), 0,0,0,0,0,0,
  0,0,0,0, s_Imm_20y3*r_Imm*b_Imm*k_Imm, 0,0,0,0,0,0,
  0,0,0,0, s_Imm_20y3*r_Imm*b_Imm*(1-k_Imm), 0,0,0,0,0,0,
  0,0,0,0,0, s_IS_20y3*r_IS*b_IS*k_IS, s_IF_20y3*r_IF*b_IF*k_IF, s_ES_20y3*r_ES*b_ES*k_ES, s_EF_20y3*r_EF*b_EF*k_EF, s_ENB_20y3*r_ENB*b_ENB*k_ENB, s_PF_20y3*r_PF*b_PF*k_PF,
  0,0,0,0,0, s_IS_20y3*r_IS*b_IS*(1-k_IS), s_IF_20y3*r_IF*b_IF*(1-k_IF), s_ES_20y3*r_ES*b_ES*(1-k_ES), s_EF_20y3*r_EF*b_EF*(1-k_EF), s_ENB_20y3*r_ENB*b_ENB*(1-k_ENB), s_PF_20y3*r_PF*b_PF*(1-k_PF),
  0,0,0,0,0, s_IS_20y3*r_IS*(1-b_IS), s_IF_20y3*r_IF*(1-b_IF), s_ES_20y3*r_ES*(1-b_ES), s_EF_20y3*r_EF*(1-b_EF), s_ENB_20y3*r_ENB*(1-b_ENB), s_PF_20y3*r_PF*(1-b_PF),
  0,0,0,0,0, s_IS_20y3*(1-r_IS), s_IF_20y3*(1-r_IF), s_ES_20y3*(1-r_ES), s_EF_20y3*(1-r_EF), s_ENB_20y3*(1-r_ENB), s_PF_20y3*(1-r_PF)
), nrow=11, byrow=TRUE)

m_20y4 <- matrix(data = c(
  0,0,0,0,0, 0.5*s_Juv_20y4,0,0.5*s_Juv_20y4,0,0,0,
  s_Juv_20y4,0,0,0,0, 0,0,0,0,0,0,
  0,s_Juv_20y4,0,0,0, 0,0,0,0,0,0,
  0,0,s_Imm_20y4,0,0, 0,0,0,0,0,0,
  0,0,0,s_Imm_20y4, s_Imm_20y4*r_Imm*(1-b_Imm)+s_Imm_20y4*(1-r_Imm), 0,0,0,0,0,0,
  0,0,0,0, s_Imm_20y4*r_Imm*b_Imm*k_Imm, 0,0,0,0,0,0,
  0,0,0,0, s_Imm_20y4*r_Imm*b_Imm*(1-k_Imm), 0,0,0,0,0,0,
  0,0,0,0,0, s_IS_20y4*r_IS*b_IS*k_IS, s_IF_20y4*r_IF*b_IF*k_IF, s_ES_20y4*r_ES*b_ES*k_ES, s_EF_20y4*r_EF*b_EF*k_EF, s_ENB_20y4*r_ENB*b_ENB*k_ENB, s_PF_20y4*r_PF*b_PF*k_PF,
  0,0,0,0,0, s_IS_20y4*r_IS*b_IS*(1-k_IS), s_IF_20y4*r_IF*b_IF*(1-k_IF), s_ES_20y4*r_ES*b_ES*(1-k_ES), s_EF_20y4*r_EF*b_EF*(1-k_EF), s_ENB_20y4*r_ENB*b_ENB*(1-k_ENB), s_PF_20y4*r_PF*b_PF*(1-k_PF),
  0,0,0,0,0, s_IS_20y4*r_IS*(1-b_IS), s_IF_20y4*r_IF*(1-b_IF), s_ES_20y4*r_ES*(1-b_ES), s_EF_20y4*r_EF*(1-b_EF), s_ENB_20y4*r_ENB*(1-b_ENB), s_PF_20y4*r_PF*(1-b_PF),
  0,0,0,0,0, s_IS_20y4*(1-r_IS), s_IF_20y4*(1-r_IF), s_ES_20y4*(1-r_ES), s_EF_20y4*(1-r_EF), s_ENB_20y4*(1-r_ENB), s_PF_20y4*(1-r_PF)
), nrow=11, byrow=TRUE)

m_20y5 <- matrix(data = c(
  0,0,0,0,0, 0.5*s_Juv_20y5,0,0.5*s_Juv_20y5,0,0,0,
  s_Juv_20y5,0,0,0,0, 0,0,0,0,0,0,
  0,s_Juv_20y5,0,0,0, 0,0,0,0,0,0,
  0,0,s_Imm_20y5,0,0, 0,0,0,0,0,0,
  0,0,0,s_Imm_20y5, s_Imm_20y5*r_Imm*(1-b_Imm)+s_Imm_20y5*(1-r_Imm), 0,0,0,0,0,0,
  0,0,0,0, s_Imm_20y5*r_Imm*b_Imm*k_Imm, 0,0,0,0,0,0,
  0,0,0,0, s_Imm_20y5*r_Imm*b_Imm*(1-k_Imm), 0,0,0,0,0,0,
  0,0,0,0,0, s_IS_20y5*r_IS*b_IS*k_IS, s_IF_20y5*r_IF*b_IF*k_IF, s_ES_20y5*r_ES*b_ES*k_ES, s_EF_20y5*r_EF*b_EF*k_EF, s_ENB_20y5*r_ENB*b_ENB*k_ENB, s_PF_20y5*r_PF*b_PF*k_PF,
  0,0,0,0,0, s_IS_20y5*r_IS*b_IS*(1-k_IS), s_IF_20y5*r_IF*b_IF*(1-k_IF), s_ES_20y5*r_ES*b_ES*(1-k_ES), s_EF_20y5*r_EF*b_EF*(1-k_EF), s_ENB_20y5*r_ENB*b_ENB*(1-k_ENB), s_PF_20y5*r_PF*b_PF*(1-k_PF),
  0,0,0,0,0, s_IS_20y5*r_IS*(1-b_IS), s_IF_20y5*r_IF*(1-b_IF), s_ES_20y5*r_ES*(1-b_ES), s_EF_20y5*r_EF*(1-b_EF), s_ENB_20y5*r_ENB*(1-b_ENB), s_PF_20y5*r_PF*(1-b_PF),
  0,0,0,0,0, s_IS_20y5*(1-r_IS), s_IF_20y5*(1-r_IF), s_ES_20y5*(1-r_ES), s_EF_20y5*(1-r_EF), s_ENB_20y5*(1-r_ENB), s_PF_20y5*(1-r_PF)
), nrow=11, byrow=TRUE)

m_20y6 <- matrix(data = c(
  0,0,0,0,0, 0.5*s_Juv_20y6,0,0.5*s_Juv_20y6,0,0,0,
  s_Juv_20y6,0,0,0,0, 0,0,0,0,0,0,
  0,s_Juv_20y6,0,0,0, 0,0,0,0,0,0,
  0,0,s_Imm_20y6,0,0, 0,0,0,0,0,0,
  0,0,0,s_Imm_20y6, s_Imm_20y6*r_Imm*(1-b_Imm)+s_Imm_20y6*(1-r_Imm), 0,0,0,0,0,0,
  0,0,0,0, s_Imm_20y6*r_Imm*b_Imm*k_Imm, 0,0,0,0,0,0,
  0,0,0,0, s_Imm_20y6*r_Imm*b_Imm*(1-k_Imm), 0,0,0,0,0,0,
  0,0,0,0,0, s_IS_20y6*r_IS*b_IS*k_IS, s_IF_20y6*r_IF*b_IF*k_IF, s_ES_20y6*r_ES*b_ES*k_ES, s_EF_20y6*r_EF*b_EF*k_EF, s_ENB_20y6*r_ENB*b_ENB*k_ENB, s_PF_20y6*r_PF*b_PF*k_PF,
  0,0,0,0,0, s_IS_20y6*r_IS*b_IS*(1-k_IS), s_IF_20y6*r_IF*b_IF*(1-k_IF), s_ES_20y6*r_ES*b_ES*(1-k_ES), s_EF_20y6*r_EF*b_EF*(1-k_EF), s_ENB_20y6*r_ENB*b_ENB*(1-k_ENB), s_PF_20y6*r_PF*b_PF*(1-k_PF),
  0,0,0,0,0, s_IS_20y6*r_IS*(1-b_IS), s_IF_20y6*r_IF*(1-b_IF), s_ES_20y6*r_ES*(1-b_ES), s_EF_20y6*r_EF*(1-b_EF), s_ENB_20y6*r_ENB*(1-b_ENB), s_PF_20y6*r_PF*(1-b_PF),
  0,0,0,0,0, s_IS_20y6*(1-r_IS), s_IF_20y6*(1-r_IF), s_ES_20y6*(1-r_ES), s_EF_20y6*(1-r_EF), s_ENB_20y6*(1-r_ENB), s_PF_20y6*(1-r_PF)
), nrow=11, byrow=TRUE)

m_20y7 <- matrix(data = c(
  0,0,0,0,0, 0.5*s_Juv_20y7,0,0.5*s_Juv_20y7,0,0,0,
  s_Juv_20y7,0,0,0,0, 0,0,0,0,0,0,
  0,s_Juv_20y7,0,0,0, 0,0,0,0,0,0,
  0,0,s_Imm_20y7,0,0, 0,0,0,0,0,0,
  0,0,0,s_Imm_20y7, s_Imm_20y7*r_Imm*(1-b_Imm)+s_Imm_20y7*(1-r_Imm), 0,0,0,0,0,0,
  0,0,0,0, s_Imm_20y7*r_Imm*b_Imm*k_Imm, 0,0,0,0,0,0,
  0,0,0,0, s_Imm_20y7*r_Imm*b_Imm*(1-k_Imm), 0,0,0,0,0,0,
  0,0,0,0,0, s_IS_20y7*r_IS*b_IS*k_IS, s_IF_20y7*r_IF*b_IF*k_IF, s_ES_20y7*r_ES*b_ES*k_ES, s_EF_20y7*r_EF*b_EF*k_EF, s_ENB_20y7*r_ENB*b_ENB*k_ENB, s_PF_20y7*r_PF*b_PF*k_PF,
  0,0,0,0,0, s_IS_20y7*r_IS*b_IS*(1-k_IS), s_IF_20y7*r_IF*b_IF*(1-k_IF), s_ES_20y7*r_ES*b_ES*(1-k_ES), s_EF_20y7*r_EF*b_EF*(1-k_EF), s_ENB_20y7*r_ENB*b_ENB*(1-k_ENB), s_PF_20y7*r_PF*b_PF*(1-k_PF),
  0,0,0,0,0, s_IS_20y7*r_IS*(1-b_IS), s_IF_20y7*r_IF*(1-b_IF), s_ES_20y7*r_ES*(1-b_ES), s_EF_20y7*r_EF*(1-b_EF), s_ENB_20y7*r_ENB*(1-b_ENB), s_PF_20y7*r_PF*(1-b_PF),
  0,0,0,0,0, s_IS_20y7*(1-r_IS), s_IF_20y7*(1-r_IF), s_ES_20y7*(1-r_ES), s_EF_20y7*(1-r_EF), s_ENB_20y7*(1-r_ENB), s_PF_20y7*(1-r_PF)
), nrow=11, byrow=TRUE)

m_20y8 <- matrix(data = c(
  0,0,0,0,0, 0.5*s_Juv_20y8,0,0.5*s_Juv_20y8,0,0,0,
  s_Juv_20y8,0,0,0,0, 0,0,0,0,0,0,
  0,s_Juv_20y8,0,0,0, 0,0,0,0,0,0,
  0,0,s_Imm_20y8,0,0, 0,0,0,0,0,0,
  0,0,0,s_Imm_20y8, s_Imm_20y8*r_Imm*(1-b_Imm)+s_Imm_20y8*(1-r_Imm), 0,0,0,0,0,0,
  0,0,0,0, s_Imm_20y8*r_Imm*b_Imm*k_Imm, 0,0,0,0,0,0,
  0,0,0,0, s_Imm_20y8*r_Imm*b_Imm*(1-k_Imm), 0,0,0,0,0,0,
  0,0,0,0,0, s_IS_20y8*r_IS*b_IS*k_IS, s_IF_20y8*r_IF*b_IF*k_IF, s_ES_20y8*r_ES*b_ES*k_ES, s_EF_20y8*r_EF*b_EF*k_EF, s_ENB_20y8*r_ENB*b_ENB*k_ENB, s_PF_20y8*r_PF*b_PF*k_PF,
  0,0,0,0,0, s_IS_20y8*r_IS*b_IS*(1-k_IS), s_IF_20y8*r_IF*b_IF*(1-k_IF), s_ES_20y8*r_ES*b_ES*(1-k_ES), s_EF_20y8*r_EF*b_EF*(1-k_EF), s_ENB_20y8*r_ENB*b_ENB*(1-k_ENB), s_PF_20y8*r_PF*b_PF*(1-k_PF),
  0,0,0,0,0, s_IS_20y8*r_IS*(1-b_IS), s_IF_20y8*r_IF*(1-b_IF), s_ES_20y8*r_ES*(1-b_ES), s_EF_20y8*r_EF*(1-b_EF), s_ENB_20y8*r_ENB*(1-b_ENB), s_PF_20y8*r_PF*(1-b_PF),
  0,0,0,0,0, s_IS_20y8*(1-r_IS), s_IF_20y8*(1-r_IF), s_ES_20y8*(1-r_ES), s_EF_20y8*(1-r_EF), s_ENB_20y8*(1-r_ENB), s_PF_20y8*(1-r_PF)
), nrow=11, byrow=TRUE)

m_20y9 <- matrix(data = c(
  0,0,0,0,0, 0.5*s_Juv_20y9,0,0.5*s_Juv_20y9,0,0,0,
  s_Juv_20y9,0,0,0,0, 0,0,0,0,0,0,
  0,s_Juv_20y9,0,0,0, 0,0,0,0,0,0,
  0,0,s_Imm_20y9,0,0, 0,0,0,0,0,0,
  0,0,0,s_Imm_20y9, s_Imm_20y9*r_Imm*(1-b_Imm)+s_Imm_20y9*(1-r_Imm), 0,0,0,0,0,0,
  0,0,0,0, s_Imm_20y9*r_Imm*b_Imm*k_Imm, 0,0,0,0,0,0,
  0,0,0,0, s_Imm_20y9*r_Imm*b_Imm*(1-k_Imm), 0,0,0,0,0,0,
  0,0,0,0,0, s_IS_20y9*r_IS*b_IS*k_IS, s_IF_20y9*r_IF*b_IF*k_IF, s_ES_20y9*r_ES*b_ES*k_ES, s_EF_20y9*r_EF*b_EF*k_EF, s_ENB_20y9*r_ENB*b_ENB*k_ENB, s_PF_20y9*r_PF*b_PF*k_PF,
  0,0,0,0,0, s_IS_20y9*r_IS*b_IS*(1-k_IS), s_IF_20y9*r_IF*b_IF*(1-k_IF), s_ES_20y9*r_ES*b_ES*(1-k_ES), s_EF_20y9*r_EF*b_EF*(1-k_EF), s_ENB_20y9*r_ENB*b_ENB*(1-k_ENB), s_PF_20y9*r_PF*b_PF*(1-k_PF),
  0,0,0,0,0, s_IS_20y9*r_IS*(1-b_IS), s_IF_20y9*r_IF*(1-b_IF), s_ES_20y9*r_ES*(1-b_ES), s_EF_20y9*r_EF*(1-b_EF), s_ENB_20y9*r_ENB*(1-b_ENB), s_PF_20y9*r_PF*(1-b_PF),
  0,0,0,0,0, s_IS_20y9*(1-r_IS), s_IF_20y9*(1-r_IF), s_ES_20y9*(1-r_ES), s_EF_20y9*(1-r_EF), s_ENB_20y9*(1-r_ENB), s_PF_20y9*(1-r_PF)
), nrow=11, byrow=TRUE)

m_20y10 <- matrix(data = c(
  0,0,0,0,0, 0.5*s_Juv_20y10,0,0.5*s_Juv_20y10,0,0,0,
  s_Juv_20y10,0,0,0,0, 0,0,0,0,0,0,
  0,s_Juv_20y10,0,0,0, 0,0,0,0,0,0,
  0,0,s_Imm_20y10,0,0, 0,0,0,0,0,0,
  0,0,0,s_Imm_20y10, s_Imm_20y10*r_Imm*(1-b_Imm)+s_Imm_20y10*(1-r_Imm), 0,0,0,0,0,0,
  0,0,0,0, s_Imm_20y10*r_Imm*b_Imm*k_Imm, 0,0,0,0,0,0,
  0,0,0,0, s_Imm_20y10*r_Imm*b_Imm*(1-k_Imm), 0,0,0,0,0,0,
  0,0,0,0,0, s_IS_20y10*r_IS*b_IS*k_IS, s_IF_20y10*r_IF*b_IF*k_IF, s_ES_20y10*r_ES*b_ES*k_ES, s_EF_20y10*r_EF*b_EF*k_EF, s_ENB_20y10*r_ENB*b_ENB*k_ENB, s_PF_20y10*r_PF*b_PF*k_PF,
  0,0,0,0,0, s_IS_20y10*r_IS*b_IS*(1-k_IS), s_IF_20y10*r_IF*b_IF*(1-k_IF), s_ES_20y10*r_ES*b_ES*(1-k_ES), s_EF_20y10*r_EF*b_EF*(1-k_EF), s_ENB_20y10*r_ENB*b_ENB*(1-k_ENB), s_PF_20y10*r_PF*b_PF*(1-k_PF),
  0,0,0,0,0, s_IS_20y10*r_IS*(1-b_IS), s_IF_20y10*r_IF*(1-b_IF), s_ES_20y10*r_ES*(1-b_ES), s_EF_20y10*r_EF*(1-b_EF), s_ENB_20y10*r_ENB*(1-b_ENB), s_PF_20y10*r_PF*(1-b_PF),
  0,0,0,0,0, s_IS_20y10*(1-r_IS), s_IF_20y10*(1-r_IF), s_ES_20y10*(1-r_ES), s_EF_20y10*(1-r_EF), s_ENB_20y10*(1-r_ENB), s_PF_20y10*(1-r_PF)
), nrow=11, byrow=TRUE)

m_20y11 <- matrix(data = c(
  0,0,0,0,0, 0.5*s_Juv_20y11,0,0.5*s_Juv_20y11,0,0,0,
  s_Juv_20y11,0,0,0,0, 0,0,0,0,0,0,
  0,s_Juv_20y11,0,0,0, 0,0,0,0,0,0,
  0,0,s_Imm_20y11,0,0, 0,0,0,0,0,0,
  0,0,0,s_Imm_20y11, s_Imm_20y11*r_Imm*(1-b_Imm)+s_Imm_20y11*(1-r_Imm), 0,0,0,0,0,0,
  0,0,0,0, s_Imm_20y11*r_Imm*b_Imm*k_Imm, 0,0,0,0,0,0,
  0,0,0,0, s_Imm_20y11*r_Imm*b_Imm*(1-k_Imm), 0,0,0,0,0,0,
  0,0,0,0,0, s_IS_20y11*r_IS*b_IS*k_IS, s_IF_20y11*r_IF*b_IF*k_IF, s_ES_20y11*r_ES*b_ES*k_ES, s_EF_20y11*r_EF*b_EF*k_EF, s_ENB_20y11*r_ENB*b_ENB*k_ENB, s_PF_20y11*r_PF*b_PF*k_PF,
  0,0,0,0,0, s_IS_20y11*r_IS*b_IS*(1-k_IS), s_IF_20y11*r_IF*b_IF*(1-k_IF), s_ES_20y11*r_ES*b_ES*(1-k_ES), s_EF_20y11*r_EF*b_EF*(1-k_EF), s_ENB_20y11*r_ENB*b_ENB*(1-k_ENB), s_PF_20y11*r_PF*b_PF*(1-k_PF),
  0,0,0,0,0, s_IS_20y11*r_IS*(1-b_IS), s_IF_20y11*r_IF*(1-b_IF), s_ES_20y11*r_ES*(1-b_ES), s_EF_20y11*r_EF*(1-b_EF), s_ENB_20y11*r_ENB*(1-b_ENB), s_PF_20y11*r_PF*(1-b_PF),
  0,0,0,0,0, s_IS_20y11*(1-r_IS), s_IF_20y11*(1-r_IF), s_ES_20y11*(1-r_ES), s_EF_20y11*(1-r_EF), s_ENB_20y11*(1-r_ENB), s_PF_20y11*(1-r_PF)
), nrow=11, byrow=TRUE)

m_20y12 <- matrix(data = c(
  0,0,0,0,0, 0.5*s_Juv_20y12,0,0.5*s_Juv_20y12,0,0,0,
  s_Juv_20y12,0,0,0,0, 0,0,0,0,0,0,
  0,s_Juv_20y12,0,0,0, 0,0,0,0,0,0,
  0,0,s_Imm_20y12,0,0, 0,0,0,0,0,0,
  0,0,0,s_Imm_20y12, s_Imm_20y12*r_Imm*(1-b_Imm)+s_Imm_20y12*(1-r_Imm), 0,0,0,0,0,0,
  0,0,0,0, s_Imm_20y12*r_Imm*b_Imm*k_Imm, 0,0,0,0,0,0,
  0,0,0,0, s_Imm_20y12*r_Imm*b_Imm*(1-k_Imm), 0,0,0,0,0,0,
  0,0,0,0,0, s_IS_20y12*r_IS*b_IS*k_IS, s_IF_20y12*r_IF*b_IF*k_IF, s_ES_20y12*r_ES*b_ES*k_ES, s_EF_20y12*r_EF*b_EF*k_EF, s_ENB_20y12*r_ENB*b_ENB*k_ENB, s_PF_20y12*r_PF*b_PF*k_PF,
  0,0,0,0,0, s_IS_20y12*r_IS*b_IS*(1-k_IS), s_IF_20y12*r_IF*b_IF*(1-k_IF), s_ES_20y12*r_ES*b_ES*(1-k_ES), s_EF_20y12*r_EF*b_EF*(1-k_EF), s_ENB_20y12*r_ENB*b_ENB*(1-k_ENB), s_PF_20y12*r_PF*b_PF*(1-k_PF),
  0,0,0,0,0, s_IS_20y12*r_IS*(1-b_IS), s_IF_20y12*r_IF*(1-b_IF), s_ES_20y12*r_ES*(1-b_ES), s_EF_20y12*r_EF*(1-b_EF), s_ENB_20y12*r_ENB*(1-b_ENB), s_PF_20y12*r_PF*(1-b_PF),
  0,0,0,0,0, s_IS_20y12*(1-r_IS), s_IF_20y12*(1-r_IF), s_ES_20y12*(1-r_ES), s_EF_20y12*(1-r_EF), s_ENB_20y12*(1-r_ENB), s_PF_20y12*(1-r_PF)
), nrow=11, byrow=TRUE)

m_20y13 <- matrix(data = c(
  0,0,0,0,0, 0.5*s_Juv_20y13,0,0.5*s_Juv_20y13,0,0,0,
  s_Juv_20y13,0,0,0,0, 0,0,0,0,0,0,
  0,s_Juv_20y13,0,0,0, 0,0,0,0,0,0,
  0,0,s_Imm_20y13,0,0, 0,0,0,0,0,0,
  0,0,0,s_Imm_20y13, s_Imm_20y13*r_Imm*(1-b_Imm)+s_Imm_20y13*(1-r_Imm), 0,0,0,0,0,0,
  0,0,0,0, s_Imm_20y13*r_Imm*b_Imm*k_Imm, 0,0,0,0,0,0,
  0,0,0,0, s_Imm_20y13*r_Imm*b_Imm*(1-k_Imm), 0,0,0,0,0,0,
  0,0,0,0,0, s_IS_20y13*r_IS*b_IS*k_IS, s_IF_20y13*r_IF*b_IF*k_IF, s_ES_20y13*r_ES*b_ES*k_ES, s_EF_20y13*r_EF*b_EF*k_EF, s_ENB_20y13*r_ENB*b_ENB*k_ENB, s_PF_20y13*r_PF*b_PF*k_PF,
  0,0,0,0,0, s_IS_20y13*r_IS*b_IS*(1-k_IS), s_IF_20y13*r_IF*b_IF*(1-k_IF), s_ES_20y13*r_ES*b_ES*(1-k_ES), s_EF_20y13*r_EF*b_EF*(1-k_EF), s_ENB_20y13*r_ENB*b_ENB*(1-k_ENB), s_PF_20y13*r_PF*b_PF*(1-k_PF),
  0,0,0,0,0, s_IS_20y13*r_IS*(1-b_IS), s_IF_20y13*r_IF*(1-b_IF), s_ES_20y13*r_ES*(1-b_ES), s_EF_20y13*r_EF*(1-b_EF), s_ENB_20y13*r_ENB*(1-b_ENB), s_PF_20y13*r_PF*(1-b_PF),
  0,0,0,0,0, s_IS_20y13*(1-r_IS), s_IF_20y13*(1-r_IF), s_ES_20y13*(1-r_ES), s_EF_20y13*(1-r_EF), s_ENB_20y13*(1-r_ENB), s_PF_20y13*(1-r_PF)
), nrow=11, byrow=TRUE)

m_20y14 <- matrix(data = c(
  0,0,0,0,0, 0.5*s_Juv_20y14,0,0.5*s_Juv_20y14,0,0,0,
  s_Juv_20y14,0,0,0,0, 0,0,0,0,0,0,
  0,s_Juv_20y14,0,0,0, 0,0,0,0,0,0,
  0,0,s_Imm_20y14,0,0, 0,0,0,0,0,0,
  0,0,0,s_Imm_20y14, s_Imm_20y14*r_Imm*(1-b_Imm)+s_Imm_20y14*(1-r_Imm), 0,0,0,0,0,0,
  0,0,0,0, s_Imm_20y14*r_Imm*b_Imm*k_Imm, 0,0,0,0,0,0,
  0,0,0,0, s_Imm_20y14*r_Imm*b_Imm*(1-k_Imm), 0,0,0,0,0,0,
  0,0,0,0,0, s_IS_20y14*r_IS*b_IS*k_IS, s_IF_20y14*r_IF*b_IF*k_IF, s_ES_20y14*r_ES*b_ES*k_ES, s_EF_20y14*r_EF*b_EF*k_EF, s_ENB_20y14*r_ENB*b_ENB*k_ENB, s_PF_20y14*r_PF*b_PF*k_PF,
  0,0,0,0,0, s_IS_20y14*r_IS*b_IS*(1-k_IS), s_IF_20y14*r_IF*b_IF*(1-k_IF), s_ES_20y14*r_ES*b_ES*(1-k_ES), s_EF_20y14*r_EF*b_EF*(1-k_EF), s_ENB_20y14*r_ENB*b_ENB*(1-k_ENB), s_PF_20y14*r_PF*b_PF*(1-k_PF),
  0,0,0,0,0, s_IS_20y14*r_IS*(1-b_IS), s_IF_20y14*r_IF*(1-b_IF), s_ES_20y14*r_ES*(1-b_ES), s_EF_20y14*r_EF*(1-b_EF), s_ENB_20y14*r_ENB*(1-b_ENB), s_PF_20y14*r_PF*(1-b_PF),
  0,0,0,0,0, s_IS_20y14*(1-r_IS), s_IF_20y14*(1-r_IF), s_ES_20y14*(1-r_ES), s_EF_20y14*(1-r_EF), s_ENB_20y14*(1-r_ENB), s_PF_20y14*(1-r_PF)
), nrow=11, byrow=TRUE)

m_20y15 <- matrix(data = c(
  0,0,0,0,0, 0.5*s_Juv_20y15,0,0.5*s_Juv_20y15,0,0,0,
  s_Juv_20y15,0,0,0,0, 0,0,0,0,0,0,
  0,s_Juv_20y15,0,0,0, 0,0,0,0,0,0,
  0,0,s_Imm_20y15,0,0, 0,0,0,0,0,0,
  0,0,0,s_Imm_20y15, s_Imm_20y15*r_Imm*(1-b_Imm)+s_Imm_20y15*(1-r_Imm), 0,0,0,0,0,0,
  0,0,0,0, s_Imm_20y15*r_Imm*b_Imm*k_Imm, 0,0,0,0,0,0,
  0,0,0,0, s_Imm_20y15*r_Imm*b_Imm*(1-k_Imm), 0,0,0,0,0,0,
  0,0,0,0,0, s_IS_20y15*r_IS*b_IS*k_IS, s_IF_20y15*r_IF*b_IF*k_IF, s_ES_20y15*r_ES*b_ES*k_ES, s_EF_20y15*r_EF*b_EF*k_EF, s_ENB_20y15*r_ENB*b_ENB*k_ENB, s_PF_20y15*r_PF*b_PF*k_PF,
  0,0,0,0,0, s_IS_20y15*r_IS*b_IS*(1-k_IS), s_IF_20y15*r_IF*b_IF*(1-k_IF), s_ES_20y15*r_ES*b_ES*(1-k_ES), s_EF_20y15*r_EF*b_EF*(1-k_EF), s_ENB_20y15*r_ENB*b_ENB*(1-k_ENB), s_PF_20y15*r_PF*b_PF*(1-k_PF),
  0,0,0,0,0, s_IS_20y15*r_IS*(1-b_IS), s_IF_20y15*r_IF*(1-b_IF), s_ES_20y15*r_ES*(1-b_ES), s_EF_20y15*r_EF*(1-b_EF), s_ENB_20y15*r_ENB*(1-b_ENB), s_PF_20y15*r_PF*(1-b_PF),
  0,0,0,0,0, s_IS_20y15*(1-r_IS), s_IF_20y15*(1-r_IF), s_ES_20y15*(1-r_ES), s_EF_20y15*(1-r_EF), s_ENB_20y15*(1-r_ENB), s_PF_20y15*(1-r_PF)
), nrow=11, byrow=TRUE)

m_20y16 <- matrix(data = c(
  0,0,0,0,0, 0.5*s_Juv_20y16,0,0.5*s_Juv_20y16,0,0,0,
  s_Juv_20y16,0,0,0,0, 0,0,0,0,0,0,
  0,s_Juv_20y16,0,0,0, 0,0,0,0,0,0,
  0,0,s_Imm_20y16,0,0, 0,0,0,0,0,0,
  0,0,0,s_Imm_20y16, s_Imm_20y16*r_Imm*(1-b_Imm)+s_Imm_20y16*(1-r_Imm), 0,0,0,0,0,0,
  0,0,0,0, s_Imm_20y16*r_Imm*b_Imm*k_Imm, 0,0,0,0,0,0,
  0,0,0,0, s_Imm_20y16*r_Imm*b_Imm*(1-k_Imm), 0,0,0,0,0,0,
  0,0,0,0,0, s_IS_20y16*r_IS*b_IS*k_IS, s_IF_20y16*r_IF*b_IF*k_IF, s_ES_20y16*r_ES*b_ES*k_ES, s_EF_20y16*r_EF*b_EF*k_EF, s_ENB_20y16*r_ENB*b_ENB*k_ENB, s_PF_20y16*r_PF*b_PF*k_PF,
  0,0,0,0,0, s_IS_20y16*r_IS*b_IS*(1-k_IS), s_IF_20y16*r_IF*b_IF*(1-k_IF), s_ES_20y16*r_ES*b_ES*(1-k_ES), s_EF_20y16*r_EF*b_EF*(1-k_EF), s_ENB_20y16*r_ENB*b_ENB*(1-k_ENB), s_PF_20y16*r_PF*b_PF*(1-k_PF),
  0,0,0,0,0, s_IS_20y16*r_IS*(1-b_IS), s_IF_20y16*r_IF*(1-b_IF), s_ES_20y16*r_ES*(1-b_ES), s_EF_20y16*r_EF*(1-b_EF), s_ENB_20y16*r_ENB*(1-b_ENB), s_PF_20y16*r_PF*(1-b_PF),
  0,0,0,0,0, s_IS_20y16*(1-r_IS), s_IF_20y16*(1-r_IF), s_ES_20y16*(1-r_ES), s_EF_20y16*(1-r_EF), s_ENB_20y16*(1-r_ENB), s_PF_20y16*(1-r_PF)
), nrow=11, byrow=TRUE)

m_20y17 <- matrix(data = c(
  0,0,0,0,0, 0.5*s_Juv_20y17,0,0.5*s_Juv_20y17,0,0,0,
  s_Juv_20y17,0,0,0,0, 0,0,0,0,0,0,
  0,s_Juv_20y17,0,0,0, 0,0,0,0,0,0,
  0,0,s_Imm_20y17,0,0, 0,0,0,0,0,0,
  0,0,0,s_Imm_20y17, s_Imm_20y17*r_Imm*(1-b_Imm)+s_Imm_20y17*(1-r_Imm), 0,0,0,0,0,0,
  0,0,0,0, s_Imm_20y17*r_Imm*b_Imm*k_Imm, 0,0,0,0,0,0,
  0,0,0,0, s_Imm_20y17*r_Imm*b_Imm*(1-k_Imm), 0,0,0,0,0,0,
  0,0,0,0,0, s_IS_20y17*r_IS*b_IS*k_IS, s_IF_20y17*r_IF*b_IF*k_IF, s_ES_20y17*r_ES*b_ES*k_ES, s_EF_20y17*r_EF*b_EF*k_EF, s_ENB_20y17*r_ENB*b_ENB*k_ENB, s_PF_20y17*r_PF*b_PF*k_PF,
  0,0,0,0,0, s_IS_20y17*r_IS*b_IS*(1-k_IS), s_IF_20y17*r_IF*b_IF*(1-k_IF), s_ES_20y17*r_ES*b_ES*(1-k_ES), s_EF_20y17*r_EF*b_EF*(1-k_EF), s_ENB_20y17*r_ENB*b_ENB*(1-k_ENB), s_PF_20y17*r_PF*b_PF*(1-k_PF),
  0,0,0,0,0, s_IS_20y17*r_IS*(1-b_IS), s_IF_20y17*r_IF*(1-b_IF), s_ES_20y17*r_ES*(1-b_ES), s_EF_20y17*r_EF*(1-b_EF), s_ENB_20y17*r_ENB*(1-b_ENB), s_PF_20y17*r_PF*(1-b_PF),
  0,0,0,0,0, s_IS_20y17*(1-r_IS), s_IF_20y17*(1-r_IF), s_ES_20y17*(1-r_ES), s_EF_20y17*(1-r_EF), s_ENB_20y17*(1-r_ENB), s_PF_20y17*(1-r_PF)
), nrow=11, byrow=TRUE)

m_20y18 <- matrix(data = c(
  0,0,0,0,0, 0.5*s_Juv_20y18,0,0.5*s_Juv_20y18,0,0,0,
  s_Juv_20y18,0,0,0,0, 0,0,0,0,0,0,
  0,s_Juv_20y18,0,0,0, 0,0,0,0,0,0,
  0,0,s_Imm_20y18,0,0, 0,0,0,0,0,0,
  0,0,0,s_Imm_20y18, s_Imm_20y18*r_Imm*(1-b_Imm)+s_Imm_20y18*(1-r_Imm), 0,0,0,0,0,0,
  0,0,0,0, s_Imm_20y18*r_Imm*b_Imm*k_Imm, 0,0,0,0,0,0,
  0,0,0,0, s_Imm_20y18*r_Imm*b_Imm*(1-k_Imm), 0,0,0,0,0,0,
  0,0,0,0,0, s_IS_20y18*r_IS*b_IS*k_IS, s_IF_20y18*r_IF*b_IF*k_IF, s_ES_20y18*r_ES*b_ES*k_ES, s_EF_20y18*r_EF*b_EF*k_EF, s_ENB_20y18*r_ENB*b_ENB*k_ENB, s_PF_20y18*r_PF*b_PF*k_PF,
  0,0,0,0,0, s_IS_20y18*r_IS*b_IS*(1-k_IS), s_IF_20y18*r_IF*b_IF*(1-k_IF), s_ES_20y18*r_ES*b_ES*(1-k_ES), s_EF_20y18*r_EF*b_EF*(1-k_EF), s_ENB_20y18*r_ENB*b_ENB*(1-k_ENB), s_PF_20y18*r_PF*b_PF*(1-k_PF),
  0,0,0,0,0, s_IS_20y18*r_IS*(1-b_IS), s_IF_20y18*r_IF*(1-b_IF), s_ES_20y18*r_ES*(1-b_ES), s_EF_20y18*r_EF*(1-b_EF), s_ENB_20y18*r_ENB*(1-b_ENB), s_PF_20y18*r_PF*(1-b_PF),
  0,0,0,0,0, s_IS_20y18*(1-r_IS), s_IF_20y18*(1-r_IF), s_ES_20y18*(1-r_ES), s_EF_20y18*(1-r_EF), s_ENB_20y18*(1-r_ENB), s_PF_20y18*(1-r_PF)
), nrow=11, byrow=TRUE)

m_20y19 <- matrix(data = c(
  0,0,0,0,0, 0.5*s_Juv_20y19,0,0.5*s_Juv_20y19,0,0,0,
  s_Juv_20y19,0,0,0,0, 0,0,0,0,0,0,
  0,s_Juv_20y19,0,0,0, 0,0,0,0,0,0,
  0,0,s_Imm_20y19,0,0, 0,0,0,0,0,0,
  0,0,0,s_Imm_20y19, s_Imm_20y19*r_Imm*(1-b_Imm)+s_Imm_20y19*(1-r_Imm), 0,0,0,0,0,0,
  0,0,0,0, s_Imm_20y19*r_Imm*b_Imm*k_Imm, 0,0,0,0,0,0,
  0,0,0,0, s_Imm_20y19*r_Imm*b_Imm*(1-k_Imm), 0,0,0,0,0,0,
  0,0,0,0,0, s_IS_20y19*r_IS*b_IS*k_IS, s_IF_20y19*r_IF*b_IF*k_IF, s_ES_20y19*r_ES*b_ES*k_ES, s_EF_20y19*r_EF*b_EF*k_EF, s_ENB_20y19*r_ENB*b_ENB*k_ENB, s_PF_20y19*r_PF*b_PF*k_PF,
  0,0,0,0,0, s_IS_20y19*r_IS*b_IS*(1-k_IS), s_IF_20y19*r_IF*b_IF*(1-k_IF), s_ES_20y19*r_ES*b_ES*(1-k_ES), s_EF_20y19*r_EF*b_EF*(1-k_EF), s_ENB_20y19*r_ENB*b_ENB*(1-k_ENB), s_PF_20y19*r_PF*b_PF*(1-k_PF),
  0,0,0,0,0, s_IS_20y19*r_IS*(1-b_IS), s_IF_20y19*r_IF*(1-b_IF), s_ES_20y19*r_ES*(1-b_ES), s_EF_20y19*r_EF*(1-b_EF), s_ENB_20y19*r_ENB*(1-b_ENB), s_PF_20y19*r_PF*(1-b_PF),
  0,0,0,0,0, s_IS_20y19*(1-r_IS), s_IF_20y19*(1-r_IF), s_ES_20y19*(1-r_ES), s_EF_20y19*(1-r_EF), s_ENB_20y19*(1-r_ENB), s_PF_20y19*(1-r_PF)
), nrow=11, byrow=TRUE)

# m_20y20: 100% improvement — should equal m_3of3
m_20y20 <- matrix(data = c(
  0,0,0,0,0, 0.5*s_Juv_20y20,0,0.5*s_Juv_20y20,0,0,0,
  s_Juv_20y20,0,0,0,0, 0,0,0,0,0,0,
  0,s_Juv_20y20,0,0,0, 0,0,0,0,0,0,
  0,0,s_Imm_20y20,0,0, 0,0,0,0,0,0,
  0,0,0,s_Imm_20y20, s_Imm_20y20*r_Imm*(1-b_Imm)+s_Imm_20y20*(1-r_Imm), 0,0,0,0,0,0,
  0,0,0,0, s_Imm_20y20*r_Imm*b_Imm*k_Imm, 0,0,0,0,0,0,
  0,0,0,0, s_Imm_20y20*r_Imm*b_Imm*(1-k_Imm), 0,0,0,0,0,0,
  0,0,0,0,0, s_IS_20y20*r_IS*b_IS*k_IS, s_IF_20y20*r_IF*b_IF*k_IF, s_ES_20y20*r_ES*b_ES*k_ES, s_EF_20y20*r_EF*b_EF*k_EF, s_ENB_20y20*r_ENB*b_ENB*k_ENB, s_PF_20y20*r_PF*b_PF*k_PF,
  0,0,0,0,0, s_IS_20y20*r_IS*b_IS*(1-k_IS), s_IF_20y20*r_IF*b_IF*(1-k_IF), s_ES_20y20*r_ES*b_ES*(1-k_ES), s_EF_20y20*r_EF*b_EF*(1-k_EF), s_ENB_20y20*r_ENB*b_ENB*(1-k_ENB), s_PF_20y20*r_PF*b_PF*(1-k_PF),
  0,0,0,0,0, s_IS_20y20*r_IS*(1-b_IS), s_IF_20y20*r_IF*(1-b_IF), s_ES_20y20*r_ES*(1-b_ES), s_EF_20y20*r_EF*(1-b_EF), s_ENB_20y20*r_ENB*(1-b_ENB), s_PF_20y20*r_PF*(1-b_PF),
  0,0,0,0,0, s_IS_20y20*(1-r_IS), s_IF_20y20*(1-r_IF), s_ES_20y20*(1-r_ES), s_EF_20y20*(1-r_EF), s_ENB_20y20*(1-r_ENB), s_PF_20y20*(1-r_PF)
), nrow=11, byrow=TRUE)

phase_list_20y <- list(m_20y1,m_20y2,m_20y3,m_20y4,m_20y5,
                       m_20y6,m_20y7,m_20y8,m_20y9,m_20y10,
                       m_20y11,m_20y12,m_20y13,m_20y14,m_20y15,
                       m_20y16,m_20y17,m_20y18,m_20y19,m_20y20)
```

---

*PART 6 — ALL PROJECTIONS AND COMPARISON PLOTS*

Matrix sequence for each scenario:
- 10-yr phase-in:            k=1-10: phase_list_10y[[k]];  k>10: m_3of3
- 20-yr phase-in:            k=1-20: phase_list_20y[[k]];  k>20: m_3of3
- 5-yr delay + immediate:    k=1-5:  m_base;               k>5:  m_3of3
- 10-yr delay + immediate:   k=1-10: m_base;               k>10: m_3of3
- 20-yr delay + immediate:   k=1-20: m_base;               k>20: m_3of3
- 5-yr delay + 5-yr phase:   k=1-5:  m_base; k=6-10: phase_list[[k-5]];  k>10: m_3of3
- 10-yr delay + 5-yr phase:  k=1-10: m_base; k=11-15: phase_list[[k-10]]; k>15: m_3of3
- 20-yr delay + 5-yr phase:  k=1-20: m_base; k=21-25: phase_list[[k-20]]; k>25: m_3of3

```{r part6-all-projections}
# Helper: run matrix projection and return total-N data frame
run_proj <- function(phase_fn, scenario_label) {
  mat <- matrix(0, nrow = years + 1, ncol = 11)
  colnames(mat) <- stage_names
  mat[1, ] <- n0_vec
  for (k in 1:years) {
    mat[k + 1, ] <- phase_fn(k) %*% mat[k, ]
  }
  as.data.frame(mat) %>%
    mutate(year = 0:years,
           N    = rowSums(across(all_of(stage_names))),
           scenario = scenario_label) %>%
    select(year, N, scenario)
}

# 10-year phase-in
proj_10y <- run_proj(
  function(k) if (k <= 10) phase_list_10y[[k]] else m_3of3,
  "10-year phase-in")

# 20-year phase-in
proj_20y <- run_proj(
  function(k) if (k <= 20) phase_list_20y[[k]] else m_3of3,
  "20-year phase-in")

# 5-year delay then immediate 3/3
proj_d5_imm <- run_proj(
  function(k) if (k <= 5)  m_base else m_3of3,
  "5-yr delay, then immediate 3/3")

# 10-year delay then immediate 3/3
proj_d10_imm <- run_proj(
  function(k) if (k <= 10) m_base else m_3of3,
  "10-yr delay, then immediate 3/3")

# 20-year delay then immediate 3/3
proj_d20_imm <- run_proj(
  function(k) if (k <= 20) m_base else m_3of3,
  "20-yr delay, then immediate 3/3")

# 5-year delay then 5-year phase-in
proj_d5_p5 <- run_proj(
  function(k) {
    if      (k <= 5)  m_base
    else if (k <= 10) phase_list[[k - 5]]
    else              m_3of3
  },
  "5-yr delay, then 5-yr phase-in")

# 10-year delay then 5-year phase-in
proj_d10_p5 <- run_proj(
  function(k) {
    if      (k <= 10) m_base
    else if (k <= 15) phase_list[[k - 10]]
    else              m_3of3
  },
  "10-yr delay, then 5-yr phase-in")

# 20-year delay then 5-year phase-in
proj_d20_p5 <- run_proj(
  function(k) {
    if      (k <= 20) m_base
    else if (k <= 25) phase_list[[k - 20]]
    else              m_3of3
  },
  "20-yr delay, then 5-yr phase-in")

# Combine all scenarios (including Part 1 scalar and Part 3 phase-in already computed)
all_scenarios <- bind_rows(
  proj_part1,           # No mitigation + Immediate 3/3
  proj_part3_total,     # 5-year phase-in
  proj_10y,
  proj_20y,
  proj_d5_imm,
  proj_d10_imm,
  proj_d20_imm,
  proj_d5_p5,
  proj_d10_p5,
  proj_d20_p5
) %>%
  mutate(scenario = factor(scenario, levels = c(
    "No mitigation (baseline)",
    "Immediate 3/3 (100% coverage)",
    "5-year phase-in",
    "10-year phase-in",
    "20-year phase-in",
    "5-yr delay, then immediate 3/3",
    "10-yr delay, then immediate 3/3",
    "20-yr delay, then immediate 3/3",
    "5-yr delay, then 5-yr phase-in",
    "10-yr delay, then 5-yr phase-in",
    "20-yr delay, then 5-yr phase-in"
  )))
```

**plot Part 6a — phase-in speed comparison**
```{r plot-part6a, fig.width = 9, fig.height = 5}
p_phasein_compare <- all_scenarios %>%
  filter(scenario %in% c("No mitigation (baseline)",
                          "Immediate 3/3 (100% coverage)",
                          "5-year phase-in",
                          "10-year phase-in",
                          "20-year phase-in")) %>%
  ggplot(aes(x = year, y = N, color = scenario, linetype = scenario)) +
  geom_line(linewidth = 0.9) +
  geom_hline(yintercept = N0, color = "grey50", linetype = "dashed", linewidth = 0.4) +
  scale_color_manual(values = c(
    "No mitigation (baseline)"      = "#d73027",
    "Immediate 3/3 (100% coverage)" = "#253494",
    "5-year phase-in"               = "#1a9641",
    "10-year phase-in"              = "#78c679",
    "20-year phase-in"              = "#c2e699")) +
  scale_linetype_manual(values = c(
    "No mitigation (baseline)"      = "dashed",
    "Immediate 3/3 (100% coverage)" = "solid",
    "5-year phase-in"               = "solid",
    "10-year phase-in"              = "solid",
    "20-year phase-in"              = "solid")) +
  scale_x_continuous(breaks = seq(0, years, 10)) +
  scale_y_continuous(labels = scales::comma, breaks = scales::pretty_breaks(n = 6)) +
  labs(x = "Years since 1990 baseline", y = "Population (N individuals)",
       color = NULL, linetype = NULL,
       title = paste0("Phase-in speed comparison — ", target_period, " baseline"),
       subtitle = "Darker green = faster phase-in") +
  themeo + theme(legend.position = "bottom")
p_phasein_compare
```

**plot Part 6b — delay then immediate comparison**
```{r plot-part6b, fig.width = 9, fig.height = 5}
p_delay_imm <- all_scenarios %>%
  filter(scenario %in% c("No mitigation (baseline)",
                          "Immediate 3/3 (100% coverage)",
                          "5-yr delay, then immediate 3/3",
                          "10-yr delay, then immediate 3/3",
                          "20-yr delay, then immediate 3/3")) %>%
  ggplot(aes(x = year, y = N, color = scenario, linetype = scenario)) +
  geom_line(linewidth = 0.9) +
  geom_hline(yintercept = N0, color = "grey50", linetype = "dashed", linewidth = 0.4) +
  scale_color_manual(values = c(
    "No mitigation (baseline)"       = "#d73027",
    "Immediate 3/3 (100% coverage)"  = "#253494",
    "5-yr delay, then immediate 3/3" = "#2166ac",
    "10-yr delay, then immediate 3/3"= "#74add1",
    "20-yr delay, then immediate 3/3"= "#abd9e9")) +
  scale_linetype_manual(values = c(
    "No mitigation (baseline)"       = "dashed",
    "Immediate 3/3 (100% coverage)"  = "solid",
    "5-yr delay, then immediate 3/3" = "solid",
    "10-yr delay, then immediate 3/3"= "solid",
    "20-yr delay, then immediate 3/3"= "solid")) +
  scale_x_continuous(breaks = seq(0, years, 10)) +
  scale_y_continuous(labels = scales::comma, breaks = scales::pretty_breaks(n = 6)) +
  labs(x = "Years since 1990 baseline", y = "Population (N individuals)",
       color = NULL, linetype = NULL,
       title = paste0("Cost of delaying immediate 3/3 — ", target_period, " baseline"),
       subtitle = "Darker blue = shorter delay before full implementation") +
  themeo + theme(legend.position = "bottom")
p_delay_imm
```

**plot Part 6c — delay then 5-year phase-in comparison**
```{r plot-part6c, fig.width = 9, fig.height = 5}
p_delay_phase <- all_scenarios %>%
  filter(scenario %in% c("No mitigation (baseline)",
                          "Immediate 3/3 (100% coverage)",
                          "5-yr delay, then 5-yr phase-in",
                          "10-yr delay, then 5-yr phase-in",
                          "20-yr delay, then 5-yr phase-in")) %>%
  ggplot(aes(x = year, y = N, color = scenario, linetype = scenario)) +
  geom_line(linewidth = 0.9) +
  geom_hline(yintercept = N0, color = "grey50", linetype = "dashed", linewidth = 0.4) +
  scale_color_manual(values = c(
    "No mitigation (baseline)"        = "#d73027",
    "Immediate 3/3 (100% coverage)"   = "#253494",
    "5-yr delay, then 5-yr phase-in"  = "#762a83",
    "10-yr delay, then 5-yr phase-in" = "#af8dc3",
    "20-yr delay, then 5-yr phase-in" = "#e7d4e8")) +
  scale_linetype_manual(values = c(
    "No mitigation (baseline)"        = "dashed",
    "Immediate 3/3 (100% coverage)"   = "solid",
    "5-yr delay, then 5-yr phase-in"  = "solid",
    "10-yr delay, then 5-yr phase-in" = "solid",
    "20-yr delay, then 5-yr phase-in" = "solid")) +
  scale_x_continuous(breaks = seq(0, years, 10)) +
  scale_y_continuous(labels = scales::comma, breaks = scales::pretty_breaks(n = 6)) +
  labs(x = "Years since 1990 baseline", y = "Population (N individuals)",
       color = NULL, linetype = NULL,
       title = paste0("Cost of delaying 5-yr phase-in — ", target_period, " baseline"),
       subtitle = "Darker purple = shorter delay before phase-in begins") +
  themeo + theme(legend.position = "bottom")
p_delay_phase
```

**plot Part 6d — all scenarios**
```{r plot-part6d, fig.width = 11, fig.height = 6}
p_all_scenarios <- ggplot(all_scenarios,
                           aes(x = year, y = N, color = scenario, linetype = scenario)) +
  geom_line(linewidth = 0.7) +
  geom_hline(yintercept = N0, color = "grey50", linetype = "dashed", linewidth = 0.4) +
  scale_color_manual(values = c(
    "No mitigation (baseline)"        = "#d73027",
    "Immediate 3/3 (100% coverage)"   = "#253494",
    "5-year phase-in"                 = "#1a9641",
    "10-year phase-in"                = "#78c679",
    "20-year phase-in"                = "#addd8e",
    "5-yr delay, then immediate 3/3"  = "#2166ac",
    "10-yr delay, then immediate 3/3" = "#74add1",
    "20-yr delay, then immediate 3/3" = "#abd9e9",
    "5-yr delay, then 5-yr phase-in"  = "#762a83",
    "10-yr delay, then 5-yr phase-in" = "#af8dc3",
    "20-yr delay, then 5-yr phase-in" = "#e7d4e8")) +
  scale_linetype_manual(values = c(
    "No mitigation (baseline)"        = "dashed",
    "Immediate 3/3 (100% coverage)"   = "solid",
    "5-year phase-in"                 = "solid",
    "10-year phase-in"                = "solid",
    "20-year phase-in"                = "solid",
    "5-yr delay, then immediate 3/3"  = "solid",
    "10-yr delay, then immediate 3/3" = "solid",
    "20-yr delay, then immediate 3/3" = "solid",
    "5-yr delay, then 5-yr phase-in"  = "solid",
    "10-yr delay, then 5-yr phase-in" = "solid",
    "20-yr delay, then 5-yr phase-in" = "solid")) +
  scale_x_continuous(breaks = seq(0, years, 10)) +
  scale_y_continuous(labels = scales::comma, breaks = scales::pretty_breaks(n = 6)) +
  labs(x = "Years since 1990 baseline", y = "Population (N individuals)",
       color = NULL, linetype = NULL,
       title = paste0("All mitigation scenarios — ", target_period, " baseline"),
       subtitle = "Green = phase-in speed; blue = delay then immediate; purple = delay then phase-in") +
  themeo +
  theme(legend.position = "right",
        legend.text = element_text(size = 7))
p_all_scenarios
```

**plot Part 6e — all scenarios with observed breeding pairs**
```{r plot-part6e, fig.width = 11, fig.height = 6}
p_all_scenarios_obs <- ggplot(all_scenarios,
                               aes(x = year, y = N, color = scenario, linetype = scenario)) +
  geom_line(linewidth = 0.7) +
  geom_hline(yintercept = N0, color = "grey50", linetype = "dashed", linewidth = 0.4) +
  geom_point(data = bird_obs,
             aes(x = year, y = N_scaled),
             inherit.aes = FALSE, color = "#e6550d", size = 3, shape = 16) +
  geom_line(data = bird_obs,
            aes(x = year, y = N_scaled),
            inherit.aes = FALSE, color = "#e6550d", linewidth = 0.5, linetype = "dotted") +
  geom_text(data = bird_obs,
            aes(x = year, y = N_scaled, label = season),
            inherit.aes = FALSE, color = "#e6550d", size = 2.5, vjust = -0.8) +
  scale_color_manual(values = c(
    "No mitigation (baseline)"        = "#d73027",
    "Immediate 3/3 (100% coverage)"   = "#253494",
    "5-year phase-in"                 = "#1a9641",
    "10-year phase-in"                = "#78c679",
    "20-year phase-in"                = "#addd8e",
    "5-yr delay, then immediate 3/3"  = "#2166ac",
    "10-yr delay, then immediate 3/3" = "#74add1",
    "20-yr delay, then immediate 3/3" = "#abd9e9",
    "5-yr delay, then 5-yr phase-in"  = "#762a83",
    "10-yr delay, then 5-yr phase-in" = "#af8dc3",
    "20-yr delay, then 5-yr phase-in" = "#e7d4e8")) +
  scale_linetype_manual(values = c(
    "No mitigation (baseline)"        = "dashed",
    "Immediate 3/3 (100% coverage)"   = "solid",
    "5-year phase-in"                 = "solid",
    "10-year phase-in"                = "solid",
    "20-year phase-in"                = "solid",
    "5-yr delay, then immediate 3/3"  = "solid",
    "10-yr delay, then immediate 3/3" = "solid",
    "20-yr delay, then immediate 3/3" = "solid",
    "5-yr delay, then 5-yr phase-in"  = "solid",
    "10-yr delay, then 5-yr phase-in" = "solid",
    "20-yr delay, then 5-yr phase-in" = "solid")) +
  scale_x_continuous(breaks = seq(-10, years, 10), limits = c(-10, years)) +
  scale_y_continuous(labels = scales::comma,
                     breaks = scales::pretty_breaks(n = 6),
                     sec.axis = sec_axis(~ . / pairs_scale,
                                         name = "Breeding pairs (Bird Island)",
                                         labels = scales::comma)) +
  labs(x = "Years since 1990 baseline", y = "Population (N individuals)",
       color = NULL, linetype = NULL,
       title = paste0("All mitigation scenarios with observed breeding pairs — ",
                      target_period, " baseline"),
       subtitle = "Orange dots = observed Bird Island breeding pairs (Poncet et al.); secondary axis scaled to total N") +
  themeo +
  theme(legend.position = "right",
        legend.text = element_text(size = 7))
p_all_scenarios_obs
```

---

*PART 7 — TRANSIENT DYNAMICS: SSD SHIFT AND REALIZED LAMBDA*

As mitigation improves survival, the population SSD shifts toward a higher proportion of
breeders. During the transition from m_base SSD to m_3of3 SSD, the realized growth rate
each year (N(t+1)/N(t)) may temporarily exceed the asymptotic lambda_3of3 as the
immature recruitment pulse moves into the breeding stages.

**SSD comparison: baseline vs full mitigation**
```{r part7-ssd-compare}
# Compute SSD for full mitigation matrix
SSD_3of3 <- as.numeric(eigen(m_3of3)$vectors[, 1] / sum(eigen(m_3of3)$vectors[, 1]))

ssd_df <- tibble(
  stage    = stage_names,
  baseline = SSD,
  full_3of3 = SSD_3of3
) %>%
  mutate(
    group = case_when(
      stage %in% c("J2", "J3")              ~ "Juveniles",
      stage %in% c("Imm4", "Imm5", "PB")   ~ "Immatures / pre-breeders",
      stage %in% c("IS", "IF", "ES", "EF") ~ "Breeders",
      stage %in% c("ENB", "Sabb")          ~ "Non-breeders"
    ),
    stage = factor(stage, levels = stage_names)
  ) %>%
  pivot_longer(cols = c(baseline, full_3of3),
               names_to  = "matrix",
               values_to = "proportion") %>%
  mutate(matrix = recode(matrix,
    baseline  = "No mitigation (baseline)",
    full_3of3 = "Full 3/3 mitigation"))

# Print proportions by broad group to see the shift
ssd_df %>%
  group_by(matrix, group) %>%
  summarise(group_pct = sum(proportion) * 100, .groups = "drop") %>%
  pivot_wider(names_from = matrix, values_from = group_pct) %>%
  print()
```

**plot Part 7a — SSD stage proportions: baseline vs full mitigation**
```{r plot-part7a-ssd, fig.width = 8, fig.height = 5}
p_ssd_compare <- ggplot(ssd_df, aes(x = stage, y = proportion * 100,
                                     fill = matrix)) +
  geom_col(position = "dodge", width = 0.7) +
  scale_fill_manual(values = c("No mitigation (baseline)" = "#d73027",
                                "Full 3/3 mitigation"      = "#253494"),
                    name = NULL) +
  scale_y_continuous(labels = function(x) paste0(round(x, 1), "%")) +
  labs(x = "Stage", y = "% of total population (SSD)",
       title = "Stable stage distribution: baseline vs full 3/3 mitigation",
       subtitle = "Shift in SSD determines direction and magnitude of transient dynamics") +
  themeo +
  theme(legend.position = "bottom",
        axis.text.x = element_text(angle = 30, hjust = 1))
p_ssd_compare
```

**plot Part 7b — realized lambda over time by scenario**
```{r plot-part7b-realized-lambda, fig.width = 11, fig.height = 5}
# Realized lambda = N(t+1)/N(t) each year — departs from asymptotic lambda during
# the transient period as population structure shifts toward the new SSD
realized_lambda_df <- all_scenarios %>%
  group_by(scenario) %>%
  arrange(year) %>%
  mutate(realized_lambda = N / lag(N)) %>%
  filter(!is.na(realized_lambda))

p_realized_lambda <- ggplot(realized_lambda_df,
                             aes(x = year, y = realized_lambda,
                                 color = scenario, linetype = scenario)) +
  geom_line(linewidth = 0.7) +
  geom_hline(yintercept = lambda_base,  color = "#d73027", linetype = "dashed",
             linewidth = 0.5, alpha = 0.6) +
  geom_hline(yintercept = lambda_3of3, color = "#253494", linetype = "dashed",
             linewidth = 0.5, alpha = 0.6) +
  annotate("text", x = years * 0.98, y = lambda_base  + 0.0005,
           label = paste0("\u03BB baseline = ", round(lambda_base, 4)),
           color = "#d73027", size = 2.8, hjust = 1) +
  annotate("text", x = years * 0.98, y = lambda_3of3 + 0.0005,
           label = paste0("\u03BB 3/3 = ", round(lambda_3of3, 4)),
           color = "#253494", size = 2.8, hjust = 1) +
  scale_color_manual(values = c(
    "No mitigation (baseline)"        = "#d73027",
    "Immediate 3/3 (100% coverage)"   = "#253494",
    "5-year phase-in"                 = "#1a9641",
    "10-year phase-in"                = "#78c679",
    "20-year phase-in"                = "#addd8e",
    "5-yr delay, then immediate 3/3"  = "#2166ac",
    "10-yr delay, then immediate 3/3" = "#74add1",
    "20-yr delay, then immediate 3/3" = "#abd9e9",
    "5-yr delay, then 5-yr phase-in"  = "#762a83",
    "10-yr delay, then 5-yr phase-in" = "#af8dc3",
    "20-yr delay, then 5-yr phase-in" = "#e7d4e8")) +
  scale_linetype_manual(values = c(
    "No mitigation (baseline)"        = "dashed",
    "Immediate 3/3 (100% coverage)"   = "solid",
    "5-year phase-in"                 = "solid",
    "10-year phase-in"                = "solid",
    "20-year phase-in"                = "solid",
    "5-yr delay, then immediate 3/3"  = "solid",
    "10-yr delay, then immediate 3/3" = "solid",
    "20-yr delay, then immediate 3/3" = "solid",
    "5-yr delay, then 5-yr phase-in"  = "solid",
    "10-yr delay, then 5-yr phase-in" = "solid",
    "20-yr delay, then 5-yr phase-in" = "solid")) +
  scale_x_continuous(breaks = seq(0, years, 10)) +
  scale_y_continuous(breaks = scales::pretty_breaks(n = 6)) +
  labs(x = "Years since 1990 baseline",
       y = "Realized \u03BB  (N(t+1) / N(t))",
       color = NULL, linetype = NULL,
       title = paste0("Realized growth rate over time — transient dynamics — ",
                      target_period, " baseline"),
       subtitle = paste0("Dashed horizontal lines = asymptotic \u03BB; ",
                         "deviations reflect transient dynamics as population structure shifts toward new SSD")) +
  themeo +
  theme(legend.position = "right",
        legend.text = element_text(size = 7))
p_realized_lambda
```

**plot Part 7c — all scenarios extended to 200 years**
```{r plot-part7c-all-200yr, fig.width = 11, fig.height = 6}
years_200 <- 200

# Re-run all scenarios for 200-year horizon using same matrix-selection logic
run_proj_200 <- function(phase_fn, scenario_label) {
  mat <- matrix(0, nrow = years_200 + 1, ncol = 11)
  colnames(mat) <- stage_names
  mat[1, ] <- n0_vec
  for (k in 1:years_200) {
    mat[k + 1, ] <- phase_fn(k) %*% mat[k, ]
  }
  as.data.frame(mat) %>%
    mutate(year = 0:years_200,
           N    = rowSums(across(all_of(stage_names))),
           scenario = scenario_label) %>%
    select(year, N, scenario)
}

all_scenarios_200 <- bind_rows(
  run_proj_200(function(k) m_base,  "No mitigation (baseline)"),
  run_proj_200(function(k) m_3of3,  "Immediate 3/3 (100% coverage)"),
  run_proj_200(function(k) if (k <= 5)  phase_list[[k]] else m_3of3,
               "5-year phase-in"),
  run_proj_200(function(k) if (k <= 10) phase_list_10y[[k]] else m_3of3,
               "10-year phase-in"),
  run_proj_200(function(k) if (k <= 20) phase_list_20y[[k]] else m_3of3,
               "20-year phase-in"),
  run_proj_200(function(k) if (k <= 5)  m_base else m_3of3,
               "5-yr delay, then immediate 3/3"),
  run_proj_200(function(k) if (k <= 10) m_base else m_3of3,
               "10-yr delay, then immediate 3/3"),
  run_proj_200(function(k) if (k <= 20) m_base else m_3of3,
               "20-yr delay, then immediate 3/3"),
  run_proj_200(function(k) {
    if      (k <= 5)  m_base
    else if (k <= 10) phase_list[[k - 5]]
    else              m_3of3 },
    "5-yr delay, then 5-yr phase-in"),
  run_proj_200(function(k) {
    if      (k <= 10) m_base
    else if (k <= 15) phase_list[[k - 10]]
    else              m_3of3 },
    "10-yr delay, then 5-yr phase-in"),
  run_proj_200(function(k) {
    if      (k <= 20) m_base
    else if (k <= 25) phase_list[[k - 20]]
    else              m_3of3 },
    "20-yr delay, then 5-yr phase-in")
) %>%
  mutate(scenario = factor(scenario, levels = levels(all_scenarios$scenario)))

p_all_200yr <- ggplot(all_scenarios_200,
                       aes(x = year, y = N, color = scenario, linetype = scenario)) +
  geom_line(linewidth = 0.7) +
  geom_hline(yintercept = N0, color = "grey50", linetype = "dashed", linewidth = 0.4) +
  geom_point(data = bird_obs,
             aes(x = year, y = N_scaled),
             inherit.aes = FALSE, color = "#e6550d", size = 3, shape = 16) +
  geom_line(data = bird_obs,
            aes(x = year, y = N_scaled),
            inherit.aes = FALSE, color = "#e6550d", linewidth = 0.5, linetype = "dotted") +
  geom_text(data = bird_obs,
            aes(x = year, y = N_scaled, label = season),
            inherit.aes = FALSE, color = "#e6550d", size = 2.5, vjust = -0.8) +
  scale_color_manual(values = c(
    "No mitigation (baseline)"        = "#d73027",
    "Immediate 3/3 (100% coverage)"   = "#253494",
    "5-year phase-in"                 = "#1a9641",
    "10-year phase-in"                = "#78c679",
    "20-year phase-in"                = "#addd8e",
    "5-yr delay, then immediate 3/3"  = "#2166ac",
    "10-yr delay, then immediate 3/3" = "#74add1",
    "20-yr delay, then immediate 3/3" = "#abd9e9",
    "5-yr delay, then 5-yr phase-in"  = "#762a83",
    "10-yr delay, then 5-yr phase-in" = "#af8dc3",
    "20-yr delay, then 5-yr phase-in" = "#e7d4e8")) +
  scale_linetype_manual(values = c(
    "No mitigation (baseline)"        = "dashed",
    "Immediate 3/3 (100% coverage)"   = "solid",
    "5-year phase-in"                 = "solid",
    "10-year phase-in"                = "solid",
    "20-year phase-in"                = "solid",
    "5-yr delay, then immediate 3/3"  = "solid",
    "10-yr delay, then immediate 3/3" = "solid",
    "20-yr delay, then immediate 3/3" = "solid",
    "5-yr delay, then 5-yr phase-in"  = "solid",
    "10-yr delay, then 5-yr phase-in" = "solid",
    "20-yr delay, then 5-yr phase-in" = "solid")) +
  scale_x_continuous(breaks = seq(-10, years_200, 20), limits = c(-10, years_200)) +
  scale_y_continuous(labels = scales::comma,
                     breaks = scales::pretty_breaks(n = 6),
                     sec.axis = sec_axis(~ . / pairs_scale,
                                         name = "Breeding pairs (Bird Island)",
                                         labels = scales::comma)) +
  labs(x = "Years since 1990 baseline (200-year horizon)",
       y = "Population (N individuals)",
       color = NULL, linetype = NULL,
       title = paste0("All scenarios — 200-year horizon — ", target_period, " baseline"),
       subtitle = "Orange dots = observed breeding pairs; secondary axis = breeding pairs equivalent") +
  themeo +
  theme(legend.position = "right",
        legend.text = element_text(size = 7))
p_all_200yr
```

---

*PART 8 — HISTORICAL REGULATORY SCENARIO*

Phased mitigation matching the approximate real regulatory timeline:
1995–2009 = 1/3 measures (tori lines only); 2010–2019 = 1/3 assumed continuing;
2020–2025 = 2/3 measures (tori lines + night setting); 2026+ = 3/3 full mitigation.
Year 0 = 1990; k=1–4 = m_base; k=5–19 = m_1of3; k=20–29 = m_1of3 (gap assumption);
k=30–35 = m_2of3; k=36+ = m_3of3.

**Part 8a — survival rates for 1/3 and 2/3 mitigation scenarios**
```{r part8a-historical-survival-rates}
# 1/3 mitigation: tori lines only
eff_1of3 <- eff_tori_lines  # 0.75

# 2/3 mitigation: tori lines + night setting (multiplicative independence assumed)
eff_2of3 <- 1 - (1 - eff_tori_lines) * (1 - eff_night_set)  # 0.90

# --- 1/3 mitigation survival rates ---
ns_Juv_1of3 <- 1 - ((dem_fm_juv + pel_fm_juv) * (1 - eff_1of3) + nm_juv)
ns_Imm_1of3 <- 1 - ((dem_fm_imm + pel_fm_imm) * (1 - eff_1of3) + nm_imm)
ns_EF_1of3  <- 1 - ((dem_fm_fb  + pel_fm_fb)  * (1 - eff_1of3) + nm_efb)
ns_ES_1of3  <- 1 - ((dem_fm_sb  + pel_fm_sb)  * (1 - eff_1of3) + nm_esb)
ns_ENB_1of3 <- 1 - ((dem_fm_nb  + pel_fm_nb)  * (1 - eff_1of3) + nm_enb)
ns_IF_1of3  <- 1 - ((dem_fm_fb  + pel_fm_fb)  * (1 - eff_1of3) + nm_ifb)
ns_IS_1of3  <- 1 - ((dem_fm_sb  + pel_fm_sb)  * (1 - eff_1of3) + nm_isb)
ns_PF_1of3  <- 1 - ((dem_fm_nb  + pel_fm_nb)  * (1 - eff_1of3) + nm_pf)

# --- 2/3 mitigation survival rates ---
ns_Juv_2of3 <- 1 - ((dem_fm_juv + pel_fm_juv) * (1 - eff_2of3) + nm_juv)
ns_Imm_2of3 <- 1 - ((dem_fm_imm + pel_fm_imm) * (1 - eff_2of3) + nm_imm)
ns_EF_2of3  <- 1 - ((dem_fm_fb  + pel_fm_fb)  * (1 - eff_2of3) + nm_efb)
ns_ES_2of3  <- 1 - ((dem_fm_sb  + pel_fm_sb)  * (1 - eff_2of3) + nm_esb)
ns_ENB_2of3 <- 1 - ((dem_fm_nb  + pel_fm_nb)  * (1 - eff_2of3) + nm_enb)
ns_IF_2of3  <- 1 - ((dem_fm_fb  + pel_fm_fb)  * (1 - eff_2of3) + nm_ifb)
ns_IS_2of3  <- 1 - ((dem_fm_sb  + pel_fm_sb)  * (1 - eff_2of3) + nm_isb)
ns_PF_2of3  <- 1 - ((dem_fm_nb  + pel_fm_nb)  * (1 - eff_2of3) + nm_pf)
```

**Part 8b — m_1of3 matrix (tori lines only, 1/3 efficacy)**
```{r part8b-m1of3}
m_1of3 <- matrix(data = c(
  0,           0,            0,            0,            0,                                                    0.5*ns_Juv_1of3,               0,                             0.5*ns_Juv_1of3,               0,                             0,                               0,
  ns_Juv_1of3, 0,            0,            0,            0,                                                    0,                             0,                             0,                             0,                             0,                               0,
  0,           ns_Juv_1of3,  0,            0,            0,                                                    0,                             0,                             0,                             0,                             0,                               0,
  0,           0,            ns_Imm_1of3,  0,            0,                                                    0,                             0,                             0,                             0,                             0,                               0,
  0,           0,            0,            ns_Imm_1of3,  ns_Imm_1of3*r_Imm*(1-b_Imm)+ns_Imm_1of3*(1-r_Imm),   0,                             0,                             0,                             0,                             0,                               0,
  0,           0,            0,            0,            ns_Imm_1of3*r_Imm*b_Imm*k_Imm,                        0,                             0,                             0,                             0,                             0,                               0,
  0,           0,            0,            0,            ns_Imm_1of3*r_Imm*b_Imm*(1-k_Imm),                    0,                             0,                             0,                             0,                             0,                               0,
  0,           0,            0,            0,            0,                                                    ns_IS_1of3*r_IS*b_IS*k_IS,     ns_IF_1of3*r_IF*b_IF*k_IF,     ns_ES_1of3*r_ES*b_ES*k_ES,     ns_EF_1of3*r_EF*b_EF*k_EF,     ns_ENB_1of3*r_ENB*b_ENB*k_ENB,   ns_PF_1of3*r_PF*b_PF*k_PF,
  0,           0,            0,            0,            0,                                                    ns_IS_1of3*r_IS*b_IS*(1-k_IS), ns_IF_1of3*r_IF*b_IF*(1-k_IF), ns_ES_1of3*r_ES*b_ES*(1-k_ES), ns_EF_1of3*r_EF*b_EF*(1-k_EF), ns_ENB_1of3*r_ENB*b_ENB*(1-k_ENB),ns_PF_1of3*r_PF*b_PF*(1-k_PF),
  0,           0,            0,            0,            0,                                                    ns_IS_1of3*r_IS*(1-b_IS),      ns_IF_1of3*r_IF*(1-b_IF),      ns_ES_1of3*r_ES*(1-b_ES),      ns_EF_1of3*r_EF*(1-b_EF),      ns_ENB_1of3*r_ENB*(1-b_ENB),     ns_PF_1of3*r_PF*(1-b_PF),
  0,           0,            0,            0,            0,                                                    ns_IS_1of3*(1-r_IS),           ns_IF_1of3*(1-r_IF),           ns_ES_1of3*(1-r_ES),           ns_EF_1of3*(1-r_EF),           ns_ENB_1of3*(1-r_ENB),           ns_PF_1of3*(1-r_PF)
), nrow = 11, byrow = TRUE)

lambda_1of3 <- as.numeric(eigen(m_1of3)$values[1])
```

**Part 8c — m_2of3 matrix (tori lines + night setting, 2/3 efficacy)**
```{r part8c-m2of3}
m_2of3 <- matrix(data = c(
  0,           0,            0,            0,            0,                                                    0.5*ns_Juv_2of3,               0,                             0.5*ns_Juv_2of3,               0,                             0,                               0,
  ns_Juv_2of3, 0,            0,            0,            0,                                                    0,                             0,                             0,                             0,                             0,                               0,
  0,           ns_Juv_2of3,  0,            0,            0,                                                    0,                             0,                             0,                             0,                             0,                               0,
  0,           0,            ns_Imm_2of3,  0,            0,                                                    0,                             0,                             0,                             0,                             0,                               0,
  0,           0,            0,            ns_Imm_2of3,  ns_Imm_2of3*r_Imm*(1-b_Imm)+ns_Imm_2of3*(1-r_Imm),   0,                             0,                             0,                             0,                             0,                               0,
  0,           0,            0,            0,            ns_Imm_2of3*r_Imm*b_Imm*k_Imm,                        0,                             0,                             0,                             0,                             0,                               0,
  0,           0,            0,            0,            ns_Imm_2of3*r_Imm*b_Imm*(1-k_Imm),                    0,                             0,                             0,                             0,                             0,                               0,
  0,           0,            0,            0,            0,                                                    ns_IS_2of3*r_IS*b_IS*k_IS,     ns_IF_2of3*r_IF*b_IF*k_IF,     ns_ES_2of3*r_ES*b_ES*k_ES,     ns_EF_2of3*r_EF*b_EF*k_EF,     ns_ENB_2of3*r_ENB*b_ENB*k_ENB,   ns_PF_2of3*r_PF*b_PF*k_PF,
  0,           0,            0,            0,            0,                                                    ns_IS_2of3*r_IS*b_IS*(1-k_IS), ns_IF_2of3*r_IF*b_IF*(1-k_IF), ns_ES_2of3*r_ES*b_ES*(1-k_ES), ns_EF_2of3*r_EF*b_EF*(1-k_EF), ns_ENB_2of3*r_ENB*b_ENB*(1-k_ENB),ns_PF_2of3*r_PF*b_PF*(1-k_PF),
  0,           0,            0,            0,            0,                                                    ns_IS_2of3*r_IS*(1-b_IS),      ns_IF_2of3*r_IF*(1-b_IF),      ns_ES_2of3*r_ES*(1-b_ES),      ns_EF_2of3*r_EF*(1-b_EF),      ns_ENB_2of3*r_ENB*(1-b_ENB),     ns_PF_2of3*r_PF*(1-b_PF),
  0,           0,            0,            0,            0,                                                    ns_IS_2of3*(1-r_IS),           ns_IF_2of3*(1-r_IF),           ns_ES_2of3*(1-r_ES),           ns_EF_2of3*(1-r_EF),           ns_ENB_2of3*(1-r_ENB),           ns_PF_2of3*(1-r_PF)
), nrow = 11, byrow = TRUE)

lambda_2of3 <- as.numeric(eigen(m_2of3)$values[1])

# Print lambdas for all four regulatory stages
tibble(
  scenario  = c("No mitigation (baseline)", "1/3 — tori lines only", "2/3 — tori + night setting", "3/3 — full mitigation"),
  lambda    = c(lambda_base, lambda_1of3, lambda_2of3, lambda_3of3),
  eff       = c(0, eff_1of3, eff_2of3, eff_3of3)
)
```

**Part 8d — historical timeline projection**
```{r part8d-historical-projection}
# Timeline (year 0 = 1990):
#   k =  1– 4 : m_base   (1990–1994, pre-regulation baseline)
#   k =  5–19 : m_1of3   (1995–2009, tori lines introduced)
#   k = 20–29 : m_1of3   (2010–2019, assumed 1/3 continues; gap in data)
#   k = 30+   : m_2of3   (2020+, current regs — tori lines + night setting — assumed to continue)

hist_fn <- function(k) {
  if (k <= 4)        m_base
  else if (k <= 29)  m_1of3
  else               m_2of3
}

# Run 60-year projection (1990–2050) for the historical scenario
years_hist <- 60

mat_hist <- matrix(0, nrow = years_hist + 1, ncol = 11)
colnames(mat_hist) <- stage_names
mat_hist[1, ] <- n0_vec

for (k in 1:years_hist) {
  mat_hist[k + 1, ] <- hist_fn(k) %*% mat_hist[k, ]
}

proj_hist <- as.data.frame(mat_hist) %>%
  mutate(year      = 0:years_hist,
         N         = rowSums(across(all_of(stage_names))),
         scenario  = "Historical regulatory timeline")

# Baselines for comparison: no mitigation and immediate 2/3 from year 0
mat_hist_base  <- matrix(0, nrow = years_hist + 1, ncol = 11)
mat_hist_23    <- matrix(0, nrow = years_hist + 1, ncol = 11)
colnames(mat_hist_base) <- colnames(mat_hist_23) <- stage_names
mat_hist_base[1, ] <- mat_hist_23[1, ] <- n0_vec

for (k in 1:years_hist) {
  mat_hist_base[k + 1, ] <- m_base  %*% mat_hist_base[k, ]
  mat_hist_23[k + 1, ]   <- m_2of3  %*% mat_hist_23[k, ]
}

proj_hist_base <- as.data.frame(mat_hist_base) %>%
  mutate(year = 0:years_hist, N = rowSums(across(all_of(stage_names))),
         scenario = "No mitigation (baseline)")

proj_hist_23 <- as.data.frame(mat_hist_23) %>%
  mutate(year = 0:years_hist, N = rowSums(across(all_of(stage_names))),
         scenario = "Immediate 2/3 from 1990 (upper bound)")

hist_combined <- bind_rows(proj_hist_base, proj_hist, proj_hist_23) %>%
  mutate(scenario = factor(scenario, levels = c(
    "No mitigation (baseline)",
    "Historical regulatory timeline",
    "Immediate 2/3 from 1990 (upper bound)"
  )))
```

**Part 8e — plot historical scenario**
```{r part8e-plot-historical, fig.width = 10, fig.height = 6}
# Shading rectangles mark each regulatory phase
phase_rects <- tibble(
  xmin  = c(0,  5, 20, 30),
  xmax  = c(5, 20, 30, years_hist),
  label = c("Baseline\n(pre-reg)", "1/3 mitigation\n(tori lines)",
            "1/3 continued\n(gap assumption)", "2/3 mitigation\n(tori+night,\ncurrent regs)"),
  fill  = c("#f7f7f7", "#fee8c8", "#fdd49e", "#fdbb84")
)

p_historical <- ggplot() +
  geom_rect(data = phase_rects,
            aes(xmin = xmin, xmax = xmax, ymin = -Inf, ymax = Inf, fill = label),
            alpha = 0.15, inherit.aes = FALSE) +
  scale_fill_manual(name = "Regulatory phase",
                    values = setNames(phase_rects$fill, phase_rects$label)) +
  geom_point(data = bird_obs, aes(x = year, y = N_scaled),
             color = "darkorange", size = 2.5, inherit.aes = FALSE) +
  geom_line(data = hist_combined,
            aes(x = year, y = N, color = scenario, linetype = scenario),
            linewidth = 0.9) +
  scale_color_manual(values = c(
    "No mitigation (baseline)"              = "black",
    "Historical regulatory timeline"        = "#e08214",
    "Immediate 2/3 from 1990 (upper bound)" = "#1b7837")) +
  scale_linetype_manual(values = c(
    "No mitigation (baseline)"              = "dashed",
    "Historical regulatory timeline"        = "solid",
    "Immediate 2/3 from 1990 (upper bound)" = "solid")) +
  scale_x_continuous(
    breaks = c(0, 5, 20, 30, 60),
    labels = c("1990", "1995", "2010", "2020", "2050"),
    limits = c(0, years_hist)) +
  scale_y_continuous(labels = scales::comma,
                     sec.axis = sec_axis(~ . / pairs_scale,
                                         name = "Breeding pairs (Bird Island)",
                                         labels = scales::comma)) +
  labs(x = "Year",
       y = "Population (N individuals)",
       color = NULL, linetype = NULL,
       title = paste0("Historical regulatory scenario — ", target_period, " baseline"),
       subtitle = "Shading = regulatory phase; orange dots = observed breeding pairs; green = immediate 2/3 from 1990 (upper bound)") +
  themeo +
  theme(legend.position = "right",
        legend.text = element_text(size = 8))
p_historical
```

---

**age class trends from dem_props (Clay 2019)**
```{r plot-age-trends, fig.width = 8, fig.height = 5}
# Use the 4 survey periods only (exclude the 1990-2009 aggregate and J1)
# Mid-year of each 5-year period used for x-axis placement
age_trends <- dem_props %>%
  filter(Dem_class != "J1",
         Period %in% c("1990-1994", "1995-1999", "2000-2004", "2005-2009")) %>%
  mutate(
    year_mid = case_when(
      Period == "1990-1994" ~ 1992,
      Period == "1995-1999" ~ 1997,
      Period == "2000-2004" ~ 2002,
      Period == "2005-2009" ~ 2007
    ),
    Dem_class = factor(Dem_class,
                       levels = c("SB", "FB", "NB", "IMM", "J2+3"),
                       labels = c("Successful breeders", "Failed breeders",
                                  "Non-breeders", "Immatures", "Juveniles (J2+3)"))
  )

p_age_trends <- ggplot(age_trends, aes(x = year_mid, y = Pop,
                                        color = Dem_class, group = Dem_class)) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 2.5) +
  scale_color_brewer(palette = "Dark2") +
  scale_x_continuous(breaks = c(1992, 1997, 2002, 2007),
                     labels = c("1990-94", "1995-99", "2000-04", "2005-09")) +
  scale_y_continuous(labels = scales::comma) +
  labs(x = NULL,
       y = "Population (N individuals)",
       color = "Age class",
       title = "Bird Island WAAL population by age class — Clay et al. 2019",
       subtitle = "5-year period averages; excludes J1") +
  themeo +
  theme(legend.position = "right")
p_age_trends
```

## Save figures

```{r save-figures}
dir.create(here("output/figures"), showWarnings = FALSE, recursive = TRUE)

fig_path <- function(name) here("output/figures",
                                 paste0(name, "_", target_period, "_", date_sfx, ".png"))

ggsave(fig_path("08_proj_part1_breeding_pairs"),       p_part1,              width = 9,  height = 5, dpi = 200)
ggsave(fig_path("08_proj_part1_age_class_obs"),       p_part1_obs,          width = 9,  height = 5, dpi = 200)
ggsave(fig_path("08_age_class_trends"),               p_age_trends,         width = 8,  height = 5, dpi = 200)
ggsave(fig_path("08_proj_part1_long_breeding_pairs"), p_part1_long,         width = 9,  height = 5, dpi = 200)
ggsave(fig_path("08_proj_part1_long_age_class_obs"),  p_part1_obs_long,     width = 9,  height = 5, dpi = 200)
ggsave(fig_path("08_proj_part2_total_validation"),    p_part2_total,        width = 9,  height = 5, dpi = 200)
ggsave(fig_path("08_proj_part2_stage_groups"),        p_part2_stages,       width = 10, height = 6, dpi = 200)
ggsave(fig_path("08_proj_part2_composition"),         p_part2_composition,  width = 10, height = 5, dpi = 200)
ggsave(fig_path("08_proj_part3_phasein"),             p_part3,              width = 9,  height = 5, dpi = 200)
ggsave(fig_path("08_proj_part6a_phasein_compare"),   p_phasein_compare,    width = 9,  height = 5, dpi = 200)
ggsave(fig_path("08_proj_part6b_delay_immediate"),   p_delay_imm,          width = 9,  height = 5, dpi = 200)
ggsave(fig_path("08_proj_part6c_delay_phasein"),     p_delay_phase,        width = 9,  height = 5, dpi = 200)
ggsave(fig_path("08_proj_part6d_all_scenarios"),     p_all_scenarios,      width = 11, height = 6, dpi = 200)
ggsave(fig_path("08_proj_part6e_all_scenarios_obs"), p_all_scenarios_obs,  width = 11, height = 6, dpi = 200)
ggsave(fig_path("08_proj_part7a_ssd_compare"),       p_ssd_compare,        width = 8,  height = 5, dpi = 200)
ggsave(fig_path("08_proj_part7b_realized_lambda"),   p_realized_lambda,    width = 11, height = 5, dpi = 200)
ggsave(fig_path("08_proj_part7c_all_200yr"),         p_all_200yr,          width = 11, height = 6, dpi = 200)
ggsave(fig_path("08_proj_part8_historical"),         p_historical,         width = 10, height = 6, dpi = 200)
message("18 figures saved to output/figures/ with suffix ", date_sfx)
```
