---
title: "Fishing Effort and Albatross Distribution Overlap"
author: "WAAL Bycatch Project"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.width = 7,
  fig.height = 10
)
```

# Overview

This script calculates spatial overlap between wandering albatross (WAAL) distributions and longline fishing effort. The analysis:

1. Loads WAAL distribution rasters by age class from Clay et al. 2019
2. Processes pelagic and demersal longline fishing effort data
3. Calculates overlap indices (bird density × fishing effort) for each age class
4. Generates publication-quality maps of distributions and overlap

# Setup

```{r libraries}
library(tidyverse)
library(terra)
library(sf)
library(here)
library(viridis)
library(RColorBrewer)
library(scales)
library(patchwork)

# Date suffix for versioned output files — prevents overwriting previous runs
date_sfx <- format(Sys.Date(), "%Y-%m-%d")
```

# Load WAAL Distributions

WAAL distributions from Clay et al. 2019, representing mean annual distribution (1990-2009) for different age classes.

```{r load-waal-distributions}
# Age class abbreviations:
# fb = failed breeders
# sb = successful breeders
# nb = non-breeders
# imm = immatures
# j1 = juveniles age 1
# j2j3 = juveniles age 2-3

# Load all age class distributions
waal_fb <- rast(here("data/WAAL_dist_Clay2019/WA_AllMonths_Both_FB_1990-2009.tif"))
waal_sb <- rast(here("data/WAAL_dist_Clay2019/WA_AllMonths_Both_SB_1990-2009.tif"))
waal_nb <- rast(here("data/WAAL_dist_Clay2019/WA_AllMonths_Both_NB_1990-2009.tif"))
waal_imm <- rast(here("data/WAAL_dist_Clay2019/WA_AllMonths_Both_IMM_1990-2009.tif"))
waal_j1 <- rast(here("data/WAAL_dist_Clay2019/WA_AllMonths_Both_J1_1990-2009.tif"))
waal_j2j3 <- rast(here("data/WAAL_dist_Clay2019/WA_AllMonths_Both_J2+3_1990-2009.tif"))
waal_all <- rast(here("data/WAAL_dist_Clay2019/WA_AllMonths_Both_1990-2009.tif"))

# Store in named list for easier processing
bird_layers <- list(
  all = waal_all,
  fb = waal_fb,
  sb = waal_sb,
  nb = waal_nb,
  imm = waal_imm,
  j1 = waal_j1,
  j2j3 = waal_j2j3
)
```

# Process Fishing Effort Data

## Pelagic Longline Effort

```{r process-pelagic-effort}
# Load pelagic longline data
pll_data <- read_csv(here("data/Pel_LL_effort.csv"))

# Remove existing WCPFC rows — replaced with updated WCPFC dataset below
pll_data <- pll_data %>%
  filter(RFMO != "WCPFC")

# Load new WCPFC pelagic longline data
wcpfc_raw <- read_csv(
  here("data/WCPFC_L_PUBLIC_BY_YY_MM_FLAG/WCPFC_L_PUBLIC_BY_YY_MM_FLAG.csv")
)

# Convert CWP 5-degree cell codes to decimal degree centroids
# CWP codes give the boundary furthest from the origin:
#   N lat: equatorward boundary → centroid = deg + 2.5  (e.g. "00N" → 2.5°N)
#   S lat: poleward boundary    → centroid = -(deg - 2.5) (e.g. "05S" → -2.5°S)
#   E lon: western boundary     → centroid = deg + 2.5  (e.g. "120E" → 122.5°E)
#   W lon: western boundary     → centroid = -(deg - 2.5) (e.g. "155W" → -152.5°W)
# Note: there is no "00S" — "05S" is the equatorward-most S cell (0–5°S)
parse_lat5 <- function(x) {
  deg <- as.numeric(stringr::str_extract(x, "\\d+"))
  hem <- stringr::str_sub(x, -1, -1)
  ifelse(hem == "S", -(deg - 2.5), deg + 2.5)
}

parse_lon5 <- function(x) {
  deg <- as.numeric(stringr::str_extract(x, "\\d+"))
  hem <- stringr::str_sub(x, -1, -1)
  ifelse(hem == "W", -(deg - 2.5), deg + 2.5)
}

# Retain relevant columns and rename to match pll_data structure
# HHOOKS is in units of hundreds of hooks — multiply by 100 to get individual hooks
wcpfc_std <- wcpfc_raw %>%
  select(YY, MM, flag_code, LAT5, LON5, HHOOKS) %>%
  mutate(
    Lat   = parse_lat5(LAT5),
    Lon   = parse_lon5(LON5),
    RFMO  = "WCPFC",
    Hooks = HHOOKS * 100
  ) %>%
  rename(
    Year  = YY,
    Month = MM,
    Flag  = flag_code
  ) %>%
  select(Year, Month, RFMO, Flag, Lon, Lat, Hooks)

# Combine with remaining pelagic data
pll_data <- bind_rows(pll_data, wcpfc_std)

# Filter to 1990-2009 to match WAAL distributions
pll_data <- pll_data %>%
  filter(Year >= 1990, Year <= 2009)

# Harmonize fleet names: standardize WCPFC ISO codes to full country names
pll_data <- pll_data %>%
  mutate(Flag = case_when(
    Flag == "JP"  ~ "Japan",
    Flag == "KR"  ~ "South Korea",
    Flag == "TW"  ~ "Taiwan",
    Flag == "CN"  ~ "China",
    Flag == "US"  ~ "USA",
    Flag == "ES"  ~ "Spain",
    Flag == "AU"  ~ "Australia",
    Flag == "NZ"  ~ "New Zealand",
    Flag == "PF"  ~ "French Polynesia",
    Flag == "VU"  ~ "Vanuatu",
    Flag == "BZ"  ~ "Belize",
    Flag == "NC"  ~ "New Caledonia",
    Flag == "FJ"  ~ "Fiji",
    Flag == "TO"  ~ "Tonga",
    Flag == "WS"  ~ "Samoa",
    Flag == "FM"  ~ "Micronesia",
    Flag == "PG"  ~ "Papua New Guinea",
    Flag == "PW"  ~ "Palau",
    Flag == "CK"  ~ "Cook Islands",
    Flag == "MH"  ~ "Marshall Islands",
    Flag == "NU"  ~ "Niue",
    TRUE          ~ Flag
  ))

# Save disaggregated PLL table for future fleet/RFMO partitioning
write_csv(pll_data, here("output/pll_effort_by_fleet.csv"))

# Calculate summary statistics
total_pel_hooks <- sum(pll_data$Hooks, na.rm = TRUE)
mean_annual_pel <- pll_data %>%
  group_by(Year) %>%
  summarize(total_hooks = sum(Hooks, na.rm = TRUE)) %>%
  pull(total_hooks) %>%
  mean()

tibble(metric = c("Total pelagic hooks", "Mean annual pelagic hooks"),
       value  = c(scales::comma(total_pel_hooks), scales::comma(mean_annual_pel)))

# Summary by RFMO
pll_summary_rfmo <- pll_data %>%
  group_by(RFMO) %>%
  summarise(total_hooks       = sum(Hooks, na.rm = TRUE),
            mean_annual_hooks = total_hooks / n_distinct(Year),
            n_years           = n_distinct(Year),
            .groups = "drop") %>%
  arrange(desc(total_hooks))

pll_summary_rfmo %>% mutate(across(c(total_hooks, mean_annual_hooks), scales::comma))

# Summary by Flag (fleet)
pll_summary_flag <- pll_data %>%
  group_by(Flag) %>%
  summarise(total_hooks       = sum(Hooks, na.rm = TRUE),
            mean_annual_hooks = total_hooks / n_distinct(Year),
            n_years           = n_distinct(Year),
            .groups = "drop") %>%
  arrange(desc(total_hooks))

pll_summary_flag %>% mutate(across(c(total_hooks, mean_annual_hooks), scales::comma))

# Southern Hemisphere summary (Lat < 0) — effort within WAAL range
pll_sh <- pll_data %>% filter(Lat < 0)

total_pel_hooks_sh <- sum(pll_sh$Hooks, na.rm = TRUE)
mean_annual_pel_sh <- pll_sh %>%
  group_by(Year) %>%
  summarize(total_hooks = sum(Hooks, na.rm = TRUE)) %>%
  pull(total_hooks) %>%
  mean()

tibble(metric = c("Total SH pelagic hooks", "Mean annual SH pelagic hooks"),
       value  = c(scales::comma(total_pel_hooks_sh), scales::comma(mean_annual_pel_sh)))

# SH summary by RFMO
pll_sh %>%
  group_by(RFMO) %>%
  summarise(total_hooks       = sum(Hooks, na.rm = TRUE),
            mean_annual_hooks = total_hooks / n_distinct(Year),
            n_years           = n_distinct(Year),
            .groups = "drop") %>%
  arrange(desc(total_hooks)) %>%
  mutate(across(c(total_hooks, mean_annual_hooks), scales::comma))

# SH summary by Flag (fleet)
pll_sh %>%
  group_by(Flag) %>%
  summarise(total_hooks       = sum(Hooks, na.rm = TRUE),
            mean_annual_hooks = total_hooks / n_distinct(Year),
            n_years           = n_distinct(Year),
            .groups = "drop") %>%
  arrange(desc(total_hooks)) %>%
  mutate(across(c(total_hooks, mean_annual_hooks), scales::comma))

# Calculate mean annual effort by location:
# Step 1 — sum hooks across months within each year to get total annual effort per cell
# Step 2 — average annual totals across years
# This correctly handles seasonal fleets (e.g. Falklands) that are only active
# certain months, which would otherwise be undercounted by a simple mean(Hooks)
pll_mean_annual <- pll_data %>%
  group_by(Lon, Lat, Year) %>%
  summarise(annual_hooks = sum(Hooks, na.rm = TRUE), .groups = 'drop') %>%
  group_by(Lon, Lat) %>%
  summarise(mean_effort = mean(annual_hooks, na.rm = TRUE), .groups = 'drop')

# Create raster directly from gridded data (data is already in 5x5 cells)
rast_pll <- rast(pll_mean_annual, type = "xyz", crs = "EPSG:4326")

# Create scaled mean annual effort (in units of 10^6 hooks)
rast_pll_mean_scaled <- rast_pll / 1e6

# Calculate cumulative effort (sum across all years) by location
pll_cumulative <- pll_data %>%
  group_by(Lon, Lat) %>%
  summarise(cumulative_effort = sum(Hooks, na.rm = TRUE), .groups = 'drop')

# Create raster directly from gridded data
rast_pll_cumulative <- rast(pll_cumulative, type = "xyz", crs = "EPSG:4326")

# Create scaled cumulative effort (in units of 10^6 hooks)
rast_pll_cumulative_scaled <- rast_pll_cumulative / 1e6
```

## Demersal Longline Effort

```{r process-demersal-effort}
# Load demersal longline data from multiple sources
ccamlr_data <- read_csv(here("data/DATA_PRODUCT_725/C2_725.csv"))
arg_data <- read_csv(here("data/DemData_Clay2019_Dryad/Arg_Dem_LL.csv"))
chile_data <- read_csv(here("data/DemData_Clay2019_Dryad/Chile_LL.csv")) %>%
  filter(FISHERY != "Pel LL SWO")  # Exclude pelagic swordfish rows (PLL, not DLL)
falk_data <- read_csv(here("data/DemData_Clay2019_Dryad/Falklands demersal LL.csv"))
namibia_data <- read_csv(here("data/DemData_Clay2019_Dryad/Namibia Dem Hks Est.csv"))
saf_data <- read_csv(here("data/DemData_Clay2019_Dryad/Saf_KKlip_Dem.csv"))
nz_data <- read_csv(here("data/Rep Log 17287 BLL effort truncated.csv"), show_col_types = FALSE)

# Standardize column names for each dataset
# fleet_col: optional column name to retain as Fleet (e.g. vessel nationality in CCAMLR)
# rfmo: RFMO affiliation string, NA for datasets not reported through an RFMO
standardize_dem_data <- function(data, dataset_name, lon_col = "Lon", lat_col = "Lat",
                                  hooks_col = "Hooks", year_col = "Year", month_col = "Month",
                                  fleet_col = NULL, rfmo = NA) {
  cols <- c(year_col, month_col, lat_col, lon_col, hooks_col)
  if (!is.null(fleet_col)) cols <- c(cols, fleet_col)

  out <- data %>%
    select(all_of(cols)) %>%
    rename(
      Year  = all_of(year_col),
      Month = all_of(month_col),
      Lat   = all_of(lat_col),
      Lon   = all_of(lon_col),
      Hooks = all_of(hooks_col)
    ) %>%
    mutate(dataset = dataset_name, RFMO = rfmo)

  if (!is.null(fleet_col)) {
    out <- out %>% rename(Fleet = all_of(fleet_col))
  } else {
    # For single-country datasets, use dataset name as fleet identifier
    out <- out %>% mutate(Fleet = dataset_name)
  }

  out
}

# Standardize each dataset
ccamlr_std <- standardize_dem_data(
  ccamlr_data, "CCAMLR",
  lon_col = "longitude_5deg", lat_col = "latitude_5deg",
  hooks_col = "hook_count", year_col = "year", month_col = "month",
  fleet_col = "vessel_nationality_code", rfmo = "CCAMLR"
)

arg_std   <- standardize_dem_data(arg_data,  "Argentina", lon_col = "Lng", hooks_col = "Effort")
chile_std <- standardize_dem_data(chile_data, "Chile",    lon_col = "Lng", hooks_col = "Effort")
falk_std  <- standardize_dem_data(falk_data,  "Falklands", lon_col = "Lng", hooks_col = "Effort")
namibia_std <- namibia_data %>% mutate(dataset = "Namibia",      RFMO = NA_character_, Fleet = "Namibia")
saf_std     <- saf_data %>% rename(Hooks = Effort) %>%
                             mutate(dataset = "South Africa", RFMO = NA_character_, Fleet = "South Africa")
nz_std <- nz_data %>%
  select(year, month_number, truncated_lat, truncated_long, total_hooks) %>%
  rename(Year = year, Month = month_number, Lat = truncated_lat,
         Lon = truncated_long, Hooks = total_hooks) %>%
  mutate(dataset = "New Zealand", RFMO = NA_character_, Fleet = "New Zealand")
# Note: NZ data uses x.0 cell centers (e.g. 170, 175); other DLL sources use x.5 (e.g. 172.5, 177.5)
# NZ is rasterized separately and resampled to the x.5 grid below to avoid checkerboard artifacts

# Diagnostic: compare coordinate grid offsets across datasets before combining
# Different 5-degree grid anchors will cause rasterization issues (checkerboard)
# Grid alignment check — first 5 unique Lon values per source (before wrapping)
tribble(
  ~source,        ~first_5_lon_values,
  "CCAMLR",       paste(head(sort(unique(ccamlr_std$Lon)), 5), collapse = ", "),
  "Argentina",    paste(head(sort(unique(arg_std$Lon)),    5), collapse = ", "),
  "Chile",        paste(head(sort(unique(chile_std$Lon)),  5), collapse = ", "),
  "Falklands",    paste(head(sort(unique(falk_std$Lon)),   5), collapse = ", "),
  "Namibia",      paste(head(sort(unique(namibia_std$Lon)),5), collapse = ", "),
  "S. Africa",    paste(head(sort(unique(saf_std$Lon)),    5), collapse = ", "),
  "New Zealand",  paste(head(sort(unique(nz_std$Lon)),     5), collapse = ", ")
)

# Combine all demersal data
dll_data <- bind_rows(
  ccamlr_std, arg_std, chile_std,
  falk_std, namibia_std, saf_std, nz_std
)

# Filter to 1990-2009 to match WAAL distributions; drop rows with missing coordinates
dll_data <- dll_data %>%
  filter(Year >= 1990, Year <= 2009) %>%
  filter(!is.na(Lon), !is.na(Lat)) %>%
  mutate(Lon = ifelse(Lon > 180, Lon - 360, Lon))

# Harmonize fleet names: standardize CCAMLR ISO codes to full country names
dll_data <- dll_data %>%
  mutate(Fleet = case_when(
    Fleet == "ARG" ~ "Argentina",
    Fleet == "AUS" ~ "Australia",
    Fleet == "BGR" ~ "Bulgaria",
    Fleet == "CHL" ~ "Chile",
    Fleet == "ESP" ~ "Spain",
    Fleet == "FRA" ~ "France",
    Fleet == "GBR" ~ "United Kingdom",
    Fleet == "JPN" ~ "Japan",
    Fleet == "KOR" ~ "South Korea",
    Fleet == "NAM" ~ "Namibia",
    Fleet == "NOR" ~ "Norway",
    Fleet == "NZL" ~ "New Zealand",
    Fleet == "RUS" ~ "Russia",
    Fleet == "SUN" ~ "USSR",
    Fleet == "UKR" ~ "Ukraine",
    Fleet == "URY" ~ "Uruguay",
    Fleet == "USA" ~ "USA",
    Fleet == "ZAF" ~ "South Africa",
    TRUE           ~ Fleet
  ))

# Save disaggregated DLL table for future fleet/RFMO partitioning
write_csv(dll_data, here("output/dll_effort_by_fleet.csv"))

# Calculate summary statistics
total_dem_hooks <- sum(dll_data$Hooks, na.rm = TRUE)
mean_annual_dem <- dll_data %>%
  group_by(Year) %>%
  summarize(total_hooks = sum(Hooks, na.rm = TRUE)) %>%
  pull(total_hooks) %>%
  mean()

tibble(metric = c("Total demersal hooks", "Mean annual demersal hooks"),
       value  = c(scales::comma(total_dem_hooks), scales::comma(mean_annual_dem)))

# Summary by dataset and RFMO
dll_summary_dataset <- dll_data %>%
  group_by(dataset, RFMO) %>%
  summarise(total_hooks       = sum(Hooks, na.rm = TRUE),
            mean_annual_hooks = total_hooks / n_distinct(Year),
            n_years           = n_distinct(Year),
            .groups = "drop") %>%
  arrange(desc(total_hooks))

dll_summary_dataset %>% mutate(across(c(total_hooks, mean_annual_hooks), scales::comma))

# Summary by Fleet (CCAMLR broken out by vessel nationality)
dll_summary_fleet <- dll_data %>%
  group_by(Fleet) %>%
  summarise(total_hooks       = sum(Hooks, na.rm = TRUE),
            mean_annual_hooks = total_hooks / n_distinct(Year),
            n_years           = n_distinct(Year),
            .groups = "drop") %>%
  arrange(desc(total_hooks))

dll_summary_fleet %>% mutate(across(c(total_hooks, mean_annual_hooks), scales::comma))

# Southern Hemisphere summary (Lat < 0) — effort within WAAL range
dll_sh <- dll_data %>% filter(Lat < 0)

total_dem_hooks_sh <- sum(dll_sh$Hooks, na.rm = TRUE)
mean_annual_dem_sh <- dll_sh %>%
  group_by(Year) %>%
  summarize(total_hooks = sum(Hooks, na.rm = TRUE)) %>%
  pull(total_hooks) %>%
  mean()

tibble(metric = c("Total SH demersal hooks", "Mean annual SH demersal hooks"),
       value  = c(scales::comma(total_dem_hooks_sh), scales::comma(mean_annual_dem_sh)))

# SH summary by dataset and RFMO
dll_sh %>%
  group_by(dataset, RFMO) %>%
  summarise(total_hooks       = sum(Hooks, na.rm = TRUE),
            mean_annual_hooks = total_hooks / n_distinct(Year),
            n_years           = n_distinct(Year),
            .groups = "drop") %>%
  arrange(desc(total_hooks)) %>%
  mutate(across(c(total_hooks, mean_annual_hooks), scales::comma))

# SH summary by Fleet
dll_sh %>%
  group_by(Fleet) %>%
  summarise(total_hooks       = sum(Hooks, na.rm = TRUE),
            mean_annual_hooks = total_hooks / n_distinct(Year),
            n_years           = n_distinct(Year),
            .groups = "drop") %>%
  arrange(desc(total_hooks)) %>%
  mutate(across(c(total_hooks, mean_annual_hooks), scales::comma))

# Process non-NZ and NZ sources separately — they use different 5° grid conventions
# Non-NZ: cell centers at x.5 (e.g. 172.5, 177.5); NZ: cell centers at x.0 (e.g. 170, 175)
# NZ is rasterized on its native x.0 grid, then bilinearly resampled to the x.5 reference grid

# Mean annual effort — non-NZ sources
# Step 1: sum hooks across months within each year; Step 2: average across years
dll_mean_no_nz <- dll_data %>%
  filter(dataset != "New Zealand") %>%
  group_by(Lon, Lat, Year) %>%
  summarise(annual_hooks = sum(Hooks, na.rm = TRUE), .groups = 'drop') %>%
  group_by(Lon, Lat) %>%
  summarise(mean_effort = mean(annual_hooks, na.rm = TRUE), .groups = 'drop')

# Mean annual effort — NZ only (same two-step approach)
dll_mean_nz <- dll_data %>%
  filter(dataset == "New Zealand") %>%
  group_by(Lon, Lat, Year) %>%
  summarise(annual_hooks = sum(Hooks, na.rm = TRUE), .groups = 'drop') %>%
  group_by(Lon, Lat) %>%
  summarise(mean_effort = mean(annual_hooks, na.rm = TRUE), .groups = 'drop')

rast_dll_no_nz      <- rast(dll_mean_no_nz, type = "xyz", crs = "EPSG:4326")
rast_dll_nz         <- rast(dll_mean_nz,    type = "xyz", crs = "EPSG:4326")
rast_dll_nz_aligned <- resample(rast_dll_nz, rast_dll_no_nz, method = "bilinear")

# NA-safe addition: treat NA as 0 when adding, then restore NA where both were NA
rast_dll <- subst(subst(rast_dll_no_nz, NA, 0) + subst(rast_dll_nz_aligned, NA, 0), 0, NA)

# Create scaled mean annual effort (in units of 10^6 hooks)
rast_dll_mean_scaled <- rast_dll / 1e6

# Cumulative effort (sum across all years) — non-NZ sources
dll_cumulative_no_nz <- dll_data %>%
  filter(dataset != "New Zealand") %>%
  group_by(Lon, Lat) %>%
  summarise(cumulative_effort = sum(Hooks, na.rm = TRUE), .groups = 'drop')

# Cumulative effort — NZ only
dll_cumulative_nz <- dll_data %>%
  filter(dataset == "New Zealand") %>%
  group_by(Lon, Lat) %>%
  summarise(cumulative_effort = sum(Hooks, na.rm = TRUE), .groups = 'drop')

rast_dll_cumulative_no_nz      <- rast(dll_cumulative_no_nz, type = "xyz", crs = "EPSG:4326")
rast_dll_cumulative_nz         <- rast(dll_cumulative_nz,    type = "xyz", crs = "EPSG:4326")
rast_dll_cumulative_nz_aligned <- resample(rast_dll_cumulative_nz, rast_dll_cumulative_no_nz, method = "bilinear")

rast_dll_cumulative <- subst(
  subst(rast_dll_cumulative_no_nz, NA, 0) + subst(rast_dll_cumulative_nz_aligned, NA, 0),
  0, NA
)

# Create scaled cumulative effort (in units of 10^6 hooks)
rast_dll_cumulative_scaled <- rast_dll_cumulative / 1e6
```

```{r}
library(dplyr)

dll_summary <- dll_data %>%
  group_by(dataset, RFMO, Fleet) %>%
  summarise(
    min_year = min(Year, na.rm = TRUE),
    max_year = max(Year, na.rm = TRUE),
    min_lat  = min(Lat,  na.rm = TRUE),
    max_lat  = max(Lat,  na.rm = TRUE),
    min_lon  = min(Lon,  na.rm = TRUE),
    max_lon  = max(Lon,  na.rm = TRUE),
    total_hooks = sum(Hooks, na.rm = TRUE),
    n_records   = n(),
    n_years     = n_distinct(Year),
    .groups = "drop"
  )

dll_summary

```


# Calculate Spatial Overlap

Overlap is calculated as the product of bird density and fishing effort within each grid cell. We calculate separate overlaps for pelagic and demersal fisheries due to different spatial extents.

```{r define-extents}
# Define spatial extents for each fishery type
# Extents chosen to match the range of each fishery and bird distributions
ext_dem <- c(-180, 180, -80, 0)   # Demersal extent
ext_pel <- c(-180, 180, -60, 0)   # Pelagic extent (updated to 60S)
ext_bird <- c(-180, 180, -90, 0)  # Full bird distribution extent
```

```{r calculate-overlap-function}
# Function to calculate overlap for all age classes with one fishery type
# Overlap = fishing effort × normalized bird density (proportion of population per cell)
# This ensures spatial patterns in the maps reflect both fishing intensity and bird distribution
calculate_overlap <- function(bird_layers, fishing_rast, extent) {
  # Crop fishing effort to study extent
  fishing_crop <- crop(fishing_rast, extent)

  # Calculate overlap for each age class
  overlaps <- map(bird_layers, function(bird_rast) {
    # Normalize against the full Clay et al. 2019 raster extent (-180 to 180,
    # -90 to 0) before cropping, so proportions reflect each cell's share of
    # the global tracked population — ensures values are comparable across age
    # classes (wide-ranging juveniles are not artificially inflated relative to
    # experienced breeders that rarely leave the study area)
    bird_norm <- bird_rast / global(bird_rast, fun = "sum", na.rm = TRUE)[1,1]
    # Crop normalized distribution to fishing raster extent
    bird_crop <- crop(bird_norm, fishing_crop)
    # Overlap index: hooks weighted by proportion of global population present
    overlap <- fishing_crop * bird_crop
    return(overlap)
  })

  return(overlaps)
}

# Calculate overlaps for both fishery types
dem_overlaps <- calculate_overlap(bird_layers, rast_dll, ext_dem)
pel_overlaps <- calculate_overlap(bird_layers, rast_pll, ext_pel)

#testing a different version that does not remove anything from any dataset, instead extending so that all cover the full southern ocean
# Full-extent overlap — resample fishing rasters to the bird raster grid as template,
# snapping to identical extent, resolution, and origin. New cells filled with NA
# (= 0 in sums). Tests whether the artificial ext_dem/ext_pel limits affect estimates.
bird_template <- bird_layers[[1]]

dll_aligned <- resample(rast_dll, bird_template, method = "bilinear")
pll_aligned <- resample(rast_pll, bird_template, method = "bilinear")

dem_overlap_simple <- map(bird_layers, ~{
  bird_norm <- .x / global(.x, fun = "sum", na.rm = TRUE)[1,1]
  dll_aligned * bird_norm
})

pel_overlap_simple <- map(bird_layers, ~{
  bird_norm <- .x / global(.x, fun = "sum", na.rm = TRUE)[1,1]
  pll_aligned * bird_norm
})
```

```{r summarize-total-hooks}
# Function to sum total hooks in overlap areas
sum_overlap_hooks <- function(overlap_list) { #takes list of the overlap rasters
  map_dbl(overlap_list, ~global(.x, "sum", na.rm = TRUE)[1,1]) #calculate total hook-weighted by bird density across all grid cells
}

# Calculate total hooks overlapping with each age class
hooks_summary <- tibble(
  age_class = names(bird_layers),
  demersal_hooks = sum_overlap_hooks(dem_overlaps),
  pelagic_hooks = sum_overlap_hooks(pel_overlaps),
  total_hooks = demersal_hooks + pelagic_hooks
)

print(hooks_summary)

# Compare cropped-extent vs full-extent overlap approaches
hooks_comparison <- tibble(
  age_class            = names(bird_layers),
  dem_cropped          = sum_overlap_hooks(dem_overlaps),
  dem_full             = sum_overlap_hooks(dem_overlap_simple),
  dem_diff_pct         = round((dem_full - dem_cropped) / dem_cropped * 100, 2),
  pel_cropped          = sum_overlap_hooks(pel_overlaps),
  pel_full             = sum_overlap_hooks(pel_overlap_simple),
  pel_diff_pct         = round((pel_full - pel_cropped) / pel_cropped * 100, 2)
)

print(hooks_comparison) # they are identical, good.

# Save summary table
write_csv(hooks_summary, here(paste0("output/overlap_hooks_summary_", date_sfx, ".csv")))
```

# Mapping

## Load base map data

```{r load-basemap}
# Load Natural Earth land polygons for mapping
ne_land <- st_read(here("data/ne_10m_land/ne_10m_land.shp"), quiet = TRUE)
```

## CCAMLR Fleet Double-Count Diagnostic

```{r diagnostic-ccamlr-fleet-overlap, fig.width=12, fig.height=8}
# Check for potential double-counting between CCAMLR fleet rows and standalone
# national datasets for all fleet codes that have a corresponding source file.
#
# NZL is handled separately using the grid-aligned raster (rast_dll_nz_aligned)
# because the standalone NZ file uses x.0 cell centers vs x.5 for all other sources.
# All other fleets are already on the x.5 grid in dll_data and can be matched directly.

# CCAMLR code → standalone dataset name
fleet_dataset_map <- tribble(
  ~ccamlr_code, ~standalone_dataset,
  "ARG",        "Argentina",
  "CHL",        "Chile",
  "ZAF",        "South Africa",
  "NAM",        "Namibia",
  "GBR",        "Falklands"
)

# Summarise mean annual effort per cell for a subset of dll_data
mean_annual_cells <- function(data) {
  data %>%
    group_by(Lon, Lat, Year) %>%
    summarise(annual_hooks = sum(Hooks, na.rm = TRUE), .groups = "drop") %>%
    group_by(Lon, Lat) %>%
    summarise(mean_effort = mean(annual_hooks, na.rm = TRUE), .groups = "drop")
}

# Build comparison list for non-NZ fleets
comparisons <- map2(fleet_dataset_map$ccamlr_code, fleet_dataset_map$standalone_dataset,
  function(code, standalone) {
    ccamlr_cells     <- mean_annual_cells(dll_data %>% filter(dataset == "CCAMLR", Fleet == code))
    standalone_cells <- mean_annual_cells(dll_data %>% filter(dataset == standalone))
    matched          <- standalone_cells %>%
                          inner_join(ccamlr_cells %>% select(Lon, Lat), by = c("Lon", "Lat"))
    list(ccamlr_code = code, standalone_name = standalone,
         ccamlr = ccamlr_cells, standalone = standalone_cells, matched = matched)
  }
)
names(comparisons) <- fleet_dataset_map$ccamlr_code

# NZL: use grid-aligned raster for standalone (converts x.0 → x.5 grid)
nz_aligned_df    <- as.data.frame(rast_dll_nz_aligned, xy = TRUE, na.rm = TRUE) %>%
                      rename(Lon = x, Lat = y)
ccamlr_nzl_cells <- mean_annual_cells(dll_data %>% filter(dataset == "CCAMLR", Fleet == "NZL"))
comparisons[["NZL"]] <- list(
  ccamlr_code     = "NZL",
  standalone_name = "New Zealand",
  ccamlr          = ccamlr_nzl_cells,
  standalone      = nz_aligned_df,
  matched         = nz_aligned_df %>%
                      inner_join(ccamlr_nzl_cells %>% select(Lon, Lat), by = c("Lon", "Lat"))
)

# Summary table: CCAMLR fleet vs standalone dataset spatial overlap
map_dfr(comparisons, function(comp) {
  tibble(
    ccamlr_code     = comp$ccamlr_code,
    standalone      = comp$standalone_name,
    ccamlr_hooks    = round(sum(comp$ccamlr$mean_effort,  na.rm = TRUE)),
    matched_hooks   = round(sum(comp$matched$mean_effort, na.rm = TRUE)),
    ccamlr_cells    = nrow(comp$ccamlr),
    matched_cells   = nrow(comp$matched)
  )
})

# Reference totals for % context — all in mean annual hooks (same units as matched_h
# from mean_annual_cells(), so comparisons are consistent across scales)
# rast_dll is the mean annual effort raster; global sum = total mean annual DLL hooks
dll_mean_annual_total <- global(rast_dll, "sum", na.rm = TRUE)[1,1]

# Southern Ocean: lat <= -30 (conventional boundary for DLL albatross overlap)
dll_so_mean_annual <- dll_data %>%
  filter(Lat <= -30) %>%
  group_by(Year) %>%
  summarise(h = sum(Hooks, na.rm = TRUE), .groups = "drop") %>%
  pull(h) %>%
  mean()

# CCAMLR total: mean annual hooks across all fleet codes
ccamlr_mean_annual <- dll_data %>%
  filter(dataset == "CCAMLR") %>%
  group_by(Year) %>%
  summarise(h = sum(Hooks, na.rm = TRUE), .groups = "drop") %>%
  pull(h) %>%
  mean()

# Standalone totals per dataset: mean annual hooks for each non-CCAMLR source
standalone_mean_annual <- dll_data %>%
  filter(dataset != "CCAMLR") %>%
  group_by(dataset, Year) %>%
  summarise(h = sum(Hooks, na.rm = TRUE), .groups = "drop") %>%
  group_by(dataset) %>%
  summarise(mean_annual = mean(h), .groups = "drop")

# % context table: matched_hooks (standalone side of overlap) as % of key reference totals
map_dfr(comparisons, function(comp) {
  standalone_ma <- standalone_mean_annual %>%
    filter(dataset == comp$standalone_name) %>%
    pull(mean_annual)
  matched_h <- sum(comp$matched$mean_effort, na.rm = TRUE)
  tibble(
    fleet              = comp$ccamlr_code,
    standalone_dataset = comp$standalone_name,
    matched_hooks      = round(matched_h),
    pct_total_dll      = round(100 * matched_h / dll_mean_annual_total, 3),
    pct_so_dll         = round(100 * matched_h / dll_so_mean_annual,    3),
    pct_ccamlr         = round(100 * matched_h / ccamlr_mean_annual,    3),
    pct_standalone     = if (length(standalone_ma) == 0 || standalone_ma == 0) NA_real_
                         else round(100 * matched_h / standalone_ma, 3)
  )
})

# Map function
make_diag_map <- function(df, title, map_extent) {
  if (nrow(df) == 0) {
    return(ggplot() +
      labs(title = paste(title, "(no matched cells)")) +
      theme_minimal() +
      theme(plot.title = element_text(size = 9, face = "bold")))
  }
  ggplot(df, aes(x = Lon, y = Lat, fill = mean_effort)) +
    geom_tile(width = 5, height = 5) +
    geom_sf(data = ne_land, inherit.aes = FALSE, fill = "grey80", color = NA) +
    coord_sf(xlim = map_extent[1:2], ylim = map_extent[3:4], expand = FALSE) +
    scale_fill_viridis_c(name = "Mean annual\nhooks", labels = scales::comma,
                         na.value = "transparent") +
    labs(title = title) +
    theme_minimal() +
    theme(plot.title  = element_text(size = 9, face = "bold"),
          legend.position = "bottom")
}

# Temporal overlap check for NZL — the only fleet with meaningful spatial matches
# NZ standalone uses x.0 grid vs CCAMLR x.5, so match within 2.5° (same 5° cell area)
# Data notes from NZ data holder confirm x.0 values are cell midpoints, not corners
nz_raw <- read_csv(here("data/Rep Log 17287 BLL effort truncated.csv"),
                   show_col_types = FALSE) %>%
  mutate(Lon = ifelse(truncated_long > 180, truncated_long - 360, truncated_long),
         Lat = truncated_lat) %>%
  filter(year >= 1990, year <= 2009)

ccamlr_nzl_raw <- dll_data %>%
  filter(dataset == "CCAMLR", Fleet == "NZL")

# Cross join and keep pairs within 2.5° (accounts for x.0 vs x.5 grid offset)
nzl_temporal <- ccamlr_nzl_raw %>%
  rename(ccamlr_lon = Lon, ccamlr_lat = Lat, ccamlr_year = Year, ccamlr_hooks = Hooks) %>%
  cross_join(
    nz_raw %>%
      select(nz_lon = Lon, nz_lat = Lat, nz_year = year, nz_hooks = total_hooks)
  ) %>%
  filter(abs(ccamlr_lon - nz_lon) <= 2.5,
         abs(ccamlr_lat - nz_lat) <= 2.5)

same_year_overlap <- nzl_temporal %>%
  filter(ccamlr_year == nz_year) %>%
  group_by(ccamlr_lon, ccamlr_lat, ccamlr_year) %>%
  summarise(ccamlr_hooks = sum(ccamlr_hooks, na.rm = TRUE),
            nz_hooks     = sum(nz_hooks,     na.rm = TRUE),
            .groups = "drop") %>%
  arrange(ccamlr_lon, ccamlr_lat, ccamlr_year)

# NZL temporal overlap summary
tibble(
  metric = c("Proximity-matched cell pairs (same 5° area)",
             "Cell-year combinations with same-year data in both",
             "Cell-year combinations with different-year data only"),
  count  = c(
    nrow(nzl_temporal %>% distinct(ccamlr_lon, ccamlr_lat, nz_lon, nz_lat)),
    nrow(same_year_overlap),
    nrow(nzl_temporal %>% filter(ccamlr_year != nz_year) %>%
           distinct(ccamlr_lon, ccamlr_lat, nz_lon, nz_lat))
  )
)

# Same-year matches — hooks per cell-year (genuine double-count risk)
print(same_year_overlap, n = Inf)

# Print one map pair per fleet
walk(comparisons, function(comp) {
  all_lons <- c(comp$ccamlr$Lon, comp$standalone$Lon)
  all_lats <- c(comp$ccamlr$Lat, comp$standalone$Lat)
  ext <- c(min(all_lons) - 5, max(all_lons) + 5,
           min(all_lats) - 5, max(all_lats) + 5)

  p1 <- make_diag_map(comp$ccamlr,  paste0("CCAMLR — ", comp$ccamlr_code), ext)
  p2 <- make_diag_map(comp$matched, paste0(comp$standalone_name, " — matched cells only"), ext)

  print(
    p1 + p2 +
      plot_annotation(title = paste("Double-count diagnostic:", comp$ccamlr_code,
                                    "vs", comp$standalone_name))
  )
})
```

## Fishing Effort Maps

```{r map-fishing-effort, fig.width=10, fig.height=6}
# Function to create fishing effort map
map_fishing_effort <- function(rast_obj, title, extent, transform = "sqrt",
                                discrete_bins = FALSE,
                                bin_breaks = c(0, 5, 10, 25, 50, 100, 500, Inf)) {
  df <- as.data.frame(rast_obj, xy = TRUE, na.rm = TRUE)
  names(df)[3] <- "effort"

  # Apply transformation for better visualization
  if (transform == "sqrt") {
    df$effort_trans <- sqrt(df$effort)
    legend_label <- "√(Hooks)"
  } else if (transform == "log") {
    df$effort_trans <- log(df$effort + 1)
    legend_label <- "log(Hooks)"
  } else {
    df$effort_trans <- df$effort
    legend_label <- "Hooks"
  }

  # Create base plot
  p <- ggplot(df, aes(x = x, y = y, fill = effort_trans)) +
    geom_raster() +
    geom_sf(data = ne_land, inherit.aes = FALSE, fill = "grey80", color = NA) +
    coord_sf(xlim = extent[1:2], ylim = extent[3:4], expand = FALSE)

  # Add appropriate fill scale
  if (discrete_bins) {
    # Use provided bin breaks (in units of 10^3 or 10^6 hooks)
    original_breaks <- bin_breaks

    # Create labels from breaks
    labels <- as.character(original_breaks)
    labels[length(labels)] <- paste0(">", original_breaks[length(original_breaks) - 1])

    # Transform breaks to match the data transformation
    if (transform == "sqrt") {
      breaks <- sqrt(original_breaks)
      breaks[1] <- 0  # sqrt(0) = 0
    } else if (transform == "log") {
      breaks <- log(original_breaks + 1)
      breaks[1] <- 0  # log(0+1) = 0
    } else {
      breaks <- original_breaks
    }

    # Create blue color palette (light to dark)
    blues <- brewer.pal(9, "Blues")[3:9]

    p <- p + scale_fill_stepsn(
      name = legend_label,
      colors = blues,
      breaks = breaks,
      labels = labels,
      na.value = "transparent",
      limits = c(0, max(df$effort_trans, na.rm = TRUE))
    ) +
    guides(fill = guide_colorbar(
      title.position = "top",
      title.hjust = 0.5,
      barwidth = unit(15, "cm"),
      barheight = unit(0.8, "cm"),
      frame.colour = "black",
      ticks.colour = "black",
      direction = "horizontal"
    ))
  } else {
    # Continuous gradient
    p <- p + scale_fill_gradientn(
      name = legend_label,
      colours = brewer.pal(9, "YlGnBu"),
      na.value = "transparent"
    )
  }

  p <- p + labs(title = title) +
    theme_minimal() +
    theme(
      plot.title = element_text(size = 14, face = "bold"),
      legend.position = "bottom",
      legend.title = element_text(size = 11, face = "bold"),
      legend.text = element_text(size = 10)
    )

  return(p)
}

# Create maps
p_dem <- map_fishing_effort(
  rast_dll,
  "Mean Annual Demersal Longline Effort (1990-2009)",
  ext_dem
)

p_pel <- map_fishing_effort(
  rast_pll,
  "Mean Annual Pelagic Longline Effort (1990-2009)",
  ext_pel
)

print(p_dem)
print(p_pel)

# Save maps
ggsave(here(paste0("output/maps/fishing_effort_demersal_", date_sfx, ".png")), p_dem,
       width = 10, height = 6, dpi = 300)
ggsave(here(paste0("output/maps/fishing_effort_pelagic_", date_sfx, ".png")), p_pel,
       width = 10, height = 6, dpi = 300)
```

## Scaled Fishing Effort Maps

```{r map-fishing-effort-scaled, fig.width=10, fig.height=12}
# Create maps for scaled mean effort - UNTRANSFORMED with discrete bins
p_dem_mean_scaled_raw <- map_fishing_effort(
  rast_dll_mean_scaled,
  "Mean Annual Demersal Longline Effort (1990-2009)\nScaled: 10^6 hooks",
  ext_dem,
  transform = "none",
  discrete_bins = TRUE
)

p_pel_mean_scaled_raw <- map_fishing_effort(
  rast_pll_mean_scaled,
  "Mean Annual Pelagic Longline Effort (1990-2009)\nScaled: 10^6 hooks",
  ext_pel,
  transform = "none",
  discrete_bins = TRUE
)

# Create maps for scaled mean effort - SQUARE ROOT with discrete bins
p_dem_mean_scaled_sqrt <- map_fishing_effort(
  rast_dll_mean_scaled,
  "Mean Annual Demersal Longline Effort (1990-2009)\nScaled: 10^6 hooks (sqrt transformed)",
  ext_dem,
  transform = "sqrt",
  discrete_bins = TRUE
)

p_pel_mean_scaled_sqrt <- map_fishing_effort(
  rast_pll_mean_scaled,
  "Mean Annual Pelagic Longline Effort (1990-2009)\nScaled: 10^6 hooks (sqrt transformed)",
  ext_pel,
  transform = "sqrt",
  discrete_bins = TRUE
)

# Create maps for scaled mean effort - LOG with discrete bins
# Demersal and pelagic both use same bins and scaling (10^6)
p_dem_mean_scaled_log <- map_fishing_effort(
  rast_dll_mean_scaled,
  "Mean Annual Demersal Longline Effort (1990-2009)\nScaled: 10^6 hooks (log transformed)",
  ext_dem,
  transform = "log",
  discrete_bins = TRUE,
  bin_breaks = c(0, 5, 10, 25, 50, 100, 500, Inf)
)

p_pel_mean_scaled_log <- map_fishing_effort(
  rast_pll_mean_scaled,
  "Mean Annual Pelagic Longline Effort (1990-2009)\nScaled: 10^6 hooks (log transformed)",
  ext_pel,
  transform = "log",
  discrete_bins = TRUE
)

# Create maps for scaled cumulative effort - UNTRANSFORMED with discrete bins
p_dem_cumulative_scaled_raw <- map_fishing_effort(
  rast_dll_cumulative_scaled,
  "Cumulative Demersal Longline Effort (1990-2009)\nScaled: 10^6 hooks",
  ext_dem,
  transform = "none",
  discrete_bins = TRUE
)

p_pel_cumulative_scaled_raw <- map_fishing_effort(
  rast_pll_cumulative_scaled,
  "Cumulative Pelagic Longline Effort (1990-2009)\nScaled: 10^6 hooks",
  ext_pel,
  transform = "none",
  discrete_bins = TRUE
)

# Create maps for scaled cumulative effort - SQUARE ROOT with discrete bins
p_dem_cumulative_scaled_sqrt <- map_fishing_effort(
  rast_dll_cumulative_scaled,
  "Cumulative Demersal Longline Effort (1990-2009)\nScaled: 10^6 hooks (sqrt transformed)",
  ext_dem,
  transform = "sqrt",
  discrete_bins = TRUE
)

p_pel_cumulative_scaled_sqrt <- map_fishing_effort(
  rast_pll_cumulative_scaled,
  "Cumulative Pelagic Longline Effort (1990-2009)\nScaled: 10^6 hooks (sqrt transformed)",
  ext_pel,
  transform = "sqrt",
  discrete_bins = TRUE
)

# Create maps for scaled cumulative effort - LOG with discrete bins
# Demersal and pelagic both use same bins and scaling (10^6)
p_dem_cumulative_scaled_log <- map_fishing_effort(
  rast_dll_cumulative_scaled,
  "Cumulative Demersal Longline Effort (1990-2009)\nScaled: 10^6 hooks (log transformed)",
  ext_dem,
  transform = "log",
  discrete_bins = TRUE,
  bin_breaks = c(0, 5, 10, 25, 50, 100, 500, Inf)
)

p_pel_cumulative_scaled_log <- map_fishing_effort(
  rast_pll_cumulative_scaled,
  "Cumulative Pelagic Longline Effort (1990-2009)\nScaled: 10^6 hooks (log transformed)",
  ext_pel,
  transform = "log",
  discrete_bins = TRUE
)

# Print all maps
print(p_dem_mean_scaled_raw)
print(p_dem_mean_scaled_sqrt)
print(p_dem_mean_scaled_log)
print(p_pel_mean_scaled_raw)
print(p_pel_mean_scaled_sqrt)
print(p_pel_mean_scaled_log)
print(p_dem_cumulative_scaled_raw)
print(p_dem_cumulative_scaled_sqrt)
print(p_dem_cumulative_scaled_log)
print(p_pel_cumulative_scaled_raw)
print(p_pel_cumulative_scaled_sqrt)
print(p_pel_cumulative_scaled_log)

# Save scaled mean effort maps - all transformations
ggsave(here(paste0("output/maps/fishing_effort_demersal_mean_scaled_raw_", date_sfx, ".png")),
       p_dem_mean_scaled_raw, width = 10, height = 6, dpi = 300)
ggsave(here(paste0("output/maps/fishing_effort_demersal_mean_scaled_sqrt_", date_sfx, ".png")),
       p_dem_mean_scaled_sqrt, width = 10, height = 6, dpi = 300)
ggsave(here(paste0("output/maps/fishing_effort_demersal_mean_scaled_log_", date_sfx, ".png")),
       p_dem_mean_scaled_log, width = 10, height = 6, dpi = 300)

ggsave(here(paste0("output/maps/fishing_effort_pelagic_mean_scaled_raw_", date_sfx, ".png")),
       p_pel_mean_scaled_raw, width = 10, height = 6, dpi = 300)
ggsave(here(paste0("output/maps/fishing_effort_pelagic_mean_scaled_sqrt_", date_sfx, ".png")),
       p_pel_mean_scaled_sqrt, width = 10, height = 6, dpi = 300)
ggsave(here(paste0("output/maps/fishing_effort_pelagic_mean_scaled_log_", date_sfx, ".png")),
       p_pel_mean_scaled_log, width = 10, height = 6, dpi = 300)

# Save scaled cumulative effort maps - all transformations
ggsave(here(paste0("output/maps/fishing_effort_demersal_cumulative_scaled_raw_", date_sfx, ".png")),
       p_dem_cumulative_scaled_raw, width = 10, height = 6, dpi = 300)
ggsave(here(paste0("output/maps/fishing_effort_demersal_cumulative_scaled_sqrt_", date_sfx, ".png")),
       p_dem_cumulative_scaled_sqrt, width = 10, height = 6, dpi = 300)
ggsave(here(paste0("output/maps/fishing_effort_demersal_cumulative_scaled_log_", date_sfx, ".png")),
       p_dem_cumulative_scaled_log, width = 10, height = 6, dpi = 300)

ggsave(here(paste0("output/maps/fishing_effort_pelagic_cumulative_scaled_raw_", date_sfx, ".png")),
       p_pel_cumulative_scaled_raw, width = 10, height = 6, dpi = 300)
ggsave(here(paste0("output/maps/fishing_effort_pelagic_cumulative_scaled_sqrt_", date_sfx, ".png")),
       p_pel_cumulative_scaled_sqrt, width = 10, height = 6, dpi = 300)
ggsave(here(paste0("output/maps/fishing_effort_pelagic_cumulative_scaled_log_", date_sfx, ".png")),
       p_pel_cumulative_scaled_log, width = 10, height = 6, dpi = 300)
```

check a mismatched section that doesn't show up in Clay et al 2019
```{r}
  # DLL sources in that box
  dll_data %>%
    filter(Lat > -40, Lon >= -90, Lon <= -70) %>%
    count(dataset, sort = TRUE)
```

## WAAL Distribution Maps

```{r map-waal-distributions, fig.width=10, fig.height=12}
# Function to create WAAL distribution map
map_waal_distribution <- function(rast_obj, age_class, extent, zlim = NULL) {
  # Convert to percent of population
  r_norm <- rast_obj / global(rast_obj, fun = "sum", na.rm = TRUE)[1,1]
  r_pct <- r_norm * 100

  # Filter out very low densities (< 0.1%)
  r_pct[r_pct < 0.1] <- NA

  df <- as.data.frame(r_pct, xy = TRUE, na.rm = TRUE)
  names(df)[3] <- "pct"

  p <- ggplot(df, aes(x = x, y = y, fill = pct)) +
    geom_raster() +
    geom_sf(data = ne_land, inherit.aes = FALSE, fill = "grey80", color = NA) +
    coord_sf(xlim = extent[1:2], ylim = extent[3:4], expand = FALSE) +
    scale_fill_gradientn(
      name = "% of Population",
      colours = brewer.pal(9, "YlGnBu"),
      na.value = "white",
      limits = zlim,
      oob = squish
    ) +
    labs(title = age_class) +
    theme_minimal() +
    theme(plot.title = element_text(size = 12, face = "bold"))

  return(p)
}

# Create maps for all age classes
dist_maps <- map2(bird_layers, names(bird_layers),
                   ~map_waal_distribution(.x, .y, ext_bird))

# Display in grid
wrap_plots(dist_maps, ncol = 2)

# Save individual maps
iwalk(dist_maps, function(p, age) {
  ggsave(
    here(paste0("output/maps/distribution_", age, "_", date_sfx, ".png")),
    p, width = 8, height = 6, dpi = 300
  )
})
```

## Overlap Maps

```{r map-overlap, fig.width=12, fig.height=10}
# Function to create overlap map with consistent color scale
map_overlap <- function(overlap_list, fishery_name, extent) {
  # Get global z-limits across all age classes for consistent coloring
  zlims <- map(overlap_list, ~global(.x, c("min", "max"), na.rm = TRUE)) %>%
    bind_rows() %>%
    summarize(min = min(min, na.rm = TRUE), max = max(max, na.rm = TRUE))

  # Create map for each age class
  maps <- imap(overlap_list, function(rast_obj, age_class) {
    df <- as.data.frame(rast_obj, xy = TRUE, na.rm = TRUE)
    names(df)[3] <- "overlap"

    ggplot(df, aes(x = x, y = y, fill = overlap)) +
      geom_raster() +
      geom_sf(data = ne_land, inherit.aes = FALSE, fill = "grey80", color = NA) +
      coord_sf(xlim = extent[1:2], ylim = extent[3:4], expand = FALSE) +
      scale_fill_viridis_c(
        name = "Hooks × Density",
        limits = c(zlims$min, zlims$max),
        na.value = "transparent",
        oob = squish
      ) +
      labs(title = age_class) +
      theme_minimal() +
      theme(plot.title = element_text(size = 10, face = "bold"))
  })

  return(maps)
}

# Create overlap maps
dem_overlap_maps <- map_overlap(dem_overlaps, "Demersal", ext_dem)
pel_overlap_maps <- map_overlap(pel_overlaps, "Pelagic", ext_pel)

# Display in grids
wrap_plots(dem_overlap_maps, ncol = 3, guides = "collect") +
  plot_annotation(title = "Demersal Longline Overlap by Age Class")

wrap_plots(pel_overlap_maps, ncol = 3, guides = "collect") +
  plot_annotation(title = "Pelagic Longline Overlap by Age Class")

# Save individual maps
iwalk(dem_overlap_maps, function(p, age) {
  ggsave(
    here(paste0("output/maps/overlap_demersal_", age, "_", date_sfx, ".png")),
    p, width = 8, height = 6, dpi = 300
  )
})

iwalk(pel_overlap_maps, function(p, age) {
  ggsave(
    here(paste0("output/maps/overlap_pelagic_", age, "_", date_sfx, ".png")),
    p, width = 8, height = 6, dpi = 300
  )
})
```

## Scaled Overlap Maps — Annual Average

```{r map-overlap-scaled, fig.width=14, fig.height=8}
# Overlap = fishing effort × normalized bird density (proportion of population per cell)
# Fixed 6-bin scale shared across all age classes and both fisheries:
#   [0, 10)        = white  (background / negligible overlap)
#   [10, 100)      = yellow
#   [100, 1000)    = light orange
#   [1000, 2000)   = orange
#   [2000, 5000)   = dark orange
#   [5000, 10000+] = red    (highest overlap; values above 10000 squished to red)

overlap_breaks <- c(10, 100, 1000, 2000, 5000)   # 5 breaks → 6 bins
overlap_upper  <- 10000
overlap_pal    <- c("white", brewer.pal(9, "YlOrRd")[4:8])  # orange (#FD8D3C) → dark red (#800026), no yellows

map_overlap_discrete <- function(overlap_list, fishery_name, extent) {
  imap(overlap_list, function(rast_obj, age_class) {
    df <- as.data.frame(rast_obj, xy = TRUE, na.rm = TRUE)
    names(df)[3] <- "overlap"

    ggplot(df, aes(x = x, y = y, fill = overlap)) +
      geom_raster() +
      geom_sf(data = ne_land, inherit.aes = FALSE, fill = "grey80", color = NA) +
      coord_sf(xlim = extent[1:2], ylim = extent[3:4], expand = FALSE) +
      scale_fill_stepsn(
        name   = "Overlap",
        breaks = overlap_breaks,
        colors = overlap_pal,
        na.value = "transparent",
        limits = c(0, overlap_upper),
        oob    = squish,
        labels = scales::label_comma(accuracy = 1)
      ) +
      guides(fill = guide_colorsteps(
        title.position = "top",
        title.hjust    = 0.5,
        barwidth       = unit(7, "cm"),
        barheight      = unit(0.4, "cm"),
        show.limits    = TRUE,
        label.theme    = element_text(size = 7)
      )) +
      labs(title = paste0(fishery_name, " \u2014 ", age_class)) +
      theme_minimal() +
      theme(
        plot.title      = element_text(size = 9, face = "bold"),
        legend.position = "bottom",
        legend.title    = element_text(size = 8, face = "bold"),
        legend.text     = element_text(size = 7),
        legend.margin   = margin(t = 1)
      )
  })
}

pel_overlap_maps_scaled <- map_overlap_discrete(pel_overlaps, "Pelagic Longline",  ext_pel)
dem_overlap_maps_scaled <- map_overlap_discrete(dem_overlaps, "Demersal Longline", ext_dem)

wrap_plots(pel_overlap_maps_scaled, ncol = 3, guides = "collect") +
  plot_annotation(title = "Pelagic Longline \u00d7 WAAL Overlap by Age Class (Annual Average)")

wrap_plots(dem_overlap_maps_scaled, ncol = 3, guides = "collect") +
  plot_annotation(title = "Demersal Longline \u00d7 WAAL Overlap by Age Class (Annual Average)")

# Save individual scaled overlap maps
iwalk(pel_overlap_maps_scaled, function(p, age) {
  ggsave(
    here(paste0("output/maps/overlap_pelagic_scaled_", age, "_", date_sfx, ".png")),
    p, width = 8, height = 6, dpi = 300
  )
})

iwalk(dem_overlap_maps_scaled, function(p, age) {
  ggsave(
    here(paste0("output/maps/overlap_demersal_scaled_", age, "_", date_sfx, ".png")),
    p, width = 8, height = 6, dpi = 300
  )
})
```

# Save Processed Data

```{r save-outputs}
# Create output directories if they don't exist
dir.create(here("output/rasters"), showWarnings = FALSE, recursive = TRUE)

# Save fishing effort rasters (unscaled)
writeRaster(rast_dll, here(paste0("output/rasters/fishing_effort_demersal_", date_sfx, ".tif")), overwrite = TRUE)
writeRaster(rast_pll, here(paste0("output/rasters/fishing_effort_pelagic_", date_sfx, ".tif")), overwrite = TRUE)

# Save scaled mean effort rasters (both scaled to 10^6 hooks)
writeRaster(rast_dll_mean_scaled, here(paste0("output/rasters/fishing_effort_demersal_mean_scaled_", date_sfx, ".tif")), overwrite = TRUE)
writeRaster(rast_pll_mean_scaled, here(paste0("output/rasters/fishing_effort_pelagic_mean_scaled_", date_sfx, ".tif")), overwrite = TRUE)

# Save cumulative effort rasters (unscaled)
writeRaster(rast_dll_cumulative, here(paste0("output/rasters/fishing_effort_demersal_cumulative_", date_sfx, ".tif")), overwrite = TRUE)
writeRaster(rast_pll_cumulative, here(paste0("output/rasters/fishing_effort_pelagic_cumulative_", date_sfx, ".tif")), overwrite = TRUE)

# Save scaled cumulative effort rasters (both scaled to 10^6 hooks)
writeRaster(rast_dll_cumulative_scaled, here(paste0("output/rasters/fishing_effort_demersal_cumulative_scaled_", date_sfx, ".tif")), overwrite = TRUE)
writeRaster(rast_pll_cumulative_scaled, here(paste0("output/rasters/fishing_effort_pelagic_cumulative_scaled_", date_sfx, ".tif")), overwrite = TRUE)

# Save overlap rasters for each age class
iwalk(dem_overlaps, function(r, age) {
  writeRaster(r, here(paste0("output/rasters/overlap_demersal_", age, "_", date_sfx, ".tif")), overwrite = TRUE)
})

iwalk(pel_overlaps, function(r, age) {
  writeRaster(r, here(paste0("output/rasters/overlap_pelagic_", age, "_", date_sfx, ".tif")), overwrite = TRUE)
})

# Save fleet/RFMO summary tables (numeric values, no comma formatting)
# Useful for downstream fleet-level bycatch partitioning
write_csv(pll_summary_rfmo,  here(paste0("output/pll_summary_by_rfmo_",     date_sfx, ".csv")))
write_csv(pll_summary_flag,  here(paste0("output/pll_summary_by_flag_",     date_sfx, ".csv")))
write_csv(dll_summary_dataset, here(paste0("output/dll_summary_by_dataset_", date_sfx, ".csv")))
write_csv(dll_summary_fleet,   here(paste0("output/dll_summary_by_fleet_",   date_sfx, ".csv")))
```

# Session Info

```{r session-info}
sessionInfo()
```
