---
title: "Wandering Albatross Bycatch - Spatial Catchability v2 (Fleet-by-Period)"
author: "WAAL Bycatch Project"
date: "`r Sys.Date()`"
output: html_document
---

note - we may want to switch the total density maps to actual numbers not normalized (decided not to as global estimates of 20100 are just MI and additional assumptions are needed to get total population)

also, may want to update the gap filling fallback calculation (mean per fishery) to be hooks-weighted (left as is too! : On hooks-weighting the mean beta — my honest opinion is no, it's not appropriate here. Beta is already a per-hook rate (BPUE divided by bird density, normalized for
  effort). Taking a hooks-weighted mean would effectively double-weight high-effort studies, since hooks already went into calculating each study's beta. The simple mean
  treats each study as an independent estimate of catchability, which is the right framing since beta is supposed to be a property of gear/practices, not of fishing
  volume. Where hooks do legitimately matter is in the B_study calculation (already hooks-weighted correctly) and potentially as a measure of statistical precision — but
  that would call for weighting by sample size or confidence interval width, not raw hook counts.)
  
  
need to ADD J1 back into calculations - DONE

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.width = 10,
  fig.height = 8
)
```

# Overview

This script implements a **spatially-varying catchability approach** using fleet-by-period
effort data and the BPUE literature review from script 04. 

## Method

**Step 1: Estimate per-study catchability**

For each BPUE study (fleet, fishery, time period, bounding box):

```
β_i = BPUE_i / (1000 × B_study_i)
```

where B_study = hooks-weighted mean TOTAL WAAL density in the study's bounding box.
This extracts "pure catchability" — the per-hook risk per unit bird density.
β is a property of the fishery's gear/practices, independent of location.

This^ builds fleet × period-specific catchability rasters spatially. 

**Step 2: Calculate bycatch using BI-specific bird density**

For each fleet × fishery × period combo, it checks which effort cells fall within each BPUE study's bounding box and assigns that study's beta to those cells. Where multiple studies overlap a cell, it averages their betas. 

It then used this equation:

```
bycatch_cell = β_fishery × hooks_cell × BI_birds_cell
```
to get bycatch per cell per fleet/fishery/time period. This is summed across each raster per fleet to get fleet bycatch, then across fishery (for each time period) to get total bycatch per fishery per time perid.

Uses Bird Island age-class density rasters directly (from script 02).
Total bycatch depends on the spatial overlap of hooks and BI birds.
Since β is calibrated from study areas (typically high total bird density)
but applied with BI-specific density, bird density genuinely matters — areas
with fewer BI birds contribute proportionally less bycatch.

# Section 1: Setup

```{r libraries}
library(tidyverse)
library(terra)
library(sf)
library(here)
library(viridis)
library(patchwork)

# Date suffix for versioned output files — prevents overwriting previous runs
date_sfx <- format(Sys.Date(), "%Y-%m-%d")
```

# Section 2: Load Base Data

## Bird densities and distributions

```{r load-bird-data}
# Total bird density (all populations combined, from script 02)
# Extent: -180..180, -90..0 — this is the master grid
total_bird_density <- rast(here("output/rasters/total_bird_density_all_pops.tif"))

# Load raw Clay age-class distributions (extent -90..0, correct)
# Raw TIFs are approximately normalized but we re-normalize explicitly below
clay_raw <- list(
  fb   = rast(here("data/WAAL_dist_Clay2019/WA_AllMonths_Both_FB_1990-2009.tif")),
  sb   = rast(here("data/WAAL_dist_Clay2019/WA_AllMonths_Both_SB_1990-2009.tif")),
  nb   = rast(here("data/WAAL_dist_Clay2019/WA_AllMonths_Both_NB_1990-2009.tif")),
  j1   = rast(here("data/WAAL_dist_Clay2019/WA_AllMonths_Both_J1_1990-2009.tif")),
  j2j3 = rast(here("data/WAAL_dist_Clay2019/WA_AllMonths_Both_J2+3_1990-2009.tif")),
  imm  = rast(here("data/WAAL_dist_Clay2019/WA_AllMonths_Both_IMM_1990-2009.tif"))
)
```

## Demographic proportions

```{r load-demographics}
dem_props <- read_csv(here("data/WAAL_dist_Clay2019/Dem_props copy.csv"),
                      show_col_types = FALSE) %>%
  mutate(Dem_class = str_replace(Dem_class, "J2\\+3", "J2J3"))

# Use only the 4 individual period rows to calculate means
# (exclude the 1990-2009 summary row to avoid double-counting)
dem_props_summary <- dem_props %>%
  filter(Period != "1990-2009") %>%
  group_by(Dem_class) %>%
  summarise(mean_prop = mean(Prop), mean_pop = mean(Pop))

age_class_props <- c(
  fb   = dem_props_summary$mean_prop[dem_props_summary$Dem_class == "FB"],
  sb   = dem_props_summary$mean_prop[dem_props_summary$Dem_class == "SB"],
  nb   = dem_props_summary$mean_prop[dem_props_summary$Dem_class == "NB"],
  j1   = dem_props_summary$mean_prop[dem_props_summary$Dem_class == "J1"],
  j2j3 = dem_props_summary$mean_prop[dem_props_summary$Dem_class == "J2J3"],
  imm  = dem_props_summary$mean_prop[dem_props_summary$Dem_class == "IMM"]
)

pop_sizes <- c(
  fb   = dem_props_summary$mean_pop[dem_props_summary$Dem_class == "FB"],
  sb   = dem_props_summary$mean_pop[dem_props_summary$Dem_class == "SB"],
  nb   = dem_props_summary$mean_pop[dem_props_summary$Dem_class == "NB"],
  j1   = dem_props_summary$mean_pop[dem_props_summary$Dem_class == "J1"],
  j2j3 = dem_props_summary$mean_pop[dem_props_summary$Dem_class == "J2J3"],
  imm  = dem_props_summary$mean_pop[dem_props_summary$Dem_class == "IMM"]
)
```

## Build BI age-class density rasters

Builds `bi_dists`, the BI bird density rasters used in bycatch calculations (Section 9).
Each cell value represents the proportion of the global WAAL population that are BI
birds of that age class. J1 is included using J2J3 survival rates in the demographic
model (no J1-specific vital rates available; assumed equal to J2J3).

Three steps applied sequentially to the raw Clay distributions:
1. Normalize each age-class distribution to sum to 1 (`clay_norm`)
2. Weight by age-class demographic proportion (`clay_weighted`)
3. Scale by BI share of global population (`bi_dists`)

After step 3, each age-class raster sums to `prop_BI × age_class_props[age]`, not 1.
The remaining code verifies these sums are correct.

Note: constructed directly from Clay distributions on the correct bird distribution
grid (-90..0). Previous versions loaded pre-built rasters from script 03_final which
were on the fishing effort grid (-50..65), cutting off South Georgia at -54°S.

```{r build-bi-density}
# Population proportions (from script 02)
# Bird Island = 60% of South Georgia = 60% × 1553/5947
breeding_pairs <- tibble(
  population = c("Crozet", "Kerguelen", "Prince Edward", "Marion", "South Georgia"),
  pairs = c(340, 354, 1800, 1900, 1553)
)
total_pairs <- sum(breeding_pairs$pairs)
prop_southg <- breeding_pairs$pairs[breeding_pairs$population == "South Georgia"] / total_pairs
prop_BI <- prop_southg * 0.6

# Normalize Clay distributions (replace NA with 0, then divide by sum)
clay_norm <- map(clay_raw, function(r) {
  r_zero <- subst(r, NA, 0)
  r_sum <- global(r_zero, "sum", na.rm = TRUE)[1, 1]
  r_zero / r_sum
})

# Verify normalization
clay_sums <- map_dbl(clay_norm, ~global(.x, "sum", na.rm = TRUE)[1, 1])

# Weight each age class by its demographic proportion
clay_weighted <- map2(clay_norm, age_class_props, function(r, prop) {
  r * prop
})

# Scale by BI proportion of global population
bi_dists <- map(clay_weighted, function(r) {
  r * prop_BI
})

# Total BI bird density (sum across age classes)
bi_density_total <- app(rast(bi_dists), sum, na.rm = TRUE)

# Verify BI density sums: each age class should sum to prop_BI * age_class_props[age]
# Total across all age classes should equal prop_BI (~0.157)
check <- tibble(
  age_class = names(bi_dists),
  sum = map_dbl(bi_dists, ~global(.x, "sum", na.rm = TRUE)[1, 1]),
  expected = prop_BI * age_class_props
) %>%
  mutate(across(where(is.numeric), ~round(.x, 6)))

# Total should equal prop_BI * sum(age_class_props)
# With J1 included, age_class_props should now sum to ~1 and total should equal prop_BI
round(global(bi_density_total, "sum", na.rm = TRUE)[1, 1], 6)
round(prop_BI * sum(age_class_props), 6)
check
round(sum(age_class_props), 6)
```

## Basemap

```{r load-basemap}
ne_land <- st_read(here("data/ne_10m_land/ne_10m_land.shp"), quiet = TRUE)
```

# Section 3: Load Fleet-by-Period Effort Data

## Pelagic longline

```{r load-pll}
# Read cleaned PLL effort from script 01 output (includes corrected WCPFC data, filtered to 1990-2009)
pll_data <- read_csv(here("output/pll_effort_by_fleet.csv"), show_col_types = FALSE)

# Standardize fleet names to match BPUE CSV naming
# Merge NA and "Other" into a single "Other" category (likely same unlabeled effort
# recorded differently across RFMOs)
pll_data <- pll_data %>%
  mutate(fleet = case_when(
    Flag == "South Korea" ~ "Korea",
    is.na(Flag) ~ "Other",
    Flag == "Other" ~ "Other",
    TRUE ~ Flag
  ))

# Assign year groups
target_year_groups <- c("1990-1994", "1995-1999", "2000-2004", "2005-2009")

pll_data <- pll_data %>%
  mutate(year_group = case_when(
    Year >= 1990 & Year <= 1994 ~ "1990-1994",
    Year >= 1995 & Year <= 1999 ~ "1995-1999",
    Year >= 2000 & Year <= 2004 ~ "2000-2004",
    Year >= 2005 & Year <= 2009 ~ "2005-2009",
    TRUE ~ NA_character_
  )) %>%
  filter(!is.na(year_group))

pll_data %>% count(fleet, sort = TRUE) %>% print(n = 30)
```

## Demersal longline

```{r load-dll}
# Read cleaned DLL effort from script 01 output (all sources combined, Chile PLL filtered,
# NZ coordinates at x.0 cell midpoints per data holder notes, filtered to 1990-2009)
dll_data <- read_csv(here("output/dll_effort_by_fleet.csv"), show_col_types = FALSE) %>%
  rename(fleet = Fleet) %>%
  mutate(year_group = case_when(
    Year >= 1990 & Year <= 1994 ~ "1990-1994",
    Year >= 1995 & Year <= 1999 ~ "1995-1999",
    Year >= 2000 & Year <= 2004 ~ "2000-2004",
    Year >= 2005 & Year <= 2009 ~ "2005-2009",
    TRUE ~ NA_character_
  )) %>%
  filter(!is.na(year_group))

dll_data %>% count(fleet, sort = TRUE) %>% print(n = 20)
```

## Rasterize effort by fleet x period

Aggregate effort to 5x5 degree mean annual hooks per cell, for each fleet x year_group.

```{r rasterize-effort}
# Use total_bird_density as the template grid
template <- total_bird_density

# Function: aggregate effort data to a raster for one fleet x period
# Uses MEAN ANNUAL hooks (total hooks / number of years in period) so that
# bycatch = beta * hooks * birds gives an annual estimate
make_effort_raster <- function(effort_data, fleet_name, yr_group, template_rast) {
  # Count distinct years with data for this fleet x period
  # Used to get true mean annual hooks (divide by years with data, not period length)
  n_years <- effort_data %>%
    filter(fleet == fleet_name, year_group == yr_group) %>%
    pull(Year) %>%
    n_distinct()
  if (n_years == 0) return(NULL)

  effort_subset <- effort_data %>%
    filter(fleet == fleet_name, year_group == yr_group, !is.na(Lon), !is.na(Lat)) %>%
    group_by(Lon, Lat) %>%
    summarise(total_hooks = sum(Hooks, na.rm = TRUE) / n_years, .groups = "drop")

  if (nrow(effort_subset) < 1) return(NULL)

  # Rasterize directly onto the template grid — places each point in its correct
  # cell without interpolation, handles single-lat or single-lon fleets correctly,
  # and matches extent automatically (no resample needed)
  effort_pts <- st_as_sf(effort_subset, coords = c("Lon", "Lat"), crs = 4326)
  rasterize(effort_pts, template_rast, field = "total_hooks", fun = "sum", background = NA)
}

# Build a nested list: effort_rasters[[fishery]][[fleet]][[year_group]]
# Also build a flat lookup table for iteration
effort_combos <- bind_rows(
  pll_data %>% distinct(fleet, year_group) %>% mutate(fishery = "PLL"),
  dll_data %>% distinct(fleet, year_group) %>% mutate(fishery = "DLL")
)

# Create effort rasters for all combos
effort_rasters <- list()

for (i in 1:nrow(effort_combos)) {
  fi <- effort_combos$fishery[i]
  fl <- effort_combos$fleet[i]
  yg <- effort_combos$year_group[i]

  src_data <- if (fi == "PLL") pll_data else dll_data
  r <- make_effort_raster(src_data, fl, yg, template)

  if (!is.null(r)) {
    key <- paste(fi, fl, yg, sep = "|")
    effort_rasters[[key]] <- r
  }
}

# Verify all expected fleet x period combos produced a raster
tibble(
  total_combos = nrow(effort_combos),
  rasters_created = length(effort_rasters),
  dropped = nrow(effort_combos) - length(effort_rasters)
)

# Helper to retrieve an effort raster
get_effort_raster <- function(fishery, fleet, year_group) {
  key <- paste(fishery, fleet, year_group, sep = "|")
  effort_rasters[[key]]
}

# Also create total effort rasters per fishery x period (summed across fleets)
# for use in gap-filling beta conversion
fishery_period_combos <- expand_grid(
  fishery = c("PLL", "DLL"),
  year_group = target_year_groups
)

total_effort_by_fp <- list()
for (i in 1:nrow(fishery_period_combos)) {
  fi <- fishery_period_combos$fishery[i]
  yg <- fishery_period_combos$year_group[i]

  # Sum all fleet rasters for this fishery x period
  matching_keys <- names(effort_rasters)[str_detect(names(effort_rasters),
                                                     paste0("^", fi, "\\|.*\\|", yg, "$"))]
  if (length(matching_keys) > 0) {
    stacked <- rast(effort_rasters[matching_keys])
    total_r <- app(stacked, sum, na.rm = TRUE)
    total_effort_by_fp[[paste(fi, yg, sep = "|")]] <- total_r
  }
}


#checking rasters
#For effort_rasters (flat list, keyed by fishery|fleet|year_group):       
  # Summary table across all rasters                                                                                                                       
  map_dfr(effort_rasters, ~global(.x, fun = c("min", "max", "sum"), na.rm = TRUE),                                                                         
          .id = "key")                                                                                                                                     
                                                                                                                                                           
  # Spot-check one raster                                                                                                                                  
 global(effort_rasters[["PLL|Japan|1990-1994"]], fun = c("min", "max", "sum"), na.rm = TRUE)                                                           
                                                                                             
  # Non-zero cell count for each                                                                                                                           
  map_dfr(effort_rasters,                                                                                                                                  
          ~tibble(n_nonzero = sum(values(.x) > 0, na.rm = TRUE),                                                                                           
                  n_na = sum(is.na(values(.x)))),                                                                                                          
          .id = "key")      
  
      # Find any rasters where sum == 0
  map_dfr(effort_rasters, ~global(.x, fun = "sum", na.rm = TRUE), .id = "key") %>%                                                                         
    filter(sum == 0)   
    
  # Do the source rows have valid hooks?                                                                                                                   
  pll_data %>% filter(fleet == "Canada", year_group == "1990-1994") %>%                                                                                    
    summarise(n = n(), n_na_hooks = sum(is.na(Hooks)), sum_hooks = sum(Hooks, na.rm = TRUE))                                                               
                                                                                                                                                           
  # Do coordinates fall within the template extent?                                                                                                        
  pll_data %>% filter(fleet == "Canada", year_group == "1990-1994") %>%                                                                                    
    summarise(lon_range = range(Lon, na.rm=TRUE), lat_range = range(Lat, na.rm=TRUE))                                                                      
  ext(template)     
                                                                                                                                                           
  #For total_effort_by_fp (keyed by fishery|year_group):                                                                                                    
  map_dfr(total_effort_by_fp, ~global(.x, fun = c("min", "max", "sum"), na.rm = TRUE),                                                                     
          .id = "key")   

     
```

# Section 4: Load BPUE Literature Data

```{r load-bpue}
bpue_std <- read_csv(here("data/BPUE_2926_standardized.csv"), show_col_types = FALSE)

# Filter to rows with valid bounding boxes and BPUE values
bpue_with_bbox <- bpue_std %>%
  filter(!is.na(lat_min) & !is.na(bpue))

# Function to create sf polygon from bbox coordinates
# Handles dateline crossings (lon_max > 180)
make_bbox_polygon <- function(lat_min, lat_max, lon_min, lon_max) {
  if (lon_max > 180) {
    coords1 <- matrix(c(lon_min, lat_min, 180, lat_min, 180, lat_max,
                         lon_min, lat_max, lon_min, lat_min),
                       ncol = 2, byrow = TRUE)
    lon_max_wrapped <- lon_max - 360
    coords2 <- matrix(c(-180, lat_min, lon_max_wrapped, lat_min,
                         lon_max_wrapped, lat_max, -180, lat_max, -180, lat_min),
                       ncol = 2, byrow = TRUE)
    st_multipolygon(list(list(coords1), list(coords2)))
  } else {
    coords <- matrix(c(lon_min, lat_min, lon_max, lat_min, lon_max, lat_max,
                        lon_min, lat_max, lon_min, lat_min),
                     ncol = 2, byrow = TRUE)
    st_polygon(list(coords))
  }
}

# Build sf object
bpue_sf <- bpue_with_bbox %>%
  rowwise() %>%
  mutate(geometry = list(make_bbox_polygon(lat_min, lat_max, lon_min, lon_max))) %>%
  ungroup() %>%
  mutate(geometry = st_sfc(geometry, crs = 4326)) %>%
  st_as_sf()

# Assign each study to ALL target periods it overlaps — uses numeric overlap logic
# so non-standard ranges (e.g. "1994-1999", "2006-2010") are handled correctly.
# Studies spanning multiple periods contribute to each, which is appropriate since
# we are averaging betas not summing them.
# Studies outside 1990-2009 (e.g. 2010-2014) return no match and are excluded.
expand_to_target_periods <- function(yg_string) {
  if (is.na(yg_string) | yg_string == "") return(character(0))

  period_bounds <- tibble(
    period = target_year_groups,
    start  = c(1990, 1995, 2000, 2005),
    end    = c(1994, 1999, 2004, 2009)
  )

  periods <- str_trim(str_split(yg_string, ",")[[1]])

  matched <- map(periods, function(p) {
    nums <- as.numeric(str_extract_all(p, "\\d{2,4}")[[1]])
    if (length(nums) < 1) return(character(0))
    start_yr <- nums[1]
    end_yr   <- if (length(nums) >= 2) nums[2] else nums[1]
    # Handle 2-digit end year (e.g. "1990-94" -> 1994)
    if (end_yr < 100) end_yr <- (start_yr %/% 100) * 100 + end_yr
    period_bounds %>%
      filter(start_yr <= end & end_yr >= start) %>%
      pull(period)
  }) %>% unlist() %>% unique()

  return(matched)
}

bpue_expanded <- bpue_sf %>%
  rowwise() %>%
  mutate(plot_year_group = list(expand_to_target_periods(year_group))) %>%
  ungroup() %>%
  unnest(plot_year_group)

# Summary: studies per fleet x fishery x period (studies spanning multiple periods
# appear in each relevant period)
bpue_expanded %>%
  st_drop_geometry() %>%
  count(fishery_type, fleet, plot_year_group) %>%
  pivot_wider(names_from = plot_year_group, values_from = n, values_fill = 0) %>%
  print(n = 50)

# Do NOT reroute fleets without effort data (Uruguay, NZ) to "Other".
# Instead, keep them as-is. They will simply be skipped in beta calculation
# (Section 6) since there's no matching effort raster, preventing their
# high BPUEs from inflating the "Other" fleet's catchability.
pll_effort_fleets <- pll_data %>% distinct(fleet) %>% pull()
dll_effort_fleets <- dll_data %>% distinct(fleet) %>% pull()

# Flag fleets with no effort data (for reporting only)
bpue_expanded <- bpue_expanded %>%
  mutate(has_effort_data = case_when(
    fishery_type == "PLL" ~ fleet %in% pll_effort_fleets,
    fishery_type == "DLL" ~ fleet %in% dll_effort_fleets,
    TRUE ~ FALSE
  ))

# Fleets WITHOUT matching effort data (will be skipped, not rerouted)
bpue_expanded %>%
  st_drop_geometry() %>%
  filter(!has_effort_data) %>%
  distinct(fishery_type, fleet) %>%
  print()

# Summary
bpue_expanded %>%
  st_drop_geometry() %>%
  count(fishery_type, fleet, plot_year_group) %>%
  pivot_wider(names_from = plot_year_group, values_from = n, values_fill = 0) %>%
  print(n = 50)
```

# Section 5: Load BPUE Summary Statistics (for gap-filling)

```{r load-bpue-summaries}
bpue_by_fleet <- read_csv(here("output/bpue_summary_by_fleet.csv"), show_col_types = FALSE)
bpue_by_region <- read_csv(here("output/bpue_summary_by_region.csv"), show_col_types = FALSE)
bpue_by_period <- read_csv(here("output/bpue_summary_by_period.csv"), show_col_types = FALSE)
bpue_by_fishery <- read_csv(here("output/bpue_summary_by_fishery.csv"), show_col_types = FALSE)

# Define ocean regions (same as script 04)
region_defs <- tribble(
  ~region,            ~lon_min, ~lon_max, ~lat_min, ~lat_max,
  "SW Atlantic",         -70,      -20,      -60,        0,
  "SE Atlantic",         -20,       20,      -60,        0,
  "SW Indian",            20,       80,      -60,        0,
  "SE Indian",            80,      140,      -60,        0,
  "SW Pacific",          140,      180,      -60,        0,
  "SE Pacific",         -180,      -70,      -60,        0,
  "Southern Ocean",     -180,      180,      -90,      -60
)

# Function to assign a point to a region
assign_region <- function(lon, lat) {
  matches <- region_defs %>%
    filter(lon >= lon_min, lon < lon_max, lat >= lat_min, lat < lat_max) %>%
    pull(region)
  if (length(matches) == 0) return(NA_character_)
  return(matches[1])
}
```

# Section 6: Calculate Catchability per BPUE Study

# Notation (Dan’s method)
# C       = bycatch mortalities
# H       = hooks
# B       = number of birds
# α_mn    = BPUE in the traditional sense: observed dead birds per 1000 hooks
#           In code: bpue_1 = BPUE / 1000 (converted to birds per hook)
# β       = catchability: probability of a bird being caught per hook per bird.
#           Analogous to the interaction coefficient in Lotka-Volterra predator-prey
#           equations. Necessary to partition total BPUE spatially.

# Derivation:
# Total bycatch can be expressed two ways:
#   (1) C_total = α_mn × H_total          (definition of BPUE)
#   (2) C_total = sum(H_i × B_i × β)      (sum across N spatial blocks i,
#                                           where H_i × β = catch prob per bird
#                                           in block i, × B_i birds)
# Setting equal and solving for β:
#   β = α_mn × H_total / sum(H_i × B_i)
#
# In the code, BPUE is in birds per 1000 hooks, so bpue_1 = BPUE / 1000:
#   β = bpue_1 × H_total / sum(H_i × B_i)
#
# H_total and sum(H_i × B_i) are computed within each study’s bounding box
# using the fleet effort raster (hooks per cell) and the total WAAL density raster.
#
# β is then applied spatially to estimate bycatch in each grid cell:
#   C_cell = β × H_cell × B_BI_cell
# where B_BI_cell is Bird Island bird density (not total WAAL density),
# since we are estimating BI-specific bycatch.
#
# Beta is applied as a per-cell catchability surface (catchability_rasters).
# Each cell's beta is the mean of all BPUE studies whose bounding box covers
# that cell. Cells not covered by any study are gap-filled hierarchically:
# fleet temporal extrapolation → regional mean → fishery mean. Bycatch per
# cell is then:
#   C_cell = beta_cell × hooks_cell × BI_birds_cell
# summed across all cells to get total mortalities per fleet x period x age class.




# This chunk calculates beta (catchability) for each BPUE study using Dan's equation:
#   beta = bpue_1 * H_total / sum(H_i * B_i)
# where bpue_1 = BPUE / 1000 (converted from birds per 1000 hooks to birds per hook).
#
# Line 612: The loop iterates over every row in bpue_betas — one row per BPUE study x period.
#
# Lines 613-621: Pulls fishery, fleet, period, and BPUE value for this study. Looks up
#   the effort raster for that fleet x period. If there is no effort raster (fleet not
#   in our data), skips to the next study.
#
# Lines 623-625: Gets the study's geographic bounding box from its sf geometry and
#   converts it to a terra extent object for cropping.
#
# Lines 627-631 — the tryCatch: Crops both the effort raster and total_bird_density to
#   the study bounding box. The tryCatch handles cases where the bounding box falls
#   entirely outside the raster extent (crop() would error). This is a legitimate edge
#   case — some BPUE studies have bounding boxes outside the Southern Ocean where we
#   have bird data. If either crop fails, skips that study.
#
# Lines 633-634: Converts BPUE from birds per 1000 hooks to birds per hook (bpue_1).
#
# Lines 636-641: Computes H_total (total hooks in box) and sum(H_i * B_i) (sum of
#   hooks x bird density across cells). Skips if either is zero or NA.
#
# Lines 643-647: Applies Dan's equation to get beta. Also stores B_study =
#   sum(H_i * B_i) / H_total (hooks-weighted mean bird density) as a diagnostic —
#   not used in the beta calculation.
#
# Lines 668+: After the loop, averages beta across all studies within each fishery
#   type (PLL/DLL). This fishery-level mean beta is used only as a last-resort
#   fallback in gap-filling (Section 7), not in the bycatch calculations themselves.

```{r calculate-betas}
# Total BI density (sum across age classes, on template grid)
bi_density_aligned <- app(rast(bi_dists), sum, na.rm = TRUE)

# Calculate beta for each BPUE study using bounding-box TOTAL WAAL density
bpue_geom <- bpue_expanded

bpue_betas <- bpue_expanded %>%
  st_drop_geometry() %>%
  mutate(beta = NA_real_,
         B_study = NA_real_,
         fleet_hooks_in_box = NA_real_,
         has_effort_match = FALSE)

for (i in 1:nrow(bpue_betas)) {
  fi <- bpue_betas$fishery_type[i]
  fl <- bpue_betas$fleet[i]
  yg <- bpue_betas$plot_year_group[i]
  bpue_val <- bpue_betas$bpue[i]

  # Get fleet effort raster for this fleet x period
  key <- paste(fi, fl, yg, sep = "|")
  effort_r <- effort_rasters[[key]]
  if (is.null(effort_r)) next

  # Get study bounding box from geometry
  bbox <- st_bbox(bpue_geom[i, ])
  bbox_ext <- ext(bbox["xmin"], bbox["xmax"], bbox["ymin"], bbox["ymax"])

  # Crop effort and TOTAL bird density to bounding box
  effort_crop <- tryCatch(crop(effort_r, bbox_ext), error = function(e) NULL)
  birds_crop <- tryCatch(crop(total_bird_density, bbox_ext), error = function(e) NULL)

  if (is.null(effort_crop) || is.null(birds_crop)) next

  # Convert BPUE from birds per 1000 hooks to birds per hook
  bpue_1 <- bpue_val / 1000

  # Compute H_total and sum(H_i x B_i) within the study bounding box
  hooks_in_box         <- global(effort_crop, "sum", na.rm = TRUE)[1, 1]
  hooks_x_birds_in_box <- global(effort_crop * birds_crop, "sum", na.rm = TRUE)[1, 1]

  if (is.na(hooks_in_box) || hooks_in_box <= 0) next
  if (is.na(hooks_x_birds_in_box) || hooks_x_birds_in_box <= 0) next

  # Dan's equation: beta = bpue_1 * H_total / sum(H_i * B_i)
  bpue_betas$fleet_hooks_in_box[i] <- hooks_in_box
  bpue_betas$beta[i]               <- bpue_1 * hooks_in_box / hooks_x_birds_in_box
  bpue_betas$B_study[i]            <- hooks_x_birds_in_box / hooks_in_box  # diagnostic: hooks-weighted mean bird density in study area
  bpue_betas$has_effort_match[i]   <- TRUE
}

# Summary of per-study beta calculations
tibble(
  total_rows = nrow(bpue_betas),
  beta_calculated = sum(!is.na(bpue_betas$beta)),
  no_effort_match = sum(is.na(bpue_betas$beta))
)

# Per-study beta values
bpue_betas %>%
  filter(!is.na(beta)) %>%
  select(fishery_type, fleet, plot_year_group, bpue, B_study, beta,
         fleet_hooks_in_box) %>%
  mutate(across(where(is.numeric), ~signif(.x, 4))) %>%
  print(n = 50)

# Compute FISHERY-LEVEL MEAN BETA
# This is the key step: average beta across all studies within each fishery type.
# Each study contributes equally (simple mean). This gives a single catchability
# value per fishery that represents "how dangerous this gear type is per hook
# per unit bird density."
beta_by_fishery_type <- bpue_betas %>%
  filter(!is.na(beta)) %>%
  group_by(fishery_type) %>%
  summarise(
    n_studies = n(),
    mean_beta = mean(beta),
    median_beta = median(beta),
    sd_beta = sd(beta),
    min_beta = min(beta),
    max_beta = max(beta),
    mean_B_study = mean(B_study),
    .groups = "drop"
  )

# Fishery-level mean beta (applied uniformly to all cells)
beta_by_fishery_type %>%
  mutate(across(where(is.numeric), ~signif(.x, 4))) %>%
  print()

# Also compute fleet-level mean beta for diagnostic comparison
beta_by_fleet_type <- bpue_betas %>%
  filter(!is.na(beta)) %>%
  group_by(fishery_type, fleet) %>%
  summarise(
    n_studies = n(),
    mean_beta = mean(beta),
    .groups = "drop"
  )

# Fleet-level mean beta (for reference)
beta_by_fleet_type %>%
  mutate(mean_beta = signif(mean_beta, 4)) %>%
  print(n = 20)
```


```{r}

# --- Supplement: compare used vs skipped BPUE values ---                                                                                                

  # Check whether existing region column is usable                                                                                                         
  bpue_betas %>% count(region, useNA = "always")
                                                                                                                                                           
  # If region looks clean, use it directly; otherwise assign from centroid:                                                                                
  # bpue_betas <- bpue_betas %>%
  #   mutate(                                                                                                                                              
  #     bbox_center_lon = (lon_min + lon_max) / 2,
  #     bbox_center_lat = (lat_min + lat_max) / 2,                                                                                                         
  #     region = map2_chr(bbox_center_lon, bbox_center_lat, assign_region)                                                                                 
  #   )                                                                                                                                                    
                                                                                                                                                           
  # Mean BPUE for studies used in analysis (has_effort_match = TRUE)                                                                                       
  bpue_used_summary <- bpue_betas %>%
    filter(has_effort_match) %>%                                                                                                                           
    group_by(fishery_type, fleet, plot_year_group, region) %>%
    summarise(                                                                                                                                             
      n_studies = n(),                                                                                                                                     
      mean_bpue = mean(bpue, na.rm = TRUE),                                                                                                                
      sd_bpue   = sd(bpue, na.rm = TRUE),                                                                                                                  
      .groups   = "drop"                                                                                                                                   
    )                                                                                                                                                      
                                                                                                                                                           
  # Overall means to compare against (fishery-level and global)                                                                                            
  bpue_ref <- bpue_betas %>%
    filter(has_effort_match) %>%                                                                                                                           
    summarise(    
      overall_mean_bpue = mean(bpue, na.rm = TRUE),
      overall_sd_bpue   = sd(bpue, na.rm = TRUE)                                                                                                           
    )
                                                                                                                                                           
  bpue_ref_by_fishery <- bpue_betas %>%                                                                                                                    
    filter(has_effort_match) %>%
    group_by(fishery_type) %>%                                                                                                                             
    summarise(                                                                                                                                             
      fishery_mean_bpue = mean(bpue, na.rm = TRUE),                                                                                                        
      fishery_sd_bpue   = sd(bpue, na.rm = TRUE),                                                                                                          
      .groups = "drop"                                                                                                                                     
    )                                                                                                                                                      
                                                                                                                                                           
  # Skipped studies with comparisons to overall and fishery means                                                                                          
  bpue_skipped_summary <- bpue_betas %>%
    filter(!has_effort_match) %>%                                                                                                                          
    left_join(bpue_ref_by_fishery, by = "fishery_type") %>%                                                                                                
    mutate(
      overall_mean  = bpue_ref$overall_mean_bpue,                                                                                                          
      diff_from_overall = bpue - overall_mean,                                                                                                             
      diff_from_fishery = bpue - fishery_mean_bpue,
      pct_diff_fishery  = (bpue - fishery_mean_bpue) / fishery_mean_bpue * 100                                                                             
    ) %>%                                                                                                                                                  
    select(fishery_type, fleet, plot_year_group, region, bpue,                                                                                             
           fishery_mean_bpue, diff_from_fishery, pct_diff_fishery,                                                                                         
           overall_mean, diff_from_overall) %>%
    mutate(across(where(is.numeric), ~signif(.x, 4)))                                                                                                      
                                                                                                                                                           
  print(bpue_used_summary, n = 100)                                                                                                                        
  print(bpue_skipped_summary, n = 100)                                                                                                                     
                                                                                                                                                           
  # Quick visual: skipped vs used BPUE distributions by fishery type                                                                                       
  bind_rows(                                                                                                                                               
    bpue_betas %>% filter(has_effort_match) %>% mutate(group = "used"),                                                                                    
    bpue_betas %>% filter(!has_effort_match) %>% mutate(group = "skipped")                                                                                 
  ) %>%
    ggplot(aes(x = bpue, fill = group)) +                                                                                                                  
    geom_histogram(alpha = 0.6, bins = 30, position = "identity") +                                                                                        
    facet_wrap(~fishery_type, scales = "free") +
    scale_fill_manual(values = c("used" = "steelblue", "skipped" = "firebrick")) +                                                                         
    labs(x = "BPUE", y = "Count", fill = NULL,
         title = "BPUE distribution: used vs skipped studies") +                                                                                           
    theme_bw()                                        
```

# Section 7: Build Catchability Surface

For each fleet × fishery × period, create a per-cell catchability map by overlaying
BPUE study bounding boxes. Where multiple studies overlap a cell, take the mean of
their betas. Cells with effort but no study coverage are marked as gaps (filled in
Section 8).

```{r build-catchability-maps}
# Get all fleet x fishery x period combos that have effort
fleet_fp_combos <- effort_combos %>%
  filter(paste(fishery, fleet, year_group, sep = "|") %in% names(effort_rasters))

# Attach betas to BPUE studies (need geometry for spatial matching)
bpue_with_betas <- bpue_expanded %>%
  left_join(
    bpue_betas %>% select(row_id, plot_year_group, beta, has_effort_match),
    by = c("row_id", "plot_year_group")
  ) %>%
  filter(!is.na(beta))

# Store results: one catchability raster per fleet x fishery x period
catchability_rasters <- list()
coverage_info <- tibble()

for (i in 1:nrow(fleet_fp_combos)) {
  fi <- fleet_fp_combos$fishery[i]
  fl <- fleet_fp_combos$fleet[i]
  yg <- fleet_fp_combos$year_group[i]
  key <- paste(fi, fl, yg, sep = "|")

  effort_r <- effort_rasters[[key]]

  # Get BPUE studies matching this fleet + fishery + period
  studies <- bpue_with_betas %>%
    filter(fishery_type == fi, fleet == fl, plot_year_group == yg)

  if (nrow(studies) == 0) {
    # No BPUE studies for this fleet x fishery x period -> all cells are gaps
    catchability_rasters[[key]] <- NA
    coverage_info <- bind_rows(coverage_info, tibble(
      fishery = fi, fleet = fl, year_group = yg,
      n_effort_cells = NA_integer_, n_covered = 0L,
      n_gap = NA_integer_, method = "all_gap"
    ))
    next
  }

  # Get cell coordinates from effort raster
  effort_df <- as.data.frame(effort_r, xy = TRUE, cells = TRUE, na.rm = TRUE)
  names(effort_df)[4] <- "effort"
  effort_df <- effort_df %>% filter(effort > 0)

  if (nrow(effort_df) == 0) next

  # Convert effort cells to sf points for spatial intersection with study bboxes
  effort_pts <- st_as_sf(effort_df, coords = c("x", "y"), crs = 4326)

  # Check which effort cells fall within each study's bounding box
  intersections <- st_intersects(effort_pts, studies)

  # For each cell, average betas of covering studies (weighted mean where overlap)
  effort_df$beta <- NA_real_
  effort_df$n_studies <- 0L

  for (j in seq_along(intersections)) {
    study_indices <- intersections[[j]]
    if (length(study_indices) > 0) {
      study_betas <- studies$beta[study_indices]
      effort_df$beta[j] <- mean(study_betas, na.rm = TRUE)
      effort_df$n_studies[j] <- length(study_indices)
    }
  }

  n_covered <- sum(!is.na(effort_df$beta))
  n_gap <- sum(is.na(effort_df$beta))

  # Create catchability raster
  catch_rast <- template * NA
  catch_rast[effort_df$cell] <- effort_df$beta

  catchability_rasters[[key]] <- catch_rast

  coverage_info <- bind_rows(coverage_info, tibble(
    fishery = fi, fleet = fl, year_group = yg,
    n_effort_cells = nrow(effort_df), n_covered = n_covered,
    n_gap = n_gap, method = "direct"
  ))
}

sum(!is.na(catchability_rasters))
coverage_info %>% print(n = 50)
```

# Section 8: Fill Gaps in Catchability Surface

Hierarchical gap-filling for cells with effort but no BPUE study coverage:

**Priority 1: Temporal extrapolation from same fleet**
If this fleet has β in another period (same fishery type), use the nearest
period's β scaled by the observed temporal trend in mean BPUE:

```
β_fill = β_nearest_period × (mean_BPUE_target_period / mean_BPUE_source_period)
```

This preserves fleet-specific fishing practices while adjusting for temporal
changes in bycatch rates (e.g., mitigation adoption over time).

**Priority 2: Regional mean β for the target period**
If the fleet has no β in any period, use the mean β from all studies in the
same region and fishery type during the target period.

**Priority 3: Fishery-type mean β (fallback)**

```{r fill-gaps}
# Pre-compute mean beta by similarity dimensions (from direct study betas)
valid_betas <- bpue_betas %>% filter(!is.na(beta))

# Assign each study to a region based on bounding box centroid
valid_betas <- valid_betas %>%
  mutate(
    bbox_center_lon = (lon_min + lon_max) / 2,
    bbox_center_lat = (lat_min + lat_max) / 2,
    study_region = map2_chr(bbox_center_lon, bbox_center_lat, assign_region)
  )

# --- Summary tables for fallback levels ---

# Mean beta by fleet x fishery x period (for temporal extrapolation)
beta_by_fleet_period <- valid_betas %>%
  group_by(fishery_type, fleet, plot_year_group) %>%
  summarise(mean_beta = mean(beta, na.rm = TRUE), n = n(), .groups = "drop")

# Mean beta by region x fishery x period (Priority 2)
beta_by_region_period <- valid_betas %>%
  filter(!is.na(study_region)) %>%
  group_by(fishery_type, study_region, plot_year_group) %>%
  summarise(mean_beta = mean(beta, na.rm = TRUE), n = n(), .groups = "drop")

# Mean beta by region x fishery (all periods pooled, for when period-specific not available)
beta_by_region <- valid_betas %>%
  filter(!is.na(study_region)) %>%
  group_by(fishery_type, study_region) %>%
  summarise(mean_beta = mean(beta, na.rm = TRUE), n = n(), .groups = "drop")

# Mean beta by fishery type (ultimate fallback)
beta_by_fishery_mean <- valid_betas %>%
  group_by(fishery_type) %>%
  summarise(mean_beta = mean(beta, na.rm = TRUE), n = n(), .groups = "drop")

# --- BPUE temporal scaling ratios ---
# From bpue_summary_by_period.csv: mean BPUE by fishery type x period
# Used to scale fleet betas across periods
bpue_period_means <- bpue_by_period %>%
  select(fishery_type, plot_year_group, mean_bpue)

# Define period ordering for "nearest" lookup
period_order <- c("1990-1994", "1995-1999", "2000-2004", "2005-2009")

# Gap-fill reference tables computed
tibble(
  fleet_period_betas   = nrow(beta_by_fleet_period),
  region_period_betas  = nrow(beta_by_region_period),
  region_betas_pooled  = nrow(beta_by_region),
  fishery_type_betas   = nrow(beta_by_fishery_mean)
)

# BPUE temporal means (for scaling)
bpue_period_means %>% print()

# Mean beta by fishery type (ultimate fallback)
beta_by_fishery_mean %>% print()

# Track gap-fill details
gap_fill_log <- tibble()

for (i in 1:nrow(fleet_fp_combos)) {
  fi <- fleet_fp_combos$fishery[i]
  fl <- fleet_fp_combos$fleet[i]
  yg <- fleet_fp_combos$year_group[i]
  key <- paste(fi, fl, yg, sep = "|")

  effort_r <- effort_rasters[[key]]

  # Get or create the catchability raster
  catch_rast <- catchability_rasters[[key]]
  if (is.logical(catch_rast) && is.na(catch_rast)) {
    catch_rast <- template * NA
  }

  # Find cells that have effort but no beta (gaps)
  effort_df <- as.data.frame(effort_r, xy = TRUE, cells = TRUE, na.rm = TRUE)
  names(effort_df)[4] <- "effort"
  effort_df <- effort_df %>% filter(effort > 0)

  if (nrow(effort_df) == 0) next

  # Get existing beta values
  existing_beta <- values(catch_rast)[effort_df$cell]
  gap_cells <- which(is.na(existing_beta))

  if (length(gap_cells) == 0) next

  # --- Priority 1: Temporal extrapolation from same fleet ---
  # Check if this fleet has beta in any other period (same fishery type)
  fleet_other_periods <- beta_by_fleet_period %>%
    filter(fishery_type == fi, fleet == fl, plot_year_group != yg, mean_beta > 0)

  if (nrow(fleet_other_periods) > 0) {
    # Find the nearest period with data
    target_idx <- match(yg, period_order)
    source_indices <- match(fleet_other_periods$plot_year_group, period_order)
    nearest_row <- fleet_other_periods[which.min(abs(source_indices - target_idx)), ]
    source_period <- nearest_row$plot_year_group
    source_beta <- nearest_row$mean_beta

    # Get BPUE means for temporal scaling
    bpue_target <- bpue_period_means %>%
      filter(fishery_type == fi, plot_year_group == yg) %>%
      pull(mean_bpue)
    bpue_source <- bpue_period_means %>%
      filter(fishery_type == fi, plot_year_group == source_period) %>%
      pull(mean_bpue)

    # Scale beta by BPUE temporal ratio
    if (length(bpue_target) > 0 && length(bpue_source) > 0 &&
        bpue_source > 0 && bpue_target > 0) {
      temporal_ratio <- bpue_target / bpue_source
      fill_beta <- source_beta * temporal_ratio
      fill_method <- paste0("fleet_temporal_", source_period)
    } else {
      # BPUE data missing for scaling — use unscaled fleet beta
      fill_beta <- source_beta
      fill_method <- paste0("fleet_unscaled_", source_period)
    }

    # Apply to all gap cells
    catch_rast[effort_df$cell[gap_cells]] <- fill_beta

    # Assign gap cells to regions for logging
    gap_df <- effort_df[gap_cells, ]
    gap_df$region <- map_chr(1:nrow(gap_df), function(j) {
      assign_region(gap_df$x[j], gap_df$y[j])
    })

    gap_fill_log <- bind_rows(gap_fill_log, tibble(
      fishery = fi, fleet = fl, year_group = yg,
      region = paste(unique(na.omit(gap_df$region)), collapse = ", "),
      n_cells = length(gap_cells),
      fill_method = fill_method,
      fill_beta = fill_beta,
      source_beta = source_beta,
      temporal_ratio = ifelse(exists("temporal_ratio"), temporal_ratio, NA_real_)
    ))

  } else {
    # --- Priority 2 & 3: No fleet data in any period ---
    # Assign each gap cell to a region, then fill by region or fishery mean
    gap_df <- effort_df[gap_cells, ]
    gap_df$region <- map_chr(1:nrow(gap_df), function(j) {
      assign_region(gap_df$x[j], gap_df$y[j])
    })

    # Fishery-type fallback beta
    fishery_beta_val <- beta_by_fishery_mean %>%
      filter(fishery_type == fi) %>%
      pull(mean_beta)

    # Process gap cells grouped by region
    gap_regions <- gap_df %>%
      filter(!is.na(region)) %>%
      distinct(region) %>%
      pull(region)

    for (reg in gap_regions) {
      reg_cells <- gap_df %>% filter(region == reg)

      # Try region x period beta first
      region_period_beta <- beta_by_region_period %>%
        filter(fishery_type == fi, study_region == reg, plot_year_group == yg) %>%
        pull(mean_beta)

      if (length(region_period_beta) > 0 && !is.na(region_period_beta[1])) {
        fill_beta <- region_period_beta[1]
        fill_method <- "region_period"
      } else {
        # Try region beta (all periods pooled)
        region_beta_val <- beta_by_region %>%
          filter(fishery_type == fi, study_region == reg) %>%
          pull(mean_beta)

        if (length(region_beta_val) > 0 && !is.na(region_beta_val[1])) {
          fill_beta <- region_beta_val[1]
          fill_method <- "region_pooled"
        } else {
          fill_beta <- fishery_beta_val[1]
          fill_method <- "fishery_mean"
        }
      }

      catch_rast[reg_cells$cell] <- fill_beta

      gap_fill_log <- bind_rows(gap_fill_log, tibble(
        fishery = fi, fleet = fl, year_group = yg,
        region = reg,
        n_cells = nrow(reg_cells),
        fill_method = fill_method,
        fill_beta = fill_beta,
        source_beta = NA_real_,
        temporal_ratio = NA_real_
      ))
    }

    # Handle gap cells with no region assignment
    no_region_cells <- gap_df %>% filter(is.na(region))
    if (nrow(no_region_cells) > 0) {
      catch_rast[no_region_cells$cell] <- fishery_beta_val[1]

      gap_fill_log <- bind_rows(gap_fill_log, tibble(
        fishery = fi, fleet = fl, year_group = yg,
        region = "no_region",
        n_cells = nrow(no_region_cells),
        fill_method = "fishery_mean",
        fill_beta = fishery_beta_val[1],
        source_beta = NA_real_,
        temporal_ratio = NA_real_
      ))
    }
  }

  # Update the catchability raster
  catchability_rasters[[key]] <- catch_rast
}

gap_fill_log %>%
  group_by(fishery, fill_method) %>%
  summarise(
    n_fleet_periods = n(),
    total_cells = sum(n_cells),
    mean_fill_beta = format(mean(fill_beta, na.rm = TRUE), scientific = TRUE),
    .groups = "drop"
  ) %>%
  print(n = 30)

# Show temporal scaling details where applied
temporal_fills <- gap_fill_log %>% filter(str_detect(fill_method, "fleet_temporal"))
if (nrow(temporal_fills) > 0) {
  temporal_fills %>%
    select(fishery, fleet, year_group, fill_method, source_beta, temporal_ratio, fill_beta) %>%
    mutate(across(where(is.numeric), ~signif(.x, 4))) %>%
    print(n = 30)
}
```

# Section 9: Calculate Bycatch

For each fleet × fishery × period, apply the per-cell catchability surface:

`bycatch_cell = β_cell × hooks_cell × BI_birds_cell`

Key: β was calibrated using GLOBAL WAAL density (all populations), but bycatch
is calculated using BI-specific bird density. This means global bird density
cancels in the total WAAL estimate (as expected: total = BPUE × hooks), but
BI's share depends on the actual spatial overlap of hooks and BI birds within
each study area. Areas where BI birds are a larger fraction of total WAAL
contribute proportionally more to BI bycatch.

```{r calculate-bycatch}
# Using per-cell catchability surface × hooks × BI bird density
# β calibrated with global WAAL, applied with BI-specific density

# Calculate bycatch for each fleet x fishery x period
bycatch_results <- tibble()

for (i in 1:nrow(fleet_fp_combos)) {
  fi <- fleet_fp_combos$fishery[i]
  fl <- fleet_fp_combos$fleet[i]
  yg <- fleet_fp_combos$year_group[i]
  key <- paste(fi, fl, yg, sep = "|")

  effort_r <- effort_rasters[[key]]
  catch_r <- catchability_rasters[[key]]

  # Skip if catchability is still NA (shouldn't happen after gap-filling)
  if (is.logical(catch_r) && is.na(catch_r)) next

  # BI bycatch per cell per age class = beta_cell * hooks_cell * BI_birds_age_cell
  byc_by_age <- map_dbl(names(bi_dists), function(age) {
    bi_age_byc <- catch_r * effort_r * bi_dists[[age]]
    global(bi_age_byc, "sum", na.rm = TRUE)[1, 1]
  })
  names(byc_by_age) <- names(bi_dists)

  bi_total <- sum(byc_by_age, na.rm = TRUE)

  bycatch_results <- bind_rows(bycatch_results, tibble(
    fishery = fi,
    fleet = fl,
    year_group = yg,
    bi_total = bi_total,
    fb = byc_by_age["fb"],
    sb = byc_by_age["sb"],
    nb = byc_by_age["nb"],
    j1 = byc_by_age["j1"],
    j2j3 = byc_by_age["j2j3"],
    imm = byc_by_age["imm"]
  ))
}

# Bycatch by fleet x fishery x period
bycatch_results %>%
  mutate(across(where(is.numeric), ~round(.x, 4))) %>%
  print(n = 100)
```

# Section 10: Summary Tables

## Annual bycatch by fishery x period

Each value is the estimated annual BI bycatch for that fishery during that period
(effort rasters use mean annual hooks).

```{r summary-by-fishery-period}
byc_by_fp <- bycatch_results %>%
  group_by(fishery, year_group) %>%
  summarise(
    n_fleets = n(),
    bi_total = sum(bi_total, na.rm = TRUE),
    fb = sum(fb, na.rm = TRUE),
    sb = sum(sb, na.rm = TRUE),
    nb = sum(nb, na.rm = TRUE),
    j1 = sum(j1, na.rm = TRUE),
    j2j3 = sum(j2j3, na.rm = TRUE),
    imm = sum(imm, na.rm = TRUE),
    .groups = "drop"
  )

# Annual BI bycatch by fishery x period
byc_by_fp %>%
  mutate(across(where(is.numeric) & !matches("n_fleets"), ~round(.x, 2))) %>%
  print(n = 20)
```

## Total annual bycatch by period (PLL + DLL combined)

```{r summary-by-period}
byc_by_period <- bycatch_results %>%
  group_by(year_group) %>%
  summarise(
    bi_total = sum(bi_total, na.rm = TRUE),
    fb = sum(fb, na.rm = TRUE),
    sb = sum(sb, na.rm = TRUE),
    nb = sum(nb, na.rm = TRUE),
    j1 = sum(j1, na.rm = TRUE),
    j2j3 = sum(j2j3, na.rm = TRUE),
    imm = sum(imm, na.rm = TRUE),
    .groups = "drop"
  )

# Total annual BI bycatch by period (PLL + DLL)
byc_by_period %>%
  mutate(across(where(is.numeric), ~round(.x, 2))) %>%
  print()

# Grand total: sum of (annual bycatch * years in period) across all periods
# All periods span 5 years
byc_cumulative <- byc_by_period %>%
  mutate(
    n_years = 5,
    period_total = bi_total * n_years
  )

# Cumulative BI bycatch over full study period (1990-2009): annual rate * years in each period
byc_cumulative %>%
  mutate(across(where(is.numeric), ~round(.x, 2))) %>%
  select(year_group, bi_total, n_years, period_total) %>%
  print()
tibble(
  total_1990_2010 = round(sum(byc_cumulative$period_total), 1),
  avg_annual_rate = round(mean(byc_by_period$bi_total), 1)
)
```

## Diagnostic: where is the bycatch coming from?

```{r diagnostic-breakdown}
# 1) Total mean annual hooks per fishery x period (summed across all fleets)
hooks_by_fp <- effort_combos %>%
  filter(paste(fishery, fleet, year_group, sep = "|") %in% names(effort_rasters)) %>%
  rowwise() %>%
  mutate(
    total_hooks = {
      r <- effort_rasters[[paste(fishery, fleet, year_group, sep = "|")]]
      if (!is.null(r)) global(r, "sum", na.rm = TRUE)[1, 1] else 0
    }
  ) %>%
  ungroup()

# Mean annual hooks by fleet x fishery x period
hooks_by_fp %>%
  arrange(fishery, desc(total_hooks)) %>%
  print(n = 60)

# Total mean annual hooks by fishery x period
hooks_by_fp %>%
  group_by(fishery, year_group) %>%
  summarise(total_hooks = sum(total_hooks), n_fleets = n(), .groups = "drop") %>%
  print(n = 20)

# 2) Bycatch breakdown by fleet
# Bycatch per fleet (top contributors)
bycatch_results %>%
  arrange(fishery, desc(bi_total)) %>%
  mutate(bi_total = round(bi_total, 4)) %>%
  select(fishery, fleet, year_group, bi_total) %>%
  print(n = 40)

# 3) Naive check: if we just applied mean BPUE to total hooks, what would we get?
#    This tells us whether the spatial method is inflating beyond what the rates imply
# Naive BPUE x hooks check
naive_check <- hooks_by_fp %>%
  group_by(fishery, year_group) %>%
  summarise(total_hooks = sum(total_hooks), .groups = "drop") %>%
  left_join(
    bpue_by_fishery %>% select(fishery_type, mean_bpue) %>% rename(fishery = fishery_type),
    by = "fishery"
  ) %>%
  mutate(
    naive_global_byc = mean_bpue / 1000 * total_hooks,
    # Compare to actual calculated bycatch
  ) %>%
  left_join(
    byc_by_fp %>% select(fishery, year_group, bi_total),
    by = c("fishery", "year_group")
  ) %>%
  mutate(
    ratio_actual_vs_naive = round(bi_total / naive_global_byc, 2)
  )

naive_check %>%
  mutate(across(where(is.numeric), ~signif(.x, 4))) %>%
  print(n = 20)

# With bird-density-corrected catchability, ratio should be < 1
# (ratio < 1 means catchability correctly accounts for hooks in low-bird areas)
# (ratio ~ 1 would mean hooks and birds perfectly overlap everywhere)

# 4) How many fleets have direct BPUE data vs are entirely gap-filled?
fleet_bpue_status <- hooks_by_fp %>%
  mutate(
    has_direct_bpue = paste(fishery, fleet, year_group, sep = "|") %in%
      paste(bpue_betas$fishery_type[!is.na(bpue_betas$beta)],
            bpue_betas$fleet[!is.na(bpue_betas$beta)],
            bpue_betas$plot_year_group[!is.na(bpue_betas$beta)], sep = "|")
  )

fleet_bpue_status %>%
  group_by(fishery, has_direct_bpue) %>%
  summarise(
    n_fleet_periods = n(),
    total_hooks = sum(total_hooks),
    .groups = "drop"
  ) %>%
  print()
```

## Mean annual bycatch by fishery (averaged across periods)

```{r summary-by-fishery}
byc_by_fishery <- byc_by_fp %>%
  group_by(fishery) %>%
  summarise(
    n_periods = n(),
    bi_annual_mean = mean(bi_total, na.rm = TRUE),
    fb = mean(fb, na.rm = TRUE),
    sb = mean(sb, na.rm = TRUE),
    nb = mean(nb, na.rm = TRUE),
    j1 = mean(j1, na.rm = TRUE),
    j2j3 = mean(j2j3, na.rm = TRUE),
    imm = mean(imm, na.rm = TRUE),
    .groups = "drop"
  )

# Mean annual BI bycatch by fishery (averaged across periods)
byc_by_fishery %>%
  mutate(across(where(is.numeric) & !matches("n_periods"), ~round(.x, 2))) %>%
  print()

round(sum(byc_by_fishery$bi_annual_mean), 2)
```

## Per capita rates

```{r per-capita}
# Average annual bycatch across all periods
n_periods <- length(target_year_groups)

byc_annual_avg <- bycatch_results %>%
  group_by(fishery) %>%
  summarise(
    across(c(fb, sb, nb, j1, j2j3, imm), ~sum(.x, na.rm = TRUE) / n_periods),
    .groups = "drop"
  )

percap_rates <- byc_annual_avg %>%
  pivot_longer(cols = c(fb, sb, nb, j1, j2j3, imm), names_to = "age_class", values_to = "annual_byc") %>%
  mutate(
    population = pop_sizes[age_class],
    percap = annual_byc / population
  )

# Per capita rates (annual average)
percap_rates %>%
  mutate(across(where(is.numeric), ~signif(.x, 4))) %>%
  print(n = 20)
```

## Per capita rates by period

Per-capita fishing mortality for each age class × 5-year period (PLL + DLL combined).
Since effort rasters already represent mean annual hooks, these are annual mortality rates.

```{r percap-by-period}
# Per-capita annual bycatch rate for each age class × period
# bycatch_results values are already mean annual (effort = mean annual hooks per period)
percap_by_period_raw <- bycatch_results %>%
  group_by(fishery, year_group) %>%
  summarise(
    across(c(fb, sb, nb, j1, j2j3, imm), ~sum(.x, na.rm = TRUE)),
    .groups = "drop"
  ) %>%
  pivot_longer(cols = c(fb, sb, nb, j1, j2j3, imm),
               names_to = "age_class", values_to = "annual_byc") %>%
  mutate(
    population = pop_sizes[age_class],
    percap     = annual_byc / population
  )

# Combine PLL + DLL: separate columns for each fishery plus a combined total
percap_by_period_table <- percap_by_period_raw %>%
  group_by(year_group, age_class) %>%
  summarise(
    pll_percap   = sum(percap[fishery == "PLL"], na.rm = TRUE),
    dll_percap   = sum(percap[fishery == "DLL"], na.rm = TRUE),
    total_percap = sum(percap, na.rm = TRUE),
    .groups = "drop"
  )

# Per capita bycatch rates by period (PLL + DLL combined): mean annual per capita fishing mortality for each age class x period
percap_by_period_table %>%
  mutate(across(where(is.numeric), ~signif(.x, 4))) %>%
  arrange(age_class, year_group) %>%
  print(n = 30)

# Save
write_csv(percap_by_period_table,
          here(paste0("output/percap_rates_by_period_04_v3_", date_sfx, ".csv")))
```

```{r percap-by-period-plot, fig.width = 10, fig.height = 6}
# Ordered age-class labels for legend
age_order  <- c("j1", "j2j3", "imm", "fb", "sb", "nb")
age_labels <- c(
  j1   = "Juveniles (J1)",
  j2j3 = "Juveniles (J2-J3)",
  imm  = "Immatures",
  fb   = "Failed breeders",
  sb   = "Successful breeders",
  nb   = "Non-breeders"
)

# Period midpoints for continuous x-axis
period_midpoints <- c(
  "1990-1994" = 1992.0,
  "1995-1999" = 1997.0,
  "2000-2004" = 2002.0,
  "2005-2009" = 2007.0
)

plot_percap_time <- percap_by_period_table %>%
  mutate(
    year_mid  = period_midpoints[year_group],
    age_label = factor(age_labels[age_class], levels = age_labels[age_order])
  )

dir.create(here("output/maps"), showWarnings = FALSE, recursive = TRUE)

p_percap_period <- ggplot(plot_percap_time,
       aes(x = year_mid, y = total_percap, color = age_label, group = age_label)) +
  geom_line(linewidth = 0.9) +
  geom_point(size = 2.5) +
  scale_x_continuous(
    breaks       = unname(period_midpoints),
    labels       = names(period_midpoints),
    minor_breaks = NULL
  ) +
  scale_color_brewer(palette = "Set1", name = "Age class") +
  labs(
    title    = "Per capita bycatch rate over time by age class",
    subtitle = "Bird Island wandering albatross — combined PLL + DLL fishing mortality",
    x        = "Period",
    y        = expression("Per capita annual fishing mortality" ~~ (yr^{-1}))
  ) +
  theme_minimal(base_size = 13) +
  theme(
    legend.position  = "right",
    axis.text.x      = element_text(angle = 30, hjust = 1),
    panel.grid.minor = element_blank()
  )
print(p_percap_period)
ggsave(here(paste0("output/maps/percap_by_period_04_v3_", date_sfx, ".png")),
       plot = p_percap_period, width = 10, height = 6, dpi = 300)
```

## Per capita rates by year

Annual per-capita fishing mortality for each age class, computed by building
year-specific effort rasters and applying each period's β catchability surface.

**Interpretation**: Annual variation reflects changes in fishing effort intensity
(hooks set per year). β (per-hook mortality risk) is fixed at the 5-year period
level since BPUE literature data are not resolved to individual years. Bird
distributions are also static (Clay et al. 1990–2009 average).

```{r percap-by-year, message = FALSE}
# NOTE: builds annual effort rasters for every fleet × year — may take several minutes

# Helper: map a calendar year to its 5-year period (for β lookup)
year_to_period <- function(year) {
  dplyr::case_when(
    year >= 1990 & year <= 1994 ~ "1990-1994",
    year >= 1995 & year <= 1999 ~ "1995-1999",
    year >= 2000 & year <= 2004 ~ "2000-2004",
    year >= 2005 & year <= 2009 ~ "2005-2009",
    TRUE ~ NA_character_
  )
}

# Helper: total-hooks raster for one fleet × fishery × calendar year
make_annual_effort_raster <- function(effort_data, fleet_name, year_val, template_rast) {
  effort_subset <- effort_data %>%
    filter(fleet == fleet_name, Year == year_val, !is.na(Lon), !is.na(Lat)) %>%
    group_by(Lon, Lat) %>%
    summarise(total_hooks = sum(Hooks, na.rm = TRUE), .groups = "drop")

  if (nrow(effort_subset) < 2) return(NULL)

  effort_rast <- tryCatch(
    rast(effort_subset, type = "xyz", crs = "EPSG:4326"),
    error = function(e) NULL
  )
  if (is.null(effort_rast)) return(NULL)

  resample(effort_rast, template_rast, method = "bilinear")
}

# All years present in either dataset
all_years <- sort(unique(c(pll_data$Year, dll_data$Year)))

annual_bycatch_results <- tibble()

for (yr in all_years) {
  period <- year_to_period(yr)
  if (is.na(period)) next

  for (fi in c("PLL", "DLL")) {
    src_data <- if (fi == "PLL") pll_data else dll_data

    fleets_in_year <- src_data %>%
      filter(Year == yr) %>%
      distinct(fleet) %>%
      pull(fleet)

    for (fl in fleets_in_year) {
      catch_key <- paste(fi, fl, period, sep = "|")
      catch_r   <- catchability_rasters[[catch_key]]

      # Skip if no catchability surface for this fleet × period
      if (is.null(catch_r) || (is.logical(catch_r) && is.na(catch_r))) next

      effort_r <- make_annual_effort_raster(src_data, fl, yr, template)
      if (is.null(effort_r)) next

      byc_by_age <- map_dbl(names(bi_dists), function(age) {
        global(catch_r * effort_r * bi_dists[[age]], "sum", na.rm = TRUE)[1, 1]
      })
      names(byc_by_age) <- names(bi_dists)

      annual_bycatch_results <- bind_rows(annual_bycatch_results, tibble(
        year     = yr,
        period   = period,
        fishery  = fi,
        fleet    = fl,
        bi_total = sum(byc_by_age, na.rm = TRUE),
        fb       = byc_by_age["fb"],
        sb       = byc_by_age["sb"],
        nb       = byc_by_age["nb"],
        j1       = byc_by_age["j1"],
        j2j3     = byc_by_age["j2j3"],
        imm      = byc_by_age["imm"]
      ))
    }
  }
}

# Sum across all fleets and fisheries, then compute per-capita by age class
annual_percap_by_age <- annual_bycatch_results %>%
  group_by(year) %>%
  summarise(
    across(c(fb, sb, nb, j1, j2j3, imm), ~sum(.x, na.rm = TRUE)),
    .groups = "drop"
  ) %>%
  pivot_longer(cols = c(fb, sb, nb, j1, j2j3, imm),
               names_to = "age_class", values_to = "annual_byc") %>%
  mutate(
    population = pop_sizes[age_class],
    percap     = annual_byc / population,
    age_label  = factor(age_labels[age_class], levels = age_labels[age_order])
  )

# Per capita bycatch rates by year (PLL + DLL combined)
annual_percap_by_age %>%
  select(year, age_class, annual_byc, percap) %>%
  mutate(across(where(is.numeric), ~signif(.x, 4))) %>%
  arrange(age_class, year) %>%
  print(n = 120)

write_csv(
  annual_percap_by_age %>% select(year, age_class, population, annual_byc, percap),
  here(paste0("output/percap_rates_by_year_04_v3_", date_sfx, ".csv"))
)
```

```{r percap-by-year-plot, fig.width = 11, fig.height = 6}
p_percap_year <- ggplot(annual_percap_by_age,
       aes(x = year, y = percap, color = age_label, group = age_label)) +
  geom_line(linewidth = 0.8, alpha = 0.9) +
  geom_point(size = 1.8) +
  scale_x_continuous(
    breaks       = seq(1990, 2010, by = 5),
    minor_breaks = seq(1990, 2010, by = 1)
  ) +
  scale_color_brewer(palette = "Set1", name = "Age class") +
  labs(
    title    = "Per capita bycatch rate by year and age class",
    subtitle = "Bird Island WAAL — PLL + DLL combined\n\u03b2 (catchability) fixed at period level; annual variation reflects effort changes only",
    x        = "Year",
    y        = expression("Per capita annual fishing mortality" ~~ (yr^{-1}))
  ) +
  theme_minimal(base_size = 13) +
  theme(
    legend.position  = "right",
    panel.grid.minor = element_line(color = "grey92")
  )
print(p_percap_year)
ggsave(here(paste0("output/maps/percap_by_year_04_v3_", date_sfx, ".png")),
       plot = p_percap_year, width = 11, height = 6, dpi = 300)
```

## Gap-fill summary

```{r gap-fill-summary}
# Coverage by fishery
gap_fill_log %>%
  count(fishery, fill_method, name = "n_entries") %>%
  print(n = 20)

# Overall: direct vs filled
coverage_overall <- coverage_info %>%
  group_by(fishery) %>%
  summarise(
    total_effort_cells = sum(n_effort_cells, na.rm = TRUE),
    direct_covered = sum(n_covered, na.rm = TRUE),
    gaps = sum(n_gap, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(
    pct_direct = round(100 * direct_covered / total_effort_cells, 1),
    pct_filled = round(100 * gaps / total_effort_cells, 1)
  )

# Direct vs gap-filled coverage
coverage_overall %>% print()
```

# Section 11: Visualization

## Catchability maps by fishery x period

```{r map-catchability, fig.width=14, fig.height=12}
map_extent <- c(-180, 180, -90, 0)

# Effort-weighted mean catchability across fleets for visualization
for (fi in c("PLL", "DLL")) {
  catch_maps <- list()

  for (yg in target_year_groups) {
    num_rast <- template * 0
    den_rast <- template * 0

    matching_keys <- names(catchability_rasters)[
      str_detect(names(catchability_rasters), paste0("^", fi, "\\|.*\\|", yg, "$"))
    ]

    for (k in matching_keys) {
      cr <- catchability_rasters[[k]]
      er <- effort_rasters[[k]]
      if (!is.logical(cr)) {
        beta_x_effort <- cr * er
        beta_x_effort[is.na(beta_x_effort)] <- 0
        er_clean <- er
        er_clean[is.na(er_clean)] <- 0
        num_rast <- num_rast + beta_x_effort
        den_rast <- den_rast + er_clean
      }
    }

    mean_catch <- num_rast / den_rast
    mean_catch[den_rast == 0] <- NA

    catch_df <- as.data.frame(mean_catch, xy = TRUE, na.rm = TRUE)
    names(catch_df)[3] <- "catchability"

    p <- ggplot(catch_df, aes(x = x, y = y, fill = catchability)) +
      geom_raster() +
      geom_sf(data = ne_land, inherit.aes = FALSE, fill = "grey80", color = NA) +
      coord_sf(xlim = map_extent[1:2], ylim = map_extent[3:4], expand = FALSE) +
      scale_fill_viridis_c(name = expression(beta), na.value = "white", option = "magma") +
      labs(title = paste(fi, yg)) +
      theme_minimal() +
      theme(plot.title = element_text(size = 11, face = "bold"), legend.position = "right")

    catch_maps[[yg]] <- p
  }

  combined <- wrap_plots(catch_maps, ncol = 2) +
    plot_annotation(title = paste(fi, "- Effort-Weighted Mean Catchability by Period"))
  print(combined)
  ggsave(here(paste0("output/maps/catchability_", tolower(fi), "_04_v3_", date_sfx, ".png")),
         plot = combined, width = 14, height = 12, dpi = 300)
}
```

## Bycatch density maps

```{r map-bycatch, fig.width=14, fig.height=12}
for (fi in c("PLL", "DLL")) {
  byc_maps <- list()

  for (yg in target_year_groups) {
    # Sum bycatch across all fleets for this fishery x period
    byc_total <- template * 0

    matching_keys <- names(catchability_rasters)[
      str_detect(names(catchability_rasters), paste0("^", fi, "\\|.*\\|", yg, "$"))
    ]

    for (k in matching_keys) {
      cr <- catchability_rasters[[k]]
      er <- effort_rasters[[k]]
      if (!is.logical(cr)) {
        # β_cell × hooks_cell × BI_birds_cell
        cell_byc <- cr * er * bi_density_aligned
        cell_byc[is.na(cell_byc)] <- 0
        byc_total <- byc_total + cell_byc
      }
    }

    byc_total[byc_total == 0] <- NA

    byc_df <- as.data.frame(byc_total, xy = TRUE, na.rm = TRUE)
    names(byc_df)[3] <- "bycatch"

    p <- ggplot(byc_df, aes(x = x, y = y, fill = bycatch)) +
      geom_raster() +
      geom_sf(data = ne_land, inherit.aes = FALSE, fill = "grey80", color = NA) +
      coord_sf(xlim = map_extent[1:2], ylim = map_extent[3:4], expand = FALSE) +
      scale_fill_viridis_c(name = "BI Bycatch", na.value = "white", direction = -1) +
      labs(title = paste(fi, yg)) +
      theme_minimal() +
      theme(plot.title = element_text(size = 11, face = "bold"), legend.position = "right")

    byc_maps[[yg]] <- p
  }

  combined <- wrap_plots(byc_maps, ncol = 2) +
    plot_annotation(title = paste(fi, "- Bird Island Bycatch Density by Period"))
  print(combined)
  ggsave(here(paste0("output/maps/bycatch_density_", tolower(fi), "_04_v3_", date_sfx, ".png")),
         plot = combined, width = 14, height = 12, dpi = 300)
}
```

## Bycatch density maps by age class

Mean annual BI bycatch density per age class, averaged across all four 5-year periods
(1990–2010). Each panel uses an independent color scale so spatial patterns are visible
even for age classes with low absolute bycatch. Shown separately per fishery, then
combined PLL + DLL.

```{r map-bycatch-by-age, fig.width = 16, fig.height = 10}
age_order_map  <- c("j1", "j2j3", "imm", "fb", "sb", "nb")
age_labels_map <- c(
  j1   = "Juveniles (J1)",
  j2j3 = "Juveniles (J2-J3)",
  imm  = "Immatures",
  fb   = "Failed breeders",
  sb   = "Successful breeders",
  nb   = "Non-breeders"
)

for (fi in c("PLL", "DLL")) {
  age_maps <- list()

  for (age in age_order_map) {
    # Accumulate mean-annual bycatch across all fleet × period combos for this fishery × age class
    byc_age_sum <- template * 0

    for (yg in target_year_groups) {
      matching_keys <- names(catchability_rasters)[
        str_detect(names(catchability_rasters), paste0("^", fi, "\\|.*\\|", yg, "$"))
      ]
      for (k in matching_keys) {
        cr <- catchability_rasters[[k]]
        er <- effort_rasters[[k]]
        if (!is.logical(cr)) {
          cell_byc <- cr * er * bi_dists[[age]]
          cell_byc[is.na(cell_byc)] <- 0
          byc_age_sum <- byc_age_sum + cell_byc
        }
      }
    }

    # Divide by n periods → mean annual bycatch density (effort rasters already mean annual)
    byc_age_mean <- byc_age_sum / length(target_year_groups)
    byc_age_mean[byc_age_mean == 0] <- NA

    byc_df <- as.data.frame(byc_age_mean, xy = TRUE, na.rm = TRUE)
    names(byc_df)[3] <- "bycatch"

    p <- ggplot(byc_df, aes(x = x, y = y, fill = bycatch)) +
      geom_raster() +
      geom_sf(data = ne_land, inherit.aes = FALSE, fill = "grey80", color = NA) +
      coord_sf(xlim = map_extent[1:2], ylim = map_extent[3:4], expand = FALSE) +
      scale_fill_viridis_c(name = "Birds/yr", na.value = "white", direction = -1) +
      labs(title = age_labels_map[age]) +
      theme_minimal() +
      theme(plot.title = element_text(size = 11, face = "bold"), legend.position = "right")

    age_maps[[age]] <- p
  }

  combined_age <- wrap_plots(age_maps[age_order_map], ncol = 3) +
    plot_annotation(
      title    = paste(fi, "\u2014 BI Bycatch Density by Age Class"),
      subtitle = "Mean annual bycatch (1990\u20132010), independent color scales per age class"
    )
  print(combined_age)
  ggsave(here(paste0("output/maps/bycatch_by_age_", tolower(fi), "_04_v3_", date_sfx, ".png")),
         plot = combined_age, width = 16, height = 10, dpi = 300)
}
```

```{r map-bycatch-by-age-combined, fig.width = 16, fig.height = 10}
# PLL + DLL combined bycatch density by age class
age_maps_comb <- list()

for (age in age_order_map) {
  byc_age_sum <- template * 0

  for (fi in c("PLL", "DLL")) {
    for (yg in target_year_groups) {
      matching_keys <- names(catchability_rasters)[
        str_detect(names(catchability_rasters), paste0("^", fi, "\\|.*\\|", yg, "$"))
      ]
      for (k in matching_keys) {
        cr <- catchability_rasters[[k]]
        er <- effort_rasters[[k]]
        if (!is.logical(cr)) {
          cell_byc <- cr * er * bi_dists[[age]]
          cell_byc[is.na(cell_byc)] <- 0
          byc_age_sum <- byc_age_sum + cell_byc
        }
      }
    }
  }

  byc_age_mean <- byc_age_sum / length(target_year_groups)
  byc_age_mean[byc_age_mean == 0] <- NA

  byc_df <- as.data.frame(byc_age_mean, xy = TRUE, na.rm = TRUE)
  names(byc_df)[3] <- "bycatch"

  p <- ggplot(byc_df, aes(x = x, y = y, fill = bycatch)) +
    geom_raster() +
    geom_sf(data = ne_land, inherit.aes = FALSE, fill = "grey80", color = NA) +
    coord_sf(xlim = map_extent[1:2], ylim = map_extent[3:4], expand = FALSE) +
    scale_fill_viridis_c(name = "Birds/yr", na.value = "white", direction = -1) +
    labs(title = age_labels_map[age]) +
    theme_minimal() +
    theme(plot.title = element_text(size = 11, face = "bold"), legend.position = "right")

  age_maps_comb[[age]] <- p
}

combined_age_all <- wrap_plots(age_maps_comb[age_order_map], ncol = 3) +
  plot_annotation(
    title    = "PLL + DLL \u2014 BI Bycatch Density by Age Class",
    subtitle = "Mean annual bycatch (1990\u20132010), independent color scales per age class"
  )
print(combined_age_all)
ggsave(here(paste0("output/maps/bycatch_by_age_combined_04_v3_", date_sfx, ".png")),
       plot = combined_age_all, width = 16, height = 10, dpi = 300)
```

# Section 12: Save Outputs

```{r save-outputs}
dir.create(here("output/rasters"), showWarnings = FALSE, recursive = TRUE)
dir.create(here("output/maps"), showWarnings = FALSE, recursive = TRUE)

# Save bycatch results (fleet-level detail)
write_csv(bycatch_results, here(paste0("output/bycatch_by_fleet_period_04_v3_", date_sfx, ".csv")))

# Save fishery x period summaries
write_csv(byc_by_fp, here(paste0("output/bycatch_by_fishery_period_04_v3_", date_sfx, ".csv")))

# Save per capita rates
write_csv(percap_rates, here(paste0("output/bycatch_percapita_04_v3_", date_sfx, ".csv")))

# Save gap-fill log
write_csv(gap_fill_log, here(paste0("output/gap_fill_log_04_v3_", date_sfx, ".csv")))

# Save coverage info
write_csv(coverage_info, here(paste0("output/coverage_info_04_v3_", date_sfx, ".csv")))

# Save beta values for each study
write_csv(
  bpue_betas %>% select(row_id, fishery_type, fleet, plot_year_group, bpue,
                         beta, B_study, fleet_hooks_in_box, has_effort_match),
  here(paste0("output/bpue_study_betas_04_v3_", date_sfx, ".csv"))
)

# Save fishery-level beta summary
write_csv(beta_by_fishery_type, here(paste0("output/beta_by_fishery_type_04_v3_", date_sfx, ".csv")))

# Save catchability rasters per fishery x period (effort-weighted mean across fleets)
for (fi in c("PLL", "DLL")) {
  for (yg in target_year_groups) {
    num_rast <- template * 0
    den_rast <- template * 0

    matching_keys <- names(catchability_rasters)[
      str_detect(names(catchability_rasters), paste0("^", fi, "\\|.*\\|", yg, "$"))
    ]

    for (k in matching_keys) {
      cr <- catchability_rasters[[k]]
      er <- effort_rasters[[k]]
      if (!is.logical(cr)) {
        beta_x_effort <- cr * er
        beta_x_effort[is.na(beta_x_effort)] <- 0
        er_clean <- er
        er_clean[is.na(er_clean)] <- 0
        num_rast <- num_rast + beta_x_effort
        den_rast <- den_rast + er_clean
      }
    }

    mean_catch <- num_rast / den_rast
    mean_catch[den_rast == 0] <- NA

    yg_clean <- str_replace(yg, "-", "_")
    writeRaster(mean_catch,
                here(paste0("output/rasters/catchability_", tolower(fi), "_", yg_clean, "_04_v3_", date_sfx, ".tif")),
                overwrite = TRUE)
  }
}

```

# Diagnostics

```{r diagnostics}
tibble(
  effort_rasters_built      = length(effort_rasters),
  catchability_rasters_built = sum(!sapply(catchability_rasters, function(x) is.logical(x) && is.na(x)))
)

# Total BI bycatch by fishery (mean annual across periods)
byc_by_fishery %>%
  mutate(across(where(is.numeric), ~round(.x, 4))) %>%
  print()

# Per-study beta value ranges
bpue_betas %>%
  filter(!is.na(beta)) %>%
  group_by(fishery_type) %>%
  summarise(
    n = n(),
    min_beta = format(min(beta), scientific = TRUE),
    mean_beta = format(mean(beta), scientific = TRUE),
    median_beta = format(median(beta), scientific = TRUE),
    max_beta = format(max(beta), scientific = TRUE),
    .groups = "drop"
  ) %>%
  print()

# Comparison: spatial method (β × hooks × BI_birds) vs naive (BPUE × hooks)
# The ratio tells us how much the spatial method adjusts for BI's actual distribution
# Ratio < 1 means BI birds are concentrated in lower-effort areas relative to total WAAL
# Spatial: actual BI bycatch from catchability surface
# Naive: mean_BPUE/1000 × total_hooks (assumes BI fraction is uniform)
naive_check %>%
  mutate(across(where(is.numeric), ~signif(.x, 4))) %>%
  print(n = 20)

# Ratio < 1 → BI birds concentrated in areas with relatively less fishing effort
# Ratio ~ 1 → BI bird distribution aligns with total WAAL distribution
```

# Section 13: Demographic Validation — Implied Natural Mortality Check

Compare 03f per-capita fishing mortality against known survival rates from the
demographic model (WAAL_bycatch_5-27-25.Rmd). If fishing mortality exceeds total
mortality (1 - survival), natural mortality would be negative — an impossibility
that flags overestimated bycatch.

Mapping from 03f age classes to demographic model survival rates:

- **j2j3** → s_Juv = 0.846 (J2, J3 stages)
- **imm** → s_Imm = 0.921 (Imm4, Imm5, PB stages)
- **fb** → breeding adults: s_EF = 0.913, s_IF = 0.920 (experienced/inexperienced failed)
- **sb** → successful breeders: s_ES = 0.895, s_IS = 0.892
- **nb** → non-breeders: s_ENB = 0.943, s_PF = 0.935

```{r demographic-validation}
# Known survival rates from demographic model (Clay et al. / Pardo et al.)
known_survival <- tibble(
  age_class = c("j1", "j2j3", "imm", "fb", "sb", "nb"),
  # Use mean survival across the stages that map to each age class
  # J1 assumed equal to J2J3 (no J1-specific vital rates available)
  s_observed = c(
    0.846,                          # s_Juv (assumed = J2J3)
    0.846,                          # s_Juv
    0.921,                          # s_Imm
    mean(c(0.913, 0.920)),          # mean(s_EF, s_IF) - breeding birds
    mean(c(0.895, 0.892)),          # mean(s_ES, s_IS) - successful breeders
    mean(c(0.943, 0.935))           # mean(s_ENB, s_PF) - non-breeders
  ),
  stage_labels = c(
    "J1 (assumed = J2J3)",
    "J2, J3",
    "Imm4, Imm5, PB",
    "EF, IF (breeding)",
    "ES, IS (successful)",
    "ENB, Sabb (non-breeding)"
  )
)

# Combine DLL + PLL per capita rates from 03f
percap_combined <- percap_rates %>%
  group_by(age_class) %>%
  summarise(
    dll_percap = sum(percap[fishery == "DLL"], na.rm = TRUE),
    pll_percap = sum(percap[fishery == "PLL"], na.rm = TRUE),
    total_fishing_mort = sum(percap, na.rm = TRUE),
    .groups = "drop"
  )

# Join and calculate implied natural mortality
validation <- known_survival %>%
  left_join(percap_combined, by = "age_class") %>%
  mutate(
    total_mortality = 1 - s_observed,
    implied_natural_mort = total_mortality - total_fishing_mort,
    fishing_pct_of_total = round(100 * total_fishing_mort / total_mortality, 1),
    flag = case_when(
      implied_natural_mort < 0 ~ "IMPOSSIBLE (negative natural mortality)",
      implied_natural_mort < 0.01 ~ "SUSPICIOUS (near-zero natural mortality)",
      fishing_pct_of_total > 50 ~ "HIGH (fishing > 50% of total mortality)",
      TRUE ~ "OK"
    )
  )

# Demographic validation: implied natural mortality
# If fishing mortality from 03f exceeds total mortality, natural mortality goes negative — meaning the bycatch estimate is too high
validation %>%
  select(age_class, stage_labels, s_observed, total_mortality,
         dll_percap, pll_percap, total_fishing_mort,
         implied_natural_mort, fishing_pct_of_total, flag) %>%
  mutate(across(where(is.numeric), ~round(.x, 4))) %>%
  print(width = Inf)

n_impossible <- sum(validation$implied_natural_mort < 0)
n_suspicious <- sum(validation$implied_natural_mort >= 0 & validation$implied_natural_mort < 0.01)

if (n_impossible > 0) {
  warning(paste(n_impossible, "age class(es) have NEGATIVE implied natural mortality — 03f fishing mortality exceeds observed total mortality; bycatch estimates are almost certainly too high."))
} else if (n_suspicious > 0) {
  warning(paste(n_suspicious, "age class(es) have near-zero natural mortality — fishing mortality accounts for nearly all observed mortality; bycatch estimates may be too high."))
}

# Fishing as % of total mortality by age class
validation %>% select(age_class, fishing_pct_of_total, flag)

# Also check per-period: rates vary by period, so check worst case
# Per-period check (worst case)
percap_by_period <- bycatch_results %>%
  group_by(fishery, year_group) %>%
  summarise(
    across(c(fb, sb, nb, j1, j2j3, imm), ~sum(.x, na.rm = TRUE)),
    .groups = "drop"
  ) %>%
  pivot_longer(cols = c(fb, sb, nb, j1, j2j3, imm), names_to = "age_class", values_to = "annual_byc") %>%
  mutate(
    population = pop_sizes[age_class],
    percap = annual_byc / population
  )

# Combine DLL+PLL per period
percap_period_combined <- percap_by_period %>%
  group_by(year_group, age_class) %>%
  summarise(total_fm = sum(percap, na.rm = TRUE), .groups = "drop") %>%
  left_join(known_survival %>% select(age_class, s_observed), by = "age_class") %>%
  mutate(
    implied_nm = (1 - s_observed) - total_fm,
    flag = ifelse(implied_nm < 0, "*** NEGATIVE ***", "ok")
  )

# Show the worst period for each age class
worst_periods <- percap_period_combined %>%
  group_by(age_class) %>%
  slice_min(implied_nm, n = 1) %>%
  ungroup()

worst_periods %>%
  select(age_class, year_group, total_fm, s_observed, implied_nm, flag) %>%
  mutate(across(where(is.numeric), ~round(.x, 4))) %>%
  print()
```

## Mortality budget in bird counts

Same validation expressed as actual numbers of birds dying per year,
which is more intuitive than per-capita rates.

- **Total deaths/yr** = (1 - s_observed) × population size
- **Bycatch deaths/yr** = annual bycatch from 03f (DLL + PLL)
- **Other deaths/yr** = total - bycatch (implied natural mortality)

The observed survival rates (from mark-recapture monitoring at Bird Island)
already include all sources of mortality. Bycatch from 03f should represent
a biologically reasonable fraction of total deaths — not all of them.

```{r mortality-budget-birds}
# Annual bycatch totals by age class (DLL + PLL combined, averaged across periods)
byc_annual_by_age <- percap_rates %>%
  group_by(age_class) %>%
  summarise(annual_bycatch = sum(annual_byc, na.rm = TRUE), .groups = "drop")

# Build the mortality budget in bird counts
mortality_budget <- validation %>%
  select(age_class, stage_labels, s_observed) %>%
  mutate(population = pop_sizes[age_class]) %>%
  left_join(byc_annual_by_age, by = "age_class") %>%
  mutate(
    total_deaths_yr = (1 - s_observed) * population,
    other_deaths_yr = total_deaths_yr - annual_bycatch,
    fishing_pct = round(100 * annual_bycatch / total_deaths_yr, 1)
  )

# Mortality budget: bird counts per year
# Total deaths = observed mortality rate × population size
# Bycatch deaths = 03f estimate (DLL + PLL, mean annual)
# Other deaths = total - bycatch (implied natural causes)
mortality_budget %>%
  mutate(across(c(population, total_deaths_yr, annual_bycatch, other_deaths_yr), ~round(.x, 0))) %>%
  select(age_class, stage_labels, population, total_deaths_yr,
         annual_bycatch, other_deaths_yr, fishing_pct) %>%
  print(width = Inf)

# Totals across all age classes
tibble(
  total_BI_population   = round(sum(mortality_budget$population), 0),
  total_deaths_yr       = round(sum(mortality_budget$total_deaths_yr), 0),
  bycatch_deaths_yr     = round(sum(mortality_budget$annual_bycatch), 0),
  other_deaths_yr       = round(sum(mortality_budget$other_deaths_yr), 0),
  fishing_pct_all_deaths = round(100 * sum(mortality_budget$annual_bycatch) /
                                   sum(mortality_budget$total_deaths_yr), 1)
)
```

## Annual implied natural mortality check

Check that per-year fishing mortality never exceeds total mortality (implied natural
mortality must remain positive). Any year-age class with implied_nm < 0 means
fishing mortality alone exceeds the observed mortality rate — biologically impossible.

```{r annual-validation}
# Join annual per-capita rates with known survival to compute implied natural mortality
annual_validation <- annual_percap_by_age %>%
  select(year, age_class, percap) %>%
  rename(total_fm = percap) %>%
  left_join(known_survival %>% select(age_class, s_observed), by = "age_class") %>%
  mutate(
    total_mortality = 1 - s_observed,
    implied_nm      = total_mortality - total_fm,
    fishing_pct     = round(100 * total_fm / total_mortality, 1),
    flag            = ifelse(implied_nm < 0, "IMPOSSIBLE — fm > total mortality", "ok")
  )

# Show any flagged years
annual_validation %>%
  filter(flag != "ok") %>%
  select(year, age_class, total_fm, total_mortality, implied_nm, flag) %>%
  mutate(across(where(is.numeric), ~round(.x, 4)))

# Worst year per age class (lowest implied natural mortality)
annual_validation %>%
  group_by(age_class) %>%
  slice_min(implied_nm, n = 1) %>%
  ungroup() %>%
  select(age_class, year, total_fm, total_mortality, implied_nm, fishing_pct, flag) %>%
  mutate(across(where(is.numeric), ~round(.x, 4))) %>%
  arrange(desc(fishing_pct))
```

```{r save-validation}
# Save validation table
write_csv(validation, here(paste0("output/demographic_validation_04_v3_", date_sfx, ".csv")))
write_csv(percap_period_combined, here(paste0("output/demographic_validation_by_period_04_v3_", date_sfx, ".csv")))
write_csv(mortality_budget, here(paste0("output/mortality_budget_04_v3_", date_sfx, ".csv")))
write_csv(annual_validation, here(paste0("output/annual_validation_04_v3_", date_sfx, ".csv")))
```

# Session Info

```{r session-info}
sessionInfo()
```
