---
---
title: "Diagnostic Figures — Effort, Catchability, and Survival by Period"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE,
                      fig.width = 10, fig.height = 6)
```

```{r libraries}
library(tidyverse)
library(here)
library(patchwork)
library(scales)
```

# Overview

This script produces diagnostic figures comparing key inputs and outcomes across the four
time periods used in the bycatch model (06a–06d):

- **Fishing effort** (mean annual hooks by fishery type)
- **Effective catchability** (mean annual BI bycatch per hook — a realized BPUE)
- **Survival rates** (s_Juv and s_Adult as set in each period script)
- **Per-capita fishing mortality** (total FM across fisheries, by age class)

The goal is to verify that all three drivers of change (effort, catchability, and
demographic conditions) are coherent across periods, and to show where changes in
survival are attributable to fishing vs. other processes.

---

# Load Data

```{r load-data}
# Most recent 03f output date
date_sfx_03f <- "2026-02-25"

# Hooks by fleet and period (03g, 2026-02-24 — the file with hooks output)
hooks_raw <- read_csv(here("output/hooks_by_fleet_period_03g_2026-02-24.csv"),
                      show_col_types = FALSE)

# Bycatch by fishery and period (03f — mean annual BI bycatch, all ages summed)
byc_period <- read_csv(here(paste0("output/bycatch_by_fishery_period_03f_", date_sfx_03f, ".csv")),
                       show_col_types = FALSE)

# Per-capita fishing mortality by period and age class (03f)
percap <- read_csv(here(paste0("output/percap_rates_by_period_03f_", date_sfx_03f, ".csv")),
                   show_col_types = FALSE)
```

---

# Prepare Data

```{r period-labels}
# Period labels used in 06a-06d (bycatch model periods)
model_periods <- c("1990-1994", "1995-1999", "2000-2004", "2005-2009")

# The hooks file (03g) uses slightly different period endpoints
# ("1990-1995" vs "1990-1994") due to raw data year groupings — mapping them here
hooks_period_map <- c(
  "1990-1995" = "1990-1994",
  "1996-2000" = "1995-1999",
  "2001-2005" = "2000-2004",
  "2006-2010" = "2005-2009"
)
```

```{r effort-data}
# Total hooks by fishery type and period (summed across fleets, excluding NAs)
# The hooks_raw total_hooks column is the 5-year total; divide by 5 for mean annual
hooks_by_period <- hooks_raw %>%
  filter(!is.na(total_hooks)) %>%
  mutate(period = recode(year_group, !!!hooks_period_map)) %>%
  filter(period %in% model_periods) %>%
  group_by(period, fishery) %>%
  summarise(mean_annual_hooks = sum(total_hooks, na.rm = TRUE) / 5,
            .groups = "drop") %>%
  mutate(
    period   = factor(period, levels = model_periods),
    fishery  = factor(fishery, levels = c("PLL", "DLL"))
  )
```

```{r catchability-data}
# Effective catchability (BPUE): mean annual BI bycatch / mean annual hooks
# bi_total in bycatch file is already mean annual
catchability_by_period <- byc_period %>%
  left_join(hooks_by_period, by = c("year_group" = "period", "fishery")) %>%
  mutate(
    period  = factor(year_group, levels = model_periods),
    fishery = factor(fishery, levels = c("PLL", "DLL")),
    # bycatch per 1000 hooks (× 1000 for readability)
    bpue_per1000 = (bi_total / mean_annual_hooks) * 1000
  )
```

```{r survival-data}
# Stage-specific survival rates (Pardo et al., shared directly, 1980-2012)
# Same values used in all four period scripts (06a-d); period variation
# is captured through the bycatch data, not the survival parameters
surv_rates <- tibble(
  stage    = c("Juvenile", "EF", "ENB", "ES", "IF", "Imm", "IS", "PF"),
  survival = c(0.846,      0.913, 0.943, 0.895, 0.920, 0.921, 0.892, 0.935)
) %>%
  mutate(
    stage_group = ifelse(stage == "Juvenile", "Juvenile", "Adult (stage-specific)"),
    stage       = factor(stage, levels = c("Juvenile", "Imm", "IS", "IF", "ES", "EF", "ENB", "PF"))
  )
```

```{r percap-data}
# Per-capita total FM by period and age class (PLL + DLL combined)
percap_clean <- percap %>%
  mutate(
    period    = factor(year_group, levels = model_periods),
    age_class = factor(age_class, levels = c("j2j3", "imm", "fb", "sb", "nb"),
                       labels = c("Juvenile", "Immature", "Breeding (F)", "Breeding (S)", "Non-breeding"))
  )
```

---

# Shared Plot Theme

```{r theme}
period_colors <- c("1990-1994" = "#D55E00",
                   "1995-1999" = "#E69F00",
                   "2000-2004" = "#009E73",
                   "2005-2009" = "#0072B2")

fishery_colors <- c("PLL" = "#CC79A7", "DLL" = "#56B4E9")

diag_theme <- theme_classic(base_size = 12) +
  theme(
    strip.background = element_blank(),
    strip.text       = element_text(face = "bold"),
    legend.position  = "bottom",
    axis.text.x      = element_text(angle = 30, hjust = 1),
    plot.title       = element_text(face = "bold", size = 11),
    legend.box       = "horizontal"
  )
```

---

# Three Dual Y-Axis Diagnostic Figures

## Plot 1 — PLL: Effort (left) + Catchability (right)

```{r p1-pll, fig.width=6, fig.height=5}
pll_data <- hooks_by_period %>%
  filter(fishery == "PLL") %>%
  left_join(
    catchability_by_period %>% filter(fishery == "PLL") %>% select(period, bpue_per1000),
    by = "period"
  )

# Scale factor maps bpue_per1000 onto the hooks_M axis for dual-axis rendering
scale_pll <- max(pll_data$mean_annual_hooks / 1e6, na.rm = TRUE) /
             max(pll_data$bpue_per1000, na.rm = TRUE)

p1 <- ggplot(pll_data, aes(x = period)) +
  geom_col(aes(y = mean_annual_hooks / 1e6), fill = "#CC79A7", alpha = 0.7, width = 0.6) +
  geom_line(aes(y = bpue_per1000 * scale_pll, group = 1),
            color = "#882255", linewidth = 1.3) +
  geom_point(aes(y = bpue_per1000 * scale_pll),
             color = "#882255", size = 3) +
  scale_y_continuous(
    name     = "Mean annual hooks (millions)",
    labels   = label_number(suffix = "M"),
    sec.axis = sec_axis(
      ~ . / scale_pll,
      name   = "BI bycatch per 1,000 hooks",
      labels = label_number(accuracy = 0.0001)
    )
  ) +
  labs(title = "Pelagic longline (PLL): effort and catchability", x = NULL) +
  diag_theme +
  theme(
    axis.title.y.left  = element_text(color = "#CC79A7"),
    axis.text.y.left   = element_text(color = "#CC79A7"),
    axis.title.y.right = element_text(color = "#882255"),
    axis.text.y.right  = element_text(color = "#882255")
  )

p1
```

## Plot 2 — DLL: Effort (left) + Catchability (right)

```{r p2-dll, fig.width=6, fig.height=5}
dll_data <- hooks_by_period %>%
  filter(fishery == "DLL") %>%
  left_join(
    catchability_by_period %>% filter(fishery == "DLL") %>% select(period, bpue_per1000),
    by = "period"
  )

scale_dll <- max(dll_data$mean_annual_hooks / 1e6, na.rm = TRUE) /
             max(dll_data$bpue_per1000, na.rm = TRUE)

p2 <- ggplot(dll_data, aes(x = period)) +
  geom_col(aes(y = mean_annual_hooks / 1e6), fill = "#56B4E9", alpha = 0.7, width = 0.6) +
  geom_line(aes(y = bpue_per1000 * scale_dll, group = 1),
            color = "#0072B2", linewidth = 1.3) +
  geom_point(aes(y = bpue_per1000 * scale_dll),
             color = "#0072B2", size = 3) +
  scale_y_continuous(
    name     = "Mean annual hooks (millions)",
    labels   = label_number(suffix = "M"),
    sec.axis = sec_axis(
      ~ . / scale_dll,
      name   = "BI bycatch per 1,000 hooks",
      labels = label_number(accuracy = 0.0001)
    )
  ) +
  labs(title = "Demersal longline (DLL): effort and catchability", x = NULL) +
  diag_theme +
  theme(
    axis.title.y.left  = element_text(color = "#56B4E9"),
    axis.text.y.left   = element_text(color = "#56B4E9"),
    axis.title.y.right = element_text(color = "#0072B2"),
    axis.text.y.right  = element_text(color = "#0072B2")
  )

p2
```

## Mean vs Median Annual Effort Diagnostic

Plots 1–2 show **mean** annual hooks per period (5-year total ÷ 5). This diagnostic loads
the raw per-year source files to check whether within-period year-to-year variability is
large enough to make the median meaningfully different. A high CV% or mean/median ratio
notably different from 1 suggests median would better represent a typical year in that period.

```{r mean-vs-median-effort}
# --- PLL: sum Hooks to annual totals per year ---
pll_annual <- read_csv(here("output/pll_effort_by_fleet.csv"), show_col_types = FALSE) %>%
  filter(Year >= 1990, Year <= 2009) %>%
  mutate(
    year_group = case_when(
      Year >= 1990 & Year <= 1994 ~ "1990-1994",
      Year >= 1995 & Year <= 1999 ~ "1995-1999",
      Year >= 2000 & Year <= 2004 ~ "2000-2004",
      Year >= 2005 & Year <= 2009 ~ "2005-2009"
    )
  ) %>%
  group_by(fishery = "PLL", year_group, year = Year) %>%
  summarise(annual_hooks = sum(Hooks, na.rm = TRUE), .groups = "drop")

# --- DLL: read cleaned output from script 01 (Chile PLL filtered, all sources combined) ---
dll_annual <- read_csv(here("output/dll_effort_by_fleet.csv"), show_col_types = FALSE) %>%
  rename(fleet = Fleet) %>%
  filter(Year >= 1990, Year <= 2009) %>%
  mutate(
    year_group = case_when(
      Year >= 1990 & Year <= 1994 ~ "1990-1994",
      Year >= 1995 & Year <= 1999 ~ "1995-1999",
      Year >= 2000 & Year <= 2004 ~ "2000-2004",
      Year >= 2005 & Year <= 2009 ~ "2005-2009"
    )
  ) %>%
  group_by(fishery = "DLL", year_group, year = Year) %>%
  summarise(annual_hooks = sum(Hooks, na.rm = TRUE), .groups = "drop")

# Combine PLL + DLL
effort_by_year <- bind_rows(pll_annual, dll_annual) %>%
  mutate(
    year_group = factor(year_group, levels = model_periods),
    fishery    = factor(fishery, levels = c("PLL", "DLL"))
  )

# Summary table: mean, median, min, max, CV%, mean/median ratio
effort_mean_median <- effort_by_year %>%
  group_by(fishery, year_group) %>%
  summarise(
    n_years        = n(),
    mean_M         = round(mean(annual_hooks,   na.rm = TRUE) / 1e6, 3),
    median_M       = round(median(annual_hooks, na.rm = TRUE) / 1e6, 3),
    min_M          = round(min(annual_hooks,    na.rm = TRUE) / 1e6, 3),
    max_M          = round(max(annual_hooks,    na.rm = TRUE) / 1e6, 3),
    cv_pct         = round(sd(annual_hooks, na.rm = TRUE) /
                             mean(annual_hooks, na.rm = TRUE) * 100, 1),
    mean_med_ratio = round(mean(annual_hooks, na.rm = TRUE) /
                             median(annual_hooks, na.rm = TRUE), 3),
    .groups = "drop"
  )

print(effort_mean_median)
```

```{r mean-vs-median-plot, fig.width=10, fig.height=5}
# Annual values as dots, mean (circle) and median (square) overlaid per period
effort_stats_long <- effort_mean_median %>%
  pivot_longer(cols = c(mean_M, median_M), names_to = "stat", values_to = "hooks_M") %>%
  mutate(stat = recode(stat, "mean_M" = "Mean", "median_M" = "Median"))

ggplot(effort_by_year, aes(x = year_group, y = annual_hooks / 1e6, color = year_group)) +
  geom_jitter(width = 0.12, size = 2.5, alpha = 0.75) +
  geom_point(data = effort_stats_long,
             aes(x = year_group, y = hooks_M, shape = stat),
             inherit.aes = FALSE, size = 5, color = "black", stroke = 1.2) +
  scale_color_manual(values = period_colors, guide = "none") +
  scale_shape_manual(values = c("Mean" = 19, "Median" = 15), name = NULL) +
  facet_wrap(~ fishery, scales = "free_y") +
  labs(
    x        = NULL,
    y        = "Annual hooks (millions)",
    title    = "Mean vs median annual fishing effort per period",
    subtitle = paste0("Coloured dots = individual years; circle = mean; square = median.\n",
                      "CV% and mean/median ratio in summary table above.")
  ) +
  theme_bw(base_size = 12) +
  theme(panel.grid.minor = element_blank(), legend.position = "bottom")
```

---

## Plot 3 — Survival (left) + Mean Per-Capita FM (right)

```{r p3-surv-fm, fig.width=6, fig.height=5}
# Mean per-capita FM across all age classes per period
mean_percap <- percap_clean %>%
  group_by(period) %>%
  summarise(mean_fm = mean(total_percap), .groups = "drop")

# Survival reference lines: constant across periods (Pardo 1980-2012)
# Show juvenile and mean adult for readability
surv_juv       <- surv_rates %>% filter(stage == "Juvenile") %>% pull(survival)
surv_adult_mean <- surv_rates %>% filter(stage_group == "Adult (stage-specific)") %>%
  summarise(m = mean(survival)) %>% pull(m)

p3 <- ggplot(mean_percap, aes(x = period, y = mean_fm, group = 1)) +
  # Constant survival reference lines (horizontal; same across all periods)
  geom_hline(yintercept = 1 - surv_juv,       color = "#E69F00", linewidth = 0.8,
             linetype = "dashed", alpha = 0.8) +
  geom_hline(yintercept = 1 - surv_adult_mean, color = "#0072B2", linewidth = 0.8,
             linetype = "dashed", alpha = 0.8) +
  # Period-varying FM
  geom_line(color = "grey30", linewidth = 1.3) +
  geom_point(color = "grey30", size = 3, shape = 17) +
  annotate("text", x = 0.6, y = 1 - surv_juv + 0.001,
           label = "Total juv. mortality", hjust = 0, size = 3, color = "#E69F00") +
  annotate("text", x = 0.6, y = 1 - surv_adult_mean + 0.001,
           label = "Total adult mortality (mean)", hjust = 0, size = 3, color = "#0072B2") +
  scale_y_continuous(
    name   = "Mean per-capita FM (all ages, PLL + DLL)",
    labels = label_number(accuracy = 0.001)
  ) +
  labs(
    title    = "Per-capita fishing mortality by period",
    subtitle = "Dashed lines = total mortality budget (1 \u2212 survival); FM is the fishing share",
    x        = NULL
  ) +
  diag_theme

p3
```

## Combined Figure

```{r combined-figure, fig.width=16, fig.height=6}
dir.create(here("output/figures"), showWarnings = FALSE, recursive = TRUE)

combined <- p1 + p2 + p3 +
  plot_annotation(
    title    = "Fishing Effort, Catchability, and Survival — Trends Across Periods",
    subtitle = "Bars = mean annual hooks; lines = bycatch per 1,000 hooks (plots 1-2) or survival/FM (plot 3)",
    theme    = theme(
      plot.title    = element_text(face = "bold", size = 13),
      plot.subtitle = element_text(size = 10, color = "grey40")
    )
  )

combined

ggsave(here("output/figures/07_diagnostic_effort_catchability_survival.png"),
       combined, width = 16, height = 6, dpi = 300)
```

---

# Fishing Effort by Fleet and Period

Mean annual hooks by fleet across the four 5-year periods, separated by fishery type.
Top fleets by total effort are shown individually; the remainder are collapsed to "Other".

```{r p-fleet-effort, fig.width=12, fig.height=8}
# Mean annual hooks per fleet × period × fishery
# total_hooks in hooks_raw is the 5-year total; divide by 5 for mean annual
fleet_effort <- hooks_raw %>%
  filter(!is.na(total_hooks)) %>%
  mutate(period = recode(year_group, !!!hooks_period_map)) %>%
  filter(period %in% model_periods) %>%
  mutate(
    mean_annual_hooks = total_hooks / 5,
    period  = factor(period,  levels = model_periods),
    fishery = factor(fishery, levels = c("PLL", "DLL"))
  )

# Identify top fleets per fishery by total hooks across all periods
top_n_fleets <- 7

top_fleets_by_fishery <- fleet_effort %>%
  group_by(fishery, fleet) %>%
  summarise(total = sum(mean_annual_hooks, na.rm = TRUE), .groups = "drop") %>%
  group_by(fishery) %>%
  slice_max(order_by = total, n = top_n_fleets) %>%
  ungroup() %>%
  select(fishery, fleet)

# Collapse remaining fleets to "Other"
fleet_effort_plot <- fleet_effort %>%
  left_join(top_fleets_by_fishery %>% mutate(is_top = TRUE),
            by = c("fishery", "fleet")) %>%
  mutate(fleet_label = ifelse(is_top == TRUE, fleet, "Other")) %>%
  group_by(period, fishery, fleet_label) %>%
  summarise(mean_annual_hooks = sum(mean_annual_hooks, na.rm = TRUE), .groups = "drop")

# Order fleet_label: top fleets by total descending, "Other" last
fleet_order <- fleet_effort_plot %>%
  group_by(fleet_label) %>%
  summarise(total = sum(mean_annual_hooks), .groups = "drop") %>%
  arrange(desc(total)) %>%
  filter(fleet_label != "Other") %>%
  pull(fleet_label) %>%
  c("Other")

fleet_effort_plot <- fleet_effort_plot %>%
  mutate(fleet_label = factor(fleet_label, levels = fleet_order))

# Color palette: distinct colors for top fleets, grey for Other
n_top <- length(fleet_order) - 1
fleet_colors <- setNames(
  c(RColorBrewer::brewer.pal(min(n_top, 12), "Paired")[seq_len(n_top)], "grey70"),
  fleet_order
)

p_fleet_effort <- ggplot(fleet_effort_plot,
                         aes(x = period, y = mean_annual_hooks / 1e6,
                             fill = fleet_label)) +
  geom_col(width = 0.7, alpha = 0.9) +
  facet_wrap(~ fishery, scales = "free_y",
             labeller = labeller(fishery = c(PLL = "Pelagic longline (PLL)",
                                             DLL = "Demersal longline (DLL)"))) +
  scale_fill_manual(values = fleet_colors, name = "Fleet") +
  scale_y_continuous(labels = scales::label_number(suffix = "M")) +
  labs(
    x        = NULL,
    y        = "Mean annual hooks (millions)",
    title    = "Fishing effort by fleet and period",
    subtitle = paste0("Top ", top_n_fleets, " fleets per fishery shown individually; remainder collapsed to \u2018Other\u2019")
  ) +
  diag_theme +
  theme(
    legend.position = "right",
    strip.text      = element_text(face = "bold", size = 11)
  )

p_fleet_effort

ggsave(here("output/figures/07_fleet_effort_by_period.png"),
       p_fleet_effort, width = 12, height = 8, dpi = 300)
```

## By Year

Annual total hooks by fleet and fishery, 1990–2009. Read directly from raw source files;
no spatial filtering applied (this is total global effort, not overlap-weighted).

```{r p-fleet-effort-year, fig.width=14, fig.height=8}
# ── PLL by year: read raw ICCAT data, filter to study area (south of 20°S) ──
# The raw PLL file covers all latitudes to 62.5°N (including tropical tuna fisheries).
# Filtering to Lat <= -20 restricts to the southern ocean study area where WAAL occur,
# matching the published Tuck 2015 figure scale (~200–400M hooks/year).
# NA Flag values are relabeled "Other" so they collapse with other minor fleets.
pll_raw_yr <- read_csv(here("output/pll_effort_by_fleet.csv"), show_col_types = FALSE) %>%
  filter(Year >= 1990, Year <= 2009, Lat <= -20) %>%
  mutate(Flag = ifelse(is.na(Flag), "Other", Flag)) %>%
  group_by(Year, fleet = Flag) %>%
  summarise(total_hooks = sum(Hooks, na.rm = TRUE), .groups = "drop") %>%
  mutate(fishery = "PLL")

# ── DLL by year: read cleaned output from script 01 (Chile PLL filtered, all sources combined) ──
dll_raw_yr <- read_csv(here("output/dll_effort_by_fleet.csv"), show_col_types = FALSE) %>%
  rename(fleet = Fleet) %>%
  filter(Year >= 1990, Year <= 2009) %>%
  group_by(Year, fleet) %>%
  summarise(total_hooks = sum(Hooks, na.rm = TRUE), .groups = "drop") %>%
  mutate(fishery = "DLL")

# ── Combine PLL + DLL ───────────────────────────────────────────────────────
fleet_effort_yr <- bind_rows(pll_raw_yr, dll_raw_yr) %>%
  mutate(fishery = factor(fishery, levels = c("PLL", "DLL")))

# ── Identify top fleets per fishery (same n as period plot) ────────────────
top_fleets_yr <- fleet_effort_yr %>%
  group_by(fishery, fleet) %>%
  summarise(total = sum(total_hooks, na.rm = TRUE), .groups = "drop") %>%
  group_by(fishery) %>%
  slice_max(order_by = total, n = top_n_fleets) %>%
  ungroup() %>%
  select(fishery, fleet)

fleet_effort_yr_plot <- fleet_effort_yr %>%
  left_join(top_fleets_yr %>% mutate(is_top = TRUE),
            by = c("fishery", "fleet")) %>%
  mutate(fleet_label = ifelse(is_top == TRUE, fleet, "Other")) %>%
  group_by(Year, fishery, fleet_label) %>%
  summarise(total_hooks = sum(total_hooks, na.rm = TRUE), .groups = "drop")

# Fleet order: top by total descending, "Other" last
fleet_order_yr <- fleet_effort_yr_plot %>%
  group_by(fleet_label) %>%
  summarise(total = sum(total_hooks), .groups = "drop") %>%
  arrange(desc(total)) %>%
  filter(fleet_label != "Other") %>%
  pull(fleet_label) %>%
  c("Other")

fleet_effort_yr_plot <- fleet_effort_yr_plot %>%
  mutate(fleet_label = factor(fleet_label, levels = fleet_order_yr))

# Color palette (same structure as period plot)
n_top_yr <- length(fleet_order_yr) - 1
fleet_colors_yr <- setNames(
  c(RColorBrewer::brewer.pal(min(n_top_yr, 12), "Paired")[seq_len(n_top_yr)], "grey70"),
  fleet_order_yr
)

p_fleet_effort_yr <- ggplot(fleet_effort_yr_plot,
                             aes(x = Year, y = total_hooks / 1e6,
                                 fill = fleet_label)) +
  geom_col(width = 0.9, alpha = 0.9) +
  facet_wrap(~ fishery, scales = "free_y",
             labeller = labeller(fishery = c(PLL = "Pelagic longline (PLL)",
                                             DLL = "Demersal longline (DLL)"))) +
  scale_fill_manual(values = fleet_colors_yr, name = "Fleet") +
  scale_x_continuous(breaks = seq(1990, 2009, by = 5)) +
  scale_y_continuous(labels = scales::label_number(suffix = "M")) +
  labs(
    x        = NULL,
    y        = "Annual hooks (millions)",
    title    = "Fishing effort by fleet and year",
    subtitle = paste0("Top ", top_n_fleets, " fleets per fishery shown individually; remainder collapsed to \u2018Other\u2019.\n",
                      "PLL filtered to south of 20\u00b0S (study area); DLL sources are already southern-ocean only.")
  ) +
  diag_theme +
  theme(
    legend.position = "right",
    strip.text      = element_text(face = "bold", size = 11)
  )

p_fleet_effort_yr

ggsave(here("output/figures/07_fleet_effort_by_year.png"),
       p_fleet_effort_yr, width = 14, height = 8, dpi = 300)
```

---

# Interpretation Notes

```{r interpretation}
# Summary table: how much did each component change from 1990-1994 to 2005-2009?

# Effort change
effort_summary <- hooks_by_period %>%
  group_by(fishery) %>%
  summarise(
    hooks_1990 = mean_annual_hooks[period == "1990-1994"],
    hooks_2005 = mean_annual_hooks[period == "2005-2009"],
    pct_change = 100 * (hooks_2005 - hooks_1990) / hooks_1990
  )

cat("=== Effort change (1990-1994 to 2005-2009) ===\n")
print(effort_summary)

# Catchability change
bpue_summary <- catchability_by_period %>%
  group_by(fishery) %>%
  summarise(
    bpue_1990 = bpue_per1000[period == "1990-1994"],
    bpue_2005 = bpue_per1000[period == "2005-2009"],
    pct_change = 100 * (bpue_2005 - bpue_1990) / bpue_1990
  )

cat("\n=== BPUE change (1990-1994 to 2005-2009) ===\n")
print(bpue_summary)

# Survival (constant across periods — Pardo 1980-2012 stage-specific values)
cat("\n=== Survival rates (constant across periods, Pardo 1980-2012) ===\n")
surv_rates %>%
  select(stage, stage_group, survival) %>%
  mutate(total_mortality = 1 - survival) %>%
  print()

# Per-capita FM change (averaged across age classes)
percap_summary <- percap_clean %>%
  group_by(period) %>%
  summarise(mean_total_percap = mean(total_percap)) %>%
  mutate(
    pct_change_from_1990 = 100 * (mean_total_percap -
      mean_total_percap[period == "1990-1994"]) /
      mean_total_percap[period == "1990-1994"]
  )

cat("\n=== Mean per-capita FM (averaged across age classes) ===\n")
print(percap_summary)
```

---

# FM Decomposition and Multi-Period Recovery Heatmaps

Three new figures extending the diagnostics above:

- **Figure P4** — stacked bar chart decomposing the FM decline each period into
  effort-driven (fewer hooks) vs catchability-driven (β) components
- **Figure P5** — the 06a coverage × efficacy lambda heatmap with period recovery
  threshold points overlaid (6b, 6c, 6d + 6a baseline)
- **Figure P6** — effort_ratio × β_ratio lambda surface with observed period points
  at their decomposed coordinates and theoretical scenario points along the right edge

---

## Lambda Helper and Behavioral Vital Rates

```{r lambda-helper}
# Mitigation scenario efficacy values (same as 06a-d; used in P6 edge points)
eff_tori_lines  <- 0.75
eff_night_set   <- 0.60
eff_line_weight <- 0.65
eff_hook_shield <- 0.35
eff_2of3 <- 1 - (1 - eff_tori_lines) * (1 - eff_night_set)
eff_3of3 <- 1 - (1 - eff_tori_lines) * (1 - eff_night_set) * (1 - eff_line_weight)

mit_scens <- tibble(
  scenario = c("Hook shielding only", "Line weighting only", "Night setting only",
               "Tori lines only", "2/3 (tori + night)", "3/3 (tori + night + weight)",
               "Complete elimination"),
  efficacy = c(eff_hook_shield, eff_line_weight, eff_night_set, eff_tori_lines,
               eff_2of3, eff_3of3, 1.0)
)

# Stage-specific behavioral vital rates (Clay 2019; same across all periods)
b_EF  <- 0.746; b_ENB <- 0.589; b_ES  <- 0.016; b_IF  <- 0.744
b_Imm <- 0.101; b_IS  <- 0.011; b_PF  <- 0.941
k_EF  <- 0.631; k_ENB <- 0.675; k_ES  <- 0.500; k_IF  <- 0.627
k_Imm <- 0.585; k_IS  <- 0.627; k_PF  <- 0.791
r_EF  <- 0.973; r_ENB <- 0.953; r_ES  <- 0.292; r_IF  <- 0.948
r_Imm <- 0.692; r_IS  <- 0.459; r_PF  <- 0.972

calc_lambda_from_survival <- function(ns_Juv, ns_Imm, ns_EF, ns_ES, ns_ENB,
                                      ns_IF, ns_IS, ns_PF) {
  n_m <- matrix(data = c(
    0,      0,       0,       0,       0,                                             0.5*ns_Juv,               0,                        0.5*ns_Juv,               0,                        0,                           0,
    ns_Juv, 0,       0,       0,       0,                                             0,                        0,                        0,                        0,                        0,                           0,
    0,      ns_Juv,  0,       0,       0,                                             0,                        0,                        0,                        0,                        0,                           0,
    0,      0,       ns_Imm,  0,       0,                                             0,                        0,                        0,                        0,                        0,                           0,
    0,      0,       0,       ns_Imm,  ns_Imm*r_Imm*(1-b_Imm)+ns_Imm*(1-r_Imm),     0,                        0,                        0,                        0,                        0,                           0,
    0,      0,       0,       0,       ns_Imm*r_Imm*b_Imm*k_Imm,                     0,                        0,                        0,                        0,                        0,                           0,
    0,      0,       0,       0,       ns_Imm*r_Imm*b_Imm*(1-k_Imm),                 0,                        0,                        0,                        0,                        0,                           0,
    0,      0,       0,       0,       0,                                             ns_IS*r_IS*b_IS*k_IS,     ns_IF*r_IF*b_IF*k_IF,     ns_ES*r_ES*b_ES*k_ES,     ns_EF*r_EF*b_EF*k_EF,     ns_ENB*r_ENB*b_ENB*k_ENB,    ns_PF*r_PF*b_PF*k_PF,
    0,      0,       0,       0,       0,                                             ns_IS*r_IS*b_IS*(1-k_IS), ns_IF*r_IF*b_IF*(1-k_IF), ns_ES*r_ES*b_ES*(1-k_ES), ns_EF*r_EF*b_EF*(1-k_EF), ns_ENB*r_ENB*b_ENB*(1-k_ENB),ns_PF*r_PF*b_PF*(1-k_PF),
    0,      0,       0,       0,       0,                                             ns_IS*r_IS*(1-b_IS),      ns_IF*r_IF*(1-b_IF),      ns_ES*r_ES*(1-b_ES),      ns_EF*r_EF*(1-b_EF),      ns_ENB*r_ENB*(1-b_ENB),     ns_PF*r_PF*(1-b_PF),
    0,      0,       0,       0,       0,                                             ns_IS*(1-r_IS),           ns_IF*(1-r_IF),           ns_ES*(1-r_ES),           ns_EF*(1-r_EF),           ns_ENB*(1-r_ENB),            ns_PF*(1-r_PF)
  ), nrow = 11, byrow = TRUE)
  as.numeric(eigen(n_m)$values[1])
}
```

---

## Load Decomposition Data (03g counterfactuals)

```{r decomp-load}
# 03g files use different period labels than 03f/Clay ("1990-1995" vs "1990-1994")
date_sfx_03g <- "2026-02-24"

# Observed bycatch from 03g (same data as 03f but with 03g period labels)
byc_obs_03g <- read_csv(
  here(paste0("output/bycatch_by_fishery_period_03g_", date_sfx_03g, ".csv")),
  show_col_types = FALSE
)

# Counterfactual: baseline beta × current effort (effort change only)
byc_eff_only <- read_csv(
  here(paste0("output/bycatch_effort_only_03g_", date_sfx_03g, ".csv")),
  show_col_types = FALSE
)

# Counterfactual: current beta × baseline effort (catchability change only)
byc_catch_only <- read_csv(
  here(paste0("output/bycatch_catchability_only_03g_", date_sfx_03g, ".csv")),
  show_col_types = FALSE
)

# Clay demographic data for period-specific population sizes
dem_props_07 <- read_csv(here("data/WAAL_dist_Clay2019/Dem_props copy.csv"),
                         show_col_types = FALSE) %>%
  mutate(Dem_class = str_replace(Dem_class, "J2\\+3", "J2J3"))

# Map 03g effort periods to Clay population periods
period_pop_map_07 <- tibble(
  year_group = c("1990-1995", "1996-2000", "2001-2005", "2006-2010"),
  pop_period = c("1990-1994", "1995-1999", "2000-2004", "2005-2009")
)

# Population sizes by age class and 03g effort period
age_cols_decomp <- c("fb", "sb", "nb", "j2j3", "imm")

pop_by_effort_period <- dem_props_07 %>%
  filter(Dem_class != "J1", Period != "1990-2009") %>%
  select(Period, Dem_class, Pop) %>%
  mutate(age_class = case_when(
    Dem_class == "J2J3" ~ "j2j3",
    Dem_class == "IMM"  ~ "imm",
    Dem_class == "FB"   ~ "fb",
    Dem_class == "SB"   ~ "sb",
    Dem_class == "NB"   ~ "nb"
  )) %>%
  left_join(period_pop_map_07, by = c("Period" = "pop_period"))

# Baseline (1990-1995 effort period) population vector — fixed denominator for all periods
pop_tmp <- pop_by_effort_period %>%
  filter(year_group == "1990-1995") %>%
  select(age_class, Pop)
pop_baseline_vec <- setNames(pop_tmp$Pop, pop_tmp$age_class)

# Helper: sum DLL + PLL bycatch, compute per-capita FM using baseline population
to_percap_03g <- function(byc_df, period_filter = NULL) {
  df <- if (!is.null(period_filter)) filter(byc_df, year_group %in% period_filter) else byc_df
  df %>%
    group_by(year_group) %>%
    summarise(across(all_of(age_cols_decomp), ~ sum(.x, na.rm = TRUE)), .groups = "drop") %>%
    pivot_longer(cols = all_of(age_cols_decomp),
                 names_to = "age_class", values_to = "bycatch") %>%
    mutate(pop = pop_baseline_vec[age_class],
           fm  = bycatch / pop)
}
```

---

## Compute FM Decomposition

```{r fm-decomposition}
baseline_03g <- "1990-1995"
later_03g    <- c("1996-2000", "2001-2005", "2006-2010")

# Observed per-capita FM for baseline period
fm_base_vec <- to_percap_03g(byc_obs_03g, period_filter = baseline_03g) %>%
  rename(fm_base = fm) %>% select(age_class, fm_base)

# Observed per-capita FM for later periods
fm_obs_later <- to_percap_03g(byc_obs_03g, period_filter = later_03g) %>%
  rename(fm_obs = fm)

# Effort-only counterfactual: baseline beta, current effort
fm_eff_only <- to_percap_03g(byc_eff_only) %>% rename(fm_eff_only = fm)

# Catchability-only counterfactual: current beta, baseline effort
fm_catch_only <- to_percap_03g(byc_catch_only) %>% rename(fm_catch_only = fm)

# Combine and compute multiplicative decomposition factors
decomp <- fm_obs_later %>%
  left_join(fm_base_vec, by = "age_class") %>%
  left_join(fm_eff_only   %>% select(year_group, age_class, fm_eff_only),
            by = c("year_group", "age_class")) %>%
  left_join(fm_catch_only %>% select(year_group, age_class, fm_catch_only),
            by = c("year_group", "age_class")) %>%
  mutate(
    effort_factor  = fm_eff_only / fm_base,      # fraction of baseline FM from effort alone
    beta_factor    = fm_obs      / fm_eff_only,  # residual fraction after effort accounted for
    overall_factor = fm_obs      / fm_base       # ≈ effort_factor × beta_factor
  )

# Bycatch-weighted period-level summary
decomp_by_period <- decomp %>%
  group_by(year_group) %>%
  summarise(
    effort_ratio  = weighted.mean(effort_factor,  w = fm_base, na.rm = TRUE),
    beta_ratio    = weighted.mean(beta_factor,    w = fm_base, na.rm = TRUE),
    overall_ratio = weighted.mean(overall_factor, w = fm_base, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  bind_rows(tibble(year_group = baseline_03g,
                   effort_ratio = 1, beta_ratio = 1, overall_ratio = 1)) %>%
  arrange(year_group)

decomp_by_period %>% mutate(across(where(is.numeric), ~ round(.x, 4))) %>% print()
```

---

## Figure P4 — FM Decomposition: Effort vs Catchability

```{r p4-fm-decomp, fig.width=7, fig.height=5}
decomp_bars <- decomp_by_period %>%
  filter(year_group != baseline_03g) %>%
  mutate(
    # Additive decomposition:
    #   effort_contrib = 1 - effort_ratio  (reduction from fewer hooks alone)
    #   beta_contrib   = effort_ratio - overall_ratio  (additional reduction from lower β)
    effort_contrib = (1 - effort_ratio)            * 100,
    beta_contrib   = (effort_ratio - overall_ratio) * 100,
    year_group = factor(year_group, levels = later_03g)
  ) %>%
  pivot_longer(cols = c(effort_contrib, beta_contrib),
               names_to = "component", values_to = "pct_reduction") %>%
  mutate(component = recode(component,
    effort_contrib = "Effort reduction (fewer hooks)",
    beta_contrib   = "Catchability reduction (\u03b2; fishing practices)"
  ))

p4 <- ggplot(decomp_bars, aes(x = year_group, y = pct_reduction, fill = component)) +
  geom_col(position = "stack", width = 0.6, alpha = 0.85) +
  geom_hline(yintercept = 0, color = "black", linewidth = 0.3) +
  scale_fill_manual(
    values = c("Effort reduction (fewer hooks)"              = "#2166ac",
               "Catchability reduction (\u03b2; fishing practices)" = "#d6604d"),
    name = "FM reduction component"
  ) +
  labs(
    x        = "Period",
    y        = "% reduction in FM relative to 1990\u20131995 baseline",
    title    = "Decomposition of fishing mortality decline by period",
    subtitle = "FM change split into effort-driven (fewer hooks) and catchability-driven (\u03b2) components"
  ) +
  diag_theme +
  theme(legend.position = "right")

p4

ggsave(here("output/figures/07_p4_fm_decomposition_bars.png"),
       p4, width = 7, height = 5, dpi = 300)
```

---

## Period-Specific FM and Natural Mortality (03f baseline)

```{r period-fm-nm}
# Total per-capita FM per period from 03f output (period labels: "1990-1994" etc.)
# Uses percap (raw, unlabeled) loaded at top of script
total_fm_03f <- percap %>%
  select(year_group, age_class, total_percap) %>%
  pivot_wider(names_from = age_class, values_from = total_percap)

# Stage-specific survival (Pardo 1980-2012, shared directly — same values as 06a-d)
s_Juv_base <- 0.846
s_EF_base  <- 0.913; s_ENB_base <- 0.943; s_ES_base <- 0.895
s_IF_base  <- 0.920; s_Imm_base <- 0.921; s_IS_base <- 0.892; s_PF_base <- 0.935

# Natural mortality for 1990-1994 baseline (used to build heatmap surface)
nm_base_03f <- total_fm_03f %>%
  filter(year_group == "1990-1994") %>%
  mutate(
    nm_juv = (1 - s_Juv_base) - j2j3,
    nm_imm = (1 - s_Imm_base) - imm,
    nm_efb = (1 - s_EF_base)  - fb,
    nm_esb = (1 - s_ES_base)  - sb,
    nm_enb = (1 - s_ENB_base) - nb,
    nm_ifb = (1 - s_IF_base)  - fb,
    nm_isb = (1 - s_IS_base)  - sb,
    nm_pf  = (1 - s_PF_base)  - nb
  )

fm_base_03f <- total_fm_03f %>% filter(year_group == "1990-1994")
```

---



---

## Figure P7 — Combined Cross-Period Heatmap with All Reference Points

Load the heatmap lambda grids, reference points (no-mitigation, historical, current),
and scenario points (hypothetical future) saved by each period script (06a–06d).
Display as a 2×2 faceted heatmap for cross-period comparison.

```{r p7-combined-heatmap, fig.width=14, fig.height=10}
date_sfx_06 <- "2026-03-26"
periods_06  <- c("1990-1994", "1995-1999", "2000-2004", "2005-2009")

# Load heatmap lambda grids (efficacy × coverage → lambda) from each period script
heatmap_all <- map_dfr(periods_06, function(p) {
  f <- here(paste0("output/heatmap_data_", p, "_", date_sfx_06, ".csv"))
  if (file.exists(f)) read_csv(f, show_col_types = FALSE) else NULL
}) %>%
  mutate(period = factor(period, levels = model_periods))

# Load reference points (no-mitigation, historical, current) from each period script
ref_all <- map_dfr(periods_06, function(p) {
  f <- here(paste0("output/heatmap_ref_pts_", p, "_", date_sfx_06, ".csv"))
  if (file.exists(f)) read_csv(f, show_col_types = FALSE) else NULL
}) %>%
  mutate(period = factor(period, levels = model_periods))

# Load scenario heatmap points (hypothetical future scenarios) from each period script
scen_all <- map_dfr(periods_06, function(p) {
  f <- here(paste0("output/scenario_heatmap_pts_", p, "_", date_sfx_06, ".csv"))
  if (file.exists(f)) read_csv(f, show_col_types = FALSE) else NULL
}) %>%
  mutate(period = factor(period, levels = model_periods))

# Shared lambda range across all periods for consistent color scale
lambda_range <- range(heatmap_all$lambda, na.rm = TRUE)

p7 <- ggplot(heatmap_all, aes(x = coverage, y = efficacy, fill = lambda)) +
  geom_tile() +
  # Lambda = 1 threshold contour
  geom_contour(aes(z = lambda), breaks = 1, color = "white", linewidth = 1) +
  # Hypothetical scenarios: compliance level progression lines and points
  geom_line(data = scen_all,
            aes(x = effective_coverage, y = effective_efficacy,
                group = interaction(period, efficacy_scenario)),
            inherit.aes = FALSE, color = "white", linewidth = 0.4, alpha = 0.5) +
  geom_point(data = scen_all %>% filter(compliance_level == "50%"),
             aes(x = effective_coverage, y = effective_efficacy),
             inherit.aes = FALSE,
             size = 2, shape = 16, color = "white", alpha = 0.8) +
  # Reference point: gold diamond at no-mitigation baseline per period
  geom_point(data = ref_all %>% filter(str_starts(ref_label, "No mitigation")),
             aes(x = effective_coverage, y = effective_efficacy),
             inherit.aes = FALSE,
             size = 5, shape = 23, fill = "gold", color = "black", stroke = 0.7) +
  geom_label(data = ref_all %>% filter(str_starts(ref_label, "No mitigation")),
             aes(x = effective_coverage, y = effective_efficacy, label = ref_label),
             inherit.aes = FALSE,
             size = 2.0, fill = "gold", alpha = 0.85,
             label.padding = unit(0.12, "lines"),
             hjust = -0.05, vjust = 0.5) +
  facet_wrap(~ period, nrow = 2) +
  scale_fill_distiller(direction = 1, palette = "RdBu",
                       limits = lambda_range,
                       name = expression(lambda)) +
  scale_x_continuous(breaks = seq(0, 1, 0.25),
                     labels = paste0(seq(0, 100, 25), "%")) +
  scale_y_continuous(breaks = seq(0, 1, 0.25),
                     labels = paste0(seq(0, 100, 25), "%")) +
  labs(
    x        = "% bycatch-weighted coverage (fisheries with mitigation)",
    y        = "Mitigation efficacy",
    title    = "Population growth rate across mitigation space \u2014 all periods",
    subtitle = paste0(
      "White contour = \u03bb = 1; gold diamonds = no-mitigation baseline per period.\n",
      "White points/lines = Bell et al. mitigation scenarios at 50% compliance."
    )
  ) +
  coord_cartesian(expand = FALSE) +
  theme_bw(base_size = 11) +
  theme(
    panel.grid      = element_blank(),
    strip.text      = element_text(face = "bold", size = 11),
    legend.position = "right"
  )

p7

ggsave(here("output/figures/07_p7_combined_heatmap_all_periods.png"),
       p7, width = 14, height = 10, dpi = 300)
```

---

```{r session-info}
sessionInfo()
```