---
title: "WAAL Population Projections — 3/3 Mitigation Scenario"
output: html_document
---

```{r setup, echo = FALSE, include = FALSE}
library(here)
library(tidyverse)
library(RColorBrewer)

# Must match the date suffix used when 06a was last run
date_sfx      <- "2026-02-25"
target_period <- "1990-1994"

themeo <- theme_classic() +
  theme(strip.background = element_blank(),
        panel.grid.major = element_line(colour = "transparent"),
        panel.grid.minor = element_line(colour = "transparent"),
        axis.line = element_blank(),
        axis.title = element_text(color = "black", size = 10),
        axis.ticks.length = unit(-0.1, "cm"),
        axis.ticks = element_line(color = "black", size = .25),
        axis.ticks.x = element_line(color = "black"),
        axis.ticks.y = element_line(color = "black"),
        panel.border = element_rect(colour = "black", fill = NA, size = .5),
        panel.spacing = unit(1, "lines"),
        legend.position = "bottom",
        legend.title = element_text(colour = "black", size = 10, face = "bold"),
        legend.text = element_text(color = "black", size = 8))
```

*POPULATION PROJECTIONS — 3/3 MITIGATION SCENARIO*

Two analyses:
  1. Immediate full 3/3 compliance — how long for the population to recover?
  2. Delayed implementation (5, 10, 20 year phase-in) — what is the cost of waiting?

Both use the scalar projection N(t+1) = lambda * N(t), following the homework structure.
For Part 1, lambda is fixed (one value per scenario).
For Part 2, lambda changes each year as coverage ramps up — computed explicitly and stored
in a coverage schedule table before the projection runs, so every input can be inspected.

---

**vital rates**
```{r vital-rates}
# Stage-specific survival rates (Pardo et al., 1980-2012)
s_Juv <- 0.846
s_EF  <- 0.913
s_ENB <- 0.943
s_ES  <- 0.895
s_IF  <- 0.920
s_Imm <- 0.921
s_IS  <- 0.892
s_PF  <- 0.935

# Stage-specific behavioral rates
b_EF  <- 0.746;  b_ENB <- 0.589;  b_ES  <- 0.016
b_IF  <- 0.744;  b_Imm <- 0.101;  b_IS  <- 0.011;  b_PF  <- 0.941

k_EF  <- 0.631;  k_ENB <- 0.675;  k_ES  <- 0.500
k_IF  <- 0.627;  k_Imm <- 0.585;  k_IS  <- 0.627;  k_PF  <- 0.791

r_EF  <- 0.973;  r_ENB <- 0.953;  r_ES  <- 0.292
r_IF  <- 0.948;  r_Imm <- 0.692;  r_IS  <- 0.459;  r_PF  <- 0.972

# Mitigation efficacy: 3/3 combined (tori lines + night setting + line weighting)
eff_tori_lines  <- 0.75
eff_night_set   <- 0.60
eff_line_weight <- 0.65
eff_3of3 <- 1 - (1 - eff_tori_lines) * (1 - eff_night_set) * (1 - eff_line_weight)

cat(sprintf("3/3 combined efficacy: %.1f%%\n", eff_3of3 * 100))
```

**load per-capita fishing mortality and natural mortality from 06a output**
```{r load-data}
# Per-capita FM and natural mortality saved by 06a (mortality_partition CSV)
mort_part <- read_csv(
  here(paste0("output/mortality_partition_", target_period, "_", date_sfx, ".csv")),
  show_col_types = FALSE)

# Extract per-capita DLL and PLL fishing mortality by stage group
dem_fm_juv <- mort_part %>% filter(stage == "Juv") %>% pull(fm_dll)
pel_fm_juv <- mort_part %>% filter(stage == "Juv") %>% pull(fm_pll)
dem_fm_imm <- mort_part %>% filter(stage == "Imm") %>% pull(fm_dll)
pel_fm_imm <- mort_part %>% filter(stage == "Imm") %>% pull(fm_pll)
dem_fm_fb  <- mort_part %>% filter(stage == "EF")  %>% pull(fm_dll)  # same for IF
pel_fm_fb  <- mort_part %>% filter(stage == "EF")  %>% pull(fm_pll)
dem_fm_sb  <- mort_part %>% filter(stage == "ES")  %>% pull(fm_dll)  # same for IS
pel_fm_sb  <- mort_part %>% filter(stage == "ES")  %>% pull(fm_pll)
dem_fm_nb  <- mort_part %>% filter(stage == "ENB") %>% pull(fm_dll)  # same for PF
pel_fm_nb  <- mort_part %>% filter(stage == "ENB") %>% pull(fm_pll)

# Natural mortality by stage (total mortality minus fishing mortality)
nm_juv <- mort_part %>% filter(stage == "Juv") %>% pull(nm)
nm_imm <- mort_part %>% filter(stage == "Imm") %>% pull(nm)
nm_efb <- mort_part %>% filter(stage == "EF")  %>% pull(nm)
nm_esb <- mort_part %>% filter(stage == "ES")  %>% pull(nm)
nm_enb <- mort_part %>% filter(stage == "ENB") %>% pull(nm)
nm_ifb <- mort_part %>% filter(stage == "IF")  %>% pull(nm)
nm_isb <- mort_part %>% filter(stage == "IS")  %>% pull(nm)
nm_pf  <- mort_part %>% filter(stage == "PF")  %>% pull(nm)

# Starting population: total Bird Island birds in the target period
dem_props <- read_csv(here("data/WAAL_dist_Clay2019/Dem_props copy.csv"),
                      show_col_types = FALSE) %>%
  mutate(Dem_class = str_replace(Dem_class, "J2\\+3", "J2J3"))

pop_period <- dem_props %>%
  filter(Period == target_period, Dem_class != "J1") %>%
  select(Dem_class, Pop)

N0 <- sum(pop_period$Pop)

cat(sprintf("N0 (%s): %.0f birds\n", target_period, N0))
```

**build matrices and extract lambdas**
```{r build-matrices}
# --- No mitigation matrix ---
# Survival = observed Pardo rates, which already encode fishing mortality
# as it existed in the baseline period (1990-1994)
# Stages: J2 J3 Imm4 Imm5 PB IS IF ES EF ENB Sabb
m_base <- matrix(data = c(
  0,     0,      0,      0,      0,                                          0.5*s_Juv,              0,                       0.5*s_Juv,              0,                       0,                         0,
  s_Juv, 0,      0,      0,      0,                                          0,                       0,                       0,                       0,                       0,                         0,
  0,     s_Juv,  0,      0,      0,                                          0,                       0,                       0,                       0,                       0,                         0,
  0,     0,      s_Imm,  0,      0,                                          0,                       0,                       0,                       0,                       0,                         0,
  0,     0,      0,      s_Imm,  s_Imm*r_Imm*(1-b_Imm)+s_Imm*(1-r_Imm),    0,                       0,                       0,                       0,                       0,                         0,
  0,     0,      0,      0,      s_Imm*r_Imm*b_Imm*k_Imm,                   0,                       0,                       0,                       0,                       0,                         0,
  0,     0,      0,      0,      s_Imm*r_Imm*b_Imm*(1-k_Imm),               0,                       0,                       0,                       0,                       0,                         0,
  0,     0,      0,      0,      0,                                          s_IS*r_IS*b_IS*k_IS,     s_IF*r_IF*b_IF*k_IF,     s_ES*r_ES*b_ES*k_ES,     s_EF*r_EF*b_EF*k_EF,     s_ENB*r_ENB*b_ENB*k_ENB,   s_PF*r_PF*b_PF*k_PF,
  0,     0,      0,      0,      0,                                          s_IS*r_IS*b_IS*(1-k_IS), s_IF*r_IF*b_IF*(1-k_IF), s_ES*r_ES*b_ES*(1-k_ES), s_EF*r_EF*b_EF*(1-k_EF), s_ENB*r_ENB*b_ENB*(1-k_ENB),s_PF*r_PF*b_PF*(1-k_PF),
  0,     0,      0,      0,      0,                                          s_IS*r_IS*(1-b_IS),      s_IF*r_IF*(1-b_IF),      s_ES*r_ES*(1-b_ES),      s_EF*r_EF*(1-b_EF),      s_ENB*r_ENB*(1-b_ENB),     s_PF*r_PF*(1-b_PF),
  0,     0,      0,      0,      0,                                          s_IS*(1-r_IS),           s_IF*(1-r_IF),           s_ES*(1-r_ES),           s_EF*(1-r_EF),           s_ENB*(1-r_ENB),           s_PF*(1-r_PF)
), nrow = 11, byrow = TRUE)

lambda_base <- as.numeric(eigen(m_base)$values[1])
SSD         <- as.numeric(eigen(m_base)$vectors[, 1] / sum(eigen(m_base)$vectors[, 1]))
n0_vec      <- SSD * N0

# --- Full 3/3 mitigation matrix ---
# Survival rates with fishing mortality reduced by eff_3of3 (tori lines + night setting + line weighting)
# ns_* = 1 - (FM_total * (1 - eff_3of3) + natural_mortality)
# FM values from 06a mortality partition output
ns_Juv <- 1 - ((dem_fm_juv + pel_fm_juv) * (1 - eff_3of3) + nm_juv)
ns_Imm <- 1 - ((dem_fm_imm + pel_fm_imm) * (1 - eff_3of3) + nm_imm)
ns_EF  <- 1 - ((dem_fm_fb  + pel_fm_fb)  * (1 - eff_3of3) + nm_efb)
ns_ES  <- 1 - ((dem_fm_sb  + pel_fm_sb)  * (1 - eff_3of3) + nm_esb)
ns_ENB <- 1 - ((dem_fm_nb  + pel_fm_nb)  * (1 - eff_3of3) + nm_enb)
ns_IF  <- 1 - ((dem_fm_fb  + pel_fm_fb)  * (1 - eff_3of3) + nm_ifb)
ns_IS  <- 1 - ((dem_fm_sb  + pel_fm_sb)  * (1 - eff_3of3) + nm_isb)
ns_PF  <- 1 - ((dem_fm_nb  + pel_fm_nb)  * (1 - eff_3of3) + nm_pf)

m_3of3 <- matrix(data = c(
  0,      0,       0,       0,       0,                                           0.5*ns_Juv,               0,                        0.5*ns_Juv,               0,                        0,                          0,
  ns_Juv, 0,       0,       0,       0,                                           0,                        0,                        0,                        0,                        0,                          0,
  0,      ns_Juv,  0,       0,       0,                                           0,                        0,                        0,                        0,                        0,                          0,
  0,      0,       ns_Imm,  0,       0,                                           0,                        0,                        0,                        0,                        0,                          0,
  0,      0,       0,       ns_Imm,  ns_Imm*r_Imm*(1-b_Imm)+ns_Imm*(1-r_Imm),   0,                        0,                        0,                        0,                        0,                          0,
  0,      0,       0,       0,       ns_Imm*r_Imm*b_Imm*k_Imm,                   0,                        0,                        0,                        0,                        0,                          0,
  0,      0,       0,       0,       ns_Imm*r_Imm*b_Imm*(1-k_Imm),               0,                        0,                        0,                        0,                        0,                          0,
  0,      0,       0,       0,       0,                                           ns_IS*r_IS*b_IS*k_IS,     ns_IF*r_IF*b_IF*k_IF,     ns_ES*r_ES*b_ES*k_ES,     ns_EF*r_EF*b_EF*k_EF,     ns_ENB*r_ENB*b_ENB*k_ENB,   ns_PF*r_PF*b_PF*k_PF,
  0,      0,       0,       0,       0,                                           ns_IS*r_IS*b_IS*(1-k_IS), ns_IF*r_IF*b_IF*(1-k_IF), ns_ES*r_ES*b_ES*(1-k_ES), ns_EF*r_EF*b_EF*(1-k_EF), ns_ENB*r_ENB*b_ENB*(1-k_ENB),ns_PF*r_PF*b_PF*(1-k_PF),
  0,      0,       0,       0,       0,                                           ns_IS*r_IS*(1-b_IS),      ns_IF*r_IF*(1-b_IF),      ns_ES*r_ES*(1-b_ES),      ns_EF*r_EF*(1-b_EF),      ns_ENB*r_ENB*(1-b_ENB),     ns_PF*r_PF*(1-b_PF),
  0,      0,       0,       0,       0,                                           ns_IS*(1-r_IS),           ns_IF*(1-r_IF),           ns_ES*(1-r_ES),           ns_EF*(1-r_EF),           ns_ENB*(1-r_ENB),           ns_PF*(1-r_PF)
), nrow = 11, byrow = TRUE)

lambda_3of3 <- as.numeric(eigen(m_3of3)$values[1])

cat(sprintf("Lambda — no mitigation: %.6f\n", lambda_base))
cat(sprintf("Lambda — full 3/3:      %.6f\n", lambda_3of3))
cat(sprintf("Difference:             %+.6f\n", lambda_3of3 - lambda_base))
cat(sprintf("N0 = %.0f birds\n", N0))


cat(sprintf("ns_Juv: %.4f  (was s_Juv: %.4f, change: %+.4f)\n", ns_Juv, s_Juv, ns_Juv - s_Juv))
  cat(sprintf("ns_Imm: %.4f  (was s_Imm: %.4f, change: %+.4f)\n", ns_Imm, s_Imm, ns_Imm - s_Imm))
  cat(sprintf("ns_EF:  %.4f  (was s_EF:  %.4f, change: %+.4f)\n", ns_EF,  s_EF,  ns_EF  - s_EF))
```

---

*PART 1 — IMMEDIATE FULL COMPLIANCE*

How long does it take the population to recover if 3/3 mitigation is implemented
across all fisheries immediately at 100% coverage?

**Step 1: set up output storage**
```{r part1-setup}
years <- 100

# One column for N, rows = years + 1 (year 0 through year 100)
output_base <- matrix(nrow = years + 1, ncol = 1) %>% as.data.frame()
output_imm  <- matrix(nrow = years + 1, ncol = 1) %>% as.data.frame()

# Year 0: starting population
output_base[1, 1] <- N0
output_imm[1, 1]  <- N0

cat(sprintf("N0 = %.0f  |  storage: %d rows x 1 col\n", N0, nrow(output_base)))
```

**Step 3: project — N(t+1) = lambda * N(t)**
```{r part1-project}
# No mitigation: same lambda every year
for (k in 1:years) {
  output_base[k + 1, 1] <- lambda_base * output_base[k, 1]
}

# Immediate full 3/3: same lambda every year
for (k in 1:years) {
  output_imm[k + 1, 1] <- lambda_3of3 * output_imm[k, 1]
}

# Combine and label
proj_part1 <- bind_rows(
  output_base %>% mutate(scenario = "No mitigation (baseline)"),
  output_imm  %>% mutate(scenario = "Immediate 3/3 (100% coverage)")
) %>%
  mutate(year     = rep(0:years, 2),
         N        = V1,
         scenario = factor(scenario,
           levels = c("No mitigation (baseline)", "Immediate 3/3 (100% coverage)"))) %>%
  select(year, N, scenario)

# Inspect first few years to verify
cat("=== First 6 years — no mitigation ===\n")
print(head(proj_part1 %>% filter(scenario == "No mitigation (baseline)"), 6))
cat("\n=== First 6 years — immediate 3/3 ===\n")
print(head(proj_part1 %>% filter(scenario == "Immediate 3/3 (100% coverage)"), 6))
```

**recovery milestones**
```{r part1-milestones}
# Years for the immediate-3/3 mean trajectory to reach growth milestones above N0
imm_traj <- proj_part1 %>% filter(scenario == "Immediate 3/3 (100% coverage)")

# helper: return NA_integer_ if milestone not reached within projection horizon
first_yr <- function(traj, threshold) {
  x <- traj %>% filter(N >= threshold) %>% slice(1) %>% pull(year)
  if (length(x) == 0) NA_integer_ else x
}

fmt_yr <- function(x) ifelse(is.na(x), paste0(">", years), as.character(x))

milestone_10  <- first_yr(imm_traj, N0 * 1.10)
milestone_25  <- first_yr(imm_traj, N0 * 1.25)
milestone_50  <- first_yr(imm_traj, N0 * 1.50)
milestone_dbl <- first_yr(imm_traj, N0 * 2.00)

cat(sprintf("Years to +10%% above N0:  %s\n", fmt_yr(milestone_10)))
cat(sprintf("Years to +25%% above N0:  %s\n", fmt_yr(milestone_25)))
cat(sprintf("Years to +50%% above N0:  %s\n", fmt_yr(milestone_50)))
cat(sprintf("Years to double N0:       %s\n", fmt_yr(milestone_dbl)))
```

**observed population totals from dem_props (validation)**
```{r dem-obs}
# 5-year period averages from Clay 2019 demographic data
# Excluding J1 to match N0 calculation; plotted at start year of each period
dem_obs <- dem_props %>%
  filter(Dem_class != "J1",
         Period %in% c("1990-1994", "1995-1999", "2000-2004", "2005-2009")) %>%
  group_by(Period) %>%
  summarise(N_obs = sum(Pop), .groups = "drop") %>%
  mutate(year = c(0, 5, 10, 15))

# Adults = SB + FB + NB;  Immatures = IMM + J2+3
dem_obs_split <- dem_props %>%
  filter(Dem_class != "J1",
         Period %in% c("1990-1994", "1995-1999", "2000-2004", "2005-2009")) %>%
  mutate(group = ifelse(Dem_class %in% c("IMM", "J2+3"), "Immatures", "Adults"),
         year  = case_when(
           Period == "1990-1994" ~ 0,
           Period == "1995-1999" ~ 5,
           Period == "2000-2004" ~ 10,
           Period == "2005-2009" ~ 15
         )) %>%
  group_by(Period, year, group) %>%
  summarise(N_obs = sum(Pop), .groups = "drop")

print(dem_obs_split)

# Bird Island breeding pair counts (Poncet et al. survey data)
# Years are end-year of austral season; x-axis is years since 1990 (start of baseline period)
# Scale factor converts breeding pairs to total N for plotting on primary y-axis:
# interpolating between 1984 (1370 pairs) and 2004 (948 pairs) gives ~1200 pairs at 1990,
# so scale_factor = N0 / 1200
pairs_scale <- N0 / 1200

bird_obs <- tibble(
  season = c("1983/84", "2003/04", "2014/15", "2023/24"),
  year   = c(-6, 14, 25, 34),       # years relative to 1990 baseline start
  pairs  = c(1370, 948, 772, 699),
  N_scaled = pairs * pairs_scale     # scaled to primary y-axis for plotting
)

cat(sprintf("Breeding pair scale factor: %.2f (N per pair)\n", pairs_scale))
print(bird_obs)
```

**plot Part 1**
```{r plot-part1, fig.width = 9, fig.height = 5}
milestones_df <- tibble(
  year  = c(milestone_10, milestone_25, milestone_50, milestone_dbl),
  label = c("+10%", "+25%", "+50%", "\u00D72"),
  N     = c(N0 * 1.10, N0 * 1.25, N0 * 1.50, N0 * 2.0)
) %>%
  filter(!is.na(year), year <= years)

p_part1 <- ggplot(proj_part1, aes(x = year, y = N,
                                   color = scenario, linetype = scenario)) +
  geom_line(linewidth = 0.9) +
  geom_hline(yintercept = N0, color = "grey50", linetype = "dashed", linewidth = 0.4) +
  geom_point(data = dem_obs,
             aes(x = year, y = N_obs),
             inherit.aes = FALSE, color = "black", size = 3, shape = 17) +
  geom_text(data = dem_obs,
            aes(x = year, y = N_obs, label = Period),
            inherit.aes = FALSE, color = "black", size = 2.5, vjust = -0.8) +
  geom_point(data = bird_obs,
             aes(x = year, y = N_scaled),
             inherit.aes = FALSE, color = "#e6550d", size = 3, shape = 16) +
  geom_line(data = bird_obs,
            aes(x = year, y = N_scaled),
            inherit.aes = FALSE, color = "#e6550d", linewidth = 0.5, linetype = "dotted") +
  geom_text(data = bird_obs,
            aes(x = year, y = N_scaled, label = season),
            inherit.aes = FALSE, color = "#e6550d", size = 2.5, vjust = -0.8) +
  geom_point(data = milestones_df,
             aes(x = year, y = N),
             inherit.aes = FALSE, color = "#253494", size = 2.5) +
  geom_text(data = milestones_df,
            aes(x = year, y = N,
                label = paste0(label, "\n(yr ", year, ")")),
            inherit.aes = FALSE,
            color = "#253494", size = 2.8, hjust = -0.15, vjust = 0.5) +
  scale_color_manual(values = c("No mitigation (baseline)"      = "#d73027",
                                 "Immediate 3/3 (100% coverage)" = "#253494")) +
  scale_linetype_manual(values = c("No mitigation (baseline)"      = "dashed",
                                    "Immediate 3/3 (100% coverage)" = "solid")) +
  scale_x_continuous(breaks = seq(-10, years, 10), limits = c(-10, years)) +
  scale_y_continuous(labels = scales::comma,
                     breaks = scales::pretty_breaks(n = 6),
                     sec.axis = sec_axis(~ . / pairs_scale,
                                         name = "Breeding pairs (Bird Island)",
                                         labels = scales::comma)) +
  labs(x = "Years since 1990 baseline",
       y = "Population (N individuals)",
       color = NULL, linetype = NULL,
       title = paste0("Population projection vs observed breeding pairs — ",
                      target_period, " baseline"),
       subtitle = paste0("\u03BB (no mitigation) = ", round(lambda_base, 5),
                         ";  \u03BB (full 3/3) = ", round(lambda_3of3, 5),
                         ";  N\u2080 = ", scales::comma(round(N0)), " birds")) +
  themeo +
  theme(legend.position = "bottom")
p_part1
```

**plot Part 1b — projection vs dem_props age class totals**
```{r plot-part1-obs, fig.width = 9, fig.height = 5}
# Adults = SB + FB + NB;  Immatures = IMM + J2+3
# Secondary y-axis scaled to breeding pairs equivalent using same pairs_scale
p_part1_obs <- ggplot(proj_part1, aes(x = year, y = N,
                                       color = scenario, linetype = scenario)) +
  geom_line(linewidth = 0.9) +
  geom_hline(yintercept = N0, color = "grey50", linetype = "dashed", linewidth = 0.4) +
  geom_line(data = dem_obs_split,
            aes(x = year, y = N_obs, group = group, linetype = group),
            inherit.aes = FALSE, color = "black", linewidth = 0.7) +
  geom_point(data = dem_obs_split,
             aes(x = year, y = N_obs, shape = group),
             inherit.aes = FALSE, color = "black", size = 3) +
  geom_point(data = dem_obs,
             aes(x = year, y = N_obs),
             inherit.aes = FALSE, color = "black", size = 3, shape = 17) +
  geom_text(data = dem_obs,
            aes(x = year, y = N_obs, label = Period),
            inherit.aes = FALSE, color = "black", size = 2.5, vjust = -0.8) +
  scale_color_manual(values = c("No mitigation (baseline)"      = "#d73027",
                                 "Immediate 3/3 (100% coverage)" = "#253494")) +
  scale_linetype_manual(values = c("No mitigation (baseline)"      = "dashed",
                                    "Immediate 3/3 (100% coverage)" = "solid",
                                    "Adults"                        = "solid",
                                    "Immatures"                     = "dashed"),
                        guide = "none") +
  scale_shape_manual(values = c("Adults" = 16, "Immatures" = 1),
                     name = "Observed age class") +
  scale_x_continuous(breaks = seq(0, years, 10), limits = c(0, years)) +
  scale_y_continuous(labels = scales::comma,
                     breaks = scales::pretty_breaks(n = 6),
                     sec.axis = sec_axis(~ . / N0 * 100,
                                         name = "% of 1990\u20131994 baseline",
                                         labels = function(x) paste0(round(x), "%"))) +
  labs(x = "Years since 1990 baseline",
       y = "Population (N individuals)",
       color = NULL,
       title = paste0("Population projection vs observed age class totals — ",
                      target_period, " baseline"),
       subtitle = "Black triangles = total; filled circles = adults; open circles = immatures (Clay 2019)") +
  themeo +
  theme(legend.position = "bottom")
p_part1_obs
```

---

*PART 2 — TO BE ADDED*

---

**age class trends from dem_props (Clay 2019)**
```{r plot-age-trends, fig.width = 8, fig.height = 5}
# Use the 4 survey periods only (exclude the 1990-2009 aggregate and J1)
# Mid-year of each 5-year period used for x-axis placement
age_trends <- dem_props %>%
  filter(Dem_class != "J1",
         Period %in% c("1990-1994", "1995-1999", "2000-2004", "2005-2009")) %>%
  mutate(
    year_mid = case_when(
      Period == "1990-1994" ~ 1992,
      Period == "1995-1999" ~ 1997,
      Period == "2000-2004" ~ 2002,
      Period == "2005-2009" ~ 2007
    ),
    Dem_class = factor(Dem_class,
                       levels = c("SB", "FB", "NB", "IMM", "J2+3"),
                       labels = c("Successful breeders", "Failed breeders",
                                  "Non-breeders", "Immatures", "Juveniles (J2+3)"))
  )

p_age_trends <- ggplot(age_trends, aes(x = year_mid, y = Pop,
                                        color = Dem_class, group = Dem_class)) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 2.5) +
  scale_color_brewer(palette = "Dark2") +
  scale_x_continuous(breaks = c(1992, 1997, 2002, 2007),
                     labels = c("1990-94", "1995-99", "2000-04", "2005-09")) +
  scale_y_continuous(labels = scales::comma) +
  labs(x = NULL,
       y = "Population (N individuals)",
       color = "Age class",
       title = "Bird Island WAAL population by age class — Clay et al. 2019",
       subtitle = "5-year period averages; excludes J1") +
  themeo +
  theme(legend.position = "right")
p_age_trends
```

## Save figures

```{r save-figures}
dir.create(here("output/figures"), showWarnings = FALSE, recursive = TRUE)

fig_path <- function(name) here("output/figures",
                                 paste0(name, "_", target_period, "_", date_sfx, ".png"))

ggsave(fig_path("08_proj_part1_breeding_pairs"),  p_part1,         width = 9,  height = 5, dpi = 200)
ggsave(fig_path("08_proj_part1_age_class_obs"),  p_part1_obs,     width = 9,  height = 5, dpi = 200)
ggsave(fig_path("08_age_class_trends"),          p_age_trends,    width = 8,  height = 5, dpi = 200)
message("3 figures saved to output/figures/ with suffix ", date_sfx)
```
