---
title: "WAAL Population Projections — Summary"
output: html_document
---

```{r setup, echo = FALSE, include = FALSE}
library(here)
library(tidyverse)
library(RColorBrewer)


input_sfx     <- "2026-02-25"                    # date 06a was last run — must match saved CSV
output_sfx    <- format(Sys.Date(), "%Y-%m-%d")  # today's date, used when saving figures
target_period <- "1990-1994" #baseline reference period

themeo <- theme_classic() + #custom plotting theme
  theme(strip.background = element_blank(),
        panel.grid.major = element_line(colour = "transparent"),
        panel.grid.minor = element_line(colour = "transparent"),
        axis.line = element_blank(),
        axis.title = element_text(color = "black", size = 10),
        axis.ticks.length = unit(-0.1, "cm"),
        axis.ticks = element_line(color = "black", size = .25),
        axis.ticks.x = element_line(color = "black"),
        axis.ticks.y = element_line(color = "black"),
        panel.border = element_rect(colour = "black", fill = NA, size = .5),
        panel.spacing = unit(1, "lines"),
        legend.position = "bottom",
        legend.title = element_text(colour = "black", size = 10, face = "bold"),
        legend.text = element_text(color = "black", size = 8))
```

*WAAL POPULATION PROJECTIONS — SUMMARY SCRIPT*

Three projection scenarios across 100-year and 500-years:
  1. Immediate full 3/3 compliance — how long for population to recover?
  2. Phase-in (5, 10, 20 years) — cost of gradual implementation?
  3. Delay then implement — cost of waiting to act?


**vital rates**
```{r vital-rates}
# Stage-specific survival rates (Pardo et al., 1980-2012)
s_Juv <- 0.846
s_EF  <- 0.913
s_ENB <- 0.943
s_ES  <- 0.895
s_IF  <- 0.920
s_Imm <- 0.921
s_IS  <- 0.892
s_PF  <- 0.935

# Stage-specific behavioral rates
b_EF  <- 0.746;  b_ENB <- 0.589;  b_ES  <- 0.016
b_IF  <- 0.744;  b_Imm <- 0.101;  b_IS  <- 0.011;  b_PF  <- 0.941

k_EF  <- 0.631;  k_ENB <- 0.675;  k_ES  <- 0.500
k_IF  <- 0.627;  k_Imm <- 0.585;  k_IS  <- 0.627;  k_PF  <- 0.791

r_EF  <- 0.973;  r_ENB <- 0.953;  r_ES  <- 0.292
r_IF  <- 0.948;  r_Imm <- 0.692;  r_IS  <- 0.459;  r_PF  <- 0.972

# Mitigation efficacy: 3/3 combined (tori lines + night setting + line weighting)
# numbers pulled from the literature and will need citations and confirmation
eff_tori_lines  <- 0.75
eff_night_set   <- 0.60
eff_line_weight <- 0.65
eff_3of3 <- 1 - (1 - eff_tori_lines) * (1 - eff_night_set) * (1 - eff_line_weight)
```

**load per-capita fishing mortality and natural mortality from 06a output**
```{r load-data}
# Per-capita FM and natural mortality saved by script 06a (mortality_partition CSV)
mort_part <- read_csv(
  here(paste0("output/mortality_partition_", target_period, "_", input_sfx, ".csv")),
  show_col_types = FALSE)

mort_part

# Extract per-capita DLL and PLL fishing mortality by stage group
dem_fm_juv <- mort_part %>% filter(stage == "Juv") %>% pull(fm_dll)
pel_fm_juv <- mort_part %>% filter(stage == "Juv") %>% pull(fm_pll)
dem_fm_imm <- mort_part %>% filter(stage == "Imm") %>% pull(fm_dll)
pel_fm_imm <- mort_part %>% filter(stage == "Imm") %>% pull(fm_pll)
dem_fm_fb  <- mort_part %>% filter(stage == "EF")  %>% pull(fm_dll)  # same for IF
pel_fm_fb  <- mort_part %>% filter(stage == "EF")  %>% pull(fm_pll)
dem_fm_sb  <- mort_part %>% filter(stage == "ES")  %>% pull(fm_dll)  # same for IS
pel_fm_sb  <- mort_part %>% filter(stage == "ES")  %>% pull(fm_pll)
dem_fm_nb  <- mort_part %>% filter(stage == "ENB") %>% pull(fm_dll)  # same for PF
pel_fm_nb  <- mort_part %>% filter(stage == "ENB") %>% pull(fm_pll)

# Natural mortality by stage (total mortality minus fishing mortality)
nm_juv <- mort_part %>% filter(stage == "Juv") %>% pull(nm)
nm_imm <- mort_part %>% filter(stage == "Imm") %>% pull(nm)
nm_efb <- mort_part %>% filter(stage == "EF")  %>% pull(nm)
nm_esb <- mort_part %>% filter(stage == "ES")  %>% pull(nm)
nm_enb <- mort_part %>% filter(stage == "ENB") %>% pull(nm)
nm_ifb <- mort_part %>% filter(stage == "IF")  %>% pull(nm)
nm_isb <- mort_part %>% filter(stage == "IS")  %>% pull(nm)
nm_pf  <- mort_part %>% filter(stage == "PF")  %>% pull(nm)

# Starting population: total Bird Island birds in the target period
dem_props <- read_csv(here("data/WAAL_dist_Clay2019/Dem_props copy.csv"),
                      show_col_types = FALSE) %>%
  mutate(Dem_class = str_replace(Dem_class, "J2\\+3", "J2J3"))

pop_period <- dem_props %>%
  filter(Period == target_period, Dem_class != "J1") %>%
  select(Dem_class, Pop)

N0 <- sum(pop_period$Pop)
```

**build matrices and extract lambdas**
```{r build-matrices}
# --- No mitigation matrix ---
# Survival = observed Pardo rates, which already encode fishing mortality
# as it existed in the baseline period (1990-1994)
# Stages: J2 J3 Imm4 Imm5 PB IS IF ES EF ENB Sabb
m_base <- matrix(data = c(
  0,     0,      0,      0,      0,                                          0.5*s_Juv,              0,                       0.5*s_Juv,              0,                       0,                         0,
  s_Juv, 0,      0,      0,      0,                                          0,                       0,                       0,                       0,                       0,                         0,
  0,     s_Juv,  0,      0,      0,                                          0,                       0,                       0,                       0,                       0,                         0,
  0,     0,      s_Imm,  0,      0,                                          0,                       0,                       0,                       0,                       0,                         0,
  0,     0,      0,      s_Imm,  s_Imm*r_Imm*(1-b_Imm)+s_Imm*(1-r_Imm),    0,                       0,                       0,                       0,                       0,                         0,
  0,     0,      0,      0,      s_Imm*r_Imm*b_Imm*k_Imm,                   0,                       0,                       0,                       0,                       0,                         0,
  0,     0,      0,      0,      s_Imm*r_Imm*b_Imm*(1-k_Imm),               0,                       0,                       0,                       0,                       0,                         0,
  0,     0,      0,      0,      0,                                          s_IS*r_IS*b_IS*k_IS,     s_IF*r_IF*b_IF*k_IF,     s_ES*r_ES*b_ES*k_ES,     s_EF*r_EF*b_EF*k_EF,     s_ENB*r_ENB*b_ENB*k_ENB,   s_PF*r_PF*b_PF*k_PF,
  0,     0,      0,      0,      0,                                          s_IS*r_IS*b_IS*(1-k_IS), s_IF*r_IF*b_IF*(1-k_IF), s_ES*r_ES*b_ES*(1-k_ES), s_EF*r_EF*b_EF*(1-k_EF), s_ENB*r_ENB*b_ENB*(1-k_ENB),s_PF*r_PF*b_PF*(1-k_PF),
  0,     0,      0,      0,      0,                                          s_IS*r_IS*(1-b_IS),      s_IF*r_IF*(1-b_IF),      s_ES*r_ES*(1-b_ES),      s_EF*r_EF*(1-b_EF),      s_ENB*r_ENB*(1-b_ENB),     s_PF*r_PF*(1-b_PF),
  0,     0,      0,      0,      0,                                          s_IS*(1-r_IS),           s_IF*(1-r_IF),           s_ES*(1-r_ES),           s_EF*(1-r_EF),           s_ENB*(1-r_ENB),           s_PF*(1-r_PF)
), nrow = 11, byrow = TRUE)

lambda_base <- as.numeric(eigen(m_base)$values[1])
SSD         <- as.numeric(eigen(m_base)$vectors[, 1] / sum(eigen(m_base)$vectors[, 1]))
n0_vec      <- SSD * N0

m_base
lambda_base
SSD
n0_vec


# --- Full 3/3 mitigation matrix ---
# ns_* = 1 - (FM_total * (1 - eff_3of3) + natural_mortality)
ns_Juv <- 1 - ((dem_fm_juv + pel_fm_juv) * (1 - eff_3of3) + nm_juv)
ns_Imm <- 1 - ((dem_fm_imm + pel_fm_imm) * (1 - eff_3of3) + nm_imm)
ns_EF  <- 1 - ((dem_fm_fb  + pel_fm_fb)  * (1 - eff_3of3) + nm_efb)
ns_ES  <- 1 - ((dem_fm_sb  + pel_fm_sb)  * (1 - eff_3of3) + nm_esb)
ns_ENB <- 1 - ((dem_fm_nb  + pel_fm_nb)  * (1 - eff_3of3) + nm_enb)
ns_IF  <- 1 - ((dem_fm_fb  + pel_fm_fb)  * (1 - eff_3of3) + nm_ifb)
ns_IS  <- 1 - ((dem_fm_sb  + pel_fm_sb)  * (1 - eff_3of3) + nm_isb)
ns_PF  <- 1 - ((dem_fm_nb  + pel_fm_nb)  * (1 - eff_3of3) + nm_pf)

m_3of3 <- matrix(data = c(
  0,      0,       0,       0,       0,                                           0.5*ns_Juv,               0,                        0.5*ns_Juv,               0,                        0,                          0,
  ns_Juv, 0,       0,       0,       0,                                           0,                        0,                        0,                        0,                        0,                          0,
  0,      ns_Juv,  0,       0,       0,                                           0,                        0,                        0,                        0,                        0,                          0,
  0,      0,       ns_Imm,  0,       0,                                           0,                        0,                        0,                        0,                        0,                          0,
  0,      0,       0,       ns_Imm,  ns_Imm*r_Imm*(1-b_Imm)+ns_Imm*(1-r_Imm),   0,                        0,                        0,                        0,                        0,                          0,
  0,      0,       0,       0,       ns_Imm*r_Imm*b_Imm*k_Imm,                   0,                        0,                        0,                        0,                        0,                          0,
  0,      0,       0,       0,       ns_Imm*r_Imm*b_Imm*(1-k_Imm),               0,                        0,                        0,                        0,                        0,                          0,
  0,      0,       0,       0,       0,                                           ns_IS*r_IS*b_IS*k_IS,     ns_IF*r_IF*b_IF*k_IF,     ns_ES*r_ES*b_ES*k_ES,     ns_EF*r_EF*b_EF*k_EF,     ns_ENB*r_ENB*b_ENB*k_ENB,   ns_PF*r_PF*b_PF*k_PF,
  0,      0,       0,       0,       0,                                           ns_IS*r_IS*b_IS*(1-k_IS), ns_IF*r_IF*b_IF*(1-k_IF), ns_ES*r_ES*b_ES*(1-k_ES), ns_EF*r_EF*b_EF*(1-k_EF), ns_ENB*r_ENB*b_ENB*(1-k_ENB),ns_PF*r_PF*b_PF*(1-k_PF),
  0,      0,       0,       0,       0,                                           ns_IS*r_IS*(1-b_IS),      ns_IF*r_IF*(1-b_IF),      ns_ES*r_ES*(1-b_ES),      ns_EF*r_EF*(1-b_EF),      ns_ENB*r_ENB*(1-b_ENB),     ns_PF*r_PF*(1-b_PF),
  0,      0,       0,       0,       0,                                           ns_IS*(1-r_IS),           ns_IF*(1-r_IF),           ns_ES*(1-r_ES),           ns_EF*(1-r_EF),           ns_ENB*(1-r_ENB),           ns_PF*(1-r_PF)
), nrow = 11, byrow = TRUE)

lambda_3of3 <- as.numeric(eigen(m_3of3)$values[1])

m_3of3
lambda_3of3

tibble(scenario  = c("No mitigation (baseline)", "Full 3/3 mitigation"),
       lambda    = c(lambda_base, lambda_3of3))
```

---

**observed population data**
```{r obs-data}
# 5-year period totals from Clay 2019 demographic data (excludes J1)
dem_obs <- dem_props %>%
  filter(Dem_class != "J1",
         Period %in% c("1990-1994", "1995-1999", "2000-2004", "2005-2009")) %>%
  group_by(Period) %>%
  summarise(N_obs = sum(Pop), .groups = "drop") %>%
  mutate(year = c(0, 5, 10, 15))

# Split into adults (SB + FB + NB) and immatures (IMM + J2+3) for stage comparison plots
dem_obs_split <- dem_props %>%
  filter(Dem_class != "J1",
         Period %in% c("1990-1994", "1995-1999", "2000-2004", "2005-2009")) %>%
  mutate(group = ifelse(Dem_class %in% c("IMM", "J2+3"), "Immatures", "Adults"),
         year  = case_when(
           Period == "1990-1994" ~ 0,
           Period == "1995-1999" ~ 5,
           Period == "2000-2004" ~ 10,
           Period == "2005-2009" ~ 15
         )) %>%
  group_by(Period, year, group) %>%
  summarise(N_obs = sum(Pop), .groups = "drop")

# Bird Island breeding pair counts (Mackley et al. census study in press, from Richard)
# Scale factor converts breeding pairs to total N for plotting / comparisons
pairs_scale <- N0 / 1200

bird_obs <- tibble(
  season   = c("1983/84", "2003/04", "2014/15", "2023/24"),
  year     = c(-6, 14, 25, 34),
  pairs    = c(1370, 948, 772, 699),
  N_scaled = pairs * pairs_scale
)

print(bird_obs)

# Validate pairs_scale against the SSD
# Active breeders counted in pair surveys = IS + IF + ES + EF (n0_vec indices 6-9)
# Divide by 2: each pair = 2 individuals
n_active_breeders <- sum(n0_vec[6:9])
pairs_from_SSD    <- n_active_breeders / 2
pct_breeding      <- n_active_breeders / N0 * 100
scale_from_SSD    <- N0 / pairs_from_SSD

tibble(
  quantity = c("SSD-implied breeding pairs at N0",
               "Assumed baseline pairs (pairs_scale denominator)",
               "Active breeders as % of N0",
               "N0 / SSD-implied pairs  (scale from SSD)",
               "N0 / 1200               (pairs_scale used)"),
  value    = c(round(pairs_from_SSD),
               1200,
               round(pct_breeding, 1),
               round(scale_from_SSD, 2),
               round(pairs_scale, 2))
) %>% print(n = Inf)
# If SSD-implied pairs is close to 1200, pairs_scale is internally consistent.
# If they differ substantially, the 1200 baseline assumption needs revisiting.
# they differ, so likely the pop is not actually at SSD, which makes sense. NOt sure if this matters for our matrix assumptions. 
#will still keep these numbesr for now as they are from actual census data and still useful for comparison
```

---

*PART 1 — USE EXISTING LAMBDA FROM PREVIOUS CODE: 100-YEARS AND 500-YEARS*

N(t+1) = lambda * N(t). Lambda is fixed per scenario 
Used to establish baseline recovery trajectory

**project**
```{r part1-project}
years     <- 100
yearslong <- 500

stage_names <- c("J2", "J3", "Imm4", "Imm5", "PB", "IS", "IF", "ES", "EF", "ENB", "Sabb")

#no mitigation vs immediate mitigation, ove 100 and 500 year time periods
output_base      <- matrix(nrow = years     + 1, ncol = 1) %>% as.data.frame()
output_imm       <- matrix(nrow = years     + 1, ncol = 1) %>% as.data.frame()
output_base_long <- matrix(nrow = yearslong + 1, ncol = 1) %>% as.data.frame()
output_imm_long  <- matrix(nrow = yearslong + 1, ncol = 1) %>% as.data.frame()

output_base[1, 1]      <- N0
output_imm[1, 1]       <- N0
output_base_long[1, 1] <- N0
output_imm_long[1, 1]  <- N0

for (k in 1:years) {
  output_base[k + 1, 1] <- lambda_base * output_base[k, 1]
  output_imm[k + 1, 1]  <- lambda_3of3 * output_imm[k, 1]
}

for (k in 1:yearslong) {
  output_base_long[k + 1, 1] <- lambda_base * output_base_long[k, 1]
  output_imm_long[k + 1, 1]  <- lambda_3of3 * output_imm_long[k, 1]
}


#combine output to use later in plotting
proj_part1 <- bind_rows(
  output_base %>% mutate(scenario = "No mitigation (baseline)"),
  output_imm  %>% mutate(scenario = "Immediate 3/3 (100% coverage)")
) %>%
  mutate(year     = rep(0:years, 2),
         N        = V1,
         scenario = factor(scenario,
           levels = c("No mitigation (baseline)", "Immediate 3/3 (100% coverage)"))) %>%
  select(year, N, scenario)

proj_part1_long <- bind_rows(
  output_base_long %>% mutate(scenario = "No mitigation (baseline)"),
  output_imm_long  %>% mutate(scenario = "Immediate 3/3 (100% coverage)")
) %>%
  mutate(year     = rep(0:yearslong, 2),
         N        = V1,
         scenario = factor(scenario,
           levels = c("No mitigation (baseline)", "Immediate 3/3 (100% coverage)"))) %>%
  select(year, N, scenario)
```


**plot 1a — 100-year projection with observed breeding pairs**
```{r plot-1a, fig.width = 9, fig.height = 5}

p_1a <- ggplot(proj_part1, aes(x = year, y = N, color = scenario, linetype = scenario)) +
  geom_line(linewidth = 0.9) +
  geom_hline(yintercept = N0, color = "grey50", linetype = "dashed", linewidth = 0.4) +
  geom_point(data = dem_obs,
             aes(x = year, y = N_obs),
             inherit.aes = FALSE, color = "black", size = 3, shape = 17) +
  geom_text(data = dem_obs,
            aes(x = year, y = N_obs, label = Period),
            inherit.aes = FALSE, color = "black", size = 2.5, vjust = -0.8) +
  geom_point(data = bird_obs,
             aes(x = year, y = N_scaled),
             inherit.aes = FALSE, color = "#e6550d", size = 3, shape = 16) +
  geom_line(data = bird_obs,
            aes(x = year, y = N_scaled),
            inherit.aes = FALSE, color = "#e6550d", linewidth = 0.5, linetype = "dotted") +
  geom_text(data = bird_obs,
            aes(x = year, y = N_scaled, label = season),
            inherit.aes = FALSE, color = "#e6550d", size = 2.5, vjust = -0.8) +
  scale_color_manual(values = c("No mitigation (baseline)"      = "#d73027",
                                 "Immediate 3/3 (100% coverage)" = "#253494")) +
  scale_linetype_manual(values = c("No mitigation (baseline)"      = "dashed",
                                    "Immediate 3/3 (100% coverage)" = "solid")) +
  scale_x_continuous(breaks = seq(-10, years, 10), limits = c(-10, years)) +
  scale_y_continuous(labels = scales::comma,
                     breaks = scales::pretty_breaks(n = 6),
                     sec.axis = sec_axis(~ . / pairs_scale,
                                         name = "Breeding pairs (Bird Island)",
                                         labels = scales::comma)) +
  labs(x = "Years since 1990 baseline",
       y = "Population (N individuals)",
       color = NULL, linetype = NULL,
       subtitle = paste0("\u03BB (no mitigation) = ", round(lambda_base, 4),
                         ";  \u03BB (full 3/3) = ", round(lambda_3of3, 4),
                         ";  N\u2080 = ", scales::comma(round(N0)),
                         " | black triangles = Clay 2019 population totals;\n orange dots = census breeding pairs")) +
  themeo + theme(legend.position = "bottom")
p_1a
```

**plot 1b — 100-year projection with observed age class lines (adults / immatures)**
```{r plot-1b, fig.width = 9, fig.height = 5}
p_1b <- ggplot(proj_part1, aes(x = year, y = N, color = scenario, linetype = scenario)) +
  geom_line(linewidth = 0.9) +
  geom_hline(yintercept = N0, color = "grey50", linetype = "dashed", linewidth = 0.4) +
  geom_line(data = dem_obs_split,
            aes(x = year, y = N_obs, group = group, linetype = group),
            inherit.aes = FALSE, color = "black", linewidth = 0.7) +
  geom_point(data = dem_obs_split,
             aes(x = year, y = N_obs, shape = group),
             inherit.aes = FALSE, color = "black", size = 3) +
  geom_point(data = dem_obs,
             aes(x = year, y = N_obs),
             inherit.aes = FALSE, color = "black", size = 3, shape = 17) +
  geom_text(data = dem_obs,
            aes(x = year, y = N_obs, label = Period),
            inherit.aes = FALSE, color = "black", size = 2.5, vjust = -0.8) +
  scale_color_manual(values = c("No mitigation (baseline)"      = "#d73027",
                                 "Immediate 3/3 (100% coverage)" = "#253494")) +
  scale_linetype_manual(values = c("No mitigation (baseline)"      = "dashed",
                                    "Immediate 3/3 (100% coverage)" = "solid",
                                    "Adults"                        = "solid",
                                    "Immatures"                     = "dashed"),
                        guide = "none") +
  scale_shape_manual(values = c("Adults" = 16, "Immatures" = 1),
                     name = "Observed age class") +
  scale_x_continuous(breaks = seq(0, years, 10), limits = c(0, years)) +
  scale_y_continuous(labels = scales::comma,
                     breaks = scales::pretty_breaks(n = 6),
                     sec.axis = sec_axis(~ . / N0 * 100,
                                         name = "% of 1990\u20131994 baseline",
                                         labels = function(x) paste0(round(x), "%"))) +
  labs(x = "Years since 1990 baseline",
       y = "Population (N individuals)",
       color = NULL,
       subtitle = "Black triangles = total population; \nfilled circles = adults; open circles = immatures (Clay 2019)") +
  themeo + theme(legend.position = "bottom")
p_1b
```

**plot 1c — 500-year projection with observed breeding pairs**
```{r plot-1c, fig.width = 9, fig.height = 5}

p_1c <- ggplot(proj_part1_long, aes(x = year, y = N, color = scenario, linetype = scenario)) +
  geom_line(linewidth = 0.9) +
  geom_hline(yintercept = N0, color = "grey50", linetype = "dashed", linewidth = 0.4) +
  geom_point(data = bird_obs,
             aes(x = year, y = N_scaled),
             inherit.aes = FALSE, color = "#e6550d", size = 3, shape = 16) +
  geom_text(data = bird_obs,
            aes(x = year, y = N_scaled, label = season),
            inherit.aes = FALSE, color = "#e6550d", size = 2.5, vjust = -0.8) +
  scale_color_manual(values = c("No mitigation (baseline)"      = "#d73027",
                                 "Immediate 3/3 (100% coverage)" = "#253494")) +
  scale_linetype_manual(values = c("No mitigation (baseline)"      = "dashed",
                                    "Immediate 3/3 (100% coverage)" = "solid")) +
  scale_x_continuous(breaks = seq(0, yearslong, 100), limits = c(-10, yearslong)) +
  scale_y_continuous(labels = scales::comma,
                     breaks = scales::pretty_breaks(n = 6),
                     sec.axis = sec_axis(~ . / pairs_scale,
                                         name = "Breeding pairs (Bird Island)",
                                         labels = scales::comma)) +
  labs(x = paste0("Years since 1990 baseline (", yearslong, "-year horizon)"),
       y = "Population (N individuals)",
       color = NULL, linetype = NULL,
       title = paste0("500-year projection vs observed breeding pairs — ", target_period, " baseline"),
       subtitle = paste0("\u03BB (no mitigation) = ", round(lambda_base, 5),
                         ";  \u03BB (full 3/3) = ", round(lambda_3of3, 5),
                         ";  orange dots = census breeding pairs")) +
  themeo + theme(legend.position = "bottom")
p_1c
```

**plot 1d — 500-year projection with observed age class lines**
```{r plot-1d, fig.width = 9, fig.height = 5}
p_1d <- ggplot(proj_part1_long, aes(x = year, y = N, color = scenario, linetype = scenario)) +
  geom_line(linewidth = 0.9) +
  geom_hline(yintercept = N0, color = "grey50", linetype = "dashed", linewidth = 0.4) +
  geom_line(data = dem_obs_split,
            aes(x = year, y = N_obs, group = group, linetype = group),
            inherit.aes = FALSE, color = "black", linewidth = 0.7) +
  geom_point(data = dem_obs_split,
             aes(x = year, y = N_obs, shape = group),
             inherit.aes = FALSE, color = "black", size = 3) +
  geom_point(data = dem_obs,
             aes(x = year, y = N_obs),
             inherit.aes = FALSE, color = "black", size = 3, shape = 17) +
  geom_text(data = dem_obs,
            aes(x = year, y = N_obs, label = Period),
            inherit.aes = FALSE, color = "black", size = 2.5, vjust = -0.8) +
  scale_color_manual(values = c("No mitigation (baseline)"      = "#d73027",
                                 "Immediate 3/3 (100% coverage)" = "#253494")) +
  scale_linetype_manual(values = c("No mitigation (baseline)"      = "dashed",
                                    "Immediate 3/3 (100% coverage)" = "solid",
                                    "Adults"                        = "solid",
                                    "Immatures"                     = "dashed"),
                        guide = "none") +
  scale_shape_manual(values = c("Adults" = 16, "Immatures" = 1),
                     name = "Observed age class") +
  scale_x_continuous(breaks = seq(0, yearslong, 100), limits = c(0, yearslong)) +
  scale_y_continuous(labels = scales::comma,
                     breaks = scales::pretty_breaks(n = 6),
                     sec.axis = sec_axis(~ . / N0 * 100,
                                         name = "% of 1990\u20131994 baseline",
                                         labels = function(x) paste0(round(x), "%"))) +
  labs(x = paste0("Years since 1990 baseline (", yearslong, "-year horizon)"),
       y = "Population (N individuals)",
       color = NULL,
       title = paste0("500-year projection vs observed age classes — ", target_period, " baseline"),
       subtitle = "Black triangles = total; filled circles = adults; \nopen circles = immatures (Clay 2019)") +
  themeo + theme(legend.position = "bottom")
p_1d
```

**plot 1e — 500-year projection, log(N) y-axis**
```{r plot-1e, fig.width = 9, fig.height = 5}
p_1e <- ggplot(proj_part1_long, aes(x = year, y = N, color = scenario, linetype = scenario)) +
  geom_line(linewidth = 0.9) +
  geom_hline(yintercept = N0, color = "grey50", linetype = "dashed", linewidth = 0.4) +
  geom_point(data = bird_obs,
             aes(x = year, y = N_scaled),
             inherit.aes = FALSE, color = "#e6550d", size = 3, shape = 16) +
  geom_text(data = bird_obs,
            aes(x = year, y = N_scaled, label = season),
            inherit.aes = FALSE, color = "#e6550d", size = 2.5, vjust = -0.8) +
  scale_color_manual(values = c("No mitigation (baseline)"      = "#d73027",
                                 "Immediate 3/3 (100% coverage)" = "#253494")) +
  scale_linetype_manual(values = c("No mitigation (baseline)"      = "dashed",
                                    "Immediate 3/3 (100% coverage)" = "solid")) +
  scale_x_continuous(breaks = seq(0, yearslong, 100), limits = c(-10, yearslong)) +
  scale_y_log10(labels = scales::comma,
                breaks = scales::trans_breaks("log10", function(x) 10^x)) +
  labs(x = paste0("Years since 1990 baseline (", yearslong, "-year horizon)"),
       y = "log\u2081\u2080(N individuals)",
       color = NULL, linetype = NULL,
       title = paste0("500-year projection, log scale — ", target_period, " baseline"),
       subtitle = paste0("\u03BB (no mitigation) = ", round(lambda_base, 5),
                         ";  \u03BB (full 3/3) = ", round(lambda_3of3, 5),
                         ";  orange dots = census breeding pairs\n"#,
                        # "Straight lines in log space reflect constant fixed lambda"
                         )) +
  themeo + theme(legend.position = "bottom")
p_1e
```

**plot 1f — annual lambda (N(t+1)/N(t)) over 500-year horizon**
```{r plot-1f, fig.width = 9, fig.height = 5}
# compute per-year realized lambda from the scalar projection
lambda_ts <- proj_part1_long %>%
  group_by(scenario) %>%
  arrange(year) %>%
  mutate(lambda_t = N / lag(N)) %>%
  filter(!is.na(lambda_t))

p_1f <- ggplot(lambda_ts, aes(x = year, y = lambda_t, color = scenario, linetype = scenario)) +
  geom_line(linewidth = 0.9) +
  geom_hline(yintercept = 1, color = "grey50", linetype = "dashed", linewidth = 0.4) +
  scale_color_manual(values = c("No mitigation (baseline)"      = "#d73027",
                                 "Immediate 3/3 (100% coverage)" = "#253494")) +
  scale_linetype_manual(values = c("No mitigation (baseline)"      = "dashed",
                                    "Immediate 3/3 (100% coverage)" = "solid")) +
  scale_x_continuous(breaks = seq(0, yearslong, 100), limits = c(0, yearslong)) +
  scale_y_continuous(breaks = scales::pretty_breaks(n = 6)) +
  labs(x = paste0("Years since 1990 baseline (", yearslong, "-year horizon)"),
       y = expression(lambda[t] ~ "(N(t+1) / N(t))"),
       color = NULL, linetype = NULL,
       title = paste0("Annual \u03BB over 500-year horizon — ", target_period, " baseline") ) + #,
       # subtitle = paste0("Fixed-lambda scalar model: lines are flat at \u03BB (no mitigation) = ",
       #                   round(lambda_base, 5), " and \u03BB (full 3/3) = ", round(lambda_3of3, 5),
       #                   "\nDashed grey line = \u03BB = 1 (population stable)")) +
  themeo + theme(legend.position = "bottom")
p_1f
```

---

*PART 2 — STAGE-STRUCTURED MATRIX PROJECTION (still with fixed lambda but using matrices directly)*

Full 11-stage matrix projection: N(t+1) = A %*% N(t).
Both scenarios start from SSD scaled to N0. Total N should match Part 1 exactly.

**project**
```{r part2-project}
mat_base  <- matrix(0, nrow = years + 1, ncol = 11)
mat_mitig <- matrix(0, nrow = years + 1, ncol = 11)
colnames(mat_base)  <- stage_names
colnames(mat_mitig) <- stage_names
mat_base[1, ]  <- n0_vec
mat_mitig[1, ] <- n0_vec

for (k in 1:years) {
  mat_base[k + 1, ]  <- m_base %*% mat_base[k, ]
  mat_mitig[k + 1, ] <- m_3of3 %*% mat_mitig[k, ]
}

# Convert to long format with broad stage groupings for plotting in ggplot
to_long <- function(mat, scenario_label) {
  as.data.frame(mat) %>%
    mutate(year = 0:years, scenario = scenario_label) %>%
    pivot_longer(cols = all_of(stage_names), names_to = "stage", values_to = "N") %>%
    mutate(
      stage = factor(stage, levels = stage_names),
      group = case_when(
        stage %in% c("J2", "J3")              ~ "Juveniles",
        stage %in% c("Imm4", "Imm5", "PB")   ~ "Immatures / pre-breeders",
        stage %in% c("IS", "IF", "ES", "EF") ~ "Breeders",
        stage %in% c("ENB", "Sabb")          ~ "Non-breeders"
      ),
      group = factor(group, levels = c("Juveniles", "Immatures / pre-breeders",
                                        "Breeders", "Non-breeders"))
    )
}

proj_part2 <- bind_rows(
  to_long(mat_base,  "No mitigation (baseline)"),
  to_long(mat_mitig, "Immediate 3/3 (100% coverage)")
) %>%
  mutate(scenario = factor(scenario,
                            levels = c("No mitigation (baseline)",
                                       "Immediate 3/3 (100% coverage)")))

proj_part2_groups <- proj_part2 %>%
  group_by(year, scenario, group) %>%
  summarise(N = sum(N), .groups = "drop")

# total N per year per scenario (for log/lambda plots)
proj_part2_total <- proj_part2 %>%
  group_by(year, scenario) %>%
  summarise(N = sum(N), .groups = "drop") %>%
  mutate(scenario = factor(scenario, levels = c("No mitigation (baseline)",
                                                 "Immediate 3/3 (100% coverage)")))
```

**plot 2a — 100-year stage-structured projection, N**
```{r plot-2a, fig.width = 9, fig.height = 5}
p_2a <- ggplot(proj_part2_total, aes(x = year, y = N, color = scenario, linetype = scenario)) +
  geom_line(linewidth = 0.9) +
  geom_hline(yintercept = N0, color = "grey50", linetype = "dashed", linewidth = 0.4) +
  geom_point(data = bird_obs, aes(x = year, y = N_scaled),
             inherit.aes = FALSE, color = "#e6550d", size = 3, shape = 16) +
  geom_line(data = bird_obs, aes(x = year, y = N_scaled),
            inherit.aes = FALSE, color = "#e6550d", linewidth = 0.5, linetype = "dotted") +
  geom_text(data = bird_obs, aes(x = year, y = N_scaled, label = season),
            inherit.aes = FALSE, color = "#e6550d", size = 2.5, vjust = -0.8) +
  scale_color_manual(values = c("No mitigation (baseline)"      = "#d73027",
                                 "Immediate 3/3 (100% coverage)" = "#253494")) +
  scale_linetype_manual(values = c("No mitigation (baseline)"      = "dashed",
                                    "Immediate 3/3 (100% coverage)" = "solid")) +
  scale_x_continuous(breaks = seq(0, years, 10), limits = c(-10, years)) +
  scale_y_continuous(labels = scales::comma,
                     breaks = scales::pretty_breaks(n = 6),
                     sec.axis = sec_axis(~ . / pairs_scale,
                                         name = "Breeding pairs (Bird Island)",
                                         labels = scales::comma)) +
  labs(x = "Years since 1990 baseline",
       y = "Population (N individuals)",
       color = NULL, linetype = NULL,
       title = paste0("100-year stage-structured projection — ", target_period, " baseline"),
       subtitle = "Orange dots = census breeding pairs") +
  themeo + theme(legend.position = "bottom")
p_2a
```

**500-year stage-structured projection (Part 2 extension)**
```{r part2-project-long}
mat_base_long  <- matrix(0, nrow = yearslong + 1, ncol = 11)
mat_mitig_long <- matrix(0, nrow = yearslong + 1, ncol = 11)
colnames(mat_base_long)  <- stage_names
colnames(mat_mitig_long) <- stage_names
mat_base_long[1, ]  <- n0_vec
mat_mitig_long[1, ] <- n0_vec

for (k in 1:yearslong) {
  mat_base_long[k + 1, ]  <- m_base %*% mat_base_long[k, ]
  mat_mitig_long[k + 1, ] <- m_3of3 %*% mat_mitig_long[k, ]
}

proj_part2_total_long <- bind_rows(
  as.data.frame(mat_base_long)  %>% mutate(year = 0:yearslong, scenario = "No mitigation (baseline)"),
  as.data.frame(mat_mitig_long) %>% mutate(year = 0:yearslong, scenario = "Immediate 3/3 (100% coverage)")
) %>%
  mutate(N        = rowSums(across(all_of(stage_names))),
         scenario = factor(scenario, levels = c("No mitigation (baseline)",
                                                 "Immediate 3/3 (100% coverage)"))) %>%
  select(year, N, scenario)
```

**plot 2c — 500-year stage-structured projection, N**
```{r plot-2c, fig.width = 9, fig.height = 5}
p_2c <- ggplot(proj_part2_total_long, aes(x = year, y = N, color = scenario, linetype = scenario)) +
  geom_line(linewidth = 0.9) +
  geom_hline(yintercept = N0, color = "grey50", linetype = "dashed", linewidth = 0.4) +
  geom_point(data = bird_obs, aes(x = year, y = N_scaled),
             inherit.aes = FALSE, color = "#e6550d", size = 3, shape = 16) +
  geom_text(data = bird_obs, aes(x = year, y = N_scaled, label = season),
            inherit.aes = FALSE, color = "#e6550d", size = 2.5, vjust = -0.8) +
  scale_color_manual(values = c("No mitigation (baseline)"      = "#d73027",
                                 "Immediate 3/3 (100% coverage)" = "#253494")) +
  scale_linetype_manual(values = c("No mitigation (baseline)"      = "dashed",
                                    "Immediate 3/3 (100% coverage)" = "solid")) +
  scale_x_continuous(breaks = seq(0, yearslong, 100), limits = c(-10, yearslong)) +
  scale_y_continuous(labels = scales::comma,
                     breaks = scales::pretty_breaks(n = 6),
                     sec.axis = sec_axis(~ . / pairs_scale,
                                         name = "Breeding pairs (Bird Island)",
                                         labels = scales::comma)) +
  labs(x = paste0("Years since 1990 baseline (", yearslong, "-year horizon)"),
       y = "Population (N individuals)",
       color = NULL, linetype = NULL,
       title = paste0("500-year stage-structured projection — ", target_period, " baseline"),
       subtitle = "Orange dots = census breeding pairs") +
  themeo + theme(legend.position = "bottom")
p_2c
```

**plot 2e — 500-year stage-structured projection, log(N) y-axis**
```{r plot-2e, fig.width = 9, fig.height = 5}
p_2e <- ggplot(proj_part2_total_long, aes(x = year, y = N, color = scenario, linetype = scenario)) +
  geom_line(linewidth = 0.9) +
  geom_hline(yintercept = N0, color = "grey50", linetype = "dashed", linewidth = 0.4) +
  scale_color_manual(values = c("No mitigation (baseline)"      = "#d73027",
                                 "Immediate 3/3 (100% coverage)" = "#253494")) +
  scale_linetype_manual(values = c("No mitigation (baseline)"      = "dashed",
                                    "Immediate 3/3 (100% coverage)" = "solid")) +
  scale_x_continuous(breaks = seq(0, yearslong, 100), limits = c(0, yearslong)) +
  scale_y_log10(labels = scales::comma,
                breaks = scales::trans_breaks("log10", function(x) 10^x)) +
  labs(x = paste0("Years since 1990 baseline (", yearslong, "-year horizon)"),
       y = "log\u2081\u2080(N individuals)",
       color = NULL, linetype = NULL,
       title = paste0("500-year stage-structured projection, log scale — ", target_period, " baseline"))+ #,
     #  subtitle = "Straight lines in log space confirm constant per-year matrix eigenvalue") +
  themeo + theme(legend.position = "bottom")
p_2e
```

**plot 2f — annual lambda (N(t+1)/N(t)) over 500-year horizon, stage-structured**
```{r plot-2f, fig.width = 9, fig.height = 5}
lambda_ts_2 <- proj_part2_total_long %>%
  group_by(scenario) %>%
  arrange(year) %>%
  mutate(lambda_t = N / lag(N)) %>%
  filter(!is.na(lambda_t))

p_2f <- ggplot(lambda_ts_2, aes(x = year, y = lambda_t, color = scenario, linetype = scenario)) +
  geom_line(linewidth = 0.9) +
  geom_hline(yintercept = 1, color = "grey50", linetype = "dashed", linewidth = 0.4) +
  scale_color_manual(values = c("No mitigation (baseline)"      = "#d73027",
                                 "Immediate 3/3 (100% coverage)" = "#253494")) +
  scale_linetype_manual(values = c("No mitigation (baseline)"      = "dashed",
                                    "Immediate 3/3 (100% coverage)" = "solid")) +
  scale_x_continuous(breaks = seq(0, yearslong, 100), limits = c(0, yearslong)) +
  scale_y_continuous(breaks = scales::pretty_breaks(n = 6)) +
  labs(x = paste0("Years since 1990 baseline (", yearslong, "-year horizon)"),
       y = expression(lambda[t] ~ "(N(t+1) / N(t))"),
       color = NULL, linetype = NULL,
       title = paste0("Annual \u03BB — stage-structured model, 500-year horizon — ", target_period, " baseline") ) + #,
       # subtitle = paste0("Fixed matrices: flat lines confirm dominant eigenvalue is constant\n",
       #                   "Dashed grey line = \u03BB = 1 (population stable)")) +
  themeo + theme(legend.position = "bottom")
p_2f
```

*PART 3 — 5-YEAR PHASE-IN*

Survival improves in equal steps over 5 years. Each year uses a new matrix until year 6,
when m_3of3 is used for the remainder. Improvement per step = (ns_* - s_*) / 5.

**5-year phase-in survival rates**
```{r part3-survival-rates}
s_Juv_p1 <- s_Juv + (1/5) * (ns_Juv - s_Juv);  s_Imm_p1 <- s_Imm + (1/5) * (ns_Imm - s_Imm)
s_EF_p1  <- s_EF  + (1/5) * (ns_EF  - s_EF);   s_ES_p1  <- s_ES  + (1/5) * (ns_ES  - s_ES)
s_ENB_p1 <- s_ENB + (1/5) * (ns_ENB - s_ENB);  s_IF_p1  <- s_IF  + (1/5) * (ns_IF  - s_IF)
s_IS_p1  <- s_IS  + (1/5) * (ns_IS  - s_IS);   s_PF_p1  <- s_PF  + (1/5) * (ns_PF  - s_PF)

s_Juv_p2 <- s_Juv + (2/5) * (ns_Juv - s_Juv);  s_Imm_p2 <- s_Imm + (2/5) * (ns_Imm - s_Imm)
s_EF_p2  <- s_EF  + (2/5) * (ns_EF  - s_EF);   s_ES_p2  <- s_ES  + (2/5) * (ns_ES  - s_ES)
s_ENB_p2 <- s_ENB + (2/5) * (ns_ENB - s_ENB);  s_IF_p2  <- s_IF  + (2/5) * (ns_IF  - s_IF)
s_IS_p2  <- s_IS  + (2/5) * (ns_IS  - s_IS);   s_PF_p2  <- s_PF  + (2/5) * (ns_PF  - s_PF)

s_Juv_p3 <- s_Juv + (3/5) * (ns_Juv - s_Juv);  s_Imm_p3 <- s_Imm + (3/5) * (ns_Imm - s_Imm)
s_EF_p3  <- s_EF  + (3/5) * (ns_EF  - s_EF);   s_ES_p3  <- s_ES  + (3/5) * (ns_ES  - s_ES)
s_ENB_p3 <- s_ENB + (3/5) * (ns_ENB - s_ENB);  s_IF_p3  <- s_IF  + (3/5) * (ns_IF  - s_IF)
s_IS_p3  <- s_IS  + (3/5) * (ns_IS  - s_IS);   s_PF_p3  <- s_PF  + (3/5) * (ns_PF  - s_PF)

s_Juv_p4 <- s_Juv + (4/5) * (ns_Juv - s_Juv);  s_Imm_p4 <- s_Imm + (4/5) * (ns_Imm - s_Imm)
s_EF_p4  <- s_EF  + (4/5) * (ns_EF  - s_EF);   s_ES_p4  <- s_ES  + (4/5) * (ns_ES  - s_ES)
s_ENB_p4 <- s_ENB + (4/5) * (ns_ENB - s_ENB);  s_IF_p4  <- s_IF  + (4/5) * (ns_IF  - s_IF)
s_IS_p4  <- s_IS  + (4/5) * (ns_IS  - s_IS);   s_PF_p4  <- s_PF  + (4/5) * (ns_PF  - s_PF)

# Step 5: 100% — equals ns_* (same as m_3of3)
s_Juv_p5 <- s_Juv + (5/5) * (ns_Juv - s_Juv);  s_Imm_p5 <- s_Imm + (5/5) * (ns_Imm - s_Imm)
s_EF_p5  <- s_EF  + (5/5) * (ns_EF  - s_EF);   s_ES_p5  <- s_ES  + (5/5) * (ns_ES  - s_ES)
s_ENB_p5 <- s_ENB + (5/5) * (ns_ENB - s_ENB);  s_IF_p5  <- s_IF  + (5/5) * (ns_IF  - s_IF)
s_IS_p5  <- s_IS  + (5/5) * (ns_IS  - s_IS);   s_PF_p5  <- s_PF  + (5/5) * (ns_PF  - s_PF)

tibble(
  stage     = c("Juv", "Imm", "EF", "ES", "ENB", "IF", "IS", "PF"),
  baseline  = c(s_Juv, s_Imm, s_EF, s_ES, s_ENB, s_IF, s_IS, s_PF),
  phase1    = c(s_Juv_p1, s_Imm_p1, s_EF_p1, s_ES_p1, s_ENB_p1, s_IF_p1, s_IS_p1, s_PF_p1),
  phase2    = c(s_Juv_p2, s_Imm_p2, s_EF_p2, s_ES_p2, s_ENB_p2, s_IF_p2, s_IS_p2, s_PF_p2),
  phase3    = c(s_Juv_p3, s_Imm_p3, s_EF_p3, s_ES_p3, s_ENB_p3, s_IF_p3, s_IS_p3, s_PF_p3),
  phase4    = c(s_Juv_p4, s_Imm_p4, s_EF_p4, s_ES_p4, s_ENB_p4, s_IF_p4, s_IS_p4, s_PF_p4),
  full_3of3 = c(s_Juv_p5, s_Imm_p5, s_EF_p5, s_ES_p5, s_ENB_p5, s_IF_p5, s_IS_p5, s_PF_p5)
) %>% print()
```

**5-year phase-in matrices**
```{r part3-matrices}
# Stages: J2 J3 Imm4 Imm5 PB IS IF ES EF ENB Sabb
m_phase1 <- matrix(data = c(
  0,        0,        0,        0,        0,                                                0.5*s_Juv_p1,               0,                           0.5*s_Juv_p1,               0,                           0,                               0,
  s_Juv_p1, 0,        0,        0,        0,                                                0,                          0,                           0,                          0,                           0,                               0,
  0,        s_Juv_p1, 0,        0,        0,                                                0,                          0,                           0,                          0,                           0,                               0,
  0,        0,        s_Imm_p1, 0,        0,                                                0,                          0,                           0,                          0,                           0,                               0,
  0,        0,        0,        s_Imm_p1, s_Imm_p1*r_Imm*(1-b_Imm)+s_Imm_p1*(1-r_Imm),    0,                          0,                           0,                          0,                           0,                               0,
  0,        0,        0,        0,        s_Imm_p1*r_Imm*b_Imm*k_Imm,                       0,                          0,                           0,                          0,                           0,                               0,
  0,        0,        0,        0,        s_Imm_p1*r_Imm*b_Imm*(1-k_Imm),                   0,                          0,                           0,                          0,                           0,                               0,
  0,        0,        0,        0,        0,                                                s_IS_p1*r_IS*b_IS*k_IS,     s_IF_p1*r_IF*b_IF*k_IF,      s_ES_p1*r_ES*b_ES*k_ES,     s_EF_p1*r_EF*b_EF*k_EF,      s_ENB_p1*r_ENB*b_ENB*k_ENB,      s_PF_p1*r_PF*b_PF*k_PF,
  0,        0,        0,        0,        0,                                                s_IS_p1*r_IS*b_IS*(1-k_IS), s_IF_p1*r_IF*b_IF*(1-k_IF),  s_ES_p1*r_ES*b_ES*(1-k_ES), s_EF_p1*r_EF*b_EF*(1-k_EF),  s_ENB_p1*r_ENB*b_ENB*(1-k_ENB),  s_PF_p1*r_PF*b_PF*(1-k_PF),
  0,        0,        0,        0,        0,                                                s_IS_p1*r_IS*(1-b_IS),      s_IF_p1*r_IF*(1-b_IF),       s_ES_p1*r_ES*(1-b_ES),      s_EF_p1*r_EF*(1-b_EF),       s_ENB_p1*r_ENB*(1-b_ENB),        s_PF_p1*r_PF*(1-b_PF),
  0,        0,        0,        0,        0,                                                s_IS_p1*(1-r_IS),           s_IF_p1*(1-r_IF),            s_ES_p1*(1-r_ES),           s_EF_p1*(1-r_EF),            s_ENB_p1*(1-r_ENB),              s_PF_p1*(1-r_PF)
), nrow = 11, byrow = TRUE)

m_phase2 <- matrix(data = c(
  0,        0,        0,        0,        0,                                                0.5*s_Juv_p2,               0,                           0.5*s_Juv_p2,               0,                           0,                               0,
  s_Juv_p2, 0,        0,        0,        0,                                                0,                          0,                           0,                          0,                           0,                               0,
  0,        s_Juv_p2, 0,        0,        0,                                                0,                          0,                           0,                          0,                           0,                               0,
  0,        0,        s_Imm_p2, 0,        0,                                                0,                          0,                           0,                          0,                           0,                               0,
  0,        0,        0,        s_Imm_p2, s_Imm_p2*r_Imm*(1-b_Imm)+s_Imm_p2*(1-r_Imm),    0,                          0,                           0,                          0,                           0,                               0,
  0,        0,        0,        0,        s_Imm_p2*r_Imm*b_Imm*k_Imm,                       0,                          0,                           0,                          0,                           0,                               0,
  0,        0,        0,        0,        s_Imm_p2*r_Imm*b_Imm*(1-k_Imm),                   0,                          0,                           0,                          0,                           0,                               0,
  0,        0,        0,        0,        0,                                                s_IS_p2*r_IS*b_IS*k_IS,     s_IF_p2*r_IF*b_IF*k_IF,      s_ES_p2*r_ES*b_ES*k_ES,     s_EF_p2*r_EF*b_EF*k_EF,      s_ENB_p2*r_ENB*b_ENB*k_ENB,      s_PF_p2*r_PF*b_PF*k_PF,
  0,        0,        0,        0,        0,                                                s_IS_p2*r_IS*b_IS*(1-k_IS), s_IF_p2*r_IF*b_IF*(1-k_IF),  s_ES_p2*r_ES*b_ES*(1-k_ES), s_EF_p2*r_EF*b_EF*(1-k_EF),  s_ENB_p2*r_ENB*b_ENB*(1-k_ENB),  s_PF_p2*r_PF*b_PF*(1-k_PF),
  0,        0,        0,        0,        0,                                                s_IS_p2*r_IS*(1-b_IS),      s_IF_p2*r_IF*(1-b_IF),       s_ES_p2*r_ES*(1-b_ES),      s_EF_p2*r_EF*(1-b_EF),       s_ENB_p2*r_ENB*(1-b_ENB),        s_PF_p2*r_PF*(1-b_PF),
  0,        0,        0,        0,        0,                                                s_IS_p2*(1-r_IS),           s_IF_p2*(1-r_IF),            s_ES_p2*(1-r_ES),           s_EF_p2*(1-r_EF),            s_ENB_p2*(1-r_ENB),              s_PF_p2*(1-r_PF)
), nrow = 11, byrow = TRUE)

m_phase3 <- matrix(data = c(
  0,        0,        0,        0,        0,                                                0.5*s_Juv_p3,               0,                           0.5*s_Juv_p3,               0,                           0,                               0,
  s_Juv_p3, 0,        0,        0,        0,                                                0,                          0,                           0,                          0,                           0,                               0,
  0,        s_Juv_p3, 0,        0,        0,                                                0,                          0,                           0,                          0,                           0,                               0,
  0,        0,        s_Imm_p3, 0,        0,                                                0,                          0,                           0,                          0,                           0,                               0,
  0,        0,        0,        s_Imm_p3, s_Imm_p3*r_Imm*(1-b_Imm)+s_Imm_p3*(1-r_Imm),    0,                          0,                           0,                          0,                           0,                               0,
  0,        0,        0,        0,        s_Imm_p3*r_Imm*b_Imm*k_Imm,                       0,                          0,                           0,                          0,                           0,                               0,
  0,        0,        0,        0,        s_Imm_p3*r_Imm*b_Imm*(1-k_Imm),                   0,                          0,                           0,                          0,                           0,                               0,
  0,        0,        0,        0,        0,                                                s_IS_p3*r_IS*b_IS*k_IS,     s_IF_p3*r_IF*b_IF*k_IF,      s_ES_p3*r_ES*b_ES*k_ES,     s_EF_p3*r_EF*b_EF*k_EF,      s_ENB_p3*r_ENB*b_ENB*k_ENB,      s_PF_p3*r_PF*b_PF*k_PF,
  0,        0,        0,        0,        0,                                                s_IS_p3*r_IS*b_IS*(1-k_IS), s_IF_p3*r_IF*b_IF*(1-k_IF),  s_ES_p3*r_ES*b_ES*(1-k_ES), s_EF_p3*r_EF*b_EF*(1-k_EF),  s_ENB_p3*r_ENB*b_ENB*(1-k_ENB),  s_PF_p3*r_PF*b_PF*(1-k_PF),
  0,        0,        0,        0,        0,                                                s_IS_p3*r_IS*(1-b_IS),      s_IF_p3*r_IF*(1-b_IF),       s_ES_p3*r_ES*(1-b_ES),      s_EF_p3*r_EF*(1-b_EF),       s_ENB_p3*r_ENB*(1-b_ENB),        s_PF_p3*r_PF*(1-b_PF),
  0,        0,        0,        0,        0,                                                s_IS_p3*(1-r_IS),           s_IF_p3*(1-r_IF),            s_ES_p3*(1-r_ES),           s_EF_p3*(1-r_EF),            s_ENB_p3*(1-r_ENB),              s_PF_p3*(1-r_PF)
), nrow = 11, byrow = TRUE)

m_phase4 <- matrix(data = c(
  0,        0,        0,        0,        0,                                                0.5*s_Juv_p4,               0,                           0.5*s_Juv_p4,               0,                           0,                               0,
  s_Juv_p4, 0,        0,        0,        0,                                                0,                          0,                           0,                          0,                           0,                               0,
  0,        s_Juv_p4, 0,        0,        0,                                                0,                          0,                           0,                          0,                           0,                               0,
  0,        0,        s_Imm_p4, 0,        0,                                                0,                          0,                           0,                          0,                           0,                               0,
  0,        0,        0,        s_Imm_p4, s_Imm_p4*r_Imm*(1-b_Imm)+s_Imm_p4*(1-r_Imm),    0,                          0,                           0,                          0,                           0,                               0,
  0,        0,        0,        0,        s_Imm_p4*r_Imm*b_Imm*k_Imm,                       0,                          0,                           0,                          0,                           0,                               0,
  0,        0,        0,        0,        s_Imm_p4*r_Imm*b_Imm*(1-k_Imm),                   0,                          0,                           0,                          0,                           0,                               0,
  0,        0,        0,        0,        0,                                                s_IS_p4*r_IS*b_IS*k_IS,     s_IF_p4*r_IF*b_IF*k_IF,      s_ES_p4*r_ES*b_ES*k_ES,     s_EF_p4*r_EF*b_EF*k_EF,      s_ENB_p4*r_ENB*b_ENB*k_ENB,      s_PF_p4*r_PF*b_PF*k_PF,
  0,        0,        0,        0,        0,                                                s_IS_p4*r_IS*b_IS*(1-k_IS), s_IF_p4*r_IF*b_IF*(1-k_IF),  s_ES_p4*r_ES*b_ES*(1-k_ES), s_EF_p4*r_EF*b_EF*(1-k_EF),  s_ENB_p4*r_ENB*b_ENB*(1-k_ENB),  s_PF_p4*r_PF*b_PF*(1-k_PF),
  0,        0,        0,        0,        0,                                                s_IS_p4*r_IS*(1-b_IS),      s_IF_p4*r_IF*(1-b_IF),       s_ES_p4*r_ES*(1-b_ES),      s_EF_p4*r_EF*(1-b_EF),       s_ENB_p4*r_ENB*(1-b_ENB),        s_PF_p4*r_PF*(1-b_PF),
  0,        0,        0,        0,        0,                                                s_IS_p4*(1-r_IS),           s_IF_p4*(1-r_IF),            s_ES_p4*(1-r_ES),           s_EF_p4*(1-r_EF),            s_ENB_p4*(1-r_ENB),              s_PF_p4*(1-r_PF)
), nrow = 11, byrow = TRUE)

# m_phase5: 100% improvement — identical to m_3of3
m_phase5 <- matrix(data = c(
  0,        0,        0,        0,        0,                                                0.5*s_Juv_p5,               0,                           0.5*s_Juv_p5,               0,                           0,                               0,
  s_Juv_p5, 0,        0,        0,        0,                                                0,                          0,                           0,                          0,                           0,                               0,
  0,        s_Juv_p5, 0,        0,        0,                                                0,                          0,                           0,                          0,                           0,                               0,
  0,        0,        s_Imm_p5, 0,        0,                                                0,                          0,                           0,                          0,                           0,                               0,
  0,        0,        0,        s_Imm_p5, s_Imm_p5*r_Imm*(1-b_Imm)+s_Imm_p5*(1-r_Imm),    0,                          0,                           0,                          0,                           0,                               0,
  0,        0,        0,        0,        s_Imm_p5*r_Imm*b_Imm*k_Imm,                       0,                          0,                           0,                          0,                           0,                               0,
  0,        0,        0,        0,        s_Imm_p5*r_Imm*b_Imm*(1-k_Imm),                   0,                          0,                           0,                          0,                           0,                               0,
  0,        0,        0,        0,        0,                                                s_IS_p5*r_IS*b_IS*k_IS,     s_IF_p5*r_IF*b_IF*k_IF,      s_ES_p5*r_ES*b_ES*k_ES,     s_EF_p5*r_EF*b_EF*k_EF,      s_ENB_p5*r_ENB*b_ENB*k_ENB,      s_PF_p5*r_PF*b_PF*k_PF,
  0,        0,        0,        0,        0,                                                s_IS_p5*r_IS*b_IS*(1-k_IS), s_IF_p5*r_IF*b_IF*(1-k_IF),  s_ES_p5*r_ES*b_ES*(1-k_ES), s_EF_p5*r_EF*b_EF*(1-k_EF),  s_ENB_p5*r_ENB*b_ENB*(1-k_ENB),  s_PF_p5*r_PF*b_PF*(1-k_PF),
  0,        0,        0,        0,        0,                                                s_IS_p5*r_IS*(1-b_IS),      s_IF_p5*r_IF*(1-b_IF),       s_ES_p5*r_ES*(1-b_ES),      s_EF_p5*r_EF*(1-b_EF),       s_ENB_p5*r_ENB*(1-b_ENB),        s_PF_p5*r_PF*(1-b_PF),
  0,        0,        0,        0,        0,                                                s_IS_p5*(1-r_IS),           s_IF_p5*(1-r_IF),            s_ES_p5*(1-r_ES),           s_EF_p5*(1-r_EF),            s_ENB_p5*(1-r_ENB),              s_PF_p5*(1-r_PF)
), nrow = 11, byrow = TRUE)

phase_list <- list(m_phase1, m_phase2, m_phase3, m_phase4, m_phase5)
```

---

*PART 4 — 10-YEAR PHASE-IN*

Each step adds 1/10 of total survival improvement. k=1-10 use phase matrices; k>10 uses m_3of3.

**10-year phase-in survival rates**
```{r part4-survival-rates}
s_Juv_10y1  <- s_Juv + (1/10)  * (ns_Juv - s_Juv);  s_Imm_10y1  <- s_Imm + (1/10)  * (ns_Imm - s_Imm)
s_EF_10y1   <- s_EF  + (1/10)  * (ns_EF  - s_EF);   s_ES_10y1   <- s_ES  + (1/10)  * (ns_ES  - s_ES)
s_ENB_10y1  <- s_ENB + (1/10)  * (ns_ENB - s_ENB);  s_IF_10y1   <- s_IF  + (1/10)  * (ns_IF  - s_IF)
s_IS_10y1   <- s_IS  + (1/10)  * (ns_IS  - s_IS);   s_PF_10y1   <- s_PF  + (1/10)  * (ns_PF  - s_PF)

s_Juv_10y2  <- s_Juv + (2/10)  * (ns_Juv - s_Juv);  s_Imm_10y2  <- s_Imm + (2/10)  * (ns_Imm - s_Imm)
s_EF_10y2   <- s_EF  + (2/10)  * (ns_EF  - s_EF);   s_ES_10y2   <- s_ES  + (2/10)  * (ns_ES  - s_ES)
s_ENB_10y2  <- s_ENB + (2/10)  * (ns_ENB - s_ENB);  s_IF_10y2   <- s_IF  + (2/10)  * (ns_IF  - s_IF)
s_IS_10y2   <- s_IS  + (2/10)  * (ns_IS  - s_IS);   s_PF_10y2   <- s_PF  + (2/10)  * (ns_PF  - s_PF)

s_Juv_10y3  <- s_Juv + (3/10)  * (ns_Juv - s_Juv);  s_Imm_10y3  <- s_Imm + (3/10)  * (ns_Imm - s_Imm)
s_EF_10y3   <- s_EF  + (3/10)  * (ns_EF  - s_EF);   s_ES_10y3   <- s_ES  + (3/10)  * (ns_ES  - s_ES)
s_ENB_10y3  <- s_ENB + (3/10)  * (ns_ENB - s_ENB);  s_IF_10y3   <- s_IF  + (3/10)  * (ns_IF  - s_IF)
s_IS_10y3   <- s_IS  + (3/10)  * (ns_IS  - s_IS);   s_PF_10y3   <- s_PF  + (3/10)  * (ns_PF  - s_PF)

s_Juv_10y4  <- s_Juv + (4/10)  * (ns_Juv - s_Juv);  s_Imm_10y4  <- s_Imm + (4/10)  * (ns_Imm - s_Imm)
s_EF_10y4   <- s_EF  + (4/10)  * (ns_EF  - s_EF);   s_ES_10y4   <- s_ES  + (4/10)  * (ns_ES  - s_ES)
s_ENB_10y4  <- s_ENB + (4/10)  * (ns_ENB - s_ENB);  s_IF_10y4   <- s_IF  + (4/10)  * (ns_IF  - s_IF)
s_IS_10y4   <- s_IS  + (4/10)  * (ns_IS  - s_IS);   s_PF_10y4   <- s_PF  + (4/10)  * (ns_PF  - s_PF)

s_Juv_10y5  <- s_Juv + (5/10)  * (ns_Juv - s_Juv);  s_Imm_10y5  <- s_Imm + (5/10)  * (ns_Imm - s_Imm)
s_EF_10y5   <- s_EF  + (5/10)  * (ns_EF  - s_EF);   s_ES_10y5   <- s_ES  + (5/10)  * (ns_ES  - s_ES)
s_ENB_10y5  <- s_ENB + (5/10)  * (ns_ENB - s_ENB);  s_IF_10y5   <- s_IF  + (5/10)  * (ns_IF  - s_IF)
s_IS_10y5   <- s_IS  + (5/10)  * (ns_IS  - s_IS);   s_PF_10y5   <- s_PF  + (5/10)  * (ns_PF  - s_PF)

s_Juv_10y6  <- s_Juv + (6/10)  * (ns_Juv - s_Juv);  s_Imm_10y6  <- s_Imm + (6/10)  * (ns_Imm - s_Imm)
s_EF_10y6   <- s_EF  + (6/10)  * (ns_EF  - s_EF);   s_ES_10y6   <- s_ES  + (6/10)  * (ns_ES  - s_ES)
s_ENB_10y6  <- s_ENB + (6/10)  * (ns_ENB - s_ENB);  s_IF_10y6   <- s_IF  + (6/10)  * (ns_IF  - s_IF)
s_IS_10y6   <- s_IS  + (6/10)  * (ns_IS  - s_IS);   s_PF_10y6   <- s_PF  + (6/10)  * (ns_PF  - s_PF)

s_Juv_10y7  <- s_Juv + (7/10)  * (ns_Juv - s_Juv);  s_Imm_10y7  <- s_Imm + (7/10)  * (ns_Imm - s_Imm)
s_EF_10y7   <- s_EF  + (7/10)  * (ns_EF  - s_EF);   s_ES_10y7   <- s_ES  + (7/10)  * (ns_ES  - s_ES)
s_ENB_10y7  <- s_ENB + (7/10)  * (ns_ENB - s_ENB);  s_IF_10y7   <- s_IF  + (7/10)  * (ns_IF  - s_IF)
s_IS_10y7   <- s_IS  + (7/10)  * (ns_IS  - s_IS);   s_PF_10y7   <- s_PF  + (7/10)  * (ns_PF  - s_PF)

s_Juv_10y8  <- s_Juv + (8/10)  * (ns_Juv - s_Juv);  s_Imm_10y8  <- s_Imm + (8/10)  * (ns_Imm - s_Imm)
s_EF_10y8   <- s_EF  + (8/10)  * (ns_EF  - s_EF);   s_ES_10y8   <- s_ES  + (8/10)  * (ns_ES  - s_ES)
s_ENB_10y8  <- s_ENB + (8/10)  * (ns_ENB - s_ENB);  s_IF_10y8   <- s_IF  + (8/10)  * (ns_IF  - s_IF)
s_IS_10y8   <- s_IS  + (8/10)  * (ns_IS  - s_IS);   s_PF_10y8   <- s_PF  + (8/10)  * (ns_PF  - s_PF)

s_Juv_10y9  <- s_Juv + (9/10)  * (ns_Juv - s_Juv);  s_Imm_10y9  <- s_Imm + (9/10)  * (ns_Imm - s_Imm)
s_EF_10y9   <- s_EF  + (9/10)  * (ns_EF  - s_EF);   s_ES_10y9   <- s_ES  + (9/10)  * (ns_ES  - s_ES)
s_ENB_10y9  <- s_ENB + (9/10)  * (ns_ENB - s_ENB);  s_IF_10y9   <- s_IF  + (9/10)  * (ns_IF  - s_IF)
s_IS_10y9   <- s_IS  + (9/10)  * (ns_IS  - s_IS);   s_PF_10y9   <- s_PF  + (9/10)  * (ns_PF  - s_PF)

# Step 10: 100% — equals ns_* (same as m_3of3)
s_Juv_10y10 <- s_Juv + (10/10) * (ns_Juv - s_Juv);  s_Imm_10y10 <- s_Imm + (10/10) * (ns_Imm - s_Imm)
s_EF_10y10  <- s_EF  + (10/10) * (ns_EF  - s_EF);   s_ES_10y10  <- s_ES  + (10/10) * (ns_ES  - s_ES)
s_ENB_10y10 <- s_ENB + (10/10) * (ns_ENB - s_ENB);  s_IF_10y10  <- s_IF  + (10/10) * (ns_IF  - s_IF)
s_IS_10y10  <- s_IS  + (10/10) * (ns_IS  - s_IS);   s_PF_10y10  <- s_PF  + (10/10) * (ns_PF  - s_PF)
```

**10-year phase-in matrices**
```{r part4-matrices}
# Stages: J2 J3 Imm4 Imm5 PB IS IF ES EF ENB Sabb
m_10y1 <- matrix(data = c(
  0,0,0,0,0, 0.5*s_Juv_10y1,0,0.5*s_Juv_10y1,0,0,0,
  s_Juv_10y1,0,0,0,0, 0,0,0,0,0,0,
  0,s_Juv_10y1,0,0,0, 0,0,0,0,0,0,
  0,0,s_Imm_10y1,0,0, 0,0,0,0,0,0,
  0,0,0,s_Imm_10y1, s_Imm_10y1*r_Imm*(1-b_Imm)+s_Imm_10y1*(1-r_Imm), 0,0,0,0,0,0,
  0,0,0,0, s_Imm_10y1*r_Imm*b_Imm*k_Imm, 0,0,0,0,0,0,
  0,0,0,0, s_Imm_10y1*r_Imm*b_Imm*(1-k_Imm), 0,0,0,0,0,0,
  0,0,0,0,0, s_IS_10y1*r_IS*b_IS*k_IS, s_IF_10y1*r_IF*b_IF*k_IF, s_ES_10y1*r_ES*b_ES*k_ES, s_EF_10y1*r_EF*b_EF*k_EF, s_ENB_10y1*r_ENB*b_ENB*k_ENB, s_PF_10y1*r_PF*b_PF*k_PF,
  0,0,0,0,0, s_IS_10y1*r_IS*b_IS*(1-k_IS), s_IF_10y1*r_IF*b_IF*(1-k_IF), s_ES_10y1*r_ES*b_ES*(1-k_ES), s_EF_10y1*r_EF*b_EF*(1-k_EF), s_ENB_10y1*r_ENB*b_ENB*(1-k_ENB), s_PF_10y1*r_PF*b_PF*(1-k_PF),
  0,0,0,0,0, s_IS_10y1*r_IS*(1-b_IS), s_IF_10y1*r_IF*(1-b_IF), s_ES_10y1*r_ES*(1-b_ES), s_EF_10y1*r_EF*(1-b_EF), s_ENB_10y1*r_ENB*(1-b_ENB), s_PF_10y1*r_PF*(1-b_PF),
  0,0,0,0,0, s_IS_10y1*(1-r_IS), s_IF_10y1*(1-r_IF), s_ES_10y1*(1-r_ES), s_EF_10y1*(1-r_EF), s_ENB_10y1*(1-r_ENB), s_PF_10y1*(1-r_PF)
), nrow=11, byrow=TRUE)

m_10y2 <- matrix(data = c(
  0,0,0,0,0, 0.5*s_Juv_10y2,0,0.5*s_Juv_10y2,0,0,0,
  s_Juv_10y2,0,0,0,0, 0,0,0,0,0,0,
  0,s_Juv_10y2,0,0,0, 0,0,0,0,0,0,
  0,0,s_Imm_10y2,0,0, 0,0,0,0,0,0,
  0,0,0,s_Imm_10y2, s_Imm_10y2*r_Imm*(1-b_Imm)+s_Imm_10y2*(1-r_Imm), 0,0,0,0,0,0,
  0,0,0,0, s_Imm_10y2*r_Imm*b_Imm*k_Imm, 0,0,0,0,0,0,
  0,0,0,0, s_Imm_10y2*r_Imm*b_Imm*(1-k_Imm), 0,0,0,0,0,0,
  0,0,0,0,0, s_IS_10y2*r_IS*b_IS*k_IS, s_IF_10y2*r_IF*b_IF*k_IF, s_ES_10y2*r_ES*b_ES*k_ES, s_EF_10y2*r_EF*b_EF*k_EF, s_ENB_10y2*r_ENB*b_ENB*k_ENB, s_PF_10y2*r_PF*b_PF*k_PF,
  0,0,0,0,0, s_IS_10y2*r_IS*b_IS*(1-k_IS), s_IF_10y2*r_IF*b_IF*(1-k_IF), s_ES_10y2*r_ES*b_ES*(1-k_ES), s_EF_10y2*r_EF*b_EF*(1-k_EF), s_ENB_10y2*r_ENB*b_ENB*(1-k_ENB), s_PF_10y2*r_PF*b_PF*(1-k_PF),
  0,0,0,0,0, s_IS_10y2*r_IS*(1-b_IS), s_IF_10y2*r_IF*(1-b_IF), s_ES_10y2*r_ES*(1-b_ES), s_EF_10y2*r_EF*(1-b_EF), s_ENB_10y2*r_ENB*(1-b_ENB), s_PF_10y2*r_PF*(1-b_PF),
  0,0,0,0,0, s_IS_10y2*(1-r_IS), s_IF_10y2*(1-r_IF), s_ES_10y2*(1-r_ES), s_EF_10y2*(1-r_EF), s_ENB_10y2*(1-r_ENB), s_PF_10y2*(1-r_PF)
), nrow=11, byrow=TRUE)

m_10y3 <- matrix(data = c(
  0,0,0,0,0, 0.5*s_Juv_10y3,0,0.5*s_Juv_10y3,0,0,0,
  s_Juv_10y3,0,0,0,0, 0,0,0,0,0,0,
  0,s_Juv_10y3,0,0,0, 0,0,0,0,0,0,
  0,0,s_Imm_10y3,0,0, 0,0,0,0,0,0,
  0,0,0,s_Imm_10y3, s_Imm_10y3*r_Imm*(1-b_Imm)+s_Imm_10y3*(1-r_Imm), 0,0,0,0,0,0,
  0,0,0,0, s_Imm_10y3*r_Imm*b_Imm*k_Imm, 0,0,0,0,0,0,
  0,0,0,0, s_Imm_10y3*r_Imm*b_Imm*(1-k_Imm), 0,0,0,0,0,0,
  0,0,0,0,0, s_IS_10y3*r_IS*b_IS*k_IS, s_IF_10y3*r_IF*b_IF*k_IF, s_ES_10y3*r_ES*b_ES*k_ES, s_EF_10y3*r_EF*b_EF*k_EF, s_ENB_10y3*r_ENB*b_ENB*k_ENB, s_PF_10y3*r_PF*b_PF*k_PF,
  0,0,0,0,0, s_IS_10y3*r_IS*b_IS*(1-k_IS), s_IF_10y3*r_IF*b_IF*(1-k_IF), s_ES_10y3*r_ES*b_ES*(1-k_ES), s_EF_10y3*r_EF*b_EF*(1-k_EF), s_ENB_10y3*r_ENB*b_ENB*(1-k_ENB), s_PF_10y3*r_PF*b_PF*(1-k_PF),
  0,0,0,0,0, s_IS_10y3*r_IS*(1-b_IS), s_IF_10y3*r_IF*(1-b_IF), s_ES_10y3*r_ES*(1-b_ES), s_EF_10y3*r_EF*(1-b_EF), s_ENB_10y3*r_ENB*(1-b_ENB), s_PF_10y3*r_PF*(1-b_PF),
  0,0,0,0,0, s_IS_10y3*(1-r_IS), s_IF_10y3*(1-r_IF), s_ES_10y3*(1-r_ES), s_EF_10y3*(1-r_EF), s_ENB_10y3*(1-r_ENB), s_PF_10y3*(1-r_PF)
), nrow=11, byrow=TRUE)

m_10y4 <- matrix(data = c(
  0,0,0,0,0, 0.5*s_Juv_10y4,0,0.5*s_Juv_10y4,0,0,0,
  s_Juv_10y4,0,0,0,0, 0,0,0,0,0,0,
  0,s_Juv_10y4,0,0,0, 0,0,0,0,0,0,
  0,0,s_Imm_10y4,0,0, 0,0,0,0,0,0,
  0,0,0,s_Imm_10y4, s_Imm_10y4*r_Imm*(1-b_Imm)+s_Imm_10y4*(1-r_Imm), 0,0,0,0,0,0,
  0,0,0,0, s_Imm_10y4*r_Imm*b_Imm*k_Imm, 0,0,0,0,0,0,
  0,0,0,0, s_Imm_10y4*r_Imm*b_Imm*(1-k_Imm), 0,0,0,0,0,0,
  0,0,0,0,0, s_IS_10y4*r_IS*b_IS*k_IS, s_IF_10y4*r_IF*b_IF*k_IF, s_ES_10y4*r_ES*b_ES*k_ES, s_EF_10y4*r_EF*b_EF*k_EF, s_ENB_10y4*r_ENB*b_ENB*k_ENB, s_PF_10y4*r_PF*b_PF*k_PF,
  0,0,0,0,0, s_IS_10y4*r_IS*b_IS*(1-k_IS), s_IF_10y4*r_IF*b_IF*(1-k_IF), s_ES_10y4*r_ES*b_ES*(1-k_ES), s_EF_10y4*r_EF*b_EF*(1-k_EF), s_ENB_10y4*r_ENB*b_ENB*(1-k_ENB), s_PF_10y4*r_PF*b_PF*(1-k_PF),
  0,0,0,0,0, s_IS_10y4*r_IS*(1-b_IS), s_IF_10y4*r_IF*(1-b_IF), s_ES_10y4*r_ES*(1-b_ES), s_EF_10y4*r_EF*(1-b_EF), s_ENB_10y4*r_ENB*(1-b_ENB), s_PF_10y4*r_PF*(1-b_PF),
  0,0,0,0,0, s_IS_10y4*(1-r_IS), s_IF_10y4*(1-r_IF), s_ES_10y4*(1-r_ES), s_EF_10y4*(1-r_EF), s_ENB_10y4*(1-r_ENB), s_PF_10y4*(1-r_PF)
), nrow=11, byrow=TRUE)

m_10y5 <- matrix(data = c(
  0,0,0,0,0, 0.5*s_Juv_10y5,0,0.5*s_Juv_10y5,0,0,0,
  s_Juv_10y5,0,0,0,0, 0,0,0,0,0,0,
  0,s_Juv_10y5,0,0,0, 0,0,0,0,0,0,
  0,0,s_Imm_10y5,0,0, 0,0,0,0,0,0,
  0,0,0,s_Imm_10y5, s_Imm_10y5*r_Imm*(1-b_Imm)+s_Imm_10y5*(1-r_Imm), 0,0,0,0,0,0,
  0,0,0,0, s_Imm_10y5*r_Imm*b_Imm*k_Imm, 0,0,0,0,0,0,
  0,0,0,0, s_Imm_10y5*r_Imm*b_Imm*(1-k_Imm), 0,0,0,0,0,0,
  0,0,0,0,0, s_IS_10y5*r_IS*b_IS*k_IS, s_IF_10y5*r_IF*b_IF*k_IF, s_ES_10y5*r_ES*b_ES*k_ES, s_EF_10y5*r_EF*b_EF*k_EF, s_ENB_10y5*r_ENB*b_ENB*k_ENB, s_PF_10y5*r_PF*b_PF*k_PF,
  0,0,0,0,0, s_IS_10y5*r_IS*b_IS*(1-k_IS), s_IF_10y5*r_IF*b_IF*(1-k_IF), s_ES_10y5*r_ES*b_ES*(1-k_ES), s_EF_10y5*r_EF*b_EF*(1-k_EF), s_ENB_10y5*r_ENB*b_ENB*(1-k_ENB), s_PF_10y5*r_PF*b_PF*(1-k_PF),
  0,0,0,0,0, s_IS_10y5*r_IS*(1-b_IS), s_IF_10y5*r_IF*(1-b_IF), s_ES_10y5*r_ES*(1-b_ES), s_EF_10y5*r_EF*(1-b_EF), s_ENB_10y5*r_ENB*(1-b_ENB), s_PF_10y5*r_PF*(1-b_PF),
  0,0,0,0,0, s_IS_10y5*(1-r_IS), s_IF_10y5*(1-r_IF), s_ES_10y5*(1-r_ES), s_EF_10y5*(1-r_EF), s_ENB_10y5*(1-r_ENB), s_PF_10y5*(1-r_PF)
), nrow=11, byrow=TRUE)

m_10y6 <- matrix(data = c(
  0,0,0,0,0, 0.5*s_Juv_10y6,0,0.5*s_Juv_10y6,0,0,0,
  s_Juv_10y6,0,0,0,0, 0,0,0,0,0,0,
  0,s_Juv_10y6,0,0,0, 0,0,0,0,0,0,
  0,0,s_Imm_10y6,0,0, 0,0,0,0,0,0,
  0,0,0,s_Imm_10y6, s_Imm_10y6*r_Imm*(1-b_Imm)+s_Imm_10y6*(1-r_Imm), 0,0,0,0,0,0,
  0,0,0,0, s_Imm_10y6*r_Imm*b_Imm*k_Imm, 0,0,0,0,0,0,
  0,0,0,0, s_Imm_10y6*r_Imm*b_Imm*(1-k_Imm), 0,0,0,0,0,0,
  0,0,0,0,0, s_IS_10y6*r_IS*b_IS*k_IS, s_IF_10y6*r_IF*b_IF*k_IF, s_ES_10y6*r_ES*b_ES*k_ES, s_EF_10y6*r_EF*b_EF*k_EF, s_ENB_10y6*r_ENB*b_ENB*k_ENB, s_PF_10y6*r_PF*b_PF*k_PF,
  0,0,0,0,0, s_IS_10y6*r_IS*b_IS*(1-k_IS), s_IF_10y6*r_IF*b_IF*(1-k_IF), s_ES_10y6*r_ES*b_ES*(1-k_ES), s_EF_10y6*r_EF*b_EF*(1-k_EF), s_ENB_10y6*r_ENB*b_ENB*(1-k_ENB), s_PF_10y6*r_PF*b_PF*(1-k_PF),
  0,0,0,0,0, s_IS_10y6*r_IS*(1-b_IS), s_IF_10y6*r_IF*(1-b_IF), s_ES_10y6*r_ES*(1-b_ES), s_EF_10y6*r_EF*(1-b_EF), s_ENB_10y6*r_ENB*(1-b_ENB), s_PF_10y6*r_PF*(1-b_PF),
  0,0,0,0,0, s_IS_10y6*(1-r_IS), s_IF_10y6*(1-r_IF), s_ES_10y6*(1-r_ES), s_EF_10y6*(1-r_EF), s_ENB_10y6*(1-r_ENB), s_PF_10y6*(1-r_PF)
), nrow=11, byrow=TRUE)

m_10y7 <- matrix(data = c(
  0,0,0,0,0, 0.5*s_Juv_10y7,0,0.5*s_Juv_10y7,0,0,0,
  s_Juv_10y7,0,0,0,0, 0,0,0,0,0,0,
  0,s_Juv_10y7,0,0,0, 0,0,0,0,0,0,
  0,0,s_Imm_10y7,0,0, 0,0,0,0,0,0,
  0,0,0,s_Imm_10y7, s_Imm_10y7*r_Imm*(1-b_Imm)+s_Imm_10y7*(1-r_Imm), 0,0,0,0,0,0,
  0,0,0,0, s_Imm_10y7*r_Imm*b_Imm*k_Imm, 0,0,0,0,0,0,
  0,0,0,0, s_Imm_10y7*r_Imm*b_Imm*(1-k_Imm), 0,0,0,0,0,0,
  0,0,0,0,0, s_IS_10y7*r_IS*b_IS*k_IS, s_IF_10y7*r_IF*b_IF*k_IF, s_ES_10y7*r_ES*b_ES*k_ES, s_EF_10y7*r_EF*b_EF*k_EF, s_ENB_10y7*r_ENB*b_ENB*k_ENB, s_PF_10y7*r_PF*b_PF*k_PF,
  0,0,0,0,0, s_IS_10y7*r_IS*b_IS*(1-k_IS), s_IF_10y7*r_IF*b_IF*(1-k_IF), s_ES_10y7*r_ES*b_ES*(1-k_ES), s_EF_10y7*r_EF*b_EF*(1-k_EF), s_ENB_10y7*r_ENB*b_ENB*(1-k_ENB), s_PF_10y7*r_PF*b_PF*(1-k_PF),
  0,0,0,0,0, s_IS_10y7*r_IS*(1-b_IS), s_IF_10y7*r_IF*(1-b_IF), s_ES_10y7*r_ES*(1-b_ES), s_EF_10y7*r_EF*(1-b_EF), s_ENB_10y7*r_ENB*(1-b_ENB), s_PF_10y7*r_PF*(1-b_PF),
  0,0,0,0,0, s_IS_10y7*(1-r_IS), s_IF_10y7*(1-r_IF), s_ES_10y7*(1-r_ES), s_EF_10y7*(1-r_EF), s_ENB_10y7*(1-r_ENB), s_PF_10y7*(1-r_PF)
), nrow=11, byrow=TRUE)

m_10y8 <- matrix(data = c(
  0,0,0,0,0, 0.5*s_Juv_10y8,0,0.5*s_Juv_10y8,0,0,0,
  s_Juv_10y8,0,0,0,0, 0,0,0,0,0,0,
  0,s_Juv_10y8,0,0,0, 0,0,0,0,0,0,
  0,0,s_Imm_10y8,0,0, 0,0,0,0,0,0,
  0,0,0,s_Imm_10y8, s_Imm_10y8*r_Imm*(1-b_Imm)+s_Imm_10y8*(1-r_Imm), 0,0,0,0,0,0,
  0,0,0,0, s_Imm_10y8*r_Imm*b_Imm*k_Imm, 0,0,0,0,0,0,
  0,0,0,0, s_Imm_10y8*r_Imm*b_Imm*(1-k_Imm), 0,0,0,0,0,0,
  0,0,0,0,0, s_IS_10y8*r_IS*b_IS*k_IS, s_IF_10y8*r_IF*b_IF*k_IF, s_ES_10y8*r_ES*b_ES*k_ES, s_EF_10y8*r_EF*b_EF*k_EF, s_ENB_10y8*r_ENB*b_ENB*k_ENB, s_PF_10y8*r_PF*b_PF*k_PF,
  0,0,0,0,0, s_IS_10y8*r_IS*b_IS*(1-k_IS), s_IF_10y8*r_IF*b_IF*(1-k_IF), s_ES_10y8*r_ES*b_ES*(1-k_ES), s_EF_10y8*r_EF*b_EF*(1-k_EF), s_ENB_10y8*r_ENB*b_ENB*(1-k_ENB), s_PF_10y8*r_PF*b_PF*(1-k_PF),
  0,0,0,0,0, s_IS_10y8*r_IS*(1-b_IS), s_IF_10y8*r_IF*(1-b_IF), s_ES_10y8*r_ES*(1-b_ES), s_EF_10y8*r_EF*(1-b_EF), s_ENB_10y8*r_ENB*(1-b_ENB), s_PF_10y8*r_PF*(1-b_PF),
  0,0,0,0,0, s_IS_10y8*(1-r_IS), s_IF_10y8*(1-r_IF), s_ES_10y8*(1-r_ES), s_EF_10y8*(1-r_EF), s_ENB_10y8*(1-r_ENB), s_PF_10y8*(1-r_PF)
), nrow=11, byrow=TRUE)

m_10y9 <- matrix(data = c(
  0,0,0,0,0, 0.5*s_Juv_10y9,0,0.5*s_Juv_10y9,0,0,0,
  s_Juv_10y9,0,0,0,0, 0,0,0,0,0,0,
  0,s_Juv_10y9,0,0,0, 0,0,0,0,0,0,
  0,0,s_Imm_10y9,0,0, 0,0,0,0,0,0,
  0,0,0,s_Imm_10y9, s_Imm_10y9*r_Imm*(1-b_Imm)+s_Imm_10y9*(1-r_Imm), 0,0,0,0,0,0,
  0,0,0,0, s_Imm_10y9*r_Imm*b_Imm*k_Imm, 0,0,0,0,0,0,
  0,0,0,0, s_Imm_10y9*r_Imm*b_Imm*(1-k_Imm), 0,0,0,0,0,0,
  0,0,0,0,0, s_IS_10y9*r_IS*b_IS*k_IS, s_IF_10y9*r_IF*b_IF*k_IF, s_ES_10y9*r_ES*b_ES*k_ES, s_EF_10y9*r_EF*b_EF*k_EF, s_ENB_10y9*r_ENB*b_ENB*k_ENB, s_PF_10y9*r_PF*b_PF*k_PF,
  0,0,0,0,0, s_IS_10y9*r_IS*b_IS*(1-k_IS), s_IF_10y9*r_IF*b_IF*(1-k_IF), s_ES_10y9*r_ES*b_ES*(1-k_ES), s_EF_10y9*r_EF*b_EF*(1-k_EF), s_ENB_10y9*r_ENB*b_ENB*(1-k_ENB), s_PF_10y9*r_PF*b_PF*(1-k_PF),
  0,0,0,0,0, s_IS_10y9*r_IS*(1-b_IS), s_IF_10y9*r_IF*(1-b_IF), s_ES_10y9*r_ES*(1-b_ES), s_EF_10y9*r_EF*(1-b_EF), s_ENB_10y9*r_ENB*(1-b_ENB), s_PF_10y9*r_PF*(1-b_PF),
  0,0,0,0,0, s_IS_10y9*(1-r_IS), s_IF_10y9*(1-r_IF), s_ES_10y9*(1-r_ES), s_EF_10y9*(1-r_EF), s_ENB_10y9*(1-r_ENB), s_PF_10y9*(1-r_PF)
), nrow=11, byrow=TRUE)

# m_10y10: 100% improvement — should equal m_3of3
m_10y10 <- matrix(data = c(
  0,0,0,0,0, 0.5*s_Juv_10y10,0,0.5*s_Juv_10y10,0,0,0,
  s_Juv_10y10,0,0,0,0, 0,0,0,0,0,0,
  0,s_Juv_10y10,0,0,0, 0,0,0,0,0,0,
  0,0,s_Imm_10y10,0,0, 0,0,0,0,0,0,
  0,0,0,s_Imm_10y10, s_Imm_10y10*r_Imm*(1-b_Imm)+s_Imm_10y10*(1-r_Imm), 0,0,0,0,0,0,
  0,0,0,0, s_Imm_10y10*r_Imm*b_Imm*k_Imm, 0,0,0,0,0,0,
  0,0,0,0, s_Imm_10y10*r_Imm*b_Imm*(1-k_Imm), 0,0,0,0,0,0,
  0,0,0,0,0, s_IS_10y10*r_IS*b_IS*k_IS, s_IF_10y10*r_IF*b_IF*k_IF, s_ES_10y10*r_ES*b_ES*k_ES, s_EF_10y10*r_EF*b_EF*k_EF, s_ENB_10y10*r_ENB*b_ENB*k_ENB, s_PF_10y10*r_PF*b_PF*k_PF,
  0,0,0,0,0, s_IS_10y10*r_IS*b_IS*(1-k_IS), s_IF_10y10*r_IF*b_IF*(1-k_IF), s_ES_10y10*r_ES*b_ES*(1-k_ES), s_EF_10y10*r_EF*b_EF*(1-k_EF), s_ENB_10y10*r_ENB*b_ENB*(1-k_ENB), s_PF_10y10*r_PF*b_PF*(1-k_PF),
  0,0,0,0,0, s_IS_10y10*r_IS*(1-b_IS), s_IF_10y10*r_IF*(1-b_IF), s_ES_10y10*r_ES*(1-b_ES), s_EF_10y10*r_EF*(1-b_EF), s_ENB_10y10*r_ENB*(1-b_ENB), s_PF_10y10*r_PF*(1-b_PF),
  0,0,0,0,0, s_IS_10y10*(1-r_IS), s_IF_10y10*(1-r_IF), s_ES_10y10*(1-r_ES), s_EF_10y10*(1-r_EF), s_ENB_10y10*(1-r_ENB), s_PF_10y10*(1-r_PF)
), nrow=11, byrow=TRUE)

phase_list_10y <- list(m_10y1, m_10y2, m_10y3, m_10y4, m_10y5,
                       m_10y6, m_10y7, m_10y8, m_10y9, m_10y10)
```

---

*PART 5 — 20-YEAR PHASE-IN*

Each step adds 1/20 of total survival improvement. k=1-20 use phase matrices; k>20 uses m_3of3.

**20-year phase-in survival rates**
```{r part5-survival-rates}
s_Juv_20y1  <- s_Juv + (1/20)  * (ns_Juv - s_Juv);  s_Imm_20y1  <- s_Imm + (1/20)  * (ns_Imm - s_Imm)
s_EF_20y1   <- s_EF  + (1/20)  * (ns_EF  - s_EF);   s_ES_20y1   <- s_ES  + (1/20)  * (ns_ES  - s_ES)
s_ENB_20y1  <- s_ENB + (1/20)  * (ns_ENB - s_ENB);  s_IF_20y1   <- s_IF  + (1/20)  * (ns_IF  - s_IF)
s_IS_20y1   <- s_IS  + (1/20)  * (ns_IS  - s_IS);   s_PF_20y1   <- s_PF  + (1/20)  * (ns_PF  - s_PF)

s_Juv_20y2  <- s_Juv + (2/20)  * (ns_Juv - s_Juv);  s_Imm_20y2  <- s_Imm + (2/20)  * (ns_Imm - s_Imm)
s_EF_20y2   <- s_EF  + (2/20)  * (ns_EF  - s_EF);   s_ES_20y2   <- s_ES  + (2/20)  * (ns_ES  - s_ES)
s_ENB_20y2  <- s_ENB + (2/20)  * (ns_ENB - s_ENB);  s_IF_20y2   <- s_IF  + (2/20)  * (ns_IF  - s_IF)
s_IS_20y2   <- s_IS  + (2/20)  * (ns_IS  - s_IS);   s_PF_20y2   <- s_PF  + (2/20)  * (ns_PF  - s_PF)

s_Juv_20y3  <- s_Juv + (3/20)  * (ns_Juv - s_Juv);  s_Imm_20y3  <- s_Imm + (3/20)  * (ns_Imm - s_Imm)
s_EF_20y3   <- s_EF  + (3/20)  * (ns_EF  - s_EF);   s_ES_20y3   <- s_ES  + (3/20)  * (ns_ES  - s_ES)
s_ENB_20y3  <- s_ENB + (3/20)  * (ns_ENB - s_ENB);  s_IF_20y3   <- s_IF  + (3/20)  * (ns_IF  - s_IF)
s_IS_20y3   <- s_IS  + (3/20)  * (ns_IS  - s_IS);   s_PF_20y3   <- s_PF  + (3/20)  * (ns_PF  - s_PF)

s_Juv_20y4  <- s_Juv + (4/20)  * (ns_Juv - s_Juv);  s_Imm_20y4  <- s_Imm + (4/20)  * (ns_Imm - s_Imm)
s_EF_20y4   <- s_EF  + (4/20)  * (ns_EF  - s_EF);   s_ES_20y4   <- s_ES  + (4/20)  * (ns_ES  - s_ES)
s_ENB_20y4  <- s_ENB + (4/20)  * (ns_ENB - s_ENB);  s_IF_20y4   <- s_IF  + (4/20)  * (ns_IF  - s_IF)
s_IS_20y4   <- s_IS  + (4/20)  * (ns_IS  - s_IS);   s_PF_20y4   <- s_PF  + (4/20)  * (ns_PF  - s_PF)

s_Juv_20y5  <- s_Juv + (5/20)  * (ns_Juv - s_Juv);  s_Imm_20y5  <- s_Imm + (5/20)  * (ns_Imm - s_Imm)
s_EF_20y5   <- s_EF  + (5/20)  * (ns_EF  - s_EF);   s_ES_20y5   <- s_ES  + (5/20)  * (ns_ES  - s_ES)
s_ENB_20y5  <- s_ENB + (5/20)  * (ns_ENB - s_ENB);  s_IF_20y5   <- s_IF  + (5/20)  * (ns_IF  - s_IF)
s_IS_20y5   <- s_IS  + (5/20)  * (ns_IS  - s_IS);   s_PF_20y5   <- s_PF  + (5/20)  * (ns_PF  - s_PF)

s_Juv_20y6  <- s_Juv + (6/20)  * (ns_Juv - s_Juv);  s_Imm_20y6  <- s_Imm + (6/20)  * (ns_Imm - s_Imm)
s_EF_20y6   <- s_EF  + (6/20)  * (ns_EF  - s_EF);   s_ES_20y6   <- s_ES  + (6/20)  * (ns_ES  - s_ES)
s_ENB_20y6  <- s_ENB + (6/20)  * (ns_ENB - s_ENB);  s_IF_20y6   <- s_IF  + (6/20)  * (ns_IF  - s_IF)
s_IS_20y6   <- s_IS  + (6/20)  * (ns_IS  - s_IS);   s_PF_20y6   <- s_PF  + (6/20)  * (ns_PF  - s_PF)

s_Juv_20y7  <- s_Juv + (7/20)  * (ns_Juv - s_Juv);  s_Imm_20y7  <- s_Imm + (7/20)  * (ns_Imm - s_Imm)
s_EF_20y7   <- s_EF  + (7/20)  * (ns_EF  - s_EF);   s_ES_20y7   <- s_ES  + (7/20)  * (ns_ES  - s_ES)
s_ENB_20y7  <- s_ENB + (7/20)  * (ns_ENB - s_ENB);  s_IF_20y7   <- s_IF  + (7/20)  * (ns_IF  - s_IF)
s_IS_20y7   <- s_IS  + (7/20)  * (ns_IS  - s_IS);   s_PF_20y7   <- s_PF  + (7/20)  * (ns_PF  - s_PF)

s_Juv_20y8  <- s_Juv + (8/20)  * (ns_Juv - s_Juv);  s_Imm_20y8  <- s_Imm + (8/20)  * (ns_Imm - s_Imm)
s_EF_20y8   <- s_EF  + (8/20)  * (ns_EF  - s_EF);   s_ES_20y8   <- s_ES  + (8/20)  * (ns_ES  - s_ES)
s_ENB_20y8  <- s_ENB + (8/20)  * (ns_ENB - s_ENB);  s_IF_20y8   <- s_IF  + (8/20)  * (ns_IF  - s_IF)
s_IS_20y8   <- s_IS  + (8/20)  * (ns_IS  - s_IS);   s_PF_20y8   <- s_PF  + (8/20)  * (ns_PF  - s_PF)

s_Juv_20y9  <- s_Juv + (9/20)  * (ns_Juv - s_Juv);  s_Imm_20y9  <- s_Imm + (9/20)  * (ns_Imm - s_Imm)
s_EF_20y9   <- s_EF  + (9/20)  * (ns_EF  - s_EF);   s_ES_20y9   <- s_ES  + (9/20)  * (ns_ES  - s_ES)
s_ENB_20y9  <- s_ENB + (9/20)  * (ns_ENB - s_ENB);  s_IF_20y9   <- s_IF  + (9/20)  * (ns_IF  - s_IF)
s_IS_20y9   <- s_IS  + (9/20)  * (ns_IS  - s_IS);   s_PF_20y9   <- s_PF  + (9/20)  * (ns_PF  - s_PF)

s_Juv_20y10 <- s_Juv + (10/20) * (ns_Juv - s_Juv);  s_Imm_20y10 <- s_Imm + (10/20) * (ns_Imm - s_Imm)
s_EF_20y10  <- s_EF  + (10/20) * (ns_EF  - s_EF);   s_ES_20y10  <- s_ES  + (10/20) * (ns_ES  - s_ES)
s_ENB_20y10 <- s_ENB + (10/20) * (ns_ENB - s_ENB);  s_IF_20y10  <- s_IF  + (10/20) * (ns_IF  - s_IF)
s_IS_20y10  <- s_IS  + (10/20) * (ns_IS  - s_IS);   s_PF_20y10  <- s_PF  + (10/20) * (ns_PF  - s_PF)

s_Juv_20y11 <- s_Juv + (11/20) * (ns_Juv - s_Juv);  s_Imm_20y11 <- s_Imm + (11/20) * (ns_Imm - s_Imm)
s_EF_20y11  <- s_EF  + (11/20) * (ns_EF  - s_EF);   s_ES_20y11  <- s_ES  + (11/20) * (ns_ES  - s_ES)
s_ENB_20y11 <- s_ENB + (11/20) * (ns_ENB - s_ENB);  s_IF_20y11  <- s_IF  + (11/20) * (ns_IF  - s_IF)
s_IS_20y11  <- s_IS  + (11/20) * (ns_IS  - s_IS);   s_PF_20y11  <- s_PF  + (11/20) * (ns_PF  - s_PF)

s_Juv_20y12 <- s_Juv + (12/20) * (ns_Juv - s_Juv);  s_Imm_20y12 <- s_Imm + (12/20) * (ns_Imm - s_Imm)
s_EF_20y12  <- s_EF  + (12/20) * (ns_EF  - s_EF);   s_ES_20y12  <- s_ES  + (12/20) * (ns_ES  - s_ES)
s_ENB_20y12 <- s_ENB + (12/20) * (ns_ENB - s_ENB);  s_IF_20y12  <- s_IF  + (12/20) * (ns_IF  - s_IF)
s_IS_20y12  <- s_IS  + (12/20) * (ns_IS  - s_IS);   s_PF_20y12  <- s_PF  + (12/20) * (ns_PF  - s_PF)

s_Juv_20y13 <- s_Juv + (13/20) * (ns_Juv - s_Juv);  s_Imm_20y13 <- s_Imm + (13/20) * (ns_Imm - s_Imm)
s_EF_20y13  <- s_EF  + (13/20) * (ns_EF  - s_EF);   s_ES_20y13  <- s_ES  + (13/20) * (ns_ES  - s_ES)
s_ENB_20y13 <- s_ENB + (13/20) * (ns_ENB - s_ENB);  s_IF_20y13  <- s_IF  + (13/20) * (ns_IF  - s_IF)
s_IS_20y13  <- s_IS  + (13/20) * (ns_IS  - s_IS);   s_PF_20y13  <- s_PF  + (13/20) * (ns_PF  - s_PF)

s_Juv_20y14 <- s_Juv + (14/20) * (ns_Juv - s_Juv);  s_Imm_20y14 <- s_Imm + (14/20) * (ns_Imm - s_Imm)
s_EF_20y14  <- s_EF  + (14/20) * (ns_EF  - s_EF);   s_ES_20y14  <- s_ES  + (14/20) * (ns_ES  - s_ES)
s_ENB_20y14 <- s_ENB + (14/20) * (ns_ENB - s_ENB);  s_IF_20y14  <- s_IF  + (14/20) * (ns_IF  - s_IF)
s_IS_20y14  <- s_IS  + (14/20) * (ns_IS  - s_IS);   s_PF_20y14  <- s_PF  + (14/20) * (ns_PF  - s_PF)

s_Juv_20y15 <- s_Juv + (15/20) * (ns_Juv - s_Juv);  s_Imm_20y15 <- s_Imm + (15/20) * (ns_Imm - s_Imm)
s_EF_20y15  <- s_EF  + (15/20) * (ns_EF  - s_EF);   s_ES_20y15  <- s_ES  + (15/20) * (ns_ES  - s_ES)
s_ENB_20y15 <- s_ENB + (15/20) * (ns_ENB - s_ENB);  s_IF_20y15  <- s_IF  + (15/20) * (ns_IF  - s_IF)
s_IS_20y15  <- s_IS  + (15/20) * (ns_IS  - s_IS);   s_PF_20y15  <- s_PF  + (15/20) * (ns_PF  - s_PF)

s_Juv_20y16 <- s_Juv + (16/20) * (ns_Juv - s_Juv);  s_Imm_20y16 <- s_Imm + (16/20) * (ns_Imm - s_Imm)
s_EF_20y16  <- s_EF  + (16/20) * (ns_EF  - s_EF);   s_ES_20y16  <- s_ES  + (16/20) * (ns_ES  - s_ES)
s_ENB_20y16 <- s_ENB + (16/20) * (ns_ENB - s_ENB);  s_IF_20y16  <- s_IF  + (16/20) * (ns_IF  - s_IF)
s_IS_20y16  <- s_IS  + (16/20) * (ns_IS  - s_IS);   s_PF_20y16  <- s_PF  + (16/20) * (ns_PF  - s_PF)

s_Juv_20y17 <- s_Juv + (17/20) * (ns_Juv - s_Juv);  s_Imm_20y17 <- s_Imm + (17/20) * (ns_Imm - s_Imm)
s_EF_20y17  <- s_EF  + (17/20) * (ns_EF  - s_EF);   s_ES_20y17  <- s_ES  + (17/20) * (ns_ES  - s_ES)
s_ENB_20y17 <- s_ENB + (17/20) * (ns_ENB - s_ENB);  s_IF_20y17  <- s_IF  + (17/20) * (ns_IF  - s_IF)
s_IS_20y17  <- s_IS  + (17/20) * (ns_IS  - s_IS);   s_PF_20y17  <- s_PF  + (17/20) * (ns_PF  - s_PF)

s_Juv_20y18 <- s_Juv + (18/20) * (ns_Juv - s_Juv);  s_Imm_20y18 <- s_Imm + (18/20) * (ns_Imm - s_Imm)
s_EF_20y18  <- s_EF  + (18/20) * (ns_EF  - s_EF);   s_ES_20y18  <- s_ES  + (18/20) * (ns_ES  - s_ES)
s_ENB_20y18 <- s_ENB + (18/20) * (ns_ENB - s_ENB);  s_IF_20y18  <- s_IF  + (18/20) * (ns_IF  - s_IF)
s_IS_20y18  <- s_IS  + (18/20) * (ns_IS  - s_IS);   s_PF_20y18  <- s_PF  + (18/20) * (ns_PF  - s_PF)

s_Juv_20y19 <- s_Juv + (19/20) * (ns_Juv - s_Juv);  s_Imm_20y19 <- s_Imm + (19/20) * (ns_Imm - s_Imm)
s_EF_20y19  <- s_EF  + (19/20) * (ns_EF  - s_EF);   s_ES_20y19  <- s_ES  + (19/20) * (ns_ES  - s_ES)
s_ENB_20y19 <- s_ENB + (19/20) * (ns_ENB - s_ENB);  s_IF_20y19  <- s_IF  + (19/20) * (ns_IF  - s_IF)
s_IS_20y19  <- s_IS  + (19/20) * (ns_IS  - s_IS);   s_PF_20y19  <- s_PF  + (19/20) * (ns_PF  - s_PF)

# Step 20: 100% — equals ns_* (same as m_3of3)
s_Juv_20y20 <- s_Juv + (20/20) * (ns_Juv - s_Juv);  s_Imm_20y20 <- s_Imm + (20/20) * (ns_Imm - s_Imm)
s_EF_20y20  <- s_EF  + (20/20) * (ns_EF  - s_EF);   s_ES_20y20  <- s_ES  + (20/20) * (ns_ES  - s_ES)
s_ENB_20y20 <- s_ENB + (20/20) * (ns_ENB - s_ENB);  s_IF_20y20  <- s_IF  + (20/20) * (ns_IF  - s_IF)
s_IS_20y20  <- s_IS  + (20/20) * (ns_IS  - s_IS);   s_PF_20y20  <- s_PF  + (20/20) * (ns_PF  - s_PF)
```

**20-year phase-in matrices**
```{r part5-matrices}
# Stages: J2 J3 Imm4 Imm5 PB IS IF ES EF ENB Sabb
m_20y1 <- matrix(data = c(
  0,0,0,0,0, 0.5*s_Juv_20y1,0,0.5*s_Juv_20y1,0,0,0,
  s_Juv_20y1,0,0,0,0, 0,0,0,0,0,0,
  0,s_Juv_20y1,0,0,0, 0,0,0,0,0,0,
  0,0,s_Imm_20y1,0,0, 0,0,0,0,0,0,
  0,0,0,s_Imm_20y1, s_Imm_20y1*r_Imm*(1-b_Imm)+s_Imm_20y1*(1-r_Imm), 0,0,0,0,0,0,
  0,0,0,0, s_Imm_20y1*r_Imm*b_Imm*k_Imm, 0,0,0,0,0,0,
  0,0,0,0, s_Imm_20y1*r_Imm*b_Imm*(1-k_Imm), 0,0,0,0,0,0,
  0,0,0,0,0, s_IS_20y1*r_IS*b_IS*k_IS, s_IF_20y1*r_IF*b_IF*k_IF, s_ES_20y1*r_ES*b_ES*k_ES, s_EF_20y1*r_EF*b_EF*k_EF, s_ENB_20y1*r_ENB*b_ENB*k_ENB, s_PF_20y1*r_PF*b_PF*k_PF,
  0,0,0,0,0, s_IS_20y1*r_IS*b_IS*(1-k_IS), s_IF_20y1*r_IF*b_IF*(1-k_IF), s_ES_20y1*r_ES*b_ES*(1-k_ES), s_EF_20y1*r_EF*b_EF*(1-k_EF), s_ENB_20y1*r_ENB*b_ENB*(1-k_ENB), s_PF_20y1*r_PF*b_PF*(1-k_PF),
  0,0,0,0,0, s_IS_20y1*r_IS*(1-b_IS), s_IF_20y1*r_IF*(1-b_IF), s_ES_20y1*r_ES*(1-b_ES), s_EF_20y1*r_EF*(1-b_EF), s_ENB_20y1*r_ENB*(1-b_ENB), s_PF_20y1*r_PF*(1-b_PF),
  0,0,0,0,0, s_IS_20y1*(1-r_IS), s_IF_20y1*(1-r_IF), s_ES_20y1*(1-r_ES), s_EF_20y1*(1-r_EF), s_ENB_20y1*(1-r_ENB), s_PF_20y1*(1-r_PF)
), nrow=11, byrow=TRUE)

m_20y2 <- matrix(data = c(
  0,0,0,0,0, 0.5*s_Juv_20y2,0,0.5*s_Juv_20y2,0,0,0,
  s_Juv_20y2,0,0,0,0, 0,0,0,0,0,0,
  0,s_Juv_20y2,0,0,0, 0,0,0,0,0,0,
  0,0,s_Imm_20y2,0,0, 0,0,0,0,0,0,
  0,0,0,s_Imm_20y2, s_Imm_20y2*r_Imm*(1-b_Imm)+s_Imm_20y2*(1-r_Imm), 0,0,0,0,0,0,
  0,0,0,0, s_Imm_20y2*r_Imm*b_Imm*k_Imm, 0,0,0,0,0,0,
  0,0,0,0, s_Imm_20y2*r_Imm*b_Imm*(1-k_Imm), 0,0,0,0,0,0,
  0,0,0,0,0, s_IS_20y2*r_IS*b_IS*k_IS, s_IF_20y2*r_IF*b_IF*k_IF, s_ES_20y2*r_ES*b_ES*k_ES, s_EF_20y2*r_EF*b_EF*k_EF, s_ENB_20y2*r_ENB*b_ENB*k_ENB, s_PF_20y2*r_PF*b_PF*k_PF,
  0,0,0,0,0, s_IS_20y2*r_IS*b_IS*(1-k_IS), s_IF_20y2*r_IF*b_IF*(1-k_IF), s_ES_20y2*r_ES*b_ES*(1-k_ES), s_EF_20y2*r_EF*b_EF*(1-k_EF), s_ENB_20y2*r_ENB*b_ENB*(1-k_ENB), s_PF_20y2*r_PF*b_PF*(1-k_PF),
  0,0,0,0,0, s_IS_20y2*r_IS*(1-b_IS), s_IF_20y2*r_IF*(1-b_IF), s_ES_20y2*r_ES*(1-b_ES), s_EF_20y2*r_EF*(1-b_EF), s_ENB_20y2*r_ENB*(1-b_ENB), s_PF_20y2*r_PF*(1-b_PF),
  0,0,0,0,0, s_IS_20y2*(1-r_IS), s_IF_20y2*(1-r_IF), s_ES_20y2*(1-r_ES), s_EF_20y2*(1-r_EF), s_ENB_20y2*(1-r_ENB), s_PF_20y2*(1-r_PF)
), nrow=11, byrow=TRUE)

m_20y3 <- matrix(data = c(
  0,0,0,0,0, 0.5*s_Juv_20y3,0,0.5*s_Juv_20y3,0,0,0,
  s_Juv_20y3,0,0,0,0, 0,0,0,0,0,0,
  0,s_Juv_20y3,0,0,0, 0,0,0,0,0,0,
  0,0,s_Imm_20y3,0,0, 0,0,0,0,0,0,
  0,0,0,s_Imm_20y3, s_Imm_20y3*r_Imm*(1-b_Imm)+s_Imm_20y3*(1-r_Imm), 0,0,0,0,0,0,
  0,0,0,0, s_Imm_20y3*r_Imm*b_Imm*k_Imm, 0,0,0,0,0,0,
  0,0,0,0, s_Imm_20y3*r_Imm*b_Imm*(1-k_Imm), 0,0,0,0,0,0,
  0,0,0,0,0, s_IS_20y3*r_IS*b_IS*k_IS, s_IF_20y3*r_IF*b_IF*k_IF, s_ES_20y3*r_ES*b_ES*k_ES, s_EF_20y3*r_EF*b_EF*k_EF, s_ENB_20y3*r_ENB*b_ENB*k_ENB, s_PF_20y3*r_PF*b_PF*k_PF,
  0,0,0,0,0, s_IS_20y3*r_IS*b_IS*(1-k_IS), s_IF_20y3*r_IF*b_IF*(1-k_IF), s_ES_20y3*r_ES*b_ES*(1-k_ES), s_EF_20y3*r_EF*b_EF*(1-k_EF), s_ENB_20y3*r_ENB*b_ENB*(1-k_ENB), s_PF_20y3*r_PF*b_PF*(1-k_PF),
  0,0,0,0,0, s_IS_20y3*r_IS*(1-b_IS), s_IF_20y3*r_IF*(1-b_IF), s_ES_20y3*r_ES*(1-b_ES), s_EF_20y3*r_EF*(1-b_EF), s_ENB_20y3*r_ENB*(1-b_ENB), s_PF_20y3*r_PF*(1-b_PF),
  0,0,0,0,0, s_IS_20y3*(1-r_IS), s_IF_20y3*(1-r_IF), s_ES_20y3*(1-r_ES), s_EF_20y3*(1-r_EF), s_ENB_20y3*(1-r_ENB), s_PF_20y3*(1-r_PF)
), nrow=11, byrow=TRUE)

m_20y4 <- matrix(data = c(
  0,0,0,0,0, 0.5*s_Juv_20y4,0,0.5*s_Juv_20y4,0,0,0,
  s_Juv_20y4,0,0,0,0, 0,0,0,0,0,0,
  0,s_Juv_20y4,0,0,0, 0,0,0,0,0,0,
  0,0,s_Imm_20y4,0,0, 0,0,0,0,0,0,
  0,0,0,s_Imm_20y4, s_Imm_20y4*r_Imm*(1-b_Imm)+s_Imm_20y4*(1-r_Imm), 0,0,0,0,0,0,
  0,0,0,0, s_Imm_20y4*r_Imm*b_Imm*k_Imm, 0,0,0,0,0,0,
  0,0,0,0, s_Imm_20y4*r_Imm*b_Imm*(1-k_Imm), 0,0,0,0,0,0,
  0,0,0,0,0, s_IS_20y4*r_IS*b_IS*k_IS, s_IF_20y4*r_IF*b_IF*k_IF, s_ES_20y4*r_ES*b_ES*k_ES, s_EF_20y4*r_EF*b_EF*k_EF, s_ENB_20y4*r_ENB*b_ENB*k_ENB, s_PF_20y4*r_PF*b_PF*k_PF,
  0,0,0,0,0, s_IS_20y4*r_IS*b_IS*(1-k_IS), s_IF_20y4*r_IF*b_IF*(1-k_IF), s_ES_20y4*r_ES*b_ES*(1-k_ES), s_EF_20y4*r_EF*b_EF*(1-k_EF), s_ENB_20y4*r_ENB*b_ENB*(1-k_ENB), s_PF_20y4*r_PF*b_PF*(1-k_PF),
  0,0,0,0,0, s_IS_20y4*r_IS*(1-b_IS), s_IF_20y4*r_IF*(1-b_IF), s_ES_20y4*r_ES*(1-b_ES), s_EF_20y4*r_EF*(1-b_EF), s_ENB_20y4*r_ENB*(1-b_ENB), s_PF_20y4*r_PF*(1-b_PF),
  0,0,0,0,0, s_IS_20y4*(1-r_IS), s_IF_20y4*(1-r_IF), s_ES_20y4*(1-r_ES), s_EF_20y4*(1-r_EF), s_ENB_20y4*(1-r_ENB), s_PF_20y4*(1-r_PF)
), nrow=11, byrow=TRUE)

m_20y5 <- matrix(data = c(
  0,0,0,0,0, 0.5*s_Juv_20y5,0,0.5*s_Juv_20y5,0,0,0,
  s_Juv_20y5,0,0,0,0, 0,0,0,0,0,0,
  0,s_Juv_20y5,0,0,0, 0,0,0,0,0,0,
  0,0,s_Imm_20y5,0,0, 0,0,0,0,0,0,
  0,0,0,s_Imm_20y5, s_Imm_20y5*r_Imm*(1-b_Imm)+s_Imm_20y5*(1-r_Imm), 0,0,0,0,0,0,
  0,0,0,0, s_Imm_20y5*r_Imm*b_Imm*k_Imm, 0,0,0,0,0,0,
  0,0,0,0, s_Imm_20y5*r_Imm*b_Imm*(1-k_Imm), 0,0,0,0,0,0,
  0,0,0,0,0, s_IS_20y5*r_IS*b_IS*k_IS, s_IF_20y5*r_IF*b_IF*k_IF, s_ES_20y5*r_ES*b_ES*k_ES, s_EF_20y5*r_EF*b_EF*k_EF, s_ENB_20y5*r_ENB*b_ENB*k_ENB, s_PF_20y5*r_PF*b_PF*k_PF,
  0,0,0,0,0, s_IS_20y5*r_IS*b_IS*(1-k_IS), s_IF_20y5*r_IF*b_IF*(1-k_IF), s_ES_20y5*r_ES*b_ES*(1-k_ES), s_EF_20y5*r_EF*b_EF*(1-k_EF), s_ENB_20y5*r_ENB*b_ENB*(1-k_ENB), s_PF_20y5*r_PF*b_PF*(1-k_PF),
  0,0,0,0,0, s_IS_20y5*r_IS*(1-b_IS), s_IF_20y5*r_IF*(1-b_IF), s_ES_20y5*r_ES*(1-b_ES), s_EF_20y5*r_EF*(1-b_EF), s_ENB_20y5*r_ENB*(1-b_ENB), s_PF_20y5*r_PF*(1-b_PF),
  0,0,0,0,0, s_IS_20y5*(1-r_IS), s_IF_20y5*(1-r_IF), s_ES_20y5*(1-r_ES), s_EF_20y5*(1-r_EF), s_ENB_20y5*(1-r_ENB), s_PF_20y5*(1-r_PF)
), nrow=11, byrow=TRUE)

m_20y6 <- matrix(data = c(
  0,0,0,0,0, 0.5*s_Juv_20y6,0,0.5*s_Juv_20y6,0,0,0,
  s_Juv_20y6,0,0,0,0, 0,0,0,0,0,0,
  0,s_Juv_20y6,0,0,0, 0,0,0,0,0,0,
  0,0,s_Imm_20y6,0,0, 0,0,0,0,0,0,
  0,0,0,s_Imm_20y6, s_Imm_20y6*r_Imm*(1-b_Imm)+s_Imm_20y6*(1-r_Imm), 0,0,0,0,0,0,
  0,0,0,0, s_Imm_20y6*r_Imm*b_Imm*k_Imm, 0,0,0,0,0,0,
  0,0,0,0, s_Imm_20y6*r_Imm*b_Imm*(1-k_Imm), 0,0,0,0,0,0,
  0,0,0,0,0, s_IS_20y6*r_IS*b_IS*k_IS, s_IF_20y6*r_IF*b_IF*k_IF, s_ES_20y6*r_ES*b_ES*k_ES, s_EF_20y6*r_EF*b_EF*k_EF, s_ENB_20y6*r_ENB*b_ENB*k_ENB, s_PF_20y6*r_PF*b_PF*k_PF,
  0,0,0,0,0, s_IS_20y6*r_IS*b_IS*(1-k_IS), s_IF_20y6*r_IF*b_IF*(1-k_IF), s_ES_20y6*r_ES*b_ES*(1-k_ES), s_EF_20y6*r_EF*b_EF*(1-k_EF), s_ENB_20y6*r_ENB*b_ENB*(1-k_ENB), s_PF_20y6*r_PF*b_PF*(1-k_PF),
  0,0,0,0,0, s_IS_20y6*r_IS*(1-b_IS), s_IF_20y6*r_IF*(1-b_IF), s_ES_20y6*r_ES*(1-b_ES), s_EF_20y6*r_EF*(1-b_EF), s_ENB_20y6*r_ENB*(1-b_ENB), s_PF_20y6*r_PF*(1-b_PF),
  0,0,0,0,0, s_IS_20y6*(1-r_IS), s_IF_20y6*(1-r_IF), s_ES_20y6*(1-r_ES), s_EF_20y6*(1-r_EF), s_ENB_20y6*(1-r_ENB), s_PF_20y6*(1-r_PF)
), nrow=11, byrow=TRUE)

m_20y7 <- matrix(data = c(
  0,0,0,0,0, 0.5*s_Juv_20y7,0,0.5*s_Juv_20y7,0,0,0,
  s_Juv_20y7,0,0,0,0, 0,0,0,0,0,0,
  0,s_Juv_20y7,0,0,0, 0,0,0,0,0,0,
  0,0,s_Imm_20y7,0,0, 0,0,0,0,0,0,
  0,0,0,s_Imm_20y7, s_Imm_20y7*r_Imm*(1-b_Imm)+s_Imm_20y7*(1-r_Imm), 0,0,0,0,0,0,
  0,0,0,0, s_Imm_20y7*r_Imm*b_Imm*k_Imm, 0,0,0,0,0,0,
  0,0,0,0, s_Imm_20y7*r_Imm*b_Imm*(1-k_Imm), 0,0,0,0,0,0,
  0,0,0,0,0, s_IS_20y7*r_IS*b_IS*k_IS, s_IF_20y7*r_IF*b_IF*k_IF, s_ES_20y7*r_ES*b_ES*k_ES, s_EF_20y7*r_EF*b_EF*k_EF, s_ENB_20y7*r_ENB*b_ENB*k_ENB, s_PF_20y7*r_PF*b_PF*k_PF,
  0,0,0,0,0, s_IS_20y7*r_IS*b_IS*(1-k_IS), s_IF_20y7*r_IF*b_IF*(1-k_IF), s_ES_20y7*r_ES*b_ES*(1-k_ES), s_EF_20y7*r_EF*b_EF*(1-k_EF), s_ENB_20y7*r_ENB*b_ENB*(1-k_ENB), s_PF_20y7*r_PF*b_PF*(1-k_PF),
  0,0,0,0,0, s_IS_20y7*r_IS*(1-b_IS), s_IF_20y7*r_IF*(1-b_IF), s_ES_20y7*r_ES*(1-b_ES), s_EF_20y7*r_EF*(1-b_EF), s_ENB_20y7*r_ENB*(1-b_ENB), s_PF_20y7*r_PF*(1-b_PF),
  0,0,0,0,0, s_IS_20y7*(1-r_IS), s_IF_20y7*(1-r_IF), s_ES_20y7*(1-r_ES), s_EF_20y7*(1-r_EF), s_ENB_20y7*(1-r_ENB), s_PF_20y7*(1-r_PF)
), nrow=11, byrow=TRUE)

m_20y8 <- matrix(data = c(
  0,0,0,0,0, 0.5*s_Juv_20y8,0,0.5*s_Juv_20y8,0,0,0,
  s_Juv_20y8,0,0,0,0, 0,0,0,0,0,0,
  0,s_Juv_20y8,0,0,0, 0,0,0,0,0,0,
  0,0,s_Imm_20y8,0,0, 0,0,0,0,0,0,
  0,0,0,s_Imm_20y8, s_Imm_20y8*r_Imm*(1-b_Imm)+s_Imm_20y8*(1-r_Imm), 0,0,0,0,0,0,
  0,0,0,0, s_Imm_20y8*r_Imm*b_Imm*k_Imm, 0,0,0,0,0,0,
  0,0,0,0, s_Imm_20y8*r_Imm*b_Imm*(1-k_Imm), 0,0,0,0,0,0,
  0,0,0,0,0, s_IS_20y8*r_IS*b_IS*k_IS, s_IF_20y8*r_IF*b_IF*k_IF, s_ES_20y8*r_ES*b_ES*k_ES, s_EF_20y8*r_EF*b_EF*k_EF, s_ENB_20y8*r_ENB*b_ENB*k_ENB, s_PF_20y8*r_PF*b_PF*k_PF,
  0,0,0,0,0, s_IS_20y8*r_IS*b_IS*(1-k_IS), s_IF_20y8*r_IF*b_IF*(1-k_IF), s_ES_20y8*r_ES*b_ES*(1-k_ES), s_EF_20y8*r_EF*b_EF*(1-k_EF), s_ENB_20y8*r_ENB*b_ENB*(1-k_ENB), s_PF_20y8*r_PF*b_PF*(1-k_PF),
  0,0,0,0,0, s_IS_20y8*r_IS*(1-b_IS), s_IF_20y8*r_IF*(1-b_IF), s_ES_20y8*r_ES*(1-b_ES), s_EF_20y8*r_EF*(1-b_EF), s_ENB_20y8*r_ENB*(1-b_ENB), s_PF_20y8*r_PF*(1-b_PF),
  0,0,0,0,0, s_IS_20y8*(1-r_IS), s_IF_20y8*(1-r_IF), s_ES_20y8*(1-r_ES), s_EF_20y8*(1-r_EF), s_ENB_20y8*(1-r_ENB), s_PF_20y8*(1-r_PF)
), nrow=11, byrow=TRUE)

m_20y9 <- matrix(data = c(
  0,0,0,0,0, 0.5*s_Juv_20y9,0,0.5*s_Juv_20y9,0,0,0,
  s_Juv_20y9,0,0,0,0, 0,0,0,0,0,0,
  0,s_Juv_20y9,0,0,0, 0,0,0,0,0,0,
  0,0,s_Imm_20y9,0,0, 0,0,0,0,0,0,
  0,0,0,s_Imm_20y9, s_Imm_20y9*r_Imm*(1-b_Imm)+s_Imm_20y9*(1-r_Imm), 0,0,0,0,0,0,
  0,0,0,0, s_Imm_20y9*r_Imm*b_Imm*k_Imm, 0,0,0,0,0,0,
  0,0,0,0, s_Imm_20y9*r_Imm*b_Imm*(1-k_Imm), 0,0,0,0,0,0,
  0,0,0,0,0, s_IS_20y9*r_IS*b_IS*k_IS, s_IF_20y9*r_IF*b_IF*k_IF, s_ES_20y9*r_ES*b_ES*k_ES, s_EF_20y9*r_EF*b_EF*k_EF, s_ENB_20y9*r_ENB*b_ENB*k_ENB, s_PF_20y9*r_PF*b_PF*k_PF,
  0,0,0,0,0, s_IS_20y9*r_IS*b_IS*(1-k_IS), s_IF_20y9*r_IF*b_IF*(1-k_IF), s_ES_20y9*r_ES*b_ES*(1-k_ES), s_EF_20y9*r_EF*b_EF*(1-k_EF), s_ENB_20y9*r_ENB*b_ENB*(1-k_ENB), s_PF_20y9*r_PF*b_PF*(1-k_PF),
  0,0,0,0,0, s_IS_20y9*r_IS*(1-b_IS), s_IF_20y9*r_IF*(1-b_IF), s_ES_20y9*r_ES*(1-b_ES), s_EF_20y9*r_EF*(1-b_EF), s_ENB_20y9*r_ENB*(1-b_ENB), s_PF_20y9*r_PF*(1-b_PF),
  0,0,0,0,0, s_IS_20y9*(1-r_IS), s_IF_20y9*(1-r_IF), s_ES_20y9*(1-r_ES), s_EF_20y9*(1-r_EF), s_ENB_20y9*(1-r_ENB), s_PF_20y9*(1-r_PF)
), nrow=11, byrow=TRUE)

m_20y10 <- matrix(data = c(
  0,0,0,0,0, 0.5*s_Juv_20y10,0,0.5*s_Juv_20y10,0,0,0,
  s_Juv_20y10,0,0,0,0, 0,0,0,0,0,0,
  0,s_Juv_20y10,0,0,0, 0,0,0,0,0,0,
  0,0,s_Imm_20y10,0,0, 0,0,0,0,0,0,
  0,0,0,s_Imm_20y10, s_Imm_20y10*r_Imm*(1-b_Imm)+s_Imm_20y10*(1-r_Imm), 0,0,0,0,0,0,
  0,0,0,0, s_Imm_20y10*r_Imm*b_Imm*k_Imm, 0,0,0,0,0,0,
  0,0,0,0, s_Imm_20y10*r_Imm*b_Imm*(1-k_Imm), 0,0,0,0,0,0,
  0,0,0,0,0, s_IS_20y10*r_IS*b_IS*k_IS, s_IF_20y10*r_IF*b_IF*k_IF, s_ES_20y10*r_ES*b_ES*k_ES, s_EF_20y10*r_EF*b_EF*k_EF, s_ENB_20y10*r_ENB*b_ENB*k_ENB, s_PF_20y10*r_PF*b_PF*k_PF,
  0,0,0,0,0, s_IS_20y10*r_IS*b_IS*(1-k_IS), s_IF_20y10*r_IF*b_IF*(1-k_IF), s_ES_20y10*r_ES*b_ES*(1-k_ES), s_EF_20y10*r_EF*b_EF*(1-k_EF), s_ENB_20y10*r_ENB*b_ENB*(1-k_ENB), s_PF_20y10*r_PF*b_PF*(1-k_PF),
  0,0,0,0,0, s_IS_20y10*r_IS*(1-b_IS), s_IF_20y10*r_IF*(1-b_IF), s_ES_20y10*r_ES*(1-b_ES), s_EF_20y10*r_EF*(1-b_EF), s_ENB_20y10*r_ENB*(1-b_ENB), s_PF_20y10*r_PF*(1-b_PF),
  0,0,0,0,0, s_IS_20y10*(1-r_IS), s_IF_20y10*(1-r_IF), s_ES_20y10*(1-r_ES), s_EF_20y10*(1-r_EF), s_ENB_20y10*(1-r_ENB), s_PF_20y10*(1-r_PF)
), nrow=11, byrow=TRUE)

m_20y11 <- matrix(data = c(
  0,0,0,0,0, 0.5*s_Juv_20y11,0,0.5*s_Juv_20y11,0,0,0,
  s_Juv_20y11,0,0,0,0, 0,0,0,0,0,0,
  0,s_Juv_20y11,0,0,0, 0,0,0,0,0,0,
  0,0,s_Imm_20y11,0,0, 0,0,0,0,0,0,
  0,0,0,s_Imm_20y11, s_Imm_20y11*r_Imm*(1-b_Imm)+s_Imm_20y11*(1-r_Imm), 0,0,0,0,0,0,
  0,0,0,0, s_Imm_20y11*r_Imm*b_Imm*k_Imm, 0,0,0,0,0,0,
  0,0,0,0, s_Imm_20y11*r_Imm*b_Imm*(1-k_Imm), 0,0,0,0,0,0,
  0,0,0,0,0, s_IS_20y11*r_IS*b_IS*k_IS, s_IF_20y11*r_IF*b_IF*k_IF, s_ES_20y11*r_ES*b_ES*k_ES, s_EF_20y11*r_EF*b_EF*k_EF, s_ENB_20y11*r_ENB*b_ENB*k_ENB, s_PF_20y11*r_PF*b_PF*k_PF,
  0,0,0,0,0, s_IS_20y11*r_IS*b_IS*(1-k_IS), s_IF_20y11*r_IF*b_IF*(1-k_IF), s_ES_20y11*r_ES*b_ES*(1-k_ES), s_EF_20y11*r_EF*b_EF*(1-k_EF), s_ENB_20y11*r_ENB*b_ENB*(1-k_ENB), s_PF_20y11*r_PF*b_PF*(1-k_PF),
  0,0,0,0,0, s_IS_20y11*r_IS*(1-b_IS), s_IF_20y11*r_IF*(1-b_IF), s_ES_20y11*r_ES*(1-b_ES), s_EF_20y11*r_EF*(1-b_EF), s_ENB_20y11*r_ENB*(1-b_ENB), s_PF_20y11*r_PF*(1-b_PF),
  0,0,0,0,0, s_IS_20y11*(1-r_IS), s_IF_20y11*(1-r_IF), s_ES_20y11*(1-r_ES), s_EF_20y11*(1-r_EF), s_ENB_20y11*(1-r_ENB), s_PF_20y11*(1-r_PF)
), nrow=11, byrow=TRUE)

m_20y12 <- matrix(data = c(
  0,0,0,0,0, 0.5*s_Juv_20y12,0,0.5*s_Juv_20y12,0,0,0,
  s_Juv_20y12,0,0,0,0, 0,0,0,0,0,0,
  0,s_Juv_20y12,0,0,0, 0,0,0,0,0,0,
  0,0,s_Imm_20y12,0,0, 0,0,0,0,0,0,
  0,0,0,s_Imm_20y12, s_Imm_20y12*r_Imm*(1-b_Imm)+s_Imm_20y12*(1-r_Imm), 0,0,0,0,0,0,
  0,0,0,0, s_Imm_20y12*r_Imm*b_Imm*k_Imm, 0,0,0,0,0,0,
  0,0,0,0, s_Imm_20y12*r_Imm*b_Imm*(1-k_Imm), 0,0,0,0,0,0,
  0,0,0,0,0, s_IS_20y12*r_IS*b_IS*k_IS, s_IF_20y12*r_IF*b_IF*k_IF, s_ES_20y12*r_ES*b_ES*k_ES, s_EF_20y12*r_EF*b_EF*k_EF, s_ENB_20y12*r_ENB*b_ENB*k_ENB, s_PF_20y12*r_PF*b_PF*k_PF,
  0,0,0,0,0, s_IS_20y12*r_IS*b_IS*(1-k_IS), s_IF_20y12*r_IF*b_IF*(1-k_IF), s_ES_20y12*r_ES*b_ES*(1-k_ES), s_EF_20y12*r_EF*b_EF*(1-k_EF), s_ENB_20y12*r_ENB*b_ENB*(1-k_ENB), s_PF_20y12*r_PF*b_PF*(1-k_PF),
  0,0,0,0,0, s_IS_20y12*r_IS*(1-b_IS), s_IF_20y12*r_IF*(1-b_IF), s_ES_20y12*r_ES*(1-b_ES), s_EF_20y12*r_EF*(1-b_EF), s_ENB_20y12*r_ENB*(1-b_ENB), s_PF_20y12*r_PF*(1-b_PF),
  0,0,0,0,0, s_IS_20y12*(1-r_IS), s_IF_20y12*(1-r_IF), s_ES_20y12*(1-r_ES), s_EF_20y12*(1-r_EF), s_ENB_20y12*(1-r_ENB), s_PF_20y12*(1-r_PF)
), nrow=11, byrow=TRUE)

m_20y13 <- matrix(data = c(
  0,0,0,0,0, 0.5*s_Juv_20y13,0,0.5*s_Juv_20y13,0,0,0,
  s_Juv_20y13,0,0,0,0, 0,0,0,0,0,0,
  0,s_Juv_20y13,0,0,0, 0,0,0,0,0,0,
  0,0,s_Imm_20y13,0,0, 0,0,0,0,0,0,
  0,0,0,s_Imm_20y13, s_Imm_20y13*r_Imm*(1-b_Imm)+s_Imm_20y13*(1-r_Imm), 0,0,0,0,0,0,
  0,0,0,0, s_Imm_20y13*r_Imm*b_Imm*k_Imm, 0,0,0,0,0,0,
  0,0,0,0, s_Imm_20y13*r_Imm*b_Imm*(1-k_Imm), 0,0,0,0,0,0,
  0,0,0,0,0, s_IS_20y13*r_IS*b_IS*k_IS, s_IF_20y13*r_IF*b_IF*k_IF, s_ES_20y13*r_ES*b_ES*k_ES, s_EF_20y13*r_EF*b_EF*k_EF, s_ENB_20y13*r_ENB*b_ENB*k_ENB, s_PF_20y13*r_PF*b_PF*k_PF,
  0,0,0,0,0, s_IS_20y13*r_IS*b_IS*(1-k_IS), s_IF_20y13*r_IF*b_IF*(1-k_IF), s_ES_20y13*r_ES*b_ES*(1-k_ES), s_EF_20y13*r_EF*b_EF*(1-k_EF), s_ENB_20y13*r_ENB*b_ENB*(1-k_ENB), s_PF_20y13*r_PF*b_PF*(1-k_PF),
  0,0,0,0,0, s_IS_20y13*r_IS*(1-b_IS), s_IF_20y13*r_IF*(1-b_IF), s_ES_20y13*r_ES*(1-b_ES), s_EF_20y13*r_EF*(1-b_EF), s_ENB_20y13*r_ENB*(1-b_ENB), s_PF_20y13*r_PF*(1-b_PF),
  0,0,0,0,0, s_IS_20y13*(1-r_IS), s_IF_20y13*(1-r_IF), s_ES_20y13*(1-r_ES), s_EF_20y13*(1-r_EF), s_ENB_20y13*(1-r_ENB), s_PF_20y13*(1-r_PF)
), nrow=11, byrow=TRUE)

m_20y14 <- matrix(data = c(
  0,0,0,0,0, 0.5*s_Juv_20y14,0,0.5*s_Juv_20y14,0,0,0,
  s_Juv_20y14,0,0,0,0, 0,0,0,0,0,0,
  0,s_Juv_20y14,0,0,0, 0,0,0,0,0,0,
  0,0,s_Imm_20y14,0,0, 0,0,0,0,0,0,
  0,0,0,s_Imm_20y14, s_Imm_20y14*r_Imm*(1-b_Imm)+s_Imm_20y14*(1-r_Imm), 0,0,0,0,0,0,
  0,0,0,0, s_Imm_20y14*r_Imm*b_Imm*k_Imm, 0,0,0,0,0,0,
  0,0,0,0, s_Imm_20y14*r_Imm*b_Imm*(1-k_Imm), 0,0,0,0,0,0,
  0,0,0,0,0, s_IS_20y14*r_IS*b_IS*k_IS, s_IF_20y14*r_IF*b_IF*k_IF, s_ES_20y14*r_ES*b_ES*k_ES, s_EF_20y14*r_EF*b_EF*k_EF, s_ENB_20y14*r_ENB*b_ENB*k_ENB, s_PF_20y14*r_PF*b_PF*k_PF,
  0,0,0,0,0, s_IS_20y14*r_IS*b_IS*(1-k_IS), s_IF_20y14*r_IF*b_IF*(1-k_IF), s_ES_20y14*r_ES*b_ES*(1-k_ES), s_EF_20y14*r_EF*b_EF*(1-k_EF), s_ENB_20y14*r_ENB*b_ENB*(1-k_ENB), s_PF_20y14*r_PF*b_PF*(1-k_PF),
  0,0,0,0,0, s_IS_20y14*r_IS*(1-b_IS), s_IF_20y14*r_IF*(1-b_IF), s_ES_20y14*r_ES*(1-b_ES), s_EF_20y14*r_EF*(1-b_EF), s_ENB_20y14*r_ENB*(1-b_ENB), s_PF_20y14*r_PF*(1-b_PF),
  0,0,0,0,0, s_IS_20y14*(1-r_IS), s_IF_20y14*(1-r_IF), s_ES_20y14*(1-r_ES), s_EF_20y14*(1-r_EF), s_ENB_20y14*(1-r_ENB), s_PF_20y14*(1-r_PF)
), nrow=11, byrow=TRUE)

m_20y15 <- matrix(data = c(
  0,0,0,0,0, 0.5*s_Juv_20y15,0,0.5*s_Juv_20y15,0,0,0,
  s_Juv_20y15,0,0,0,0, 0,0,0,0,0,0,
  0,s_Juv_20y15,0,0,0, 0,0,0,0,0,0,
  0,0,s_Imm_20y15,0,0, 0,0,0,0,0,0,
  0,0,0,s_Imm_20y15, s_Imm_20y15*r_Imm*(1-b_Imm)+s_Imm_20y15*(1-r_Imm), 0,0,0,0,0,0,
  0,0,0,0, s_Imm_20y15*r_Imm*b_Imm*k_Imm, 0,0,0,0,0,0,
  0,0,0,0, s_Imm_20y15*r_Imm*b_Imm*(1-k_Imm), 0,0,0,0,0,0,
  0,0,0,0,0, s_IS_20y15*r_IS*b_IS*k_IS, s_IF_20y15*r_IF*b_IF*k_IF, s_ES_20y15*r_ES*b_ES*k_ES, s_EF_20y15*r_EF*b_EF*k_EF, s_ENB_20y15*r_ENB*b_ENB*k_ENB, s_PF_20y15*r_PF*b_PF*k_PF,
  0,0,0,0,0, s_IS_20y15*r_IS*b_IS*(1-k_IS), s_IF_20y15*r_IF*b_IF*(1-k_IF), s_ES_20y15*r_ES*b_ES*(1-k_ES), s_EF_20y15*r_EF*b_EF*(1-k_EF), s_ENB_20y15*r_ENB*b_ENB*(1-k_ENB), s_PF_20y15*r_PF*b_PF*(1-k_PF),
  0,0,0,0,0, s_IS_20y15*r_IS*(1-b_IS), s_IF_20y15*r_IF*(1-b_IF), s_ES_20y15*r_ES*(1-b_ES), s_EF_20y15*r_EF*(1-b_EF), s_ENB_20y15*r_ENB*(1-b_ENB), s_PF_20y15*r_PF*(1-b_PF),
  0,0,0,0,0, s_IS_20y15*(1-r_IS), s_IF_20y15*(1-r_IF), s_ES_20y15*(1-r_ES), s_EF_20y15*(1-r_EF), s_ENB_20y15*(1-r_ENB), s_PF_20y15*(1-r_PF)
), nrow=11, byrow=TRUE)

m_20y16 <- matrix(data = c(
  0,0,0,0,0, 0.5*s_Juv_20y16,0,0.5*s_Juv_20y16,0,0,0,
  s_Juv_20y16,0,0,0,0, 0,0,0,0,0,0,
  0,s_Juv_20y16,0,0,0, 0,0,0,0,0,0,
  0,0,s_Imm_20y16,0,0, 0,0,0,0,0,0,
  0,0,0,s_Imm_20y16, s_Imm_20y16*r_Imm*(1-b_Imm)+s_Imm_20y16*(1-r_Imm), 0,0,0,0,0,0,
  0,0,0,0, s_Imm_20y16*r_Imm*b_Imm*k_Imm, 0,0,0,0,0,0,
  0,0,0,0, s_Imm_20y16*r_Imm*b_Imm*(1-k_Imm), 0,0,0,0,0,0,
  0,0,0,0,0, s_IS_20y16*r_IS*b_IS*k_IS, s_IF_20y16*r_IF*b_IF*k_IF, s_ES_20y16*r_ES*b_ES*k_ES, s_EF_20y16*r_EF*b_EF*k_EF, s_ENB_20y16*r_ENB*b_ENB*k_ENB, s_PF_20y16*r_PF*b_PF*k_PF,
  0,0,0,0,0, s_IS_20y16*r_IS*b_IS*(1-k_IS), s_IF_20y16*r_IF*b_IF*(1-k_IF), s_ES_20y16*r_ES*b_ES*(1-k_ES), s_EF_20y16*r_EF*b_EF*(1-k_EF), s_ENB_20y16*r_ENB*b_ENB*(1-k_ENB), s_PF_20y16*r_PF*b_PF*(1-k_PF),
  0,0,0,0,0, s_IS_20y16*r_IS*(1-b_IS), s_IF_20y16*r_IF*(1-b_IF), s_ES_20y16*r_ES*(1-b_ES), s_EF_20y16*r_EF*(1-b_EF), s_ENB_20y16*r_ENB*(1-b_ENB), s_PF_20y16*r_PF*(1-b_PF),
  0,0,0,0,0, s_IS_20y16*(1-r_IS), s_IF_20y16*(1-r_IF), s_ES_20y16*(1-r_ES), s_EF_20y16*(1-r_EF), s_ENB_20y16*(1-r_ENB), s_PF_20y16*(1-r_PF)
), nrow=11, byrow=TRUE)

m_20y17 <- matrix(data = c(
  0,0,0,0,0, 0.5*s_Juv_20y17,0,0.5*s_Juv_20y17,0,0,0,
  s_Juv_20y17,0,0,0,0, 0,0,0,0,0,0,
  0,s_Juv_20y17,0,0,0, 0,0,0,0,0,0,
  0,0,s_Imm_20y17,0,0, 0,0,0,0,0,0,
  0,0,0,s_Imm_20y17, s_Imm_20y17*r_Imm*(1-b_Imm)+s_Imm_20y17*(1-r_Imm), 0,0,0,0,0,0,
  0,0,0,0, s_Imm_20y17*r_Imm*b_Imm*k_Imm, 0,0,0,0,0,0,
  0,0,0,0, s_Imm_20y17*r_Imm*b_Imm*(1-k_Imm), 0,0,0,0,0,0,
  0,0,0,0,0, s_IS_20y17*r_IS*b_IS*k_IS, s_IF_20y17*r_IF*b_IF*k_IF, s_ES_20y17*r_ES*b_ES*k_ES, s_EF_20y17*r_EF*b_EF*k_EF, s_ENB_20y17*r_ENB*b_ENB*k_ENB, s_PF_20y17*r_PF*b_PF*k_PF,
  0,0,0,0,0, s_IS_20y17*r_IS*b_IS*(1-k_IS), s_IF_20y17*r_IF*b_IF*(1-k_IF), s_ES_20y17*r_ES*b_ES*(1-k_ES), s_EF_20y17*r_EF*b_EF*(1-k_EF), s_ENB_20y17*r_ENB*b_ENB*(1-k_ENB), s_PF_20y17*r_PF*b_PF*(1-k_PF),
  0,0,0,0,0, s_IS_20y17*r_IS*(1-b_IS), s_IF_20y17*r_IF*(1-b_IF), s_ES_20y17*r_ES*(1-b_ES), s_EF_20y17*r_EF*(1-b_EF), s_ENB_20y17*r_ENB*(1-b_ENB), s_PF_20y17*r_PF*(1-b_PF),
  0,0,0,0,0, s_IS_20y17*(1-r_IS), s_IF_20y17*(1-r_IF), s_ES_20y17*(1-r_ES), s_EF_20y17*(1-r_EF), s_ENB_20y17*(1-r_ENB), s_PF_20y17*(1-r_PF)
), nrow=11, byrow=TRUE)

m_20y18 <- matrix(data = c(
  0,0,0,0,0, 0.5*s_Juv_20y18,0,0.5*s_Juv_20y18,0,0,0,
  s_Juv_20y18,0,0,0,0, 0,0,0,0,0,0,
  0,s_Juv_20y18,0,0,0, 0,0,0,0,0,0,
  0,0,s_Imm_20y18,0,0, 0,0,0,0,0,0,
  0,0,0,s_Imm_20y18, s_Imm_20y18*r_Imm*(1-b_Imm)+s_Imm_20y18*(1-r_Imm), 0,0,0,0,0,0,
  0,0,0,0, s_Imm_20y18*r_Imm*b_Imm*k_Imm, 0,0,0,0,0,0,
  0,0,0,0, s_Imm_20y18*r_Imm*b_Imm*(1-k_Imm), 0,0,0,0,0,0,
  0,0,0,0,0, s_IS_20y18*r_IS*b_IS*k_IS, s_IF_20y18*r_IF*b_IF*k_IF, s_ES_20y18*r_ES*b_ES*k_ES, s_EF_20y18*r_EF*b_EF*k_EF, s_ENB_20y18*r_ENB*b_ENB*k_ENB, s_PF_20y18*r_PF*b_PF*k_PF,
  0,0,0,0,0, s_IS_20y18*r_IS*b_IS*(1-k_IS), s_IF_20y18*r_IF*b_IF*(1-k_IF), s_ES_20y18*r_ES*b_ES*(1-k_ES), s_EF_20y18*r_EF*b_EF*(1-k_EF), s_ENB_20y18*r_ENB*b_ENB*(1-k_ENB), s_PF_20y18*r_PF*b_PF*(1-k_PF),
  0,0,0,0,0, s_IS_20y18*r_IS*(1-b_IS), s_IF_20y18*r_IF*(1-b_IF), s_ES_20y18*r_ES*(1-b_ES), s_EF_20y18*r_EF*(1-b_EF), s_ENB_20y18*r_ENB*(1-b_ENB), s_PF_20y18*r_PF*(1-b_PF),
  0,0,0,0,0, s_IS_20y18*(1-r_IS), s_IF_20y18*(1-r_IF), s_ES_20y18*(1-r_ES), s_EF_20y18*(1-r_EF), s_ENB_20y18*(1-r_ENB), s_PF_20y18*(1-r_PF)
), nrow=11, byrow=TRUE)

m_20y19 <- matrix(data = c(
  0,0,0,0,0, 0.5*s_Juv_20y19,0,0.5*s_Juv_20y19,0,0,0,
  s_Juv_20y19,0,0,0,0, 0,0,0,0,0,0,
  0,s_Juv_20y19,0,0,0, 0,0,0,0,0,0,
  0,0,s_Imm_20y19,0,0, 0,0,0,0,0,0,
  0,0,0,s_Imm_20y19, s_Imm_20y19*r_Imm*(1-b_Imm)+s_Imm_20y19*(1-r_Imm), 0,0,0,0,0,0,
  0,0,0,0, s_Imm_20y19*r_Imm*b_Imm*k_Imm, 0,0,0,0,0,0,
  0,0,0,0, s_Imm_20y19*r_Imm*b_Imm*(1-k_Imm), 0,0,0,0,0,0,
  0,0,0,0,0, s_IS_20y19*r_IS*b_IS*k_IS, s_IF_20y19*r_IF*b_IF*k_IF, s_ES_20y19*r_ES*b_ES*k_ES, s_EF_20y19*r_EF*b_EF*k_EF, s_ENB_20y19*r_ENB*b_ENB*k_ENB, s_PF_20y19*r_PF*b_PF*k_PF,
  0,0,0,0,0, s_IS_20y19*r_IS*b_IS*(1-k_IS), s_IF_20y19*r_IF*b_IF*(1-k_IF), s_ES_20y19*r_ES*b_ES*(1-k_ES), s_EF_20y19*r_EF*b_EF*(1-k_EF), s_ENB_20y19*r_ENB*b_ENB*(1-k_ENB), s_PF_20y19*r_PF*b_PF*(1-k_PF),
  0,0,0,0,0, s_IS_20y19*r_IS*(1-b_IS), s_IF_20y19*r_IF*(1-b_IF), s_ES_20y19*r_ES*(1-b_ES), s_EF_20y19*r_EF*(1-b_EF), s_ENB_20y19*r_ENB*(1-b_ENB), s_PF_20y19*r_PF*(1-b_PF),
  0,0,0,0,0, s_IS_20y19*(1-r_IS), s_IF_20y19*(1-r_IF), s_ES_20y19*(1-r_ES), s_EF_20y19*(1-r_EF), s_ENB_20y19*(1-r_ENB), s_PF_20y19*(1-r_PF)
), nrow=11, byrow=TRUE)

# m_20y20: 100% improvement — should equal m_3of3
m_20y20 <- matrix(data = c(
  0,0,0,0,0, 0.5*s_Juv_20y20,0,0.5*s_Juv_20y20,0,0,0,
  s_Juv_20y20,0,0,0,0, 0,0,0,0,0,0,
  0,s_Juv_20y20,0,0,0, 0,0,0,0,0,0,
  0,0,s_Imm_20y20,0,0, 0,0,0,0,0,0,
  0,0,0,s_Imm_20y20, s_Imm_20y20*r_Imm*(1-b_Imm)+s_Imm_20y20*(1-r_Imm), 0,0,0,0,0,0,
  0,0,0,0, s_Imm_20y20*r_Imm*b_Imm*k_Imm, 0,0,0,0,0,0,
  0,0,0,0, s_Imm_20y20*r_Imm*b_Imm*(1-k_Imm), 0,0,0,0,0,0,
  0,0,0,0,0, s_IS_20y20*r_IS*b_IS*k_IS, s_IF_20y20*r_IF*b_IF*k_IF, s_ES_20y20*r_ES*b_ES*k_ES, s_EF_20y20*r_EF*b_EF*k_EF, s_ENB_20y20*r_ENB*b_ENB*k_ENB, s_PF_20y20*r_PF*b_PF*k_PF,
  0,0,0,0,0, s_IS_20y20*r_IS*b_IS*(1-k_IS), s_IF_20y20*r_IF*b_IF*(1-k_IF), s_ES_20y20*r_ES*b_ES*(1-k_ES), s_EF_20y20*r_EF*b_EF*(1-k_EF), s_ENB_20y20*r_ENB*b_ENB*(1-k_ENB), s_PF_20y20*r_PF*b_PF*(1-k_PF),
  0,0,0,0,0, s_IS_20y20*r_IS*(1-b_IS), s_IF_20y20*r_IF*(1-b_IF), s_ES_20y20*r_ES*(1-b_ES), s_EF_20y20*r_EF*(1-b_EF), s_ENB_20y20*r_ENB*(1-b_ENB), s_PF_20y20*r_PF*(1-b_PF),
  0,0,0,0,0, s_IS_20y20*(1-r_IS), s_IF_20y20*(1-r_IF), s_ES_20y20*(1-r_ES), s_EF_20y20*(1-r_EF), s_ENB_20y20*(1-r_ENB), s_PF_20y20*(1-r_PF)
), nrow=11, byrow=TRUE)

phase_list_20y <- list(m_20y1, m_20y2, m_20y3, m_20y4, m_20y5,
                       m_20y6, m_20y7, m_20y8, m_20y9, m_20y10,
                       m_20y11, m_20y12, m_20y13, m_20y14, m_20y15,
                       m_20y16, m_20y17, m_20y18, m_20y19, m_20y20)
```

**project with 5-year phase-in**
```{r part3-project}
mat_phasein <- matrix(0, nrow = years + 1, ncol = 11)
colnames(mat_phasein) <- stage_names
mat_phasein[1, ] <- n0_vec

for (k in 1:years) {
  A <- if (k <= 5) phase_list[[k]] else m_3of3
  mat_phasein[k + 1, ] <- A %*% mat_phasein[k, ]
}

proj_part3_total <- as.data.frame(mat_phasein) %>%
  mutate(year     = 0:years,
         N        = rowSums(across(all_of(stage_names))),
         scenario = "5-year phase-in") %>%
  select(year, N, scenario)
```

**plot 3a — 100-year 5-yr phase-in projection, N**
```{r plot-3a, fig.width = 9, fig.height = 5}
# combine 100-year baseline/immediate with 5-yr phase-in for comparison
proj_part3_compare <- bind_rows(
  proj_part2_total,
  proj_part3_total
) %>%
  mutate(scenario = factor(scenario, levels = c("No mitigation (baseline)",
                                                 "Immediate 3/3 (100% coverage)",
                                                 "5-year phase-in")))

p_3a <- ggplot(proj_part3_compare, aes(x = year, y = N, color = scenario, linetype = scenario)) +
  geom_line(linewidth = 0.9) +
  geom_hline(yintercept = N0, color = "grey50", linetype = "dashed", linewidth = 0.4) +
  geom_point(data = bird_obs, aes(x = year, y = N_scaled),
             inherit.aes = FALSE, color = "#e6550d", size = 3, shape = 16) +
  geom_line(data = bird_obs, aes(x = year, y = N_scaled),
            inherit.aes = FALSE, color = "#e6550d", linewidth = 0.5, linetype = "dotted") +
  geom_text(data = bird_obs, aes(x = year, y = N_scaled, label = season),
            inherit.aes = FALSE, color = "#e6550d", size = 2.5, vjust = -0.8) +
  scale_color_manual(values = c("No mitigation (baseline)"      = "#d73027",
                                 "Immediate 3/3 (100% coverage)" = "#253494",
                                 "5-year phase-in"               = "#1a9641")) +
  scale_linetype_manual(values = c("No mitigation (baseline)"      = "dashed",
                                    "Immediate 3/3 (100% coverage)" = "solid",
                                    "5-year phase-in"               = "solid")) +
  scale_x_continuous(breaks = seq(0, years, 10), limits = c(-10, years)) +
  scale_y_continuous(labels = scales::comma,
                     breaks = scales::pretty_breaks(n = 6),
                     sec.axis = sec_axis(~ . / pairs_scale,
                                         name = "Breeding pairs (Bird Island)",
                                         labels = scales::comma)) +
  labs(x = "Years since 1990 baseline",
       y = "Population (N individuals)",
       color = NULL, linetype = NULL,
       title = paste0("100-year projection with 5-yr phase-in — ", target_period, " baseline"),
       subtitle = "Orange dots = census breeding pairs") +
  themeo + theme(legend.position = "bottom")
p_3a
```

**500-year 5-year phase-in projection (Part 3 extension)**
```{r part3-project-long}
mat_phasein_long <- matrix(0, nrow = yearslong + 1, ncol = 11)
colnames(mat_phasein_long) <- stage_names
mat_phasein_long[1, ] <- n0_vec

for (k in 1:yearslong) {
  A <- if (k <= 5) phase_list[[k]] else m_3of3
  mat_phasein_long[k + 1, ] <- A %*% mat_phasein_long[k, ]
}

proj_part3_total_long <- as.data.frame(mat_phasein_long) %>%
  mutate(year     = 0:yearslong,
         N        = rowSums(across(all_of(stage_names))),
         scenario = "5-year phase-in") %>%
  select(year, N, scenario)

# combine with baseline and immediate 3/3 for context in plots
proj_part3_compare_long <- bind_rows(
  proj_part2_total_long,
  proj_part3_total_long
) %>%
  mutate(scenario = factor(scenario, levels = c("No mitigation (baseline)",
                                                 "Immediate 3/3 (100% coverage)",
                                                 "5-year phase-in")))
```

**plot 3c — 500-year 5-yr phase-in projection, N**
```{r plot-3c, fig.width = 9, fig.height = 5}
p_3c <- ggplot(proj_part3_compare_long, aes(x = year, y = N, color = scenario, linetype = scenario)) +
  geom_line(linewidth = 0.9) +
  geom_hline(yintercept = N0, color = "grey50", linetype = "dashed", linewidth = 0.4) +
  geom_point(data = bird_obs, aes(x = year, y = N_scaled),
             inherit.aes = FALSE, color = "#e6550d", size = 3, shape = 16) +
  geom_text(data = bird_obs, aes(x = year, y = N_scaled, label = season),
            inherit.aes = FALSE, color = "#e6550d", size = 2.5, vjust = -0.8) +
  scale_color_manual(values = c("No mitigation (baseline)"      = "#d73027",
                                 "Immediate 3/3 (100% coverage)" = "#253494",
                                 "5-year phase-in"               = "#1a9641")) +
  scale_linetype_manual(values = c("No mitigation (baseline)"      = "dashed",
                                    "Immediate 3/3 (100% coverage)" = "solid",
                                    "5-year phase-in"               = "solid")) +
  scale_x_continuous(breaks = seq(0, yearslong, 100), limits = c(-10, yearslong)) +
  scale_y_continuous(labels = scales::comma,
                     breaks = scales::pretty_breaks(n = 6),
                     sec.axis = sec_axis(~ . / pairs_scale,
                                         name = "Breeding pairs (Bird Island)",
                                         labels = scales::comma)) +
  labs(x = paste0("Years since 1990 baseline (", yearslong, "-year horizon)"),
       y = "Population (N individuals)",
       color = NULL, linetype = NULL,
       title = paste0("500-year projection with 5-yr phase-in — ", target_period, " baseline"),
       subtitle = "Orange dots = census breeding pairs") +
  themeo + theme(legend.position = "bottom")
p_3c
```

**plot 3e — 500-year 5-yr phase-in projection, log(N) y-axis**
```{r plot-3e, fig.width = 9, fig.height = 5}
p_3e <- ggplot(proj_part3_compare_long, aes(x = year, y = N, color = scenario, linetype = scenario)) +
  geom_line(linewidth = 0.9) +
  geom_hline(yintercept = N0, color = "grey50", linetype = "dashed", linewidth = 0.4) +
  scale_color_manual(values = c("No mitigation (baseline)"      = "#d73027",
                                 "Immediate 3/3 (100% coverage)" = "#253494",
                                 "5-year phase-in"               = "#1a9641")) +
  scale_linetype_manual(values = c("No mitigation (baseline)"      = "dashed",
                                    "Immediate 3/3 (100% coverage)" = "solid",
                                    "5-year phase-in"               = "solid")) +
  scale_x_continuous(breaks = seq(0, yearslong, 100), limits = c(0, yearslong)) +
  scale_y_log10(labels = scales::comma,
                breaks = scales::trans_breaks("log10", function(x) 10^x)) +
  labs(x = paste0("Years since 1990 baseline (", yearslong, "-year horizon)"),
       y = "log\u2081\u2080(N individuals)",
       color = NULL, linetype = NULL,
       title = paste0("500-year projection with 5-yr phase-in, log scale — ", target_period, " baseline"))+ #,
      # subtitle = "Green = 5-yr phase-in; converges to immediate 3/3 trajectory after year 5") +
  themeo + theme(legend.position = "bottom")
p_3e
```

**plot 3f — annual lambda (N(t+1)/N(t)) over 500 years, 5-yr phase-in**
```{r plot-3f, fig.width = 9, fig.height = 5}
lambda_ts_3 <- proj_part3_compare_long %>%
  group_by(scenario) %>%
  arrange(year) %>%
  mutate(lambda_t = N / lag(N)) %>%
  filter(!is.na(lambda_t))

p_3f <- ggplot(lambda_ts_3, aes(x = year, y = lambda_t, color = scenario, linetype = scenario)) +
  geom_line(linewidth = 0.9) +
  geom_hline(yintercept = 1, color = "grey50", linetype = "dashed", linewidth = 0.4) +
  scale_color_manual(values = c("No mitigation (baseline)"      = "#d73027",
                                 "Immediate 3/3 (100% coverage)" = "#253494",
                                 "5-year phase-in"               = "#1a9641")) +
  scale_linetype_manual(values = c("No mitigation (baseline)"      = "dashed",
                                    "Immediate 3/3 (100% coverage)" = "solid",
                                    "5-year phase-in"               = "solid")) +
  scale_x_continuous(breaks = seq(0, yearslong, 100), limits = c(0, yearslong)) +
  scale_y_continuous(breaks = scales::pretty_breaks(n = 6)) +
  labs(x = paste0("Years since 1990 baseline (", yearslong, "-year horizon)"),
       y = expression(lambda[t] ~ "(N(t+1) / N(t))"),
       color = NULL, linetype = NULL,
       title = paste0("Annual \u03BB — 5-yr phase-in scenario, 500-year horizon — ", target_period, " baseline"),
       subtitle = paste0("Green line ramps up over years 1\u20135 as survival improves step-by-step,\n",
                         "then levels off at \u03BB (full 3/3); dashed grey = \u03BB = 1")) +
  themeo + theme(legend.position = "bottom")
p_3f
```

---

*PART 6 — ALL PROJECTIONS AND COMPARISON PLOTS*

Matrix selection logic per scenario:
- 10-yr phase-in:           k=1-10: phase_list_10y[[k]];  k>10: m_3of3
- 20-yr phase-in:           k=1-20: phase_list_20y[[k]];  k>20: m_3of3
- 5-yr delay + immediate:   k=1-5:  m_base;               k>5:  m_3of3
- 10-yr delay + immediate:  k=1-10: m_base;               k>10: m_3of3
- 20-yr delay + immediate:  k=1-20: m_base;               k>20: m_3of3
- 5-yr delay + 5-yr phase:  k=1-5:  m_base; k=6-10:  phase_list[[k-5]];  k>10: m_3of3
- 10-yr delay + 5-yr phase: k=1-10: m_base; k=11-15: phase_list[[k-10]]; k>15: m_3of3
- 20-yr delay + 5-yr phase: k=1-20: m_base; k=21-25: phase_list[[k-20]]; k>25: m_3of3

**run all projections**
```{r part6-all-projections}
# Helper: run matrix projection and return total-N data frame
run_proj <- function(phase_fn, scenario_label) {
  mat <- matrix(0, nrow = years + 1, ncol = 11)
  colnames(mat) <- stage_names
  mat[1, ] <- n0_vec
  for (k in 1:years) {
    mat[k + 1, ] <- phase_fn(k) %*% mat[k, ]
  }
  as.data.frame(mat) %>%
    mutate(year     = 0:years,
           N        = rowSums(across(all_of(stage_names))),
           scenario = scenario_label) %>%
    select(year, N, scenario)
}

proj_10y <- run_proj(
  function(k) if (k <= 10) phase_list_10y[[k]] else m_3of3,
  "10-year phase-in")

proj_20y <- run_proj(
  function(k) if (k <= 20) phase_list_20y[[k]] else m_3of3,
  "20-year phase-in")

proj_d5_imm <- run_proj(
  function(k) if (k <= 5)  m_base else m_3of3,
  "5-yr delay, then immediate 3/3")

proj_d10_imm <- run_proj(
  function(k) if (k <= 10) m_base else m_3of3,
  "10-yr delay, then immediate 3/3")

proj_d20_imm <- run_proj(
  function(k) if (k <= 20) m_base else m_3of3,
  "20-yr delay, then immediate 3/3")

proj_d5_p5 <- run_proj(
  function(k) {
    if      (k <= 5)  m_base
    else if (k <= 10) phase_list[[k - 5]]
    else              m_3of3
  },
  "5-yr delay, then 5-yr phase-in")

proj_d10_p5 <- run_proj(
  function(k) {
    if      (k <= 10) m_base
    else if (k <= 15) phase_list[[k - 10]]
    else              m_3of3
  },
  "10-yr delay, then 5-yr phase-in")

proj_d20_p5 <- run_proj(
  function(k) {
    if      (k <= 20) m_base
    else if (k <= 25) phase_list[[k - 20]]
    else              m_3of3
  },
  "20-yr delay, then 5-yr phase-in")

all_scenarios <- bind_rows(
  proj_part1,
  proj_part3_total,
  proj_10y,
  proj_20y,
  proj_d5_imm,
  proj_d10_imm,
  proj_d20_imm,
  proj_d5_p5,
  proj_d10_p5,
  proj_d20_p5
) %>%
  mutate(scenario = factor(scenario, levels = c(
    "No mitigation (baseline)",
    "Immediate 3/3 (100% coverage)",
    "5-year phase-in",
    "10-year phase-in",
    "20-year phase-in",
    "5-yr delay, then immediate 3/3",
    "10-yr delay, then immediate 3/3",
    "20-yr delay, then immediate 3/3",
    "5-yr delay, then 5-yr phase-in",
    "10-yr delay, then 5-yr phase-in",
    "20-yr delay, then 5-yr phase-in"
  )))
```

**plot 6a — phase-in speed comparison**
```{r plot-6a, fig.width = 9, fig.height = 5}
p_6a <- all_scenarios %>%
  filter(scenario %in% c("No mitigation (baseline)",
                          "Immediate 3/3 (100% coverage)",
                          "5-year phase-in",
                          "10-year phase-in",
                          "20-year phase-in")) %>%
  ggplot(aes(x = year, y = N, color = scenario, linetype = scenario)) +
  geom_line(linewidth = 0.9) +
  geom_hline(yintercept = N0, color = "grey50", linetype = "dashed", linewidth = 0.4) +
  geom_point(data = bird_obs, aes(x = year, y = N_scaled),
             inherit.aes = FALSE, color = "#e6550d", size = 3, shape = 16) +
  geom_line(data = bird_obs, aes(x = year, y = N_scaled),
            inherit.aes = FALSE, color = "#e6550d", linewidth = 0.5, linetype = "dotted") +
  geom_text(data = bird_obs, aes(x = year, y = N_scaled, label = season),
            inherit.aes = FALSE, color = "#e6550d", size = 2.5, vjust = -0.8) +
  scale_color_manual(values = c(
    "No mitigation (baseline)"      = "#d73027",
    "Immediate 3/3 (100% coverage)" = "#253494",
    "5-year phase-in"               = "#1a9641",
    "10-year phase-in"              = "#78c679",
    "20-year phase-in"              = "#c2e699")) +
  scale_linetype_manual(values = c(
    "No mitigation (baseline)"      = "dashed",
    "Immediate 3/3 (100% coverage)" = "solid",
    "5-year phase-in"               = "solid",
    "10-year phase-in"              = "solid",
    "20-year phase-in"              = "solid")) +
  scale_x_continuous(breaks = seq(-10, years, 10), limits = c(-10, years)) +
  scale_y_continuous(labels = scales::comma,
                     breaks = scales::pretty_breaks(n = 6),
                     sec.axis = sec_axis(~ . / pairs_scale,
                                         name = "Breeding pairs (Bird Island)",
                                         labels = scales::comma)) +
  labs(x = "Years since 1990 baseline", y = "Population (N individuals)",
       color = NULL, linetype = NULL,
       title = paste0("Phase-in speed comparison — ", target_period, " baseline"),
       subtitle = "Darker green = faster phase-in; orange dots = observed breeding pairs") +
  themeo + theme(legend.position = "bottom")
p_6a
```

**plot 6b — cost of delaying then implementing immediately**
```{r plot-6b, fig.width = 9, fig.height = 5}
p_6b <- all_scenarios %>%
  filter(scenario %in% c("No mitigation (baseline)",
                          "Immediate 3/3 (100% coverage)",
                          "5-yr delay, then immediate 3/3",
                          "10-yr delay, then immediate 3/3",
                          "20-yr delay, then immediate 3/3")) %>%
  ggplot(aes(x = year, y = N, color = scenario, linetype = scenario)) +
  geom_line(linewidth = 0.9) +
  geom_hline(yintercept = N0, color = "grey50", linetype = "dashed", linewidth = 0.4) +
  geom_point(data = bird_obs, aes(x = year, y = N_scaled),
             inherit.aes = FALSE, color = "#e6550d", size = 3, shape = 16) +
  geom_line(data = bird_obs, aes(x = year, y = N_scaled),
            inherit.aes = FALSE, color = "#e6550d", linewidth = 0.5, linetype = "dotted") +
  geom_text(data = bird_obs, aes(x = year, y = N_scaled, label = season),
            inherit.aes = FALSE, color = "#e6550d", size = 2.5, vjust = -0.8) +
  scale_color_manual(values = c(
    "No mitigation (baseline)"        = "#d73027",
    "Immediate 3/3 (100% coverage)"   = "#253494",
    "5-yr delay, then immediate 3/3"  = "#2166ac",
    "10-yr delay, then immediate 3/3" = "#74add1",
    "20-yr delay, then immediate 3/3" = "#abd9e9")) +
  scale_linetype_manual(values = c(
    "No mitigation (baseline)"        = "dashed",
    "Immediate 3/3 (100% coverage)"   = "solid",
    "5-yr delay, then immediate 3/3"  = "solid",
    "10-yr delay, then immediate 3/3" = "solid",
    "20-yr delay, then immediate 3/3" = "solid")) +
  scale_x_continuous(breaks = seq(-10, years, 10), limits = c(-10, years)) +
  scale_y_continuous(labels = scales::comma,
                     breaks = scales::pretty_breaks(n = 6),
                     sec.axis = sec_axis(~ . / pairs_scale,
                                         name = "Breeding pairs (Bird Island)",
                                         labels = scales::comma)) +
  labs(x = "Years since 1990 baseline", y = "Population (N individuals)",
       color = NULL, linetype = NULL,
       title = paste0("Cost of delaying immediate 3/3 — ", target_period, " baseline"),
       subtitle = "Darker blue = shorter delay; orange dots = observed breeding pairs") +
  themeo + theme(legend.position = "bottom")
p_6b
```

**plot 6c — cost of delaying then using 5-year phase-in**
```{r plot-6c, fig.width = 9, fig.height = 5}
p_6c <- all_scenarios %>%
  filter(scenario %in% c("No mitigation (baseline)",
                          "Immediate 3/3 (100% coverage)",
                          "5-yr delay, then 5-yr phase-in",
                          "10-yr delay, then 5-yr phase-in",
                          "20-yr delay, then 5-yr phase-in")) %>%
  ggplot(aes(x = year, y = N, color = scenario, linetype = scenario)) +
  geom_line(linewidth = 0.9) +
  geom_hline(yintercept = N0, color = "grey50", linetype = "dashed", linewidth = 0.4) +
  geom_point(data = bird_obs, aes(x = year, y = N_scaled),
             inherit.aes = FALSE, color = "#e6550d", size = 3, shape = 16) +
  geom_line(data = bird_obs, aes(x = year, y = N_scaled),
            inherit.aes = FALSE, color = "#e6550d", linewidth = 0.5, linetype = "dotted") +
  geom_text(data = bird_obs, aes(x = year, y = N_scaled, label = season),
            inherit.aes = FALSE, color = "#e6550d", size = 2.5, vjust = -0.8) +
  scale_color_manual(values = c(
    "No mitigation (baseline)"        = "#d73027",
    "Immediate 3/3 (100% coverage)"   = "#253494",
    "5-yr delay, then 5-yr phase-in"  = "#762a83",
    "10-yr delay, then 5-yr phase-in" = "#af8dc3",
    "20-yr delay, then 5-yr phase-in" = "#e7d4e8")) +
  scale_linetype_manual(values = c(
    "No mitigation (baseline)"        = "dashed",
    "Immediate 3/3 (100% coverage)"   = "solid",
    "5-yr delay, then 5-yr phase-in"  = "solid",
    "10-yr delay, then 5-yr phase-in" = "solid",
    "20-yr delay, then 5-yr phase-in" = "solid")) +
  scale_x_continuous(breaks = seq(-10, years, 10), limits = c(-10, years)) +
  scale_y_continuous(labels = scales::comma,
                     breaks = scales::pretty_breaks(n = 6),
                     sec.axis = sec_axis(~ . / pairs_scale,
                                         name = "Breeding pairs (Bird Island)",
                                         labels = scales::comma)) +
  labs(x = "Years since 1990 baseline", y = "Population (N individuals)",
       color = NULL, linetype = NULL,
       title = paste0("Cost of delaying 5-yr phase-in — ", target_period, " baseline"),
       subtitle = "Darker purple = shorter delay; orange dots = observed breeding pairs") +
  themeo + theme(legend.position = "bottom")
p_6c
```

**plot 6d — all scenarios with observed breeding pairs**
```{r plot-6d, fig.width = 11, fig.height = 6}
p_6d <- ggplot(all_scenarios, aes(x = year, y = N, color = scenario, linetype = scenario)) +
  geom_line(linewidth = 0.7) +
  geom_hline(yintercept = N0, color = "grey50", linetype = "dashed", linewidth = 0.4) +
  geom_point(data = bird_obs, aes(x = year, y = N_scaled),
             inherit.aes = FALSE, color = "#e6550d", size = 3, shape = 16) +
  geom_line(data = bird_obs, aes(x = year, y = N_scaled),
            inherit.aes = FALSE, color = "#e6550d", linewidth = 0.5, linetype = "dotted") +
  geom_text(data = bird_obs, aes(x = year, y = N_scaled, label = season),
            inherit.aes = FALSE, color = "#e6550d", size = 2.5, vjust = -0.8) +
  scale_color_manual(values = c(
    "No mitigation (baseline)"        = "#d73027",
    "Immediate 3/3 (100% coverage)"   = "#253494",
    "5-year phase-in"                 = "#1a9641",
    "10-year phase-in"                = "#78c679",
    "20-year phase-in"                = "#addd8e",
    "5-yr delay, then immediate 3/3"  = "#2166ac",
    "10-yr delay, then immediate 3/3" = "#74add1",
    "20-yr delay, then immediate 3/3" = "#abd9e9",
    "5-yr delay, then 5-yr phase-in"  = "#762a83",
    "10-yr delay, then 5-yr phase-in" = "#af8dc3",
    "20-yr delay, then 5-yr phase-in" = "#e7d4e8")) +
  scale_linetype_manual(values = c(
    "No mitigation (baseline)"        = "dashed",
    "Immediate 3/3 (100% coverage)"   = "solid",
    "5-year phase-in"                 = "solid",
    "10-year phase-in"                = "solid",
    "20-year phase-in"                = "solid",
    "5-yr delay, then immediate 3/3"  = "solid",
    "10-yr delay, then immediate 3/3" = "solid",
    "20-yr delay, then immediate 3/3" = "solid",
    "5-yr delay, then 5-yr phase-in"  = "solid",
    "10-yr delay, then 5-yr phase-in" = "solid",
    "20-yr delay, then 5-yr phase-in" = "solid")) +
  scale_x_continuous(breaks = seq(-10, years, 10), limits = c(-10, years)) +
  scale_y_continuous(labels = scales::comma,
                     breaks = scales::pretty_breaks(n = 6),
                     sec.axis = sec_axis(~ . / pairs_scale,
                                         name = "Breeding pairs (Bird Island)",
                                         labels = scales::comma)) +
  labs(x = "Years since 1990 baseline", y = "Population (N individuals)",
       color = NULL, linetype = NULL,
       title = paste0("All mitigation scenarios — ", target_period, " baseline"),
       subtitle = "Green = phase-in speed; blue = delay then immediate; purple = delay then phase-in; orange dots = observed") +
  themeo +
  theme(legend.position = "right", legend.text = element_text(size = 7))
p_6d
```

**plot 6e — annual lambda, delay-then-immediate scenarios**
```{r plot-6e, fig.width = 9, fig.height = 5}
lambda_delay_imm <- all_scenarios %>%
  filter(scenario %in% c("No mitigation (baseline)",
                          "Immediate 3/3 (100% coverage)",
                          "5-yr delay, then immediate 3/3",
                          "10-yr delay, then immediate 3/3",
                          "20-yr delay, then immediate 3/3")) %>%
  group_by(scenario) %>%
  arrange(year) %>%
  mutate(lambda_t = N / lag(N)) %>%
  filter(!is.na(lambda_t))

p_6e <- ggplot(lambda_delay_imm, aes(x = year, y = lambda_t, color = scenario, linetype = scenario)) +
  geom_line(linewidth = 0.9) +
  geom_hline(yintercept = 1, color = "grey50", linetype = "dashed", linewidth = 0.4) +
  scale_color_manual(values = c(
    "No mitigation (baseline)"        = "#d73027",
    "Immediate 3/3 (100% coverage)"   = "#253494",
    "5-yr delay, then immediate 3/3"  = "#2166ac",
    "10-yr delay, then immediate 3/3" = "#74add1",
    "20-yr delay, then immediate 3/3" = "#abd9e9")) +
  scale_linetype_manual(values = c(
    "No mitigation (baseline)"        = "dashed",
    "Immediate 3/3 (100% coverage)"   = "solid",
    "5-yr delay, then immediate 3/3"  = "solid",
    "10-yr delay, then immediate 3/3" = "solid",
    "20-yr delay, then immediate 3/3" = "solid")) +
  scale_x_continuous(breaks = seq(0, years, 10), limits = c(0, years)) +
  scale_y_continuous(breaks = scales::pretty_breaks(n = 6)) +
  labs(x = "Years since 1990 baseline",
       y = expression(lambda[t] ~ "(N(t+1) / N(t))"),
       color = NULL, linetype = NULL,
       title = paste0("Annual \u03BB — cost of delaying immediate 3/3 — ", target_period, " baseline"),
       subtitle = "Step jump in \u03BB at end of delay period; darker blue = shorter delay\nDashed grey = \u03BB = 1") +
  themeo + theme(legend.position = "bottom")
p_6e
```

**plot 6f — annual lambda, delay-then-phase-in scenarios**
```{r plot-6f, fig.width = 9, fig.height = 5}
lambda_delay_phasein <- all_scenarios %>%
  filter(scenario %in% c("No mitigation (baseline)",
                          "Immediate 3/3 (100% coverage)",
                          "5-yr delay, then 5-yr phase-in",
                          "10-yr delay, then 5-yr phase-in",
                          "20-yr delay, then 5-yr phase-in")) %>%
  group_by(scenario) %>%
  arrange(year) %>%
  mutate(lambda_t = N / lag(N)) %>%
  filter(!is.na(lambda_t))

p_6f <- ggplot(lambda_delay_phasein, aes(x = year, y = lambda_t, color = scenario, linetype = scenario)) +
  geom_line(linewidth = 0.9) +
  geom_hline(yintercept = 1, color = "grey50", linetype = "dashed", linewidth = 0.4) +
  scale_color_manual(values = c(
    "No mitigation (baseline)"        = "#d73027",
    "Immediate 3/3 (100% coverage)"   = "#253494",
    "5-yr delay, then 5-yr phase-in"  = "#762a83",
    "10-yr delay, then 5-yr phase-in" = "#af8dc3",
    "20-yr delay, then 5-yr phase-in" = "#e7d4e8")) +
  scale_linetype_manual(values = c(
    "No mitigation (baseline)"        = "dashed",
    "Immediate 3/3 (100% coverage)"   = "solid",
    "5-yr delay, then 5-yr phase-in"  = "solid",
    "10-yr delay, then 5-yr phase-in" = "solid",
    "20-yr delay, then 5-yr phase-in" = "solid")) +
  scale_x_continuous(breaks = seq(0, years, 10), limits = c(0, years)) +
  scale_y_continuous(breaks = scales::pretty_breaks(n = 6)) +
  labs(x = "Years since 1990 baseline",
       y = expression(lambda[t] ~ "(N(t+1) / N(t))"),
       color = NULL, linetype = NULL,
       title = paste0("Annual \u03BB — cost of delaying 5-yr phase-in — ", target_period, " baseline"),
       subtitle = "Ramp in \u03BB over 5-yr phase-in after delay ends; darker purple = shorter delay\nDashed grey = \u03BB = 1") +
  themeo + theme(legend.position = "bottom")
p_6f
```

**plot 6g — annual lambda, all 6 delay scenarios**
```{r plot-6g, fig.width = 11, fig.height = 5}
lambda_all_delays <- all_scenarios %>%
  filter(scenario %in% c("No mitigation (baseline)",
                          "Immediate 3/3 (100% coverage)",
                          "5-yr delay, then immediate 3/3",
                          "10-yr delay, then immediate 3/3",
                          "20-yr delay, then immediate 3/3",
                          "5-yr delay, then 5-yr phase-in",
                          "10-yr delay, then 5-yr phase-in",
                          "20-yr delay, then 5-yr phase-in")) %>%
  group_by(scenario) %>%
  arrange(year) %>%
  mutate(lambda_t = N / lag(N)) %>%
  filter(!is.na(lambda_t))

p_6g <- ggplot(lambda_all_delays, aes(x = year, y = lambda_t, color = scenario, linetype = scenario)) +
  geom_line(linewidth = 0.9) +
  geom_hline(yintercept = 1, color = "grey50", linetype = "dashed", linewidth = 0.4) +
  scale_color_manual(values = c(
    "No mitigation (baseline)"        = "#d73027",
    "Immediate 3/3 (100% coverage)"   = "#253494",
    "5-yr delay, then immediate 3/3"  = "#2166ac",
    "10-yr delay, then immediate 3/3" = "#74add1",
    "20-yr delay, then immediate 3/3" = "#abd9e9",
    "5-yr delay, then 5-yr phase-in"  = "#762a83",
    "10-yr delay, then 5-yr phase-in" = "#af8dc3",
    "20-yr delay, then 5-yr phase-in" = "#e7d4e8")) +
  scale_linetype_manual(values = c(
    "No mitigation (baseline)"        = "dashed",
    "Immediate 3/3 (100% coverage)"   = "solid",
    "5-yr delay, then immediate 3/3"  = "solid",
    "10-yr delay, then immediate 3/3" = "solid",
    "20-yr delay, then immediate 3/3" = "solid",
    "5-yr delay, then 5-yr phase-in"  = "dashed",
    "10-yr delay, then 5-yr phase-in" = "dashed",
    "20-yr delay, then 5-yr phase-in" = "dashed")) +
  scale_x_continuous(breaks = seq(0, years, 10), limits = c(0, years)) +
  scale_y_continuous(breaks = scales::pretty_breaks(n = 6)) +
  labs(x = "Years since 1990 baseline",
       y = expression(lambda[t] ~ "(N(t+1) / N(t))"),
       color = NULL, linetype = NULL,
       title = paste0("Annual \u03BB — all delay scenarios — ", target_period, " baseline")) + #,
       #subtitle = "Blue = delay then immediate (solid); purple = delay then 5-yr phase-in (dashed); darker = shorter delay\nDashed grey = \u03BB = 1") +
  themeo + theme(legend.position = "right", legend.text = element_text(size = 7))
p_6g
```

---
```{r}
# dominant eigenvector of m_3of3 = SSD
  ev <- eigen(m_3of3)
  ssd <- Re(ev$vectors[, 1])
  ssd <- ssd / sum(ssd)  # normalize

  # compare proportions to n0_vec / sum(n0_vec)
  round(cbind(ssd, n0_vec / sum(n0_vec)), 4)

```

## Save figures

```{r save-figures}
dir.create(here("output/figures"), showWarnings = FALSE, recursive = TRUE)

fig_path <- function(name) here("output/figures",
                                 paste0(name, "_", target_period, "_", output_sfx, ".png"))

ggsave(fig_path("09_1a_100yr_breeding_pairs"),     p_1a,  width = 9,  height = 5, dpi = 200)
ggsave(fig_path("09_1b_100yr_age_classes"),        p_1b,  width = 9,  height = 5, dpi = 200)
ggsave(fig_path("09_1c_500yr_breeding_pairs"),     p_1c,  width = 9,  height = 5, dpi = 200)
ggsave(fig_path("09_1d_500yr_age_classes"),        p_1d,  width = 9,  height = 5, dpi = 200)
ggsave(fig_path("09_1e_500yr_logN"),               p_1e,  width = 9,  height = 5, dpi = 200)
ggsave(fig_path("09_1f_500yr_annual_lambda"),      p_1f,  width = 9,  height = 5, dpi = 200)
ggsave(fig_path("09_2a_100yr_matrix_N"),            p_2a,  width = 9,  height = 5, dpi = 200)
ggsave(fig_path("09_2c_500yr_matrix_N"),            p_2c,  width = 9,  height = 5, dpi = 200)
ggsave(fig_path("09_2e_500yr_matrix_logN"),        p_2e,  width = 9,  height = 5, dpi = 200)
ggsave(fig_path("09_2f_500yr_matrix_lambda"),      p_2f,  width = 9,  height = 5, dpi = 200)
ggsave(fig_path("09_3a_100yr_phasein_N"),           p_3a,  width = 9,  height = 5, dpi = 200)
ggsave(fig_path("09_3c_500yr_phasein_N"),           p_3c,  width = 9,  height = 5, dpi = 200)
ggsave(fig_path("09_3e_500yr_phasein_logN"),       p_3e,  width = 9,  height = 5, dpi = 200)
ggsave(fig_path("09_3f_500yr_phasein_lambda"),     p_3f,  width = 9,  height = 5, dpi = 200)
ggsave(fig_path("09_6a_phasein_compare"),          p_6a,  width = 9,  height = 5, dpi = 200)
ggsave(fig_path("09_6b_delay_immediate"),          p_6b,  width = 9,  height = 5, dpi = 200)
ggsave(fig_path("09_6c_delay_phasein"),            p_6c,  width = 9,  height = 5, dpi = 200)
ggsave(fig_path("09_6d_all_scenarios"),            p_6d,  width = 11, height = 6, dpi = 200)
ggsave(fig_path("09_6e_lambda_delay_immediate"),   p_6e,  width = 9,  height = 5, dpi = 200)
ggsave(fig_path("09_6f_lambda_delay_phasein"),     p_6f,  width = 9,  height = 5, dpi = 200)
ggsave(fig_path("09_6g_lambda_all_delays"),        p_6g,  width = 11, height = 5, dpi = 200)
message("9 figures saved to output/figures/ with suffix ", output_sfx)
```
