---
title: "Wandering albatross bycatch mitigation impacts on demography"
author: "Sarah Becker"
date: "4/13/2022"
output: html_document
---


```{r, echo= FALSE, include =FALSE}
library(here)
library(tidyverse)
```

**set vital rates. Try first without widowing**
```{r}
#survival rates - drawn from Weimerskirch 1998
s_c <- 0.64 #chick+ fledgling survival year 0-1
s_j <- 0.8 #juvenile survival year 1-3
s_i <- 0.92 #immature survival year 3-7
s_s <- 0.94 #subadult survival year 7-10
s_a <- 0.94 #adult survival year 10+
#males show senescence but females don't, so not including it here, though perhaps I should since it impacts breeding success?

#fecundity building blocks
#breeding success - drawn from Weimerskirch 1998 &Rackete 2021
bs_s <- 0.7 #subadult breeding success
bs_a <- 0.8 #adult breeding success

#breeding probability (biennial breeders) - adult rate drawn from Rackete 2021, subadult downscaled slightly from that
bp_s <- 0.4 #subadult breeding probability (range of ages at which individuals recruit)
bp_a <- 0.45 #adult breeding probablity (biennial breeding, most breeding by age 10)

c <- 0.5# # of female chicks per year

#fecundity 
f_s <- bs_s*bp_s*c
f_a <- bs_a*bp_a*c

```

**create matrix. first without widowing, try to add that later**
```{r}
#set up age-based matrix based on vital rates
#Age classes (years)  0-1     1-2    2-3    3-4    4-5    5-6    6-7    7-8        8-9        9-10       10+        
m <- matrix(data = c( 0,      0,    0,      0,     0,     0,     0,     s_s*f_s,   s_s*f_s,   s_s*f_s,   s_a*f_a, #0-1
                      s_c,    0,    0,      0,     0,     0,     0,     0,         0,         0,         0,       #1-2
                      0,      s_j,  0,      0,     0,     0,     0,     0,         0,         0,         0,       #2-3
                      0,      0,    s_j,    0,     0,     0,     0,     0,         0,         0,         0,       #3-4
                      0,      0,    0,      s_i,   0,     0,     0,     0,         0,         0,         0,       #4-5
                      0,      0,    0,      0,     s_i,   0,     0,     0,         0,         0,         0,       #5-6
                      0,      0,    0,      0,     0,     s_i,   0,     0,         0,         0,         0,       #6-7
                      0,      0,    0,      0,     0,     0,     s_i,   0,         0,         0,         0,       #7-8
                      0,      0,    0,      0,     0,     0,     0,     s_s,       0,         0,         0,       #8-9
                      0,      0,    0,      0,     0,     0,     0,     0,         s_s,       0,         0,       #9-10
                      0,      0,    0,      0,     0,     0,     0,     0,         0,         s_s,       s_a),    #10+
                   nrow = 11, byrow = TRUE) 
m
```

**extract lambda, SSDs, and RVs from eigenvalues**
```{r}
#deterministic growth rate (lambda)
#extract eigenvalues and lambda from the matrix 
m_eig <- eigen(m)
lambda <- as.numeric(m_eig$values[1])
lambda # this matches the ideal population growth rate in the table

#Stable state distribution
#extract and scale the right eigenvector 
SSD <- round(as.numeric(m_eig$vectors[,1]/sum(m_eig$vectors[,1])),3)
SSD

#extract reproductive values
RV <- as.numeric(eigen(t(m))$vectors[,1]/eigen(t(m))$vectors[1,1])
RV
```

**identify mortality rate, mitigation efficacy**
```{r}
#bycatch mortality rates
# ~25 - 35% of mortality has been attributed to bycatch (Cleeland et al 2021), 
# so I calculated the baseline mortality and divided by 4 (testing 33% now - /3)
m_j <- (1-s_j)/3
m_i <- (1-s_i)/3
m_s <- (1-s_s)/3
m_a <- (1-s_a)/3

mitigation_eff <- 0.8
```

**test two examples of mortality adjustment**
```{r}
#zero mitigation
  n_s_j <- 1-((m_j*(1-(mitigation_eff*0)))+(m_j*2))
  n_s_i <- 1-((m_i*(1-(mitigation_eff*0)))+(m_i*2))
  n_s_s <- 1-((m_s*(1-(mitigation_eff*0)))+(m_s*2))
  n_s_a <- 1-((m_a*(1-(mitigation_eff*0)))+(m_s*2))
  
  #Age classes (years)    0-1     1-2     2-3     3-4     4-5    5-6    6-7    7-8          8-9          9-10         10+        
  n_m <- matrix(data = c( 0,      0,      0,      0,      0,     0,     0,     n_s_s*f_s,   n_s_s*f_s,   n_s_s*f_s,   n_s_a*f_a, #0-1
                          s_c,    0,      0,      0,      0,     0,     0,     0,           0,           0,           0,         #1-2
                          0,      n_s_j,  0,      0,      0,     0,     0,     0,           0,           0,           0,         #2-3
                          0,      0,      n_s_j,  0,      0,     0,     0,     0,           0,           0,           0,         #3-4
                          0,      0,      0,      n_s_i,  0,     0,     0,     0,           0,           0,           0,         #4-5
                          0,      0,      0,      0,      n_s_i, 0,     0,     0,           0,           0,           0,         #5-6
                          0,      0,      0,      0,      0,     n_s_i, 0,     0,           0,           0,           0,         #6-7
                          0,      0,      0,      0,      0,     0,     n_s_i, 0,           0,           0,           0,         #7-8
                          0,      0,      0,      0,      0,     0,     0,     n_s_s,       0,           0,           0,         #8-9
                          0,      0,      0,      0,      0,     0,     0,     0,           n_s_s,       0,           0,         #9-10
                          0,      0,      0,      0,      0,     0,     0,     0,           0,           n_s_s,       n_s_a),    #10+
                   nrow = 11, byrow = TRUE) 
  
  
  #extract eigenvalues and lambda from the matrix 
  n_m_eig <- eigen(n_m)
  lambda_n <- as.numeric(n_m_eig$values[1]) # 0.9792239
  lambda #0.9792239

#50% mitigation
  n_s_j <- 1-((m_j*(1-(mitigation_eff*0.5)))+(m_j*2))
  n_s_i <- 1-((m_i*(1-(mitigation_eff*0.5)))+(m_i*2))
  n_s_s <- 1-((m_s*(1-(mitigation_eff*0.5)))+(m_s*2))
  n_s_a <- 1-((m_a*(1-(mitigation_eff*0.5)))+(m_s*2))
  
  #Age classes (years)    0-1     1-2     2-3     3-4     4-5    5-6    6-7    7-8          8-9          9-10         10+        
  n_m <- matrix(data = c( 0,      0,      0,      0,      0,     0,     0,     n_s_s*f_s,   n_s_s*f_s,   n_s_s*f_s,   n_s_a*f_a, #0-1
                          s_c,    0,      0,      0,      0,     0,     0,     0,           0,           0,           0,         #1-2
                          0,      n_s_j,  0,      0,      0,     0,     0,     0,           0,           0,           0,         #2-3
                          0,      0,      n_s_j,  0,      0,     0,     0,     0,           0,           0,           0,         #3-4
                          0,      0,      0,      n_s_i,  0,     0,     0,     0,           0,           0,           0,         #4-5
                          0,      0,      0,      0,      n_s_i, 0,     0,     0,           0,           0,           0,         #5-6
                          0,      0,      0,      0,      0,     n_s_i, 0,     0,           0,           0,           0,         #6-7
                          0,      0,      0,      0,      0,     0,     n_s_i, 0,           0,           0,           0,         #7-8
                          0,      0,      0,      0,      0,     0,     0,     n_s_s,       0,           0,           0,         #8-9
                          0,      0,      0,      0,      0,     0,     0,     0,           n_s_s,       0,           0,         #9-10
                          0,      0,      0,      0,      0,     0,     0,     0,           0,           n_s_s,       n_s_a),    #10+
                   nrow = 11, byrow = TRUE) 
  
  
  #extract eigenvalues and lambda from the matrix 
  n_m_eig <- eigen(n_m)
  lambda_n <- as.numeric(n_m_eig$values[1]) # 0.9891705
  lambda #  0.9792239
  
  
 #100% mitigation
  n_s_j <- 1-((m_j*(1-(mitigation_eff*1)))+(m_j*2))
  n_s_i <- 1-((m_i*(1-(mitigation_eff*1)))+(m_i*2))
  n_s_s <- 1-((m_s*(1-(mitigation_eff*1)))+(m_s*2))
  n_s_a <- 1-((m_a*(1-(mitigation_eff*1)))+(m_s*2))
  
  #Age classes (years)    0-1     1-2     2-3     3-4     4-5    5-6    6-7    7-8          8-9          9-10         10+        
  n_m <- matrix(data = c( 0,      0,      0,      0,      0,     0,     0,     n_s_s*f_s,   n_s_s*f_s,   n_s_s*f_s,   n_s_a*f_a, #0-1
                          s_c,    0,      0,      0,      0,     0,     0,     0,           0,           0,           0,         #1-2
                          0,      n_s_j,  0,      0,      0,     0,     0,     0,           0,           0,           0,         #2-3
                          0,      0,      n_s_j,  0,      0,     0,     0,     0,           0,           0,           0,         #3-4
                          0,      0,      0,      n_s_i,  0,     0,     0,     0,           0,           0,           0,         #4-5
                          0,      0,      0,      0,      n_s_i, 0,     0,     0,           0,           0,           0,         #5-6
                          0,      0,      0,      0,      0,     n_s_i, 0,     0,           0,           0,           0,         #6-7
                          0,      0,      0,      0,      0,     0,     n_s_i, 0,           0,           0,           0,         #7-8
                          0,      0,      0,      0,      0,     0,     0,     n_s_s,       0,           0,           0,         #8-9
                          0,      0,      0,      0,      0,     0,     0,     0,           n_s_s,       0,           0,         #9-10
                          0,      0,      0,      0,      0,     0,     0,     0,           0,           n_s_s,       n_s_a),    #10+
                   nrow = 11, byrow = TRUE) 
  
  
  #extract eigenvalues and lambda from the matrix 
  n_m_eig <- eigen(n_m)
  lambda_n <- as.numeric(n_m_eig$values[1]) # 0.9991264
  lambda #  0.9792239 
```


**loop through different bycatch reductions scenarios, find lambda for each**
```{r}
bycatch_reductions <- data.frame(reduction = seq(0, 1, by =0.01),
                        lambda = NA)

for(i in 1:nrow(bycatch_reductions)){
  
  n_s_j <- 1-((m_j*(1-(mitigation_eff*bycatch_reductions$reduction[i])))+(m_j*2))
  n_s_i <- 1-((m_i*(1-(mitigation_eff*bycatch_reductions$reduction[i])))+(m_i*2))
  n_s_s <- 1-((m_s*(1-(mitigation_eff*bycatch_reductions$reduction[i])))+(m_s*2))
  n_s_a <- 1-((m_a*(1-(mitigation_eff*bycatch_reductions$reduction[i])))+(m_s*2))
  
  #Age classes (years)    0-1     1-2     2-3     3-4     4-5    5-6    6-7    7-8          8-9          9-10         10+        
  n_m <- matrix(data = c( 0,      0,      0,      0,      0,     0,     0,     n_s_s*f_s,   n_s_s*f_s,   n_s_s*f_s,   n_s_a*f_a, #0-1
                          s_c,    0,      0,      0,      0,     0,     0,     0,           0,           0,           0,         #1-2
                          0,      n_s_j,  0,      0,      0,     0,     0,     0,           0,           0,           0,         #2-3
                          0,      0,      n_s_j,  0,      0,     0,     0,     0,           0,           0,           0,         #3-4
                          0,      0,      0,      n_s_i,  0,     0,     0,     0,           0,           0,           0,         #4-5
                          0,      0,      0,      0,      n_s_i, 0,     0,     0,           0,           0,           0,         #5-6
                          0,      0,      0,      0,      0,     n_s_i, 0,     0,           0,           0,           0,         #6-7
                          0,      0,      0,      0,      0,     0,     n_s_i, 0,           0,           0,           0,         #7-8
                          0,      0,      0,      0,      0,     0,     0,     n_s_s,       0,           0,           0,         #8-9
                          0,      0,      0,      0,      0,     0,     0,     0,           n_s_s,       0,           0,         #9-10
                          0,      0,      0,      0,      0,     0,     0,     0,           0,           n_s_s,       n_s_a),    #10+
                   nrow = 11, byrow = TRUE) 
  
  
  #extract eigenvalues and lambda from the matrix 
  n_m_eig <- eigen(n_m)
  lambda_n <- as.numeric(n_m_eig$values[1])
  bycatch_reductions$lambda[i] <- lambda_n # this matches the ideal population growth rate in the table

}
```

**plot bycatch reduction against changes to lambda**
```{r}
ggplot(bycatch_reductions)+
  geom_line(aes(x=reduction, y=as.numeric(lambda)))+
  geom_hline(aes(yintercept = 1), color = "blue", alpha =0.5)+ 
  geom_vline(aes(xintercept = 0.66), color = "red", alpha =0.5)+
   geom_vline(aes(xintercept = 0.33), color = "red", alpha =0.5)+
  labs(x="% bycatch reduction", y="lambda")+
  theme_classic()
```