title: "albatross demo with widowing"
author: "Sarah Becker"
date: "4/19/2022"
output: html_document
---

```{r, echo= FALSE, include =FALSE}
library(here)
library(tidyverse)
```

**set vital rates. Try first without widowing**
```{r}
#survival rates - drawn from Weimerskirch 1998
s_c <- 0.64 #chick+ fledgling survival year 0-1
s_j <- 0.85 #juvenile survival year 1-3
s_i <- 0.92 #immature survival year 3-7
s_s <- 0.94 #subadult survival year 7-10
s_a <- 0.94 #adult survival year 10+
#males show senescence but femailes don't, so not including it here, though perhaps I should since it impacts breeding success?

#fecundity building blocks
#breeding success - drawn from Weimerskirch 1998 &Rackete 2021
bs_a <- 0.76 #adult breeding success

#breeding probability (biennial breeders) - adult rate drawn from Rackete 2021, subadult downscaled slightly from that
bp_a <- 0.45 #adult breeding probablity (biennial breeding, most breeding by age 10)

c <- 0.5# # of female chicks per year

#fecundity 
f_a <- bs_a*bp_a*c
  
#add widowing
#assume 25% prob year1 -> 100% prob year 4?
#Sun et al 2022 - takes 4 years to reach next breeding attempt, breeding prob with new mate after widowhood = ~0.25 (fig5)
w <- 0.25
w4 <- 1
```

**create matrix**
```{r}
#set up age-based matrix based on vital rates

#Age classes (years)  0-1     1-2    2-3    3-4    4-5    5-6    6-7    7-8   8-9     9-10     10+         nw,        w1,             w2,              w3,             w4 
m <- matrix(data = c(0,     0,    0,      0,     0,     0,     0,     0,      0,      0,       s_a*f_a,    s_a*f_a,    s_a*f_a*w,    s_a*f_a*w,     s_a*f_a*w,     s_a*f_a*w4,      #0-1
                    s_c,    0,    0,      0,     0,     0,     0,     0,      0,      0,       0,          0,          0,             0,              0,              0,              #1-2
                    0,      s_j,  0,      0,     0,     0,     0,     0,      0,      0,       0,          0,          0,             0,              0,              0,              #2-3
                    0,      0,    s_j,    0,     0,     0,     0,     0,      0,      0,       0,          0,          0,             0,              0,              0,              #3-4
                    0,      0,    0,      s_i,   0,     0,     0,     0,      0,      0,       0,          0,          0,             0,              0,              0,              #4-5
                    0,      0,    0,      0,     s_i,   0,     0,     0,      0,      0,       0,          0,          0,             0,              0,              0,              #5-6
                    0,      0,    0,      0,     0,     s_i,   0,     0,      0,      0,       0,          0,          0,             0,              0,              0,              #6-7
                    0,      0,    0,      0,     0,     0,     s_i,   0,      0,      0,       0,          0,          0,             0,              0,              0,              #7-8
                    0,      0,    0,      0,     0,     0,     0,     s_s,    0,      0,       0,          0,          0,             0,              0,              0,              #8-9
                    0,      0,    0,      0,     0,     0,     0,     0,      s_s,    0,       0,          0,          0,             0,              0,              0,              #9-10
                    0,      0,    0,      0,     0,     0,     0,     0,      0,      s_s,     0,          0,          0,             0,              0,              0,              #10
                    0,      0,    0,      0,     0,     0,     0,     0,      0,      0,       s_a*s_a,    s_a*s_a,    s_a*s_a*w,     s_a*s_a*w,      s_a*s_a*w,      s_a*s_a*w4,     #nw
                    0,      0,    0,      0,     0,     0,     0,     0,      0,      0,       s_a*(1-s_a),s_a*(1-s_a),s_a*(1-s_a)*w, s_a*(1-s_a)*w,  s_a*(1-s_a)*w,  s_a*(1-s_a)*w4, #w1
                    0,      0,    0,      0,     0,     0,     0,     0,      0,      0,       0,          0,          s_a*(1-w),     0,              0,              0,              #w2 
                    0,      0,    0,      0,     0,     0,     0,     0,      0,      0,       0,          0,          0,             s_a*(1-w),      0,              0,              #w3 
                    0,      0,    0,      0,     0,     0,     0,     0,      0,      0,       0,          0,          0,             0,              s_a*(1-w),      0),             #w4 
                   nrow = 16, byrow = TRUE) 
m
```

**extract lambda, SSDs, and RVs from eigenvalues**
```{r}
#deterministic growth rate (lambda)
#extract eigenvalues and lambda from the matrix 
m_eig <- eigen(m)
lambda <- as.numeric(m_eig$values[1])
lambda # this matches the ideal population growth rate in the table

#Stable state distribution
#extract and scale the right eigenvector 
SSD <- round(as.numeric(m_eig$vectors[,1]/sum(m_eig$vectors[,1])),3)
SSD

#extract reproductive values
RV <- as.numeric(eigen(t(m))$vectors[,1]/eigen(t(m))$vectors[1,1])
RV
```

**identify mortality rate, mitigation efficacy**
```{r}
#bycatch mortality rates
# ~25 - 35% of mortality has been attributed to bycatch (Cleeland et al 2021), 
# so I calculated the baseline mortality and divided by 4 (testing 33% now - /3)
m_j <- (1-s_j)/3 
m_i <- (1-s_i)/3
m_s <- (1-s_s)/3
m_a <- (1-s_a)/3

mitigation_eff <- 0.8
```

**loop through different bycatch reductions scenarios, find lambda for each**
```{r}
bycatch_reductions <- data.frame(reduction = seq(0, 1, by =0.01),
                        lambda = NA)

for(i in 1:nrow(bycatch_reductions)){
  
  n_s_j <- 1-((m_j*(1-(mitigation_eff*bycatch_reductions$reduction[i])))+(m_j*2))
  n_s_i <- 1-((m_i*(1-(mitigation_eff*bycatch_reductions$reduction[i])))+(m_i*2))
  n_s_s <- 1-((m_s*(1-(mitigation_eff*bycatch_reductions$reduction[i])))+(m_s*2))
  n_s_a <- 1-((m_a*(1-(mitigation_eff*bycatch_reductions$reduction[i])))+(m_s*2))
  
  #Age classes (years)  0-1   1-2    2-3     3-4    4-5    5-6    6-7    7-8     8-9     9-10     10+              nw,              w1,                w2,                 w3,                 w4 
  n_m <- matrix(data = c(0,   0,     0,      0,     0,     0,     0,     0,      0,      0,       n_s_a*f_a,       n_s_a*f_a,       n_s_a*f_a*w,      n_s_a*f_a*w,       n_s_a*f_a*w,       n_s_a*f_a*1,        #0-1
                       s_c,   0,     0,      0,     0,     0,     0,     0,      0,      0,       0,               0,               0,                 0,                  0,                  0,                  #1-2
                       0,     n_s_j, 0,      0,     0,     0,     0,     0,      0,      0,       0,               0,               0,                 0,                  0,                  0,                  #2-3
                       0,     0,     n_s_j,  0,     0,     0,     0,     0,      0,      0,       0,               0,               0,                 0,                  0,                  0,                  #3-4
                       0,     0,     0,      n_s_i, 0,     0,     0,     0,      0,      0,       0,               0,               0,                 0,                  0,                  0,                  #4-5
                       0,     0,     0,      0,     n_s_i, 0,     0,     0,      0,      0,       0,               0,               0,                 0,                  0,                  0,                  #5-6
                       0,     0,     0,      0,     0,     n_s_i, 0,     0,      0,      0,       0,               0,               0,                 0,                  0,                  0,                  #6-7
                       0,     0,     0,      0,     0,     0,     n_s_i, 0,      0,      0,       0,               0,               0,                 0,                  0,                  0,                  #7-8
                       0,     0,     0,      0,     0,     0,     0,     n_s_s,  0,      0,       0,               0,               0,                 0,                  0,                  0,                  #8-9
                       0,     0,     0,      0,     0,     0,     0,     0,      n_s_s,  0,       0,               0,               0,                 0,                  0,                  0,                  #9-10
                       0,     0,     0,      0,     0,     0,     0,     0,      0,      n_s_s,   0,               0,               0,                 0,                  0,                  0,                  #10
                       0,     0,     0,      0,     0,     0,     0,     0,      0,      0,       n_s_a*n_s_a,     n_s_a*n_s_a,     n_s_a*n_s_a*w,     n_s_a*n_s_a*w,      n_s_a*n_s_a*w,      n_s_a*n_s_a*w4,     #nw
                       0,     0,     0,      0,     0,     0,     0,     0,      0,      0,       n_s_a*(1-n_s_a), n_s_a*(1-n_s_a), n_s_a*(1-n_s_a)*w, n_s_a*(1-n_s_a)*w,  n_s_a*(1-n_s_a)*w,  n_s_a*(1-n_s_a)*w4, #w1
                       0,     0,     0,      0,     0,     0,     0,     0,      0,      0,       0,               0,               n_s_a*(1-w),       0,                  0,                  0,                  #w2 
                       0,     0,     0,      0,     0,     0,     0,     0,      0,      0,       0,               0,               0,                 n_s_a*(1-w),        0,                  0,                  #w3 
                       0,     0,     0,      0,     0,     0,     0,     0,      0,      0,       0,               0,               0,                 0,                  n_s_a*(1-w),        0),                 #w4 
                      nrow = 16, byrow = TRUE) 
  
  #extract eigenvalues and lambda from the matrix 
  n_m_eig <- eigen(n_m)
  lambda_n <- as.numeric(n_m_eig$values[1])
  bycatch_reductions$lambda[i] <- lambda_n # this matches the ideal population growth rate in the table

}
```

**plot bycatch reduction against changes to lambda**
```{r}
ggplot(bycatch_reductions)+
  geom_line(aes(x=reduction, y=as.numeric(lambda)))+
   geom_hline(aes(yintercept = 1), color = "blue", alpha =0.5)+ 
  # geom_vline(aes(xintercept = 0.66), color = "red", alpha =0.5)+
  #  geom_vline(aes(xintercept = 0.33), color = "red", alpha =0.5)+
 labs(x="% fisheries with bycatch mitigation", y="lambda")+
  theme_classic()
```

**add uncertainty - vary mitigation efficacy**
```{r}
bycatch_reductions2 <- seq(0, 1, by =0.01)
# mitigation_eff_range <- seq(0.45,0.95, by  = 0.01)
mitigation_eff_range <- seq(0.5,1, by  = 0.1)
output <- matrix(nrow = length(bycatch_reductions2), ncol = length(mitigation_eff_range)) %>% as.data.frame() 

for(k in 1:length(mitigation_eff_range)){
  mit_eff <- mitigation_eff_range[k]
  
  for(i in 1:length(bycatch_reductions2)){
  
  n_s_j <- 1-((m_j*(1-(mit_eff*bycatch_reductions2[i])))+(m_j*2))
  n_s_i <- 1-((m_i*(1-(mit_eff*bycatch_reductions2[i])))+(m_i*2))
  n_s_s <- 1-((m_s*(1-(mit_eff*bycatch_reductions2[i])))+(m_s*2))
  n_s_a <- 1-((m_a*(1-(mit_eff*bycatch_reductions2[i])))+(m_s*2))
  
  #Age classes (years)  0-1   1-2    2-3     3-4    4-5    5-6    6-7    7-8     8-9     9-10     10+              nw,              w1,                w2,                 w3,                 w4 
  n_m <- matrix(data = c(0,   0,     0,      0,     0,     0,     0,     0,      0,      0,       n_s_a*f_a,       n_s_a*f_a,       n_s_a*f_a*w,      n_s_a*f_a*w,       n_s_a*f_a*w,       n_s_a*f_a*w4,        #0-1
                       s_c,   0,     0,      0,     0,     0,     0,     0,      0,      0,       0,               0,               0,                 0,                  0,                  0,                  #1-2
                       0,     n_s_j, 0,      0,     0,     0,     0,     0,      0,      0,       0,               0,               0,                 0,                  0,                  0,                  #2-3
                       0,     0,     n_s_j,  0,     0,     0,     0,     0,      0,      0,       0,               0,               0,                 0,                  0,                  0,                  #3-4
                       0,     0,     0,      n_s_i, 0,     0,     0,     0,      0,      0,       0,               0,               0,                 0,                  0,                  0,                  #4-5
                       0,     0,     0,      0,     n_s_i, 0,     0,     0,      0,      0,       0,               0,               0,                 0,                  0,                  0,                  #5-6
                       0,     0,     0,      0,     0,     n_s_i, 0,     0,      0,      0,       0,               0,               0,                 0,                  0,                  0,                  #6-7
                       0,     0,     0,      0,     0,     0,     n_s_i, 0,      0,      0,       0,               0,               0,                 0,                  0,                  0,                  #7-8
                       0,     0,     0,      0,     0,     0,     0,     n_s_s,  0,      0,       0,               0,               0,                 0,                  0,                  0,                  #8-9
                       0,     0,     0,      0,     0,     0,     0,     0,      n_s_s,  0,       0,               0,               0,                 0,                  0,                  0,                  #9-10
                       0,     0,     0,      0,     0,     0,     0,     0,      0,      n_s_s,   0,               0,               0,                 0,                  0,                  0,                  #10
                       0,     0,     0,      0,     0,     0,     0,     0,      0,      0,       n_s_a*n_s_a,     n_s_a*n_s_a,     n_s_a*n_s_a*w,     n_s_a*n_s_a*w,      n_s_a*n_s_a*w,      n_s_a*n_s_a*w4,     #nw
                       0,     0,     0,      0,     0,     0,     0,     0,      0,      0,       n_s_a*(1-n_s_a), n_s_a*(1-n_s_a), n_s_a*(1-n_s_a)*w, n_s_a*(1-n_s_a)*w,  n_s_a*(1-n_s_a)*w,  n_s_a*(1-n_s_a)*w4, #w1
                       0,     0,     0,      0,     0,     0,     0,     0,      0,      0,       0,               0,               n_s_a*(1-w),       0,                  0,                  0,                  #w2 
                       0,     0,     0,      0,     0,     0,     0,     0,      0,      0,       0,               0,               0,                 n_s_a*(1-w),        0,                  0,                  #w3 
                       0,     0,     0,      0,     0,     0,     0,     0,      0,      0,       0,               0,               0,                 0,                  n_s_a*(1-w),        0),                 #w4 
                      nrow = 16, byrow = TRUE) 
  
  #extract eigenvalues and lambda from the matrix 
  n_m_eig <- eigen(n_m)
  lambda_n <- as.numeric(n_m_eig$values[1])
  output[i,k] <- lambda_n # this matches the ideal population growth rate in the table

}
}

```

```{r}

# oldnames = c("V1","V2","V3","V4" ,"V5" , "V6" , "V7"  ,"V8" , "V9" , "V10", "V11" ,"V12", "V13" ,"V14" ,"V15" ,"V16" ,"V17" ,"V18" ,"V19" ,"V20" ,"V21" ,"V22" ,"V23", "V24" ,"V25", "V26" ,"V27" ,"V28" ,"V29", "V30" ,"V31" ,"V32", "V33", "V34", "V35" ,"V36", "V37", "V38", "V39" ,"V40" ,"V41" ,"V42" ,"V43", "V44" ,"V45" ,"V46" ,"V47" ,"V48", "V49", "V50" ,"V51")
# newnames = c("0.45", "0.46", "0.47", "0.48", "0.49", "0.50", "0.51" ,"0.52" ,"0.53" ,"0.54" ,"0.55", "0.56" ,"0.57" ,"0.58" ,"0.59" ,"0.60" ,"0.61", "0.62", "0.63" ,"0.64" ,"0.65" ,"0.66" ,"0.67" ,"0.68" ,"0.69" ,"0.70", "0.71" ,"0.72", "0.73", "0.74", "0.75" ,"0.76" ,"0.77" ,"0.78" ,"0.79", "0.80" ,"0.81" ,"0.82" ,"0.83" ,"0.84", "0.85", "0.86", "0.87", "0.88" ,"0.89", "0.90", "0.91" ,"0.92", "0.93", "0.94" ,"0.95")

oldnames = c("V1","V2","V3","V4" ,"V5", "V6")
newnames = c("0.5", "0.6", "0.7", "0.8", "0.9", "1")

output<- output %>% 
  rename_with(~ newnames[which(oldnames == .x)], .cols = oldnames)

output$bycatch_reductions <- bycatch_reductions2

output_l <- output %>% 
  pivot_longer(cols = !bycatch_reductions,names_to = "mit_eff", values_to = "lambda")

output_l$mit_eff <- as.numeric(output_l$mit_eff)
```

```{r}
ggplot(output_l)+
  geom_line(aes(x=bycatch_reductions, y=lambda, group = factor(mit_eff), color = factor(mit_eff)))+
   geom_hline(aes(yintercept = 1), color = "darkgrey", alpha =0.5)+ 
  # geom_vline(aes(xintercept = 0.66), color = "red", alpha =0.5)+
  #  geom_vline(aes(xintercept = 0.33), color = "red", alpha =0.5)+
  labs(x="% fisheries with bycatch mitigation", y="lambda", color = "% mitigation efficacy")+
  theme_classic()
```

**add uncertainty, simulate 1000 runs with varying range of vital rates (randomly draw with replacement from values found in the literature)** 
**mortality = 33% bycatch, 67% other causes**
```{r}
#full_output = matrix(nrow = length(bycatch_reductions2)*length(mitigation_eff_range), ncol = length(sims)) %>% as.data.frame() 
full_output <- vector("list", 6)
bycatch_reductions2 <- seq(0, 1, by =0.01)
mitigation_eff_range <- seq(0.5,1, by  = 0.1)
sims = 1:100

for(k in 1:length(mitigation_eff_range)){
  mit_eff <- mitigation_eff_range[k]
  
  output2 <- matrix(nrow = length(bycatch_reductions2), ncol = length(sims)) %>% as.data.frame() 
  
for(j in 1:length(sims)){
    
    s_c2 <- sample(seq(0.63,0.65, by =0.01), size =1, replace = TRUE) #chick+ fledgling survival year 0-1 - 0.64 
    s_j2 <- sample(seq(0.8,0.85, by = 0.01), size =1, replace = TRUE) #juvenile survival year 1-3 -  0.8 
    s_i2 <- sample(seq(0.91,0.94, by =0.01), size =1, replace = TRUE) #immature survival year 3-7 -  0.92
    s_s2 <- sample(seq(0.94,0.95, by =0.01), size =1, replace =TRUE) #subadult survival year 7-10 - 0.94
    s_a2 <- sample(seq(0.94,0.95, by =0.01), size=1, replace=TRUE) #adult survival year 10+ - 0.94

    bs_a2 <- sample(seq(0.68,0.8, by =0.01), size =1, replace = TRUE) #adult breeding success - 0.8 
    bp_a2 <- sample(seq(0.4,0.7, by =0.01), size=1, replace = TRUE)  #adult breeding probablity (biennial breeding, most breeding by age 10) - 0.45
    c2 <- 0.5# # of female chicks per year
    f_a2 <- bs_a*bp_a*c #fecundity
  
    w2 <- sample(seq(0.4,0.5, by =0.01), size=1, replace = TRUE)
    w42 <- 1
    
    m_j2 <- (1-s_j2)/3
    m_i2 <- (1-s_i2)/3
    m_s2 <- (1-s_s2)/3
    m_a2 <- (1-s_a2)/3
    
for(i in 1:length(bycatch_reductions2)){
  
  #calculate new survival rates = 1-(bycatch mortality*(1-efficacy of intervention action * proportion of fishery implementing intervention actions))+mortality remainder)
  n_s_j2 <- 1-((m_j2*(1-(mit_eff*bycatch_reductions2[i])))+(m_j2*2)) 
  n_s_i2 <- 1-((m_i2*(1-(mit_eff*bycatch_reductions2[i])))+(m_i2*2))
  n_s_s2 <- 1-((m_s2*(1-(mit_eff*bycatch_reductions2[i])))+(m_s2*2))
  n_s_a2 <- 1-((m_a2*(1-(mit_eff*bycatch_reductions2[i])))+(m_a2*2))
  
  #Age classes (years)  0-1   1-2    2-3     3-4    4-5    5-6    6-7    7-8     8-9     9-10     10+              nw,              w1,                w2,                 w3,                 w4 
  n_m2 <- matrix(data = c(0,   0,     0,      0,     0,     0,     0,     0,      0,      0,       n_s_a2*f_a2,       n_s_a2*f_a2,       n_s_a2*f_a2*w2,      n_s_a2*f_a2*w2,       n_s_a2*f_a2*w2,       n_s_a2*f_a2*w42,        #0-1
                       s_c2,  0,     0,      0,     0,     0,     0,     0,      0,      0,       0,               0,               0,                 0,                  0,                  0,                  #1-2
                       0,     n_s_j2, 0,      0,     0,     0,     0,     0,      0,      0,       0,               0,               0,                 0,                  0,                  0,                  #2-3
                       0,     0,     n_s_j2,  0,     0,     0,     0,     0,      0,      0,       0,               0,               0,                 0,                  0,                  0,                  #3-4
                       0,     0,     0,      n_s_i2, 0,     0,     0,     0,      0,      0,       0,               0,               0,                 0,                  0,                  0,                  #4-5
                       0,     0,     0,      0,     n_s_i2, 0,     0,     0,      0,      0,       0,               0,               0,                 0,                  0,                  0,                  #5-6
                       0,     0,     0,      0,     0,     n_s_i2, 0,     0,      0,      0,       0,               0,               0,                 0,                  0,                  0,                  #6-7
                       0,     0,     0,      0,     0,     0,     n_s_i2, 0,      0,      0,       0,               0,               0,                 0,                  0,                  0,                  #7-8
                       0,     0,     0,      0,     0,     0,     0,     n_s_s2,  0,      0,       0,               0,               0,                 0,                  0,                  0,                  #8-9
                       0,     0,     0,      0,     0,     0,     0,     0,      n_s_s2,  0,       0,               0,               0,                 0,                  0,                  0,                  #9-10
                       0,     0,     0,      0,     0,     0,     0,     0,      0,      n_s_s2,   0,               0,               0,                 0,                  0,                  0,                  #10
                       0,     0,     0,      0,     0,     0,     0,     0,      0,      0,       n_s_a2*n_s_a2,     n_s_a2*n_s_a2,     n_s_a2*n_s_a2*w2,     n_s_a2*n_s_a2*w2,      n_s_a2*n_s_a2*w2,      n_s_a2*n_s_a2*w42,     #nw
                       0,     0,     0,      0,     0,     0,     0,     0,      0,      0,       n_s_a2*(1-n_s_a2), n_s_a2*(1-n_s_a2), n_s_a2*(1-n_s_a2)*w2, n_s_a2*(1-n_s_a2)*w2,  n_s_a2*(1-n_s_a2)*w2,  n_s_a2*(1-n_s_a2)*w42, #w1
                       0,     0,     0,      0,     0,     0,     0,     0,      0,      0,       0,               0,               n_s_a2*(1-w2),       0,                  0,                  0,                  #w2 
                       0,     0,     0,      0,     0,     0,     0,     0,      0,      0,       0,               0,               0,                 n_s_a2*(1-w2),        0,                  0,                  #w3 
                       0,     0,     0,      0,     0,     0,     0,     0,      0,      0,       0,               0,               0,                 0,                  n_s_a2*(1-w2),        0),                 #w4 
                      nrow = 16, byrow = TRUE) 
  
  #extract eigenvalues and lambda from the matrix 
  n_m_eig2 <- eigen(n_m2)
  lambda_n2 <- as.numeric(n_m_eig2$values[1])
  output2[i,j] <- lambda_n2 # this matches the ideal population growth rate in the table
  output2 <- as.data.frame(output2)
  output2$mitigation_efficacy <- mit_eff
  output2$bycatch_reduction <- bycatch_reductions2
    
 full_output[[k]] <- output2

}
}
}

full_output[[1]]

# sim_results <- full_output[[1]]
# sim_results$quantile5 <-  apply(sim_results[1:100], 1, quantile, (probs =c(0.05)))
# sim_results$quantile95 <-  apply(sim_results[1:100], 1, quantile, (probs =c(0.95)))
```

```{r}
mitigation_eff_range #0.5 0.6 0.7 0.8 0.9 1.0

mit05_l <- full_output[[1]] %>% 
  pivot_longer(cols = starts_with("V"), names_to = "simrun", values_to = "lambda")
mit06_l <- full_output[[2]] %>% 
  pivot_longer(cols = starts_with("V"), names_to = "simrun", values_to = "lambda")
mit07_l <- full_output[[3]] %>% 
  pivot_longer(cols = starts_with("V"), names_to = "simrun", values_to = "lambda")
mit08_l <- full_output[[4]] %>% 
  pivot_longer(cols = starts_with("V"), names_to = "simrun", values_to = "lambda")
mit09_l <- full_output[[5]] %>% 
  pivot_longer(cols = starts_with("V"), names_to = "simrun", values_to = "lambda")
mit1_l <- full_output[[6]] %>% 
  pivot_longer(cols = starts_with("V"), names_to = "simrun", values_to = "lambda")

fulldat <- rbind(mit05_l,mit06_l,mit07_l,mit08_l,mit09_l,mit1_l)
fulldat$mitsim <- paste(fulldat$mitigation_efficacy, fulldat$simrun)

mit05 <- full_output[[1]] 
mit06 <- full_output[[2]] 
mit07 <- full_output[[3]] 
mit08 <- full_output[[4]] 
mit09 <- full_output[[5]] 
mit1 <- full_output[[6]] 

full_dat_w <- rbind(mit05,mit06,mit07,mit08,mit09,mit1)
full_dat_w$quantile5 <-  apply(full_dat_w[1:100], 1, quantile, (probs =c(0.05)))
full_dat_w$quantile95 <-  apply(full_dat_w[1:100], 1, quantile, (probs =c(0.95)))
full_dat_w$mitsim <- paste(full_dat_w$mitigation_efficacy, full_dat_w$simrun)

```

```{r}
ggplot(fulldat)+
  geom_line(aes(x=bycatch_reduction, y=lambda, group = simrun), color = "grey")+
   geom_hline(aes(yintercept = 1), color = "darkgrey", alpha =0.5)+ 
  # geom_vline(aes(xintercept = 0.66), color = "red", alpha =0.5)+
  #  geom_vline(aes(xintercept = 0.33), color = "red", alpha =0.5)+
  labs(x="% fisheries with bycatch mitigation", y="lambda")+
  facet_wrap(~mitigation_efficacy)+
  theme_classic()
```
```{r}
ggplot(full_dat_w)+
  geom_line(aes(x=bycatch_reduction, y=quantile5), color = "grey")+
   geom_line(aes(x=bycatch_reduction, y=quantile95), color = "grey")+
   geom_hline(aes(yintercept = 1), color = "darkgrey", alpha =0.5)+ 
  # geom_vline(aes(xintercept = 0.66), color = "red", alpha =0.5)+
  #  geom_vline(aes(xintercept = 0.33), color = "red", alpha =0.5)+
  labs(x="% fisheries with bycatch mitigation", y="lambda")+
  facet_wrap(~mitigation_efficacy)+
  theme_classic()
```


```{r}
ggplot(full_dat_w)+
  #geom_line(aes(x=bycatch_reduction, y=lambda, group = mitsim), color = "grey", alpha =0.25)+
   geom_line(aes(x=bycatch_reduction, y=quantile5, group = mitsim, color = factor(mitigation_efficacy)), alpha =0.5)+
  geom_line(aes(x=bycatch_reduction, y=quantile95, group = mitsim, color = factor(mitigation_efficacy)), alpha =0.5)+
  geom_line(data = bycatch_reductions, aes(x=reduction, y=as.numeric(lambda)))+
  geom_hline(aes(yintercept = 1), color = "darkgrey", alpha =0.5)+ 
  # geom_vline(aes(xintercept = 0.66), color = "red", alpha =0.5)+
  #  geom_vline(aes(xintercept = 0.33), color = "red", alpha =0.5)+
  labs(x="% fisheries with bycatch mitigation", y="lambda")+
  theme_classic()
```
```{r}
ggplot(fulldat)+
  #geom_line(aes(x=bycatch_reduction, y=lambda, group = mitsim), color = "grey", alpha =0.25)+
   geom_line(aes(x=bycatch_reduction, y=lambda, group = mitsim, color = factor(mitigation_efficacy)), alpha =0.5)+
  geom_line(data = bycatch_reductions, aes(x=reduction, y=as.numeric(lambda)))+
  geom_hline(aes(yintercept = 1), color = "darkgrey", alpha =0.5)+ 
  # geom_vline(aes(xintercept = 0.66), color = "red", alpha =0.5)+
  #  geom_vline(aes(xintercept = 0.33), color = "red", alpha =0.5)+
  labs(x="% fisheries with bycatch mitigation", y="lambda")+
  theme_classic()
```


```

